### 2023-11-03 paper/main.tex
Add paper draft

\documentclass[10pt]{article} % For LaTeX2e
\usepackage{tmlr}
% If accepted, instead use the following line for the camera-ready submission:
%\usepackage[accepted]{tmlr}
% To de-anonymize and remove mentions to TMLR (for example for posting to preprint servers), instead use the following:
%\usepackage[preprint]{tmlr}

% Optional math commands from https://github.com/goodfeli/dlbook_notation.
\input{math_commands.tex}

\usepackage{hyperref}
\usepackage{url}

\title{Learning associations between astronomical data and natural language with multi-modal models}

% Authors must not appear in the submitted version. They should be hidden
% as long as the tmlr package is used without the [accepted] or [preprint] options.
% Non-anonymous submissions will be rejected without review.

\author{\name Kyunghyun Cho \email kyunghyun.cho@nyu.edu \\
      \addr Department of Computer Science\\
      University of New York
      \AND
      \name Raia Hadsell \email raia@google.com \\
      \addr DeepMind
      \AND
      \name Hugo Larochelle \email hugolarochelle@google.com\\
      \addr Mila, Universit\'e de Montr\'eal \\
      Google Research\\
      CIFAR Fellow}

% The \author macro works with any number of authors. Use \AND 
% to separate the names and addresses of multiple authors.

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

\def\month{MM}  % Insert correct month for camera-ready version
\def\year{YYYY} % Insert correct year for camera-ready version
\def\openreview{\url{https://openreview.net/forum?id=XXXX}} % Insert correct link to OpenReview for camera-ready version

\begin{document}

\maketitle

\begin{abstract}
Abstract
\end{abstract}

\section{Introduction}

\subsubsection*{Broader Impact Statement}
In this optional section, TMLR encourages authors to discuss possible repercussions of their work,
notably any potential negative impact that a user of this research should be aware of. 
Authors should consult the TMLR Ethics Guidelines available on the TMLR website
for guidance on how to approach this subject.

\subsubsection*{Author Contributions}
If you'd like to, you may include a section for author contributions as is done
in many journals. This is optional and at the discretion of the authors. Only add
this information once your submission is accepted and deanonymized.

\subsubsection*{Acknowledgments}
Use unnumbered third level headings for the acknowledgments. All
acknowledgments, including those to funding agencies, go at the end of the paper.
Only add this information once your submission is accepted and deanonymized.

\bibliography{main}
\bibliographystyle{tmlr}

\appendix
\section{Appendix}
You may include other additional sections here.

\end{document}

### 2023-11-03 configs/base.py
Debug SigLIP loss; no issues found

@@ -12,7 +12,7 @@ def get_config():
     wandb.group = "proposals"
     wandb.job_type = "training"
     wandb.name = None
-    wandb.log_train = True
+    wandb.log_train = False
     wandb.workdir = "./logging/"
 
     # Text
@@ -73,16 +73,16 @@ def get_config():
     data.augment_crop = True
     data.augment_subsample_text = True
     data.max_length_words = 77
-    data.tfrecords_dir = "tfrecords_v1"
-    data.shuffle_within_batch = True
+    data.tfrecords_dir = "tfrecords_v2"
+    data.shuffle_within_batch = False
 
     # Training
     config.training = training = ml_collections.ConfigDict()
     training.half_precision = False
     training.train_fraction = 0.95
-    training.batch_size = 32  # Must be divisible by number of devices; this is the total batch size, not per-device
+    training.batch_size = 16  # Must be divisible by number of devices; this is the total batch size, not per-device
     training.n_train_steps = 101_000
-    training.warmup_steps = 5_000
+    training.warmup_steps = 1_000
     training.log_every_steps = 100
     training.eval_every_steps = 200
     training.save_every_steps = 20_000
@@ -90,7 +90,7 @@ def get_config():
     # Optimizer (AdamW)
     config.optim = optim = ml_collections.ConfigDict()
     optim.learning_rate = 1e-4
-    optim.weight_decay = 1e-4
+    optim.weight_decay = 1e-3
 
     # Seed
     config.seed = 42

### 2023-11-03 models/losses.py

@@ -7,9 +7,9 @@ jax.config.update("jax_enable_x64", True)
 def mini_batch_sigmoid_loss(text_embeds, image_embeds, logit_scale, logit_bias, negative_samples):
     """Positive samples are on the diagonal"""
     bs = text_embeds.shape[0]
-    if negative_samples:
+    if negative_samples:  # All negative samples 
         labels = -np.ones((bs, bs))
-    else:
+    else:  # Positive samples are on the diagonal
         labels = 2 * np.eye(bs) - np.ones((bs, bs))
     logits = np.matmul(text_embeds, image_embeds.T) * logit_scale + logit_bias
     
@@ -18,12 +18,17 @@ def mini_batch_sigmoid_loss(text_embeds, image_embeds, logit_scale, logit_bias, 
     return -np.mean(jax.nn.log_sigmoid(labels * logits))
 
 def sigmoid_loss(outputs):
+    """ SigLIP loss (https://arxiv.org/abs/2303.15343); 
+        Adapted from https://github.com/borisdayma/clip-jax/blob/main/training/train.py
+    """
+
+    # Get outputs
     text_embeds = outputs["text_embeds"]
     image_embeds = outputs["image_embeds"]
     logit_scale = outputs["logit_scale"]
     logit_bias = outputs.get("logit_bias", 0.)
 
-    # Number of chunks
+    # Number of chunks (devices)
     axis_size = jax.lax.psum(1, axis_name="batch")
 
     # Calculate local device loss

### 2023-11-03 scripts/submit_train.sh

@@ -2,11 +2,11 @@
 
 #SBATCH --job-name=train
 #SBATCH --nodes=1
-#SBATCH --mem=160GB
+#SBATCH --mem=200GB
 #SBATCH --time=48:00:00
 #SBATCH --gres=gpu:4
 #SBATCH --account=iaifi_lab
-#SBATCH -p gpu_requeue
+#SBATCH -p iaifi_gpu
 
 export TF_CPP_MIN_LOG_LEVEL="2"
 

### 2023-11-21 configs/base.py
Starting to make plots

@@ -87,7 +87,8 @@ def get_config():
     training.log_every_steps = 100
     training.eval_every_steps = 200
     training.save_every_steps = 10_000
-    training.loss_type = "softmax"  # "sigmoid" or "softmax"
+    training.loss_type = "sigmoid"  # "sigmoid" or "softmax"
+    training.n_eval_batches = 10  # How many batches to use for evaluation
 
     # Optimizer (AdamW)
     config.optim = optim = ml_collections.ConfigDict()

### 2023-11-21 train.py

@@ -194,11 +194,15 @@ def train(config: ConfigDict, workdir: str = "./logging/") -> train_state.TrainS
             # Evaluate periodically
             # Evaluate before starting, hence no `and (step != 0)`
             if (step % config.training.eval_every_steps == 0) and (jax.process_index() == 0):
+
+                # Log step at which evaluating 
+                logging.info(f"Evaluating at step {step}")
+                
                 val_metrics = []
                 val_batches = iter(val_ds)
 
                 # Validate on 10 batches
-                for _ in range(10):
+                for _ in range(config.training.n_eval_batches):
                     images, captions = next(val_batches)
                     images = np.array(images)
                         

### 2023-11-21 configs/base.py
Add retrieval acc and image to text retrieval plots

@@ -74,7 +74,7 @@ def get_config():
     data.augment_subsample_text = True
     data.max_length_words = 77
     data.tfrecords_dir = "tfrecords_v3"
-    data.shuffle_within_batch = False
+    data.shuffle_within_batch = True
 
     # Training
     config.training = training = ml_collections.ConfigDict()
@@ -87,12 +87,12 @@ def get_config():
     training.log_every_steps = 100
     training.eval_every_steps = 200
     training.save_every_steps = 10_000
-    training.loss_type = "sigmoid"  # "sigmoid" or "softmax"
+    training.loss_type = "softmax"  # "sigmoid" or "softmax"
     training.n_eval_batches = 10  # How many batches to use for evaluation
 
     # Optimizer (AdamW)
     config.optim = optim = ml_collections.ConfigDict()
-    optim.learning_rate = 1e-4
+    optim.learning_rate = 1e-5
     optim.weight_decay = 1e-3
 
     # Seed

### 2023-11-21 scripts/submit_train.sh

@@ -4,9 +4,9 @@
 #SBATCH --nodes=1
 #SBATCH --mem=200GB
 #SBATCH --time=48:00:00
-#SBATCH --gres=gpu:4
+#SBATCH --gres=gpu:nvidia_a100-sxm4-40gb:4
 #SBATCH --account=iaifi_lab
-#SBATCH -p iaifi_gpu
+#SBATCH -p gpu_requeue
 
 export TF_CPP_MIN_LOG_LEVEL="2"
 

### 2023-11-21 scripts/submit_train.sh
Sync remote changes

@@ -4,7 +4,7 @@
 #SBATCH --nodes=1
 #SBATCH --mem=200GB
 #SBATCH --time=48:00:00
-#SBATCH --gres=gpu:nvidia_a100-sxm4-40gb:4
+#SBATCH --gres=gpu:nvidia_a100-sxm4-80gb:4
 #SBATCH --account=iaifi_lab
 #SBATCH -p gpu_requeue
 

### 2023-11-21 train.py

@@ -135,6 +135,9 @@ def train(config: ConfigDict, workdir: str = "./logging/") -> train_state.TrainS
     logging.info(f"Augment rotate: {config.data.augment_rotate}")
     logging.info(f"Subsample text: {config.data.augment_subsample_text}. Max length: {max_length_words} words")
 
+    if config.data.shuffle_within_batch:
+        logging.info(f"Shuffling images within batch")
+
     logging.info("Starting training...")
 
     train_metrics = []
@@ -167,7 +170,6 @@ def train(config: ConfigDict, workdir: str = "./logging/") -> train_state.TrainS
             
             # Optionally shuffle "pixel_values" within batch
             if config.data.shuffle_within_batch:
-                logging.info(f"Shuffling images within batch")
                 batch["pixel_values"] = jax.random.permutation(rng, batch["pixel_values"], axis=0)
 
             # Split batch across devices

### 2023-11-21 paper/main.tex

@@
 \documentclass[10pt]{article} % For LaTeX2e
-\usepackage{tmlr}
+\usepackage[preprint]{tmlr}
 % If accepted, instead use the following line for the camera-ready submission:
@@
 \usepackage{url}
+\usepackage{graphicx}
 
-\title{Learning associations between astronomical data and natural language with multi-modal models}
+\title{\textsc{HubbleCLIP}: Learning associations between astronomical data and natural language with multi-modal models}
 
@@
 
-\author{\name Kyunghyun Cho \email kyunghyun.cho@nyu.edu \\
-      \addr Department of Computer Science\\
-      University of New York
+\author{\name Siddharth Mishra-Sharma \email smsharma@mit.edu \\
+      \addr The NSF AI Institute for Artificial Intelligence and Fundamental Interactions\\
+      Center for Theoretical Physics, Massachusetts Institute of Technology, Cambridge, MA 02139, USA \\
+      Department of Physics, Harvard University, Cambridge, MA 02138, USA
       \AND
-      \name Raia Hadsell \email raia@google.com \\
-      \addr DeepMind
+      \name Yiding Song \email ydsong@mit.edu \\
+      \addr The NSF AI Institute for Artificial Intelligence and Fundamental Interactions\\
+      Department of Physics, Massachusetts Institute of Technology, Cambridge, MA 02139, USA \\
       \AND
-      \name Hugo Larochelle \email hugolarochelle@google.com\\
-      \addr Mila, Universit\'e de Montr\'eal \\
-      Google Research\\
-      CIFAR Fellow}
-
+      \name Jesse Thaler \email jthaler@mit.edu \\
+      \addr The NSF AI Institute for Artificial Intelligence and Fundamental Interactions\\
+      Center for Theoretical Physics, Massachusetts Institute of Technology, Cambridge, MA 02139, USA \\
+}
 % The \author macro works with any number of authors. Use \AND 
@@
 \section{Introduction}
+\label{sec:intro}
+
+\section{Dataset and Processing}
+\label{sec:dataset}
+
+\section{Methodology}
+\label{sec:methodology}
+
+\subsection{Language-Image Pre-training}
+\label{sec:pretraining}
+
+\subsection{Pre-trained CLIP Model}
+\label{sec:clip}
+
+\subsection{Fine-tuning Objectives}
+\label{sec:finetuning}
+
+\section{Results and Discussion}
+\label{sec:results}
+
+\subsection{Fine-tuned Retrieval Accuracy}
+\label{sec:retrieval_acc}
+
+\subsection{`Zero-shot' Hypothesis and Object Retrieval}
+\label{sec:zero_shot}
+
+\subsection{Text-to-Image Retrieval}
+\label{sec:tti}
+
+\cite{Hinton06}
+
+\begin{figure*}[!h]
+\includegraphics[width=0.95\textwidth]{plots/retrieval_acc.pdf}
+\caption{Retrieval accuracy}
+\label{fig:retrieval_acc}
+\end{figure*}
+
+\begin{figure*}[!h]
+\includegraphics[width=0.95\textwidth]{plots/itt.pdf}
+\caption{Retrieval accuracy}
+\label{fig:itt}
+\end{figure*}
+
+\begin{figure*}[!h]
+\includegraphics[width=0.95\textwidth]{plots/tti_base.pdf}
+\caption{Retrieval accuracy}
+\label{fig:tti}
+\end{figure*}
+
+\begin{figure*}[!h]
+\includegraphics[width=0.95\textwidth]{plots/tti.pdf}
+\caption{Retrieval accuracy}
+\label{fig:tti_base}
+\end{figure*}
 

### 2023-11-21 paper/main.tex
Text updates

@@
 
-\title{\textsc{HubbleCLIP}: Learning associations between astronomical data and natural language with multi-modal models}
+\title{\textsc{HubbleCLIP}: Learning associations between astronomical observations and natural language with multi-modal models}
 
@@
 
-\author{\name Siddharth Mishra-Sharma \email smsharma@mit.edu \\
+\author{\name Siddharth Mishra-Sharma \email \href{mailto:smsharma@mit.edu}{smsharma@mit.edu} \\
       \addr The NSF AI Institute for Artificial Intelligence and Fundamental Interactions\\
@@
       \AND
-      \name Yiding Song \email ydsong@mit.edu \\
+      \name Yiding Song \email \href{mailto:ydsong@mit.edu}{ydsong@mit.edu} \\
       \addr The NSF AI Institute for Artificial Intelligence and Fundamental Interactions\\
@@
       \AND
-      \name Jesse Thaler \email jthaler@mit.edu \\
+      \name Jesse Thaler \email \href{mailto:jthaler@mit.edu}{jthaler@mit.edu} \\
       \addr The NSF AI Institute for Artificial Intelligence and Fundamental Interactions\\

### 2023-12-23 configs/base.py
Create dataset with summarized caption; tested training; works better than raw abstracts

@@ -18,8 +18,8 @@ def get_config():
     config.clip = clip = ml_collections.ConfigDict()
     clip.dtype = "float32"
     clip.use_pretrained = True
-    # clip.pretrained_model_name = "openai/clip-vit-base-patch16" 
-    clip.pretrained_model_name = "openai/clip-vit-large-patch14"
+    clip.pretrained_model_name = "openai/clip-vit-base-patch16" 
+    # clip.pretrained_model_name = "openai/clip-vit-large-patch14"
     clip.random_init_vision = False
     clip.random_init_text = False
 
@@ -27,9 +27,10 @@ def get_config():
     config.data = data = ml_collections.ConfigDict()
     data.augment_rotate = True
     data.augment_crop = True
-    data.augment_subsample_text = True
+    data.augment_subsample_text = False
     data.max_length_words = 77
-    data.tfrecords_dir = "tfrecords_v3"
+    data.tfrecords_dir = "tfrecords_v4"
+    data.caption_type = "summary"  # "abstract" or "summary"
     data.shuffle_within_batch = False
     data.data_dir = "/n/holyscratch01/iaifi_lab/smsharma/hubble_data/"
 

### 2023-12-23 summarize.py

@@ -87,7 +87,7 @@ def download(
     df.to_csv(f"{data_dir}/{save_filename}.csv", index=False)
 
 if __name__ == "__main__":
-    parser = argparse.ArgumentParser(description="Download Data Based on Command Line Arguments")
+    parser = argparse.ArgumentParser(description="Summarize abstracts.")
     parser.add_argument("--data_dir", type=str, default="./data/", help="Directory to save the downloaded data.")
     parser.add_argument("--observations_dir", type=str, default="observations_v2", help="Observations dir within the `data_dir`.")
     parser.add_argument("--n_max_abstracts", type=int, default=9999999, help="Maximum number of abstracts to summarize.")

### 2023-12-23 train.py

@@ -18,7 +18,7 @@ import flax
 from flax.core import FrozenDict
 from flax.training import common_utils, train_state, orbax_utils
 import orbax
-from dm_pix import rotate, random_crop
+from dm_pix import rotate, random_crop, random_flip_up_down, random_flip_left_right
 import tensorflow as tf
 from tqdm import trange
 
@@ -64,17 +64,17 @@ def train(config: ConfigDict, workdir: str = "./logging/") -> train_state.TrainS
 
     # Find all TFRecord files and make datasets
     files_train = tf.io.gfile.glob(f"{config.data.data_dir}/{config.data.tfrecords_dir}/*train*.tfrecord")
-    train_ds = make_dataloader(files_train, batch_size=config.training.batch_size, seed=config.seed, split='train', shuffle=True)
+    train_ds = make_dataloader(files_train, batch_size=config.training.batch_size, seed=config.seed, split='train', caption_type=config.data.caption_type, shuffle=True)
 
     files_val = tf.io.gfile.glob(f"{config.data.data_dir}/{config.data.tfrecords_dir}/*val*.tfrecord")
-    val_ds = make_dataloader(files_val, batch_size=config.training.batch_size_val, seed=config.seed, split='val', shuffle=False)
+    val_ds = make_dataloader(files_val, batch_size=config.training.batch_size_val, seed=config.seed, split='val', caption_type=config.data.caption_type, shuffle=False)
 
     batches = iter(train_ds)
 
     logging.info("Loaded the dataset")
 
     # # Model configs
-    # # NOTE: From legacy code; not used currently, but potentially if options are fed to create custom CLIP model
+    # # NOTE: From legacy code; not used currently, but potentially useful if options are fed to create custom CLIP model
     # text_config = FrozenDict(config.text_config)
     # vision_config = FrozenDict(config.vision_config)
     # clip_config = FrozenDict(config.clip)
@@ -184,11 +184,20 @@ def train(config: ConfigDict, workdir: str = "./logging/") -> train_state.TrainS
             images, captions = next(batches)
             images = np.array(images)
 
-            # Augment images through random rotations
+            # Augment images through random rotations and flips
             if config.data.augment_rotate:
+
+                # Rotations
                 rng_aug, _ = jax.random.split(rng_aug)
                 rotation_angles = jax.random.choice(rng_aug, rot_angles_90, shape=(images.shape[0],))
[…]

### 2023-12-23 utils/dataset_utils.py

@@ -1,13 +1,16 @@
+from functools import partial
+
 import tensorflow as tf
 import jax
 import flax
 from absl import logging
 
 # Parse function
-def parse_function(example):
+def parse_function(example, caption_type="abstract"):
     features = {
         "image": tf.io.FixedLenFeature([], tf.string),
         "abstract": tf.io.FixedLenFeature([], tf.string),
+        "summary": tf.io.FixedLenFeature([], tf.string),
         "image_height": tf.io.FixedLenFeature([], tf.int64),
         "image_width": tf.io.FixedLenFeature([], tf.int64),
     }
@@ -15,7 +18,7 @@ def parse_function(example):
     parsed_features = tf.io.parse_single_example(example, features)
 
     image = parsed_features["image"]
-    caption = parsed_features["abstract"]
+    caption = parsed_features[caption_type]
     image_height = parsed_features["image_height"]
     image_width = parsed_features["image_width"]
 
@@ -31,7 +34,7 @@ def _normalize(image, abstract):
     return image, abstract
 
 
-def make_dataloader(files, batch_size, seed, split="train", shuffle=True):
+def make_dataloader(files, batch_size, seed, split="train", shuffle=True, caption_type="abstract"):
     ds = tf.data.TFRecordDataset(files)
 
     # Count total examples 
@@ -39,8 +42,8 @@ def make_dataloader(files, batch_size, seed, split="train", shuffle=True):
 
     # Log number of examples
     logging.info(f"Using {num_total} examples in dataset with split {split}")
-
-    ds = ds.map(parse_function)
+    
+    ds = ds.map(partial(parse_function, caption_type=caption_type))
     ds = ds.map(_normalize)
     ds = ds.cache()
 

### 2023-12-23 configs/base.py
Launch jobs with summarized data

@@ -21,7 +21,7 @@ def get_config():
     clip.pretrained_model_name = "openai/clip-vit-base-patch16" 
     # clip.pretrained_model_name = "openai/clip-vit-large-patch14"
     clip.random_init_vision = False
-    clip.random_init_text = False
+    clip.random_init_text = True
 
     # Data
     config.data = data = ml_collections.ConfigDict()

### 2023-12-23 scripts/submit_train.sh

@@ -7,7 +7,7 @@
 # #SBATCH --gres=gpu:nvidia_a100-sxm4-80gb:4
 #SBATCH --gres=gpu:4
 #SBATCH --account=iaifi_lab
-#SBATCH -p iaifi_gpu
+#SBATCH -p iaifi_gpu_priority
 
 export TF_CPP_MIN_LOG_LEVEL="2"
 

### 2023-12-25 paper/main.tex
Intro draft

@@
 
-\usepackage{hyperref}
 \usepackage{url}
 \usepackage{graphicx}
+\usepackage{array}
+\usepackage{booktabs}
+\usepackage{lipsum}
 
-\title{\textsc{HubbleCLIP}: Learning associations between astronomical observations and natural language with multi-modal models}
+\definecolor{xlinkcolor}{rgb}{0.7752941176470588, 0.22078431372549023, 0.2262745098039215}
+
+\usepackage[
+pdfnewwindow=true,      % links in new window
+colorlinks=true,    % false: boxed links; true: colored links
+linkcolor=xlinkcolor,     % color of internal links
+citecolor=xlinkcolor,     % color of links to bibliography
+filecolor=xlinkcolor,  % color of file links
+urlcolor=xlinkcolor,      % color of external links
+final=true,
+]{hyperref}
+
+\title{\textsc{HubbleCLIP}: Associating Astronomical Observations and Natural Language with Multi-Modal Models}
 
@@
 
+\newcommand{\datafolder}[1]{\def\thedatafolder{#1}}
+
 \begin{document}
@@
 \begin{abstract}
-Abstract
+We present a multi-modal model which associates astronomical observations imaged by the \emph{Hubble} Space Telescope (HST) with natural language.
+The model is fine-tuned from a base CLIP model using summarized proposal abstracts corresponding to HST observations.
+We show that the model embodies a meaningful joint representation between observations and text through experiments targeting observation retrieval (i.e., retrieving most relevant set of observations using natural language queries) and description retrieval (i.e., querying the astrophysical object classes and science use cases most relevant to a given observation).
+The model demonstrates the potential for using generalist rather than task-specific models for astrophysics research, in particular by leveraging text as an interface.
 \end{abstract}
@@
 
+Machine learning (ML) is starting to have a significant impact on the sciences, with astrophysics being no exception.
+Machine learning methods have demonstrated promise when applied to every part of the astrophysics pipeline, from instrument design, to data acquisition, to its analysis.
+Until recently, most applications of ML within astrophysics have focused on replacing or augmenting traditional techniques with ML counterparts in order to improve performance on specific tasks.
+
+The \emph{Foundation Model} paradigm, in contrast, seeks to develop generalist models which can be deployed on a wide range of tasks.
+The paradigm has been highly successful in domains like computer vision and natural language processing, as demonstrated by the widespread adoption of tools like CLIP, ChatGPT, Dall-E, and Stable Diffusion.
+These models are typically pre-trained on massive amounts of unlabeled data using self-supervised learning techniques, enabling them to learn powerful representations which can be optionally fine-tuned to address domain-specific tasks.
+At the heart of the paradigm lies the triumph of scale -- scaling up model size, dataset size, and compute.
+However, foundation models can often be fine-tuned using only small amounts of domain-specific data, increasing their usefulness when applied to those specific domains.
+
+There is considerable interest in developing custom foundation models for the sciences, with astrophysics being ripe for such an effort due to several reasons.
+The first is the availability of large amounts of publicly-available data as a contingency of publicy-funded data-taking efforts.
+The second is the multi-modality inherent astrophysical observations, with different types of data (e.g., images, spectra, light curves, textual descriptions) often available for each observation.
+This multi-modality was recently exploited to train \textsc{AstroCLIP} -- a joint representation between multi-band images and optical spectra from the Dark Energy Spectroscopic Instrument (DESI).
+\textsc{AstroLLaMA} is another recent effort to fine-tune a publicly-available model (Llama-2) on astrophysics-specific textual data from the arXiv.
+
+The success of the foundation model paradigm partly relies on the ability to leverage text as a \emph{universal interface}.
+In this work, we take this outlook and train a joint representation between observations taken by the \emph{Hubble} Space Telescope (HST) and natural language.
+We do so by using the associations between observation proposals and corresponding downstream observations.
+We show that fine-tuning a CLIP (Contrastive Language-Image Pre-training) model on this data enables learning meaningful joint representations.
+
+The paper is organized as follows.
+In Sec.~\ref{sec:dataset}, we describe the dataset used in this work, including its curation and processing.
+In Sec.~\ref{sec:methodology}, we describe the methodology used to train and evaluate the model.
+In Sec.~\ref{sec:results}, we present the results of our experiments on image and text retrieval tasks.
+We discuss future prospects and conclude in Sec.~\ref{sec:conclusion}.
+
 \section{Dataset and Processing}
 \label{sec:dataset}
+
+\subsection{Summarization via guided generation}
+\label{sec:summarization}
 
@@
 
+\section{Conclusion}
+\label{sec:conclusion}
+
 \cite{Hinton06}
+
+% \begin{figure*}[!h]
+% \includegraphics[width=0.95\textwidth]{example-image-a}
+% \caption{Overview of dataset}
+% \label{fig:dataset}
+% \end{figure*}
+
+\datafolder{./plots/data/} % Set your folder path here
+
+\begin{table}[h!]
+      \centering
+      \begin{tabular}{m{0.21\textwidth} p{1.9cm} p{1.9cm} m{8cm}}
+          \toprule
+          \centering \bfseries Image & \centering \bfseries Obs. cycle & \centering \bfseries Prop. ID & \centering \bfseries Summarized abstract \tabularnewline
+          \midrule
+          \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_0.pdf} & \centering \input{\thedatafolder/cycle_0.txt} & \centering \input{\thedatafolder/id_0.txt} &  {\scriptsize \input{\thedatafolder/sum_0.txt}} \tabularnewline
+          \midrule
+          \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_1.pdf} & \centering \input{\thedatafolder/cycle_1.txt} & \centering \input{\thedatafolder/id_1.txt} &  {\scriptsize \input{\thedatafolder/sum_1.txt}} \tabularnewline
+          \midrule
+          \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_2.pdf} & \centering \input{\thedatafolder/cycle_2.txt} & \centering \input{\thedatafolder/id_2.txt} &  {\scriptsize \input{\thedatafolder/sum_2.txt}} \tabularnewline
+          \midrule
+          \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_3.pdf} & \centering \input{\thedatafolder/cycle_3.txt} & \centering \input{\thedatafolder/id_3.txt} &  {\scriptsize \input{\thedatafolder/sum_3.txt}} \tabularnewline
+          \bottomrule
+      \end{tabular}
+      \caption{Data overview}
+  \end{table}
+
+\begin{figure*}[!h]
+      \includegraphics[width=0.95\textwidth]{plots/val_metrics.pdf}
+      \caption{Retrieval accuracy}
+      \label{fig:retrieval_acc}
+      \end{figure*}
 

### 2023-12-31 configs/base.py
Add capability to just fine tune an added head

@@ -22,6 +22,8 @@ def get_config():
     # clip.pretrained_model_name = "openai/clip-vit-large-patch14"
     clip.random_init_vision = False
     clip.random_init_text = False
+    clip.transfer_head = False
+    clip.d_transfer_head = 1024
 
     # Data
     config.data = data = ml_collections.ConfigDict()
@@ -53,7 +55,7 @@ def get_config():
     # Optimizer (AdamW)
     config.optim = optim = ml_collections.ConfigDict()
     optim.schedule = "cosine"
-    optim.learning_rate = 1e-6
+    optim.learning_rate = 1e-5
     optim.weight_decay = 1e-3
 
     # Seed

### 2023-12-31 train.py

@@ -17,6 +17,7 @@ import optax
 import flax
 from flax.core import FrozenDict
 from flax.training import common_utils, train_state, orbax_utils
+from flax import traverse_util
 import orbax
 from dm_pix import rotate, random_crop, random_flip_up_down, random_flip_left_right
 import tensorflow as tf
@@ -27,6 +28,7 @@ from transformers import AutoProcessor, FlaxCLIPModel
 from utils.dataset_utils import make_dataloader
 from utils.text_utils import process_truncate_captions
 from models.train_utils import param_count, train_step, eval_step, to_wandb_config
+from models.clip_transfer import FlaxCLIPModelTransfer
 
 replicate = flax.jax_utils.replicate
 unreplicate = flax.jax_utils.unreplicate
@@ -104,32 +106,44 @@ def train(config: ConfigDict, workdir: str = "./logging/") -> train_state.TrainS
             params_init = model.module.init(rng, input_ids=np.zeros((1, model.config.text_config.max_length)), 
                             attention_mask=np.zeros((1, model.config.text_config.max_length)),
                             pixel_values=np.zeros((1, model.config.vision_config.image_size, model.config.vision_config.image_size, 3)),
-                            position_ids=np.zeros((1, model.config.text_config.max_length)))
+                            position_ids=np.zeros((1, model.config.text_config.max_length)))['params']
             
             if config.clip.random_init_vision:
-                model.params['vision_model'] = params_init['params']['vision_model']
-                model.params['visual_projection'] = params_init['params']['visual_projection']
+                model.params['vision_model'] = params_init['vision_model']
+                model.params['visual_projection'] = params_init['visual_projection']
 
                 logging.info("Randomly initialized vision model")
             
             if config.clip.random_init_text:
-                model.params['text_model'] = params_init['params']['text_model']
-                model.params['text_projection'] = params_init['params']['text_projection']
+                model.params['text_model'] = params_init['text_model']
+                model.params['text_projection'] = params_init['text_projection']
 
                 logging.info("Randomly initialized text model")
 
-        params = FrozenDict(model.params)
         logging.info(f"Loaded pretrained model {config.clip.pretrained_model_name}")
 
     else:
         raise NotImplementedError
 
-    logging.info(f"Number of parameters: {param_count(params)}")
[…]

### 2023-12-31 configs/base.py
Added continuous rotations option; made evals deterministic; by default no crops

@@ -28,7 +28,8 @@ def get_config():
     # Data
     config.data = data = ml_collections.ConfigDict()
     data.augment_rotate = True
-    data.augment_crop = True
+    data.augment_rotate_type = "discrete"  # "continuous" rotation angles or "discrete" 90-deg flips
+    data.augment_crop = False
     data.augment_subsample_text = False
     data.max_length_words = 77
     data.tfrecords_dir = "tfrecords_v4"

### 2023-12-31 train.py

@@ -190,7 +190,7 @@ def train(config: ConfigDict, workdir: str = "./logging/") -> train_state.TrainS
 
     # Log info about augmentations
     logging.info(f"Augment crop: {config.data.augment_crop}")
-    logging.info(f"Augment rotate: {config.data.augment_rotate}")
+    logging.info(f"Augment rotate: {config.data.augment_rotate}, {config.data.augment_rotate_type}")
     logging.info(f"Subsample text: {config.data.augment_subsample_text}. Max length: {max_length_words} words")
 
     if config.data.shuffle_within_batch:
@@ -210,7 +210,12 @@ def train(config: ConfigDict, workdir: str = "./logging/") -> train_state.TrainS
 
                 # Rotations
                 rng_aug, _ = jax.random.split(rng_aug)
-                rotation_angles = jax.random.choice(rng_aug, rot_angles_90, shape=(images.shape[0],))
+                if config.data.augment_rotate_type == "continuous":
+                    rotation_angles = jax.random.choice(rng_aug, rot_angles_90, shape=(images.shape[0],))  # Angles in radians
+                elif config.data.augment_rotate_type == "discrete":
+                    rotation_angles = jax.random.uniform(rng_aug, shape=(images.shape[0],)) * 2 * np.pi  # Angles in radians
+                else:
+                    raise ValueError(f"Invalid augment_rotate_type: {config.data.augment_rotate_type}")
                 images = jax.vmap(partial(rotate, mode='constant', cval=1.))(images, rotation_angles)
 
                 # Flips
@@ -226,13 +231,10 @@ def train(config: ConfigDict, workdir: str = "./logging/") -> train_state.TrainS
                 rng_aug, _ = jax.random.split(rng_aug)
                 images = jax.vmap(random_crop, in_axes=(None,0,None))(rng_aug, images, (model.config.vision_config.image_size, model.config.vision_config.image_size, 3))
 
-            if config.clip.use_pretrained:
-                # NOTE: Image arrays should be ints in the range [0, 255]
-                captions = process_truncate_captions(captions, rng_aug, max_length_words=max_length_words)
-                inputs = processor(text=captions, images=images * 255.,  return_tensors="np", padding="max_length", truncation=True, max_length=model.config.text_config.max_length)
-                batch = inputs.data
-            else:
-                raise NotImplementedError
+            # NOTE: Image arrays should be ints in the range [0, 255] here
+            captions = process_truncate_captions(captions, rng_aug, max_length_words=max_length_words)
[…]

### 2023-12-31 utils/summarize_utils.py

@@ -10,6 +10,7 @@ class ConstrainedResponseHST(BaseModel):
 
 @outlines.prompt
 def prompt_fn(abstract):
+     # TODO: Add start and end tags as options
      """<|im_start|>system
 You are an expert astrophysicist, with broad expertise across observational and theoretical astrophysics.
 You are able to extract core information from astrophysical texts.

### 2023-12-31 scripts/submit_train.sh
Command line config specification

@@ -27,4 +27,4 @@ alias python=/n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python
 alias jupyter=/n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/jupyter
 
 cd /n/holystore01/LABS/iaifi_lab/Users/smsharma/multimodal-data/
-/n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py
+/n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py # --config.data.augment_rotate_type="continuous"

### 2023-12-31 summarize.py
Prompt engineering; edits summarize to work with mixtral

@@ -1,30 +1,64 @@
 import sys, os
 sys.path.append("../")
 
+import outlines
 import outlines.models as models
 import outlines.text as text
 
 import argparse
+from functools import partial
 from tqdm import tqdm
 import pandas as pd
 
+import transformers
+from transformers import BitsAndBytesConfig
+import torch
 from utils.summarize_utils import prompt_fn, ConstrainedResponseHST
 from utils.abstract_utils import read_abstracts_file
 
-def download(
+def summarize(
     data_dir: str = "./data/",
     observations_dir: str = "observations_v1",
+    i_start: int = 0,
     n_max_abstracts: int = 9999999,
     abstract_filename: str = "abstracts.cat",
-    model_name_awq: str = "TheBloke/OpenHermes-2.5-Mistral-7B-AWQ",
+    model_name: str = "mistralai/Mixtral-8x7B-Instruct-v0.1",  # "TheBloke/OpenHermes-2.5-Mistral-7B-AWQ",
     save_filename: str = "summary_v1",
     seed: int = 42,
     verbose: bool = False,
     n_max_tries: int = 5,
 ):
     
-    # Load the model
-    model = models.awq(model_name_awq)
+    if model_name == "mistralai/Mixtral-8x7B-Instruct-v0.1":
+
+        config = transformers.AutoConfig.from_pretrained(
+            model_name, trust_remote_code=True, asd=True,
+        )
+
+        bnb_config = BitsAndBytesConfig(
+            load_in_4bit=True,
+            bnb_4bit_use_double_quant=True,
+            bnb_4bit_quant_type="nf4",
+            bnb_4bit_compute_dtype=torch.bfloat16
+        )
+
+        model = models.transformers(
+            
+            model_name=model_name,
+            model_kwargs={
+                "config": config,
+                "quantization_config": bnb_config,
+                "trust_remote_code": True,
+                "device_map": "auto",
+                "load_in_4bit": True,
+                "cache_dir": "/n/holystore01/LABS/iaifi_lab/Users/smsharma/hf_cache/"
+            },
+        )
+        
+    elif model_name == "TheBloke/OpenHermes-2.5-Mistral-7B-AWQ":
+        model = models.awq(model_name)  
+    else:
+        raise ValueError("Invalid model name.")
 
     # Load the abstracts
     abstract_file = f"{data_dir}/{abstract_filename}"
@@ -50,7 +84,7 @@ def download(
[…]

### 2023-12-31 utils/summarize_utils.py

@@ -1,40 +1,56 @@
 import outlines
 import outlines.models as models
 
-from pydantic import BaseModel, Field, constr, conlist
+from pydantic import BaseModel, Field, constr, conlist, validator
 from enum import Enum
 
+import re
+
 class ConstrainedResponseHST(BaseModel):
+    # objects_and_phenomena: conlist(str, min_length=1, max_length=5)
+    # science_use_cases: conlist(str, min_length=1, max_length=5)
     objects_and_phenomena: str
     science_use_cases: str
 
+    # @validator('*', pre=True, each_item=True)
+    # def clean_and_validate_strings(cls, v):
+    #     if isinstance(v, str):
+    #         # Strip leading unwanted characters
+    #         v = v.lstrip(':, ')
+
+    #         # # Regular expression to allow alphanumeric, spaces, and specific symbols
+    #         # allowed = re.compile(r"^[a-zA-Z0-9\s\.\-\+\(\)\*\/]*$")
+    #         # if not allowed.match(v) or not v.strip():
+    #         #     raise ValueError('Invalid string. Allowed: alphanumeric, spaces, ".", "-", "+", "(", ")", "*", "/"')
+
+    #     return v
+
 @outlines.prompt
 def prompt_fn(abstract):
-     # TODO: Add start and end tags as options
-     """<|im_start|>system
-You are an expert astrophysicist, with broad expertise across observational and theoretical astrophysics.
-You are able to extract core information from astrophysical texts.
+        
+    """<s>[INST] You are an expert astrophysicist, with broad expertise across observational and theoretical astrophysics. You are able to extract core information from astrophysical texts.
 
-<|im_start|>user
 Abstract: "{{abstract}}"
 
-The above is an abstract for a proposed observation taken by the Hubble Space Telescope.
-
-Your will summarize the nature of eventual observation taken by the Telescope, mentioning the astrophysical objects imaged and possible downstream science use cases described in the abstract.
+Based on the above observational proposal abstract, your task is to summarize the nature of the eventual observations. You will identify the astrophysical objects and phenomena, as well as the potential science use cases described in the abstract.
 
 Follow these instructions exactly:
-- You will never mention the Hubble Space Telescope, HST, or the HST archive.
-- You will only write text and numbers.
-- You will not return empty values.
-- You will only write in English.
-- You will mention the class (e.g., barred spiral galaxy) and not just the specific instance (e.g., Andromeda).
[…]

### 2024-01-01 scripts/submit_summarize.sh
Submit summarize with mixtral

@@ -1,12 +1,12 @@
 #!/bin/bash
 
-#SBATCH --job-name=summarize
+#SBATCH --job-name=summarize_mixtral
 #SBATCH --nodes=1
 #SBATCH --mem=80GB
 #SBATCH --time=48:00:00
 #SBATCH --gres=gpu:4
 #SBATCH --account=iaifi_lab
-#SBATCH -p iaifi_gpu
+#SBATCH -p iaifi_gpu_priority
 
 export TF_CPP_MIN_LOG_LEVEL="2"
 
@@ -26,4 +26,4 @@ alias python=/n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python
 alias jupyter=/n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/jupyter
 
 cd /n/holystore01/LABS/iaifi_lab/Users/smsharma/multimodal-data/
-/n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python summarize.py --data_dir /n/holystore01/LABS/iaifi_lab/Users/smsharma/multimodal-data/data --observations_dir observations_v1 --save_filename summary_v1
+/n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python summarize.py --data_dir /n/holystore01/LABS/iaifi_lab/Users/smsharma/multimodal-data/data --observations_dir observations_v1 --save_filename summary_v2 --verbose

### 2024-01-01 summarize.py
New summaries

@@ -83,6 +83,8 @@ def summarize(
                                for d in dirs
                                if d.startswith("proposal_") and not d.endswith('.ipynb_checkpoints')]
 
+    generator = outlines.generate.json(model, ConstrainedResponseHST)
+    
     # Walk through data folder
     for directory in tqdm(directories_with_images[i_start:i_start + n_max_abstracts]):
         proposal_id = directory.split("proposal_")[-1]  # Extract proposal id from the directory name
@@ -92,8 +94,7 @@ def summarize(
 
         # Generate summary with constrained response
         prompt = prompt_fn(abstract)
-        generator = outlines.generate.json(model, ConstrainedResponseHST)
-
+        
         # Try up to n_max_tries times; if it fails, skip and go to the next abstract
         for _ in range(n_max_tries):
             result = None

### 2024-01-01 paper/main.tex
Text edits

@@
 \usepackage{lipsum}
+\usepackage{xspace}
+\usepackage{pdflscape} 
+\usepackage{multirow} 
+\usepackage{listings}
+
+% Custom colors
+\usepackage{color}
+\definecolor{deepblue}{rgb}{0,0,0.5}
+\definecolor{deepred}{rgb}{0.6,0,0}
+\definecolor{deepgreen}{rgb}{0,0.5,0}
 
 \definecolor{xlinkcolor}{rgb}{0.7752941176470588, 0.22078431372549023, 0.2262745098039215}
+\definecolor{codegreen}{rgb}{0,0.6,0}
+\definecolor{codegray}{rgb}{0.5,0.5,0.5}
+\definecolor{codepurple}{rgb}{0.58,0,0.82}
+\definecolor{backcolour}{rgb}{0.95,0.95,0.92}
+
+\newcommand{\hubble}{\emph{Hubble}\xspace}
 
@@
 ]{hyperref}
+
+\lstdefinestyle{mystyle}{
+  backgroundcolor=\color{backcolour},
+  commentstyle=\color{codegreen},
+  keywordstyle=\color{magenta},
+  numberstyle=\tiny\color{codegray},
+  stringstyle=\color{codegreen},
+  basicstyle=\ttfamily\footnotesize,
+  breakatwhitespace=false,
+  breaklines=true,
+  captionpos=b,
+  keepspaces=true,
+  numbers=left,
+  numbersep=5pt,
+  showspaces=false,
+  showstringspaces=false,
+  showtabs=false,
+  frame=single, % adds a frame around the code
+  rulecolor=\color{black}, % if you want to change the frame color
+  tabsize=2
+}
+
+\lstset{style=mystyle}
 
@@
 \begin{abstract}
-We present a multi-modal model which associates astronomical observations imaged by the \emph{Hubble} Space Telescope (HST) with natural language.
-The model is fine-tuned from a base CLIP model using summarized proposal abstracts corresponding to HST observations.
-We show that the model embodies a meaningful joint representation between observations and text through experiments targeting observation retrieval (i.e., retrieving most relevant set of observations using natural language queries) and description retrieval (i.e., querying the astrophysical object classes and science use cases most relevant to a given observation).
+We present a multi-modal model which associates astronomical observations imaged by the \hubble Space Telescope (HST) with natural language.
+The model is fine-tuned from a base CLIP model using successful proposal abstracts corresponding to HST observations, summarized via guided large language (LLM) model generation.
+We show that the model embodies a meaningful joint representation between observations and text through experiments targeting observation retrieval (i.e., retrieving the most relevant observations from a set using natural language queries) and description retrieval (i.e., querying the astrophysical object classes and science use cases most relevant to a given observation).
 The model demonstrates the potential for using generalist rather than task-specific models for astrophysics research, in particular by leveraging text as an interface.
@@
 At the heart of the paradigm lies the triumph of scale -- scaling up model size, dataset size, and compute.
-However, foundation models can often be fine-tuned using only small amounts of domain-specific data, increasing their usefulness when applied to those specific domains.
+However, foundation models often benefit from fine tuning using a relatively small amounts of domain-specific data, increasing their usefulness when applied to those specific domains.
 
@@
 The second is the multi-modality inherent astrophysical observations, with different types of data (e.g., images, spectra, light curves, textual descriptions) often available for each observation.
-This multi-modality was recently exploited to train \textsc{AstroCLIP} -- a joint representation between multi-band images and optical spectra from the Dark Energy Spectroscopic Instrument (DESI).
-\textsc{AstroLLaMA} is another recent effort to fine-tune a publicly-available model (Llama-2) on astrophysics-specific textual data from the arXiv.
+This multi-modality was recently exploited to train \textsc{AstroCLIP}~\citep{lanusse2023astroclip} -- a joint representation between multi-band images and optical spectra from the Dark Energy Spectroscopic Instrument (DESI).
+\textsc{AstroLLaMA}~\citep{nguyen2023astrollama} is another recent effort to fine-tune a publicly-available model (\textsc{Llama-2}) on astrophysics-specific textual data from the arXiv.
 
-The success of the foundation model paradigm partly relies on the ability to leverage text as a \emph{universal interface}.
-In this work, we take this outlook and train a joint representation between observations taken by the \emph{Hubble} Space Telescope (HST) and natural language.
+The CLIP~\citep{radford2021learning} family of models has shown strong performance on a variety of downstream tasks including zero-shot classification and image retrieval.
+
+\textsc{GeoCLIP}~\citep{cepeda2023geoclip}.
+
+The success of the foundation model paradigm partly relies on the ability to flexibly leverage text as a \emph{universal interface}.
+In this work, we take this outlook and train a joint representation between observations taken by the \hubble Space Telescope (HST) and natural language.
 We do so by using the associations between observation proposals and corresponding downstream observations.
@@
 
-\section{Dataset and Processing}
+\section{Dataset Construction}
 \label{sec:dataset}
 
-\subsection{Summarization via guided generation}
+We curate a dataset of images of \hubble observations and corresponding text descriptions.
+We rely on summarized versions of proposal abstracts from the Proposal Abstracts Catalog\footnote{\url{https://archive.stsci.edu/hst/proposal_abstracts.html}} -- a catalog of all accepted \hubble proposals -- as captions for the observations.
+The HST has been operational for 33 years, having been launched on April 24, 1990, and we use proposals and observations up to Cycle 31.
+
+Examples of images and corresponding captions are shown in Tab.~\ref{tab:dataset}.
+It can be seen that these images have specific characteristics as well as artifacts particular to the nature of data-taking which distinguish them from the distribution of natural images typically used for large-scale pre-training of foundation models.
+This further motivates the need for fine-tuning on domain-specific data.
+
+\subsection{Data Selection and Pre-Processing}
+
+Observations corresponding to individual proposal IDs are queried through the Mikulski Archive for Space Telescopes (MAST) via the \texttt{astroquery} interface.
+Products of type \texttt{PREVIEW} are filtered in, corresponding to preview postcard images.
+Note that these are not science-grade observations, but rather low-resolution images used for quick-look purposes; given the nature of associations we aim to learn, we deem this adequate for our current purposes.
+A maximum of 20 images are downloaded per proposal ID, selected at random, in order to avoid biasing the model towards proposals with a larger number of observations.
+Images are resized to a maximum resolution per side of 512 pixels before saving.
+Color previews (i.e., observations taken with multiple wavelength filters assigned to individual RGB channels) are manually excluded via a filename filter in order to maintain consistency across the dataset; models trained on datasets with color images included were observed to show worse performance on generalization metrics.
+If no appropriate images corresponding to an abstract are found, the abstract is excluded from the dataset.
+
+In total 31,859 images corresponding to 4,438 abstracts are included in the fine-tuning dataset.
+3,194 images are held out for validation, with no abstract being common between training and validation sets in order to ensure an independent set of text-image pairs for testing.
+
+\subsection{Summarization via Guided Generation}
 \label{sec:summarization}
+
+Raw abstracts summarize the corresponding successful HST observing proposal, which intends to make the case for allocating \hubble telescope time towards a particular set of observations.
+These abstracts are written in a diversity of styles, formats, and lengths, being highly variable in the nature of content as well.
+Although the abstracts can be used as-is as image captions, we explore the use of summarization via guided LLM generation to standardize the captions used for fine-tuning the CLIP model.
+The goal is to summarize the objects and phenomena, as well as potential downstream science use cases corresponding to the HST observations in order to increase the signal between text and images.
+
+The method from \cite{willard2023efficient} is used to produce an LLM-generated summary of the abstract conforming to a particular schema, specified in JSON format.
+The schema is designed to represent a list of the objects (e.g., `Type Ia supernova') and phenomena (e.g., `gravitational lensing'), as well as potential downstream science uses cases (e.g., `set constraints on supernova explosion models') that could correspond to the eventual imaged observation given the abstract text.
+
+The procedure guides the generation of LLM outputs while ensuring that the schema is respected at every point in the generation by masking out tokens that would violate the intended format.
+By framing the problem in terms of transitions between a set of finite-state machines, \cite{willard2023efficient} showed that guided generation can be performed with negligible overhead compared to unconstrained generation.
+This ensures that the output of the LLM strictly conforms to the format of the following example:
+\begin{lstlisting}[language=Python]
+{
+  'objects_and_phenomena': ['star forming galaxy', 'lensed galaxy', ...], 
+  'science_use_cases': ['measure lensing magnification', 'probe spectral energy distributions', ...]
+}
+\end{lstlisting}
+which is then used to construct the summarized caption by combining the two key elements.
+Examples of raw abstracts and corresponding LLM-generated summaries are shown in Tab.~\ref{tab:datasetsumm}.
+Further details on the summarization procedure, including the prompts used and a more detailed description of guided generation, are provided in App.~\ref{app:summarization}.
+
+The open-weights, instruction-tuned model \textsc{Mixtral-8x7B}\footnote{\url{https://huggingface.co/mistralai/Mixtral-8x7B-Instruct-v0.1}} is used to generate the summaries, with guided generation performed using the \texttt{outlines}\footnote{\url{https://github.com/outlines-dev/outlines}} package.
+
+The goal if summarization-via-guided-generation is to increase the signal between text and images by standardizing the captions used for fine-tuning the CLIP model.
+
+\datafolder{./plots/data/}
+
+\begin{table}[h!]
+      \centering
+      \begin{tabular}{m{0.21\textwidth} p{1.9cm} p{1.9cm} m{8cm}}
+          \toprule
+          \centering \bfseries Image & \centering \bfseries Obs. cycle & \centering \bfseries Prop. ID & \centering \bfseries LLM-extracted summary \tabularnewline
+          \midrule
+          \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_0.pdf} & \centering \input{\thedatafolder/cycle_0.txt} & \centering \input{\thedatafolder/id_0.txt} &  {\scriptsize \input{\thedatafolder/sum_0.txt}} \tabularnewline
+          \midrule
+          \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_1.pdf} & \centering \input{\thedatafolder/cycle_1.txt} & \centering \input{\thedatafolder/id_1.txt} &  {\scriptsize \input{\thedatafolder/sum_1.txt}} \tabularnewline
+          \midrule
+          \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_2.pdf} & \centering \input{\thedatafolder/cycle_2.txt} & \centering \input{\thedatafolder/id_2.txt} &  {\scriptsize \input{\thedatafolder/sum_2.txt}} \tabularnewline
+          \midrule
+          \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_3.pdf} & \centering \input{\thedatafolder/cycle_3.txt} & \centering \input{\thedatafolder/id_3.txt} &  {\scriptsize \input{\thedatafolder/sum_3.txt}} \tabularnewline
+          \bottomrule
+      \end{tabular}
+      \caption{Examples of images and corresponding captions, constructed using the LLM-extracted summaries. The CLIP model is fine-tuned on these text-image associations.}
+      \label{tab:dataset}
+  \end{table}
+
+\begin{landscape}
+      \begin{table}[h!]
+          \centering
+          \begin{tabular}{m{1.8cm} m{8cm} m{5cm} m{6cm}}
+              \toprule
+              \bfseries Prop. ID & \centering\arraybackslash \bfseries Proposal abstract & \multicolumn{2}{c}{\bfseries LLM-extracted summary} \tabularnewline
+              \cmidrule(r){3-4}
+              & & \centering\arraybackslash \bfseries Objects and phenomena & \centering\arraybackslash \bfseries Science use cases \tabularnewline
+              \midrule
+              \input{\thedatafolder/id1_0.txt} & {\scriptsize \input{\thedatafolder/abs1_0.txt}} & {\scriptsize \input{\thedatafolder/obj1_0.txt}} & {\scriptsize \input{\thedatafolder/sci1_0.txt}} \tabularnewline
+              \midrule
+              \input{\thedatafolder/id1_1.txt} & {\scriptsize \input{\thedatafolder/abs1_1.txt}} & {\scriptsize \input{\thedatafolder/obj1_1.txt}} & {\scriptsize \input{\thedatafolder/sci1_1.txt}} \tabularnewline
+              \midrule
+              \input{\thedatafolder/id1_2.txt} & {\scriptsize \input{\thedatafolder/abs1_2.txt}} & {\scriptsize \input{\thedatafolder/obj1_2.txt}} & {\scriptsize \input{\thedatafolder/sci1_2.txt}} \tabularnewline
+              \midrule
+              \input{\thedatafolder/id1_3.txt} & {\scriptsize \input{\thedatafolder/abs1_3.txt}} & {\scriptsize \input{\thedatafolder/obj1_3.txt}} & {\scriptsize \input{\thedatafolder/sci1_3.txt}} \tabularnewline
+              \bottomrule
+          \end{tabular}
+          \caption{Examples of the beginning of raw proposal abstracts (second column) and LLM (\textsc{Mixtral-8x7B})-extracted summaries (rightmost two columns), separately extracting objects and phenomena as well as potential downstream science use cases.
+The LLM-extracted summaries are used for associating text with observations.}
+          \label{tab:datasetsumm}
+      \end{table}
+  \end{landscape}
 
@@
 
+Our goal is to learn a semantically  meaningful joint representation of HST image observations and natural language, with the intention of using it for a variety of downstream tasks.
+We leverage the strong generalization capabilities demonstrated by CLIP (Contrastive Language-Image Pretraining) and adapt these to work with domain-specific \hubble data via fine-tuning; we describe these below.
+
 \subsection{Language-Image Pre-training}
@@
 
-\subsection{Pre-trained CLIP Model}
-\label{sec:clip}
+CLIP \citep[Contrastive Language-Image Pretraining;][]{radford2021learning} is a multi-modal model pre-trained on a large corpus of image-text pairs via weak supervision using a contrastive loss.
+Given a minibatch $\mathcal{B}$ of $|\mathcal{B}|$ image-text pairs $\{(I_i, T_i)\}$, the goal is to align the learned representations of corresponding (positive) pairs $(I_i, T_i)$ while repelling the representations of unaligned (negative) pairs $(I_i, T_{j\neq i})$.
+Image and text encoders $f(\cdot)$ and $g(\cdot)$ are used to map images and text to a common embedding space, and the standard softmax-based InfoNCE~\citep{oord2018representation} contrastive loss function is given by
+\begin{equation}
+\mathcal L(\mathcal B) = -\frac{1}{2|\mathcal{B}|} \sum_{i=1}^{|\mathcal{B}|}\left(\overbrace{\log \frac{e^{t {x}_i \cdot {y}_i}}{\sum_{j=1}^{|\mathcal{B}|} e^{t {x}_i \cdot {y}_j}}}^{\text {image } \rightarrow \text { text softmax }}+\overbrace{\log \frac{e^{t {x}_i \cdot {y}_i}}{\sum_{j=1}^{|\mathcal{B}|} e^{t {x}_j \cdot {y}_i}}}^{\text {text } \rightarrow \text { image softmax }}\right)
+\end{equation}
+where ${x}_i=\frac{f\left(I_i\right)}{\left\|f\left(I_i\right)\right\|_2}$ and ${y}_i=\frac{g\left(T_i\right)}{\left\|g\left(T_i\right)\right\|_2}$ are the normalized representations of the $i$-th image and text, respectively, and $t$ is a (typically learnable) temperature hyperparameter.
 
-\subsection{Fine-tuning Objectives}
+As the base model, we use the CLIP-ViT-B/16 \citep{radford2021learning} variant trained by OpenAI.
+This model uses a (12-layer, 12-head, 768-embedding dimension) vision transformer as the image encoder and a (12-layer, 8-head, 512-embedding dimension) sequence transformer as the text backbone.
+Linear projection layers map the outputs of the image and text encoders to a common embedding space of dimension 512.
+The model was trained on 400 million image-text pairs from the internet.
+% , and the CLIP family of models has shown strong performance on a variety of downstream tasks including zero-shot classification and image retrieval.
+
+\subsection{Fine-tuning Procedure}
 \label{sec:finetuning}
@@
 \label{sec:results}
+
+% Effect of using summarization
+% Effect of using a base model vs fine-tuning from scratch
 
@@
 
-\section{Conclusion}
+\section{Outlook and Conclusions}
 \label{sec:conclusion}
-
-\cite{Hinton06}
 
@@
 % \end{figure*}
-
-\datafolder{./plots/data/} % Set your folder path here
-
-\begin{table}[h!]
-      \centering
-      \begin{tabular}{m{0.21\textwidth} p{1.9cm} p{1.9cm} m{8cm}}
-          \toprule
-          \centering \bfseries Image & \centering \bfseries Obs. cycle & \centering \bfseries Prop. ID & \centering \bfseries Summarized abstract \tabularnewline
-          \midrule
-          \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_0.pdf} & \centering \input{\thedatafolder/cycle_0.txt} & \centering \input{\thedatafolder/id_0.txt} &  {\scriptsize \input{\thedatafolder/sum_0.txt}} \tabularnewline
-          \midrule
-          \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_1.pdf} & \centering \input{\thedatafolder/cycle_1.txt} & \centering \input{\thedatafolder/id_1.txt} &  {\scriptsize \input{\thedatafolder/sum_1.txt}} \tabularnewline
-          \midrule
-          \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_2.pdf} & \centering \input{\thedatafolder/cycle_2.txt} & \centering \input{\thedatafolder/id_2.txt} &  {\scriptsize \input{\thedatafolder/sum_2.txt}} \tabularnewline
-          \midrule
-          \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_3.pdf} & \centering \input{\thedatafolder/cycle_3.txt} & \centering \input{\thedatafolder/id_3.txt} &  {\scriptsize \input{\thedatafolder/sum_3.txt}} \tabularnewline
-          \bottomrule
-      \end{tabular}
-      \caption{Data overview}
-  \end{table}
 
@@
 
-\subsubsection*{Broader Impact Statement}
-In this optional section, TMLR encourages authors to discuss possible repercussions of their work,
-notably any potential negative impact that a user of this research should be aware of. 
-Authors should consult the TMLR Ethics Guidelines available on the TMLR website
-for guidance on how to approach this subject.
+% \subsubsection*{Broader Impact Statement}
+% In this optional section, TMLR encourages authors to discuss possible repercussions of their work,
+% notably any potential negative impact that a user of this research should be aware of. 
+% Authors should consult the TMLR Ethics Guidelines available on the TMLR website
+% for guidance on how to approach this subject.
 
-\subsubsection*{Author Contributions}
-If you'd like to, you may include a section for author contributions as is done
-in many journals. This is optional and at the discretion of the authors. Only add
-this information once your submission is accepted and deanonymized.
+% \subsubsection*{Author Contributions}
+% If you'd like to, you may include a section for author contributions as is done
+% in many journals. This is optional and at the discretion of the authors. Only add
+% this information once your submission is accepted and deanonymized.
+
+\subsubsection*{Code}
 
 \subsubsection*{Acknowledgments}
-Use unnumbered third level headings for the acknowledgments. All
-acknowledgments, including those to funding agencies, go at the end of the paper.
-Only add this information once your submission is accepted and deanonymized.
+This work is supported by the National Science Foundation under Cooperative Agreement PHY-2019786 (The NSF AI Institute for Artificial Intelligence and Fundamental Interactions, \url{http://iaifi.org/}).
+This material is based upon work supported by the U.S.
+Department of Energy, Office of Science, Office of High Energy Physics of U.S.
+Department of Energy under grant Contract Number  DE-SC0012567.
+The computations in this paper were run on the FASRC Cannon cluster supported by the FAS Division of Science Research Computing Group at Harvard University.
+
+This research is based on observations made with the NASA/ESA Hubble Space Telescope obtained from the Space Telescope Science Institute, which is operated by the Association of Universities for Research in Astronomy, Inc., under NASA contract NAS 5-26555.
+
+Based on observations made with the NASA/ESA Hubble Space Telescope, and obtained from the Hubble Legacy Archive, which is a collaboration between the Space Telescope Science Institute (STScI/NASA), the Space Telescope European Coordinating Facility (ST-ECF/ESAC/ESA) and the Canadian Astronomy Data Centre (CADC/NRC/CSA).
 
@@
 \appendix
-\section{Appendix}
-You may include other additional sections here.
+\section{Summarization via regex-guided generation}
+\label{app:summarization}
+
+The following prompt is used to summarize the abstracts.
+
+\begin{lstlisting}[language=Python]
+@outlines.prompt
+def prompt_fn(abstract):
+
+      """<s>[INST] You are an expert astrophysicist, with broad expertise across observational and theoretical astrophysics. You are able to extract core information from astrophysical texts.
+
+Abstract: "{{abstract}}"
+
+Based on the above observational proposal abstract, your task is to summarize the nature of the eventual observations.
+You will identify the astrophysical objects and phenomena, as well as the potential science use cases described in the abstract.
+
+Follow these instructions exactly:
+- Mention up to 5 items for both categories; do not mention more than 5 items in either category. 
+- Choose the most relevant ones if there are more than 5 items in a category.
+- Never mention the Hubble Space Telescope, HST, or the HST archive.
+- Mention the class (e.g., barred spiral galaxy) and not just the specific instance (e.g., Andromeda).
+- Name the objects in the science use cases, if appropriate.
+- Write out full names of objects in addition to acronyms.
+- Do not list irrelevant objects which do not describe the eventual observation, such as units or proposal Cycle numbers. List fewer but more relevant objects, if in doubt.
+- Each science case listed must be self-contained but succinct.
+- Only write in English.
+- Do not list items that are too generic (e.g., galaxy, faint object, kinematics)
+- The total length of text should not exceed 80 words.
+- Present your lists in a comma-separated format; no dashed or numbered lists.
+
+Example output: {'objects_and_phenomena':'spiral galaxies, galaxy clusters, supernova remnants', 'science_use_cases':'model galactic structure and evolution, characterize dark matter distribution in clusters, analyze expansion rates of supernova remnants'}
+
+Answer in JSON format. The JSON should be a dictionary with keys "objects_and_phenomena" and "science_use_cases".
+
+[/INST]
+"""
+\end{lstlisting}
 

### 2024-01-01 paper/main.tex
More text edits

@@
 
-\subsection{Data Selection and Pre-Processing}
+\subsection{Data Selection and Pre-processing}
 
@@
 A maximum of 20 images are downloaded per proposal ID, selected at random, in order to avoid biasing the model towards proposals with a larger number of observations.
-Images are resized to a maximum resolution per side of 512 pixels before saving.
+Images are centered and resized to a resolution per side of 512 pixels before saving.
 Color previews (i.e., observations taken with multiple wavelength filters assigned to individual RGB channels) are manually excluded via a filename filter in order to maintain consistency across the dataset; models trained on datasets with color images included were observed to show worse performance on generalization metrics.
@@
 
-Raw abstracts summarize the corresponding successful HST observing proposal, which intends to make the case for allocating \hubble telescope time towards a particular set of observations.
+Raw abstracts summarize the corresponding successful HST observing proposals, which intend to make the case for allocating \hubble telescope time towards a particular set of observations.
 These abstracts are written in a diversity of styles, formats, and lengths, being highly variable in the nature of content as well.
@@
 This model uses a (12-layer, 12-head, 768-embedding dimension) vision transformer as the image encoder and a (12-layer, 8-head, 512-embedding dimension) sequence transformer as the text backbone.
+The text encoder has a maximum length of 77-tokens and the image encoder a native resolution of $224\times224$.
 Linear projection layers map the outputs of the image and text encoders to a common embedding space of dimension 512.
+In total, the model has 149,620,737 trainable parameters.
 The model was trained on 400 million image-text pairs from the internet.
-% , and the CLIP family of models has shown strong performance on a variety of downstream tasks including zero-shot classification and image retrieval.
 
-\subsection{Fine-tuning Procedure}
+\subsection{Fine-tuning Procedure and Evaluation}
 \label{sec:finetuning}
+
+The base CLIP model is fine-tuned using the dataset described in Sec.~\ref{sec:dataset}, using either the LLM-summarized data or the raw proposal abstracts.
+When using raw proposal abstracts, random chunks of the text delimited by periods are selected on the fly to fit within the maximum length of the text encoder.
+Images are downsampled to the native resolution of the image encoder and randomly rotated over training.
+Given the relatively modest size of the fine-tuning set, a batch size $|\mathcal B| = 32$ is used throughout.
+We note that positive and negative image-text association is noisy and imperfect, since multiple images can be associated with the same abstract.
+
+We explore three different methods of training the model on our domain dataset: \emph{(1)} Fine-tuning the entire network, starting from the pre-trained base model; \emph{(2)} Freezing the base image and text encoders, and training a small projection head on top of these; and \emph{(3)} Training the entire model from scratch.
+
+The model is trained over 50,000 steps with 5000 linear warmup steps and cosine decay using the AdamW optimizer \citep{DBLP:conf/iclr/LoshchilovH19,DBLP:journals/corr/KingmaB14} with peak learning rate of either $10^{-5}$ or $10^{-6}$ and weight decay $10^{-3}$.
+Training takes approximately 6 hours on 4 Nvidia A100 GPUs.
 

### 2024-01-01 scripts/submit_train.sh
Start new runs

@@ -4,10 +4,9 @@
 #SBATCH --nodes=1
 #SBATCH --mem=200GB
 #SBATCH --time=48:00:00
-# #SBATCH --gres=gpu:nvidia_a100-sxm4-80gb:4
 #SBATCH --gres=gpu:4
 #SBATCH --account=iaifi_lab
-#SBATCH -p iaifi_gpu
+#SBATCH -p iaifi_gpu_requeue
 
 export TF_CPP_MIN_LOG_LEVEL="2"
 
@@ -27,4 +26,10 @@ alias python=/n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python
 alias jupyter=/n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/jupyter
 
 cd /n/holystore01/LABS/iaifi_lab/Users/smsharma/multimodal-data/
-/n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py # --config.data.augment_rotate_type="continuous"
+
+/n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py  # Base config, with summary captions
+# /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.training.n_train_steps=10001 --config.training.warmup_steps=500 --config.optim.learning_rate=1e-6  # Fewer steps
+# /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.optim.learning_rate=1e-6  # Lower LR
+# /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.data.caption_type="abstract"  --config.data.augment_subsample_text=True  # Full abstract
+# /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.clip.random_init_text=True --config.clip.random_init_image=True  # From scratch
+# /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.clip.transfer_head=True  # Fine-tune just head

### 2024-01-02 train.py
Get restore working in eval notebook; reinstate augmentations for eval

@@ -277,31 +277,31 @@ def train(config: ConfigDict, workdir: str = "./logging/") -> train_state.TrainS
                     images, captions = next(val_batches)
                     images = np.array(images)
 
-                    # # NOTE: Don't augment images for eval!  
-                    # # Augment images through random rotations and flips
-                    # if config.data.augment_rotate:
+                    # NOTE: Don't augment images for eval! NOTE: Changed my mind.
+                    # Augment images through random rotations and flips
+                    if config.data.augment_rotate:
 
-                    #     # Rotations
-                    #     rng_eval, _ = jax.random.split(rng_eval)
-                    #     if config.data.augment_rotate_type == "continuous":
-                    #         rotation_angles = jax.random.choice(rng_eval, rot_angles_90, shape=(images.shape[0],))  # Angles in radians
-                    #     elif config.data.augment_rotate_type == "discrete":
-                    #         rotation_angles = jax.random.uniform(rng_eval, shape=(images.shape[0],)) * 2 * np.pi  # Angles in radians
-                    #     else:
-                    #         raise ValueError(f"Invalid augment_rotate_type: {config.data.augment_rotate_type}")
-                    #     images = jax.vmap(partial(rotate, mode='constant', cval=1.))(images, rotation_angles)
+                        # Rotations
+                        rng_eval, _ = jax.random.split(rng_eval)
+                        if config.data.augment_rotate_type == "continuous":
+                            rotation_angles = jax.random.choice(rng_eval, rot_angles_90, shape=(images.shape[0],))  # Angles in radians
+                        elif config.data.augment_rotate_type == "discrete":
+                            rotation_angles = jax.random.uniform(rng_eval, shape=(images.shape[0],)) * 2 * np.pi  # Angles in radians
+                        else:
+                            raise ValueError(f"Invalid augment_rotate_type: {config.data.augment_rotate_type}")
+                        images = jax.vmap(partial(rotate, mode='constant', cval=1.))(images, rotation_angles)
                         
-                    #     # Flips
-                    #     rng_eval, _ = jax.random.split(rng_eval)
-                    #     images = jax.vmap(partial(random_flip_up_down, key=rng_eval))(image=images)
+                        # Flips
+                        rng_eval, _ = jax.random.split(rng_eval)
+                        images = jax.vmap(partial(random_flip_up_down, key=rng_eval))(image=images)
 
-                    #     rng_eval, _ = jax.random.split(rng_eval)
[…]

### 2024-01-02 configs/base.py
Simplify ckpt manager to use only one metric; switched to const schedule by default

@@ -48,13 +48,13 @@ def get_config():
     training.eval_every_steps = 200
     training.loss_type = "softmax"  # "sigmoid" or "softmax"
     training.n_eval_batches = 20  # How many batches to use for evaluation
-    training.ckpt_best_metric = ("top_10_accuracy", "loss")
-    training.ckpt_best_metric_best_mode = ("max", "min")  # "max" or "min" for each metric in `ckpt_best_metric`
+    training.ckpt_best_metric = "loss"  # "loss" or "top_xx_accuracy"
+    training.ckpt_best_metric_best_mode = "min"  # "max" or "min" 
     training.ckpt_keep_top_n = 3  # Save the top `ckpt_keep_top_n` checkpoints based on `ckpt_best_metric`
 
     # Optimizer (AdamW)
     config.optim = optim = ml_collections.ConfigDict()
-    optim.schedule = "cosine"
+    optim.schedule = "linear"
     optim.learning_rate = 1e-5
     optim.weight_decay = 1e-3
 

### 2024-01-02 train.py

@@ -147,15 +147,13 @@ def train(config: ConfigDict, workdir: str = "./logging/") -> train_state.TrainS
 
     ## Training config and loop
     
-    # Create checkpoint managers, potentially based on multiple metrics
+    # Create checkpoint manager
 
-    ckpt_mgrs = []
-    for metric, best_mode in zip(config.training.ckpt_best_metric, config.training.ckpt_best_metric_best_mode):
-        best_fn = lambda metrics: metrics[f"val/{metric}"]
-        
-        mgr_options = orbax.checkpoint.CheckpointManagerOptions(create=True, step_prefix=f'step_{metric}', max_to_keep=config.training.ckpt_keep_top_n, best_fn=best_fn, best_mode=best_mode)
+    best_fn = lambda metrics: metrics[f"val/{config.training.ckpt_best_metric}"]
+    
+    mgr_options = orbax.checkpoint.CheckpointManagerOptions(create=True, step_prefix=f'step_{config.training.ckpt_best_metric}', max_to_keep=config.training.ckpt_keep_top_n, best_fn=best_fn, best_mode=config.training.ckpt_best_metric_best_mode)
 
-        ckpt_mgrs.append(orbax.checkpoint.CheckpointManager(f"{workdir}/ckpts/", orbax.checkpoint.Checkpointer(orbax.checkpoint.PyTreeCheckpointHandler()), mgr_options))
+    ckpt_mgr = orbax.checkpoint.CheckpointManager(f"{workdir}/ckpts/", orbax.checkpoint.Checkpointer(orbax.checkpoint.PyTreeCheckpointHandler()), mgr_options)
 
     # Optimizer and schedule
 
@@ -332,8 +330,7 @@ def train(config: ConfigDict, workdir: str = "./logging/") -> train_state.TrainS
                 state_ckpt = unreplicate(pstate)
                 save_args = orbax_utils.save_args_from_target(state_ckpt)
 
-                for ckpt_mgr in ckpt_mgrs:
-                    ckpt_mgr.save(step, state_ckpt, save_kwargs={'save_args': save_args}, metrics=summary)
+                ckpt_mgr.save(step, state_ckpt, save_kwargs={'save_args': save_args}, metrics=summary)
 
     logging.info("All done! Have a great day.")
 

### 2024-01-02 configs/base.py
Move eval to before train step

@@ -54,7 +54,7 @@ def get_config():
 
     # Optimizer (AdamW)
     config.optim = optim = ml_collections.ConfigDict()
-    optim.schedule = "linear"
+    optim.schedule = "constant"
     optim.learning_rate = 1e-5
     optim.weight_decay = 1e-3
 

### 2024-01-02 train.py

@@ -199,68 +199,12 @@ def train(config: ConfigDict, workdir: str = "./logging/") -> train_state.TrainS
     train_metrics = []
     with trange(config.training.n_train_steps) as steps:
         for step in steps:
-            rng, rng_aug = jax.random.split(rng)
-            images, captions = next(batches)
-            images = np.array(images)
-
-            # Augment images through random rotations and flips
-            if config.data.augment_rotate:
-
-                # Rotations
-                rng_aug, _ = jax.random.split(rng_aug)
-                if config.data.augment_rotate_type == "continuous":
-                    rotation_angles = jax.random.choice(rng_aug, rot_angles_90, shape=(images.shape[0],))  # Angles in radians
-                elif config.data.augment_rotate_type == "discrete":
-                    rotation_angles = jax.random.uniform(rng_aug, shape=(images.shape[0],)) * 2 * np.pi  # Angles in radians
-                else:
-                    raise ValueError(f"Invalid augment_rotate_type: {config.data.augment_rotate_type}")
-                images = jax.vmap(partial(rotate, mode='constant', cval=1.))(images, rotation_angles)
-
-                # Flips
-                rng_aug, _ = jax.random.split(rng_aug)
-                images = jax.vmap(partial(random_flip_up_down, key=rng_aug))(image=images)
-
-                rng_aug, _ = jax.random.split(rng_aug)
-                images = jax.vmap(partial(random_flip_left_right, key=rng_aug))(image=images)
-
-            # Augment images through random crops
-            # Otherwise, they'll be downsampled to the vision model's image size
-            if config.data.augment_crop:
-                rng_aug, _ = jax.random.split(rng_aug)
-                images = jax.vmap(random_crop, in_axes=(None,0,None))(rng_aug, images, (model.config.vision_config.image_size, model.config.vision_config.image_size, 3))
-
-            # NOTE: Image arrays should be ints in the range [0, 255] here
-            captions = process_truncate_captions(captions, rng_aug, max_length_words=max_length_words)
-            inputs = processor(text=captions, images=(images * 255.).astype(np.uint8),  return_tensors="np", padding="max_length", truncation=True, max_length=model.config.text_config.max_length)
-            batch = inputs.data
-            
-            # Optionally shuffle "pixel_values" within batch
-            if config.data.shuffle_within_batch:
-                batch["pixel_values"] = jax.random.permutation(rng, batch["pixel_values"], axis=0)
-
[…]

### 2024-01-02 train.py
Remove prefix from ckpy manager

@@ -147,11 +147,10 @@ def train(config: ConfigDict, workdir: str = "./logging/") -> train_state.TrainS
 
     ## Training config and loop
     
-    # Create checkpoint manager
+    # Checkpoint manager
 
     best_fn = lambda metrics: metrics[f"val/{config.training.ckpt_best_metric}"]
-    
-    mgr_options = orbax.checkpoint.CheckpointManagerOptions(create=True, step_prefix=f'step_{config.training.ckpt_best_metric}', max_to_keep=config.training.ckpt_keep_top_n, best_fn=best_fn, best_mode=config.training.ckpt_best_metric_best_mode)
+    mgr_options = orbax.checkpoint.CheckpointManagerOptions(create=True, step_prefix=f'step', max_to_keep=config.training.ckpt_keep_top_n, best_fn=best_fn, best_mode=config.training.ckpt_best_metric_best_mode)
 
     ckpt_mgr = orbax.checkpoint.CheckpointManager(f"{workdir}/ckpts/", orbax.checkpoint.Checkpointer(orbax.checkpoint.PyTreeCheckpointHandler()), mgr_options)
 
@@ -204,7 +203,7 @@ def train(config: ConfigDict, workdir: str = "./logging/") -> train_state.TrainS
             rng_eval = jax.random.PRNGKey(config.seed)
 
             # Eval portion
-            
+
             # Evaluate before starting, hence no `and (step != 0)`
             if (step % config.training.eval_every_steps == 0) and (jax.process_index() == 0):
 

### 2024-01-02 configs/base.py
Update envs and launch new jobs

@@ -42,7 +42,7 @@ def get_config():
     training.train_fraction = 0.95
     training.batch_size = 32  # Must be divisible by number of devices; this is the total batch size, not per-device
     training.batch_size_val = 100
-    training.n_train_steps = 40_001
+    training.n_train_steps = 20_001
     training.warmup_steps = 2000
     training.log_every_steps = 100
     training.eval_every_steps = 200

### 2024-01-02 scripts/submit_train.sh

@@ -6,7 +6,7 @@
 #SBATCH --time=48:00:00
 #SBATCH --gres=gpu:4
 #SBATCH --account=iaifi_lab
-#SBATCH -p iaifi_gpu_requeue
+#SBATCH -p iaifi_gpu_priority
 
 export TF_CPP_MIN_LOG_LEVEL="2"
 
@@ -27,9 +27,15 @@ alias jupyter=/n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/jupyter
 
 cd /n/holystore01/LABS/iaifi_lab/Users/smsharma/multimodal-data/
 
-/n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py  # Base config, with summary captions
-# /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.training.n_train_steps=10001 --config.training.warmup_steps=500 --config.optim.learning_rate=1e-6  # Fewer steps
+# /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py  # Base config, with summary captions
+# /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.training.n_train_steps=5001 --config.optim.learning_rate=1e-5  # Fewer steps
+
 # /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.optim.learning_rate=1e-6  # Lower LR
+# /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.training.n_train_steps=5001 --config.optim.learning_rate=1e-6  # Fewer steps
+
+/n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.clip.transfer_head=True  # Fine-tune just head
+/n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.optim.schedule="cosine" # Cosine lr schedule
+
+# # For when things are working
[…]

### 2024-01-02 scripts/submit_train.sh
Launch addl jobs to requeue queue

@@ -6,7 +6,7 @@
 #SBATCH --time=48:00:00
 #SBATCH --gres=gpu:4
 #SBATCH --account=iaifi_lab
-#SBATCH -p iaifi_gpu_priority
+#SBATCH -p iaifi_gpu_requeue
 
 export TF_CPP_MIN_LOG_LEVEL="2"
 
@@ -27,15 +27,15 @@ alias jupyter=/n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/jupyter
 
 cd /n/holystore01/LABS/iaifi_lab/Users/smsharma/multimodal-data/
 
-# /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py  # Base config, with summary captions
+/n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py  # Base config, with summary captions
 # /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.training.n_train_steps=5001 --config.optim.learning_rate=1e-5  # Fewer steps
 
 # /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.optim.learning_rate=1e-6  # Lower LR
 # /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.training.n_train_steps=5001 --config.optim.learning_rate=1e-6  # Fewer steps
 
-/n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.clip.transfer_head=True  # Fine-tune just head
-/n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.optim.schedule="cosine" # Cosine lr schedule
+# /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.clip.transfer_head=True  # Fine-tune just head
+# /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.optim.schedule="cosine" # Cosine lr schedule
 
-# # For when things are working
[…]

### 2024-01-02 scripts/submit_train.sh
Launch more run configs

@@ -6,7 +6,7 @@
 #SBATCH --time=48:00:00
 #SBATCH --gres=gpu:4
 #SBATCH --account=iaifi_lab
-#SBATCH -p iaifi_gpu_requeue
+#SBATCH -p iaifi_gpu_priority
 
 export TF_CPP_MIN_LOG_LEVEL="2"
 
@@ -28,14 +28,10 @@ alias jupyter=/n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/jupyter
 cd /n/holystore01/LABS/iaifi_lab/Users/smsharma/multimodal-data/
 
 /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py  # Base config, with summary captions
-# /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.training.n_train_steps=5001 --config.optim.learning_rate=1e-5  # Fewer steps
-
 # /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.optim.learning_rate=1e-6  # Lower LR
-# /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.training.n_train_steps=5001 --config.optim.learning_rate=1e-6  # Fewer steps
 
 # /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.clip.transfer_head=True  # Fine-tune just head
 # /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.optim.schedule="cosine" # Cosine lr schedule
-
-# For when things are working
 # /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.data.caption_type="abstract"  --config.data.augment_subsample_text=True  # Full abstract
-# /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.clip.random_init_text=True --config.clip.random_init_image=True  # From scratch
[…]

### 2024-01-02 paper/main.tex
Add some comments

@@
 
+\newcommand{\package}[1]{\textsl{#1}\xspace}
+\newcommand{\eg}{{e.\,g.}\xspace}
+\newcommand{\ie}{{i.\,e.}\xspace}
+\newcommand{\SM}[1]{\textcolor{blue}{[SM: #1]}}
 \newcommand{\hubble}{\emph{Hubble}\xspace}
+
+\def\preprintno{XXXX} % Insert correct preprint number
 
@@
 ]{hyperref}
+
+% Define a new fancy page style
+\fancypagestyle{firstpage}{
+    \rhead{MIT-CTP/\preprintno}
+    % Define other header and footer elements if necessary
+}
 
@@
 
+\thispagestyle{firstpage}
+
 \begin{abstract}
@@
 We curate a dataset of images of \hubble observations and corresponding text descriptions.
-We rely on summarized versions of proposal abstracts from the Proposal Abstracts Catalog\footnote{\url{https://archive.stsci.edu/hst/proposal_abstracts.html}} -- a catalog of all accepted \hubble proposals -- as captions for the observations.
-The HST has been operational for 33 years, having been launched on April 24, 1990, and we use proposals and observations up to Cycle 31.
+We rely on summarized versions of proposal abstracts from the Proposal Abstracts Catalog\footnote{\url{https://archive.stsci.edu/hst/proposal_abstracts.html}} -- a catalog of all accepted \hubble proposals -- to derive captions for the observations.
+The HST has been operational for 33 years, having been launched on April 24, 1990.
+We use available proposals and observations up to Cycle 30, which commenced data-taking in 2022.
 
@@
 3,194 images are held out for validation, with no abstract being common between training and validation sets in order to ensure an independent set of text-image pairs for testing.
+
+% Mention that the signal is noisy!
 
@@
 The goal if summarization-via-guided-generation is to increase the signal between text and images by standardizing the captions used for fine-tuning the CLIP model.
+
+% Note that the signal is very imperfect!
 
@@
           \end{tabular}
-          \caption{Examples of the beginning of raw proposal abstracts (second column) and LLM (\textsc{Mixtral-8x7B})-extracted summaries (rightmost two columns), separately extracting objects and phenomena as well as potential downstream science use cases.
+          \caption{Examples of the initial parts of raw proposal abstracts (second column) and LLM (\textsc{Mixtral-8x7B})-extracted summaries (rightmost two columns), separately extracting objects and phenomena as well as potential downstream science use cases.
 The LLM-extracted summaries are used for associating text with observations.}
@@
 Given a minibatch $\mathcal{B}$ of $|\mathcal{B}|$ image-text pairs $\{(I_i, T_i)\}$, the goal is to align the learned representations of corresponding (positive) pairs $(I_i, T_i)$ while repelling the representations of unaligned (negative) pairs $(I_i, T_{j\neq i})$.
-Image and text encoders $f(\cdot)$ and $g(\cdot)$ are used to map images and text to a common embedding space, and the standard softmax-based InfoNCE~\citep{oord2018representation} contrastive loss function is given by
+Image and text encoders $f(\cdot)$ and $g(\cdot)$ are used to map images and text to a common embedding space.
+The standard softmax-based bidirectional variation of the InfoNCE~\citep{oord2018representation} contrastive loss function, as used by CLIP, is particularly effective for multimodal learning.
+This bidirectionality is crucial as it ensures the model learns to map both images to text and text to images with equal importance.
+This symmetry in learning is essential for tasks that require a mutual understanding and interchangeability between visual and textual representations, such as image captioning, text-to-image synthesis, and cross-modal retrieval.
+The bidirectional loss is given by \citep{radford2021learning}
 \begin{equation}
-\mathcal L(\mathcal B) = -\frac{1}{2|\mathcal{B}|} \sum_{i=1}^{|\mathcal{B}|}\left(\overbrace{\log \frac{e^{t {x}_i \cdot {y}_i}}{\sum_{j=1}^{|\mathcal{B}|} e^{t {x}_i \cdot {y}_j}}}^{\text {image } \rightarrow \text { text softmax }}+\overbrace{\log \frac{e^{t {x}_i \cdot {y}_i}}{\sum_{j=1}^{|\mathcal{B}|} e^{t {x}_j \cdot {y}_i}}}^{\text {text } \rightarrow \text { image softmax }}\right)
+\mathcal L(\mathcal B) = -\frac{1}{2|\mathcal{B}|} \sum_{i=1}^{|\mathcal{B}|}\left(\overbrace{\log \frac{e^{{x}_i \cdot {y}_i / \tau}}{\sum_{j=1}^{|\mathcal{B}|} e^{{x}_i \cdot {y}_j / \tau}}}^{\text {image } \rightarrow \text { text softmax }}+\overbrace{\log \frac{e^{{x}_i \cdot {y}_i / \tau}}{\sum_{j=1}^{|\mathcal{B}|} e^{{x}_j \cdot {y}_i / \tau}}}^{\text {text } \rightarrow \text { image softmax }}\right)
 \end{equation}
-where ${x}_i=\frac{f\left(I_i\right)}{\left\|f\left(I_i\right)\right\|_2}$ and ${y}_i=\frac{g\left(T_i\right)}{\left\|g\left(T_i\right)\right\|_2}$ are the normalized representations of the $i$-th image and text, respectively, and $t$ is a (typically learnable) temperature hyperparameter.
+where ${x}_i=\frac{f\left(I_i\right)}{\left\|f\left(I_i\right)\right\|_2}$ and ${y}_i=\frac{g\left(T_i\right)}{\left\|g\left(T_i\right)\right\|_2}$ are the normalized representations of the $i$-th image and text, respectively, and $\tau$ is a (typically learnable) temperature hyperparameter.
+This bidirectional approach not only enhances the model's ability to create robust and versatile embeddings but also aligns with the intuitive understanding of how humans interact with and interpret multimodal information, further bridging the gap between artificial and natural intelligence in multimodal contexts.
 
@@
 The base CLIP model is fine-tuned using the dataset described in Sec.~\ref{sec:dataset}, using either the LLM-summarized data or the raw proposal abstracts.
-When using raw proposal abstracts, random chunks of the text delimited by periods are selected on the fly to fit within the maximum length of the text encoder.
-Images are downsampled to the native resolution of the image encoder and randomly rotated over training.
+When using raw proposal abstracts, random chunks of the text delimited by periods are selected on the fly to fit within the maximum token length of the text encoder.
+Images are randomly cropped to the native resolution of the image encoder and randomly rotated at each training step.
 Given the relatively modest size of the fine-tuning set, a batch size $|\mathcal B| = 32$ is used throughout.
-We note that positive and negative image-text association is noisy and imperfect, since multiple images can be associated with the same abstract.
+We note that the positive and negative image-text association is noisy and imperfect, since multiple images can be associated with the same abstract.
 
@@
 
+% Should write BIS!
+
 % \subsubsection*{Author Contributions}
@@
 \subsubsection*{Code}
+
+% Code repo and package cites.
+% Dataset availability
 

### 2024-01-02 scripts/submit_train.sh
More run configs

@@ -35,3 +35,5 @@ cd /n/holystore01/LABS/iaifi_lab/Users/smsharma/multimodal-data/
 # /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.data.caption_type="abstract"  --config.data.augment_subsample_text=True  # Full abstract
 # /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.clip.random_init_text=True --config.clip.random_init_vision=True  # From scratch
 # /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.training.loss_type="sigmoid"
+# /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.data.shuffle_within_batch=True  # Shuffle within batch
+# /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.training.batch_size=64  # Larger batch size

### 2024-01-03 paper/main.tex
Bib

@@
 
-\subsection{Data Selection and Pre-processing}
+\subsection{Data Selection and Pre-Processing}
 
@@
 where ${x}_i=\frac{f\left(I_i\right)}{\left\|f\left(I_i\right)\right\|_2}$ and ${y}_i=\frac{g\left(T_i\right)}{\left\|g\left(T_i\right)\right\|_2}$ are the normalized representations of the $i$-th image and text, respectively, and $\tau$ is a (typically learnable) temperature hyperparameter.
-This bidirectional approach not only enhances the model's ability to create robust and versatile embeddings but also aligns with the intuitive understanding of how humans interact with and interpret multimodal information, further bridging the gap between artificial and natural intelligence in multimodal contexts.
+% This bidirectional approach not only enhances the model's ability to create robust and versatile embeddings but also aligns with the intuitive understanding of how humans interact with and interpret multimodal information, further bridging the gap between artificial and natural intelligence in multimodal contexts.
 
@@
 This model uses a (12-layer, 12-head, 768-embedding dimension) vision transformer as the image encoder and a (12-layer, 8-head, 512-embedding dimension) sequence transformer as the text backbone.
-The text encoder has a maximum length of 77-tokens and the image encoder a native resolution of $224\times224$.
+The text encoder has a maximum length of 77 tokens and the image encoder a native resolution of $224\times224$.
 Linear projection layers map the outputs of the image and text encoders to a common embedding space of dimension 512.
@@
 
+Models were instantiated and trained using the \package{Jax} \citep{jax2018github} ecosystem.
 The model is trained over 50,000 steps with 5000 linear warmup steps and cosine decay using the AdamW optimizer \citep{DBLP:conf/iclr/LoshchilovH19,DBLP:journals/corr/KingmaB14} with peak learning rate of either $10^{-5}$ or $10^{-6}$ and weight decay $10^{-3}$.
@@
 
+The code, dataset, and models used in this work is available at \url{https://www.github.com/smsharma/HubbleCLIP}.
+
+This work relied on the \package{Astroquery} \citep{2019AJ....157...98G}, \package{bitsandbytes} \citep{dettmers2022llmint8}, \package{Flax} \citep{flax2020github}, \package{Jax} \citep{jax2018github}, \package{Jupyter} \citep{Kluyver2016jupyter}, \package{Matplotlib} \citep{Hunter:2007}, \package{Numpy} \citep{harris2020array}, \package{Optax} \citep{deepmind2020jax}, \package{Outlines}, \package{Pandas}, \package{Pydantic}, \package{PyTorch} \citep{paszke2019pytorch}, \package{SciPy} \citep{2020SciPy-NMeth}, \package{Transformers} \citep{wolf2019huggingface}, and \package{wandb} \citep{wandb} software packages.
+
 % Code repo and package cites.
@@
 Department of Energy under grant Contract Number  DE-SC0012567.
+YS was supported by the Research Science Institute (RSI) program at MIT.
 The computations in this paper were run on the FASRC Cannon cluster supported by the FAS Division of Science Research Computing Group at Harvard University.
@@
 \appendix
-\section{Summarization via regex-guided generation}
+\section{Summarization via Regex-Guided Generation}
 \label{app:summarization}
@@
 
+\section{Additional Evaluation Metrics and Ablations}
+\label{app:ablations}
+
 \end{document}

### 2024-01-03 paper/main.tex
Paper update

@@
 \begin{equation}
-\mathcal L(\mathcal B) = -\frac{1}{2|\mathcal{B}|} \sum_{i=1}^{|\mathcal{B}|}\left(\overbrace{\log \frac{e^{{x}_i \cdot {y}_i / \tau}}{\sum_{j=1}^{|\mathcal{B}|} e^{{x}_i \cdot {y}_j / \tau}}}^{\text {image } \rightarrow \text { text softmax }}+\overbrace{\log \frac{e^{{x}_i \cdot {y}_i / \tau}}{\sum_{j=1}^{|\mathcal{B}|} e^{{x}_j \cdot {y}_i / \tau}}}^{\text {text } \rightarrow \text { image softmax }}\right)
+\mathcal L(\mathcal B) = -\frac{1}{2|\mathcal{B}|} \sum_{i=1}^{|\mathcal{B}|}\left(\underbrace{\log \frac{e^{{x}_i \cdot {y}_i / \tau}}{\sum_{j=1}^{|\mathcal{B}|} e^{{x}_i \cdot {y}_j / \tau}}}_{\text {image } \rightarrow \text { text softmax }}+\underbrace{\log \frac{e^{{x}_i \cdot {y}_i / \tau}}{\sum_{j=1}^{|\mathcal{B}|} e^{{x}_j \cdot {y}_i / \tau}}}_{\text {text } \rightarrow \text { image softmax }}\right)
 \end{equation}
-where ${x}_i=\frac{f\left(I_i\right)}{\left\|f\left(I_i\right)\right\|_2}$ and ${y}_i=\frac{g\left(T_i\right)}{\left\|g\left(T_i\right)\right\|_2}$ are the normalized representations of the $i$-th image and text, respectively, and $\tau$ is a (typically learnable) temperature hyperparameter.
+where ${x}_i=\frac{f\left(I_i\right)}{\left\|f\left(I_i\right)\right\|_2}$ and ${y}_i=\frac{g\left(T_i\right)}{\left\|g\left(T_i\right)\right\|_2}$ are the normalized representations of the $i$-th image and text, respectively, and $\tau$ is a learnable temperature hyperparameter.
 % This bidirectional approach not only enhances the model's ability to create robust and versatile embeddings but also aligns with the intuitive understanding of how humans interact with and interpret multimodal information, further bridging the gap between artificial and natural intelligence in multimodal contexts.
@@
 Models were instantiated and trained using the \package{Jax} \citep{jax2018github} ecosystem.
-The model is trained over 50,000 steps with 5000 linear warmup steps and cosine decay using the AdamW optimizer \citep{DBLP:conf/iclr/LoshchilovH19,DBLP:journals/corr/KingmaB14} with peak learning rate of either $10^{-5}$ or $10^{-6}$ and weight decay $10^{-3}$.
-Training takes approximately 6 hours on 4 Nvidia A100 GPUs.
+The model is trained over 20,000 steps with 5000 linear warmup steps and cosine decay using the AdamW optimizer \citep{DBLP:conf/iclr/LoshchilovH19,DBLP:journals/corr/KingmaB14} with peak learning rate of $10^{-5}$ and weight decay $10^{-3}$.
+Training takes approximately 3 hours on 4 Nvidia A100 GPUs.
+
+% Describe evals
 
@@
 \includegraphics[width=0.95\textwidth]{plots/val_metrics.pdf}
-      \caption{Retrieval accuracy}
-      \label{fig:retrieval_acc}
-      \end{figure*}
-
-\begin{figure*}[!h]
-\includegraphics[width=0.95\textwidth]{plots/retrieval_acc.pdf}
 \caption{Retrieval accuracy}
@@
 
+\begin{figure*}[!h]
+\includegraphics[width=0.95\textwidth]{plots/retrieval_acc.pdf}
+\caption{Retrieval accuracy}
+\label{fig:retrieval_acc}
+\end{figure*}
+
 \end{document}

### 2024-01-03 paper/main.tex
Draft updates

@@
 \definecolor{xlinkcolor}{rgb}{0.7752941176470588, 0.22078431372549023, 0.2262745098039215}
-\definecolor{codegreen}{rgb}{0,0.6,0}
+\definecolor{codegreen}{rgb}{0,0.4,0}
 \definecolor{codegray}{rgb}{0.5,0.5,0.5}
@@
 We present a multi-modal model which associates astronomical observations imaged by the \hubble Space Telescope (HST) with natural language.
-The model is fine-tuned from a base CLIP model using successful proposal abstracts corresponding to HST observations, summarized via guided large language (LLM) model generation.
+The model is fine-tuned from a base Contrastive Language–Image Pre-training (CLIP) model using successful proposal abstracts corresponding to HST observations, summarized via guided large language model (LLM) generation.
 We show that the model embodies a meaningful joint representation between observations and text through experiments targeting observation retrieval (i.e., retrieving the most relevant observations from a set using natural language queries) and description retrieval (i.e., querying the astrophysical object classes and science use cases most relevant to a given observation).
@@
 
-Observations corresponding to individual proposal IDs are queried through the Mikulski Archive for Space Telescopes (MAST) via the \texttt{astroquery} interface.
+Observations corresponding to individual proposal IDs are queried through the Mikulski Archive for Space Telescopes (MAST) via the \package{Astroquery} interface.
 Products of type \texttt{PREVIEW} are filtered in, corresponding to preview postcard images.
-Note that these are not science-grade observations, but rather low-resolution images used for quick-look purposes; given the nature of associations we aim to learn, we deem this adequate for our current purposes.
+Note that these are not science-grade observations, but rather lower-resolution images used for quick-look purposes; given the nature of associations we aim to learn, we deem this adequate for our current purposes.
 A maximum of 20 images are downloaded per proposal ID, selected at random, in order to avoid biasing the model towards proposals with a larger number of observations.
@@
 
-The open-weights, instruction-tuned model \textsc{Mixtral-8x7B}\footnote{\url{https://huggingface.co/mistralai/Mixtral-8x7B-Instruct-v0.1}} is used to generate the summaries, with guided generation performed using the \texttt{outlines}\footnote{\url{https://github.com/outlines-dev/outlines}} package.
+The open-weights, instruction-tuned model \textsc{Mixtral-8x7B-Instruct}\footnote{\url{https://huggingface.co/mistralai/Mixtral-8x7B-Instruct-v0.1}} is used to generate the summaries, with guided generation performed using the \package{Outlines}\footnote{\url{https://github.com/outlines-dev/outlines}} package.
 
@@
 \mathcal L(\mathcal B) = -\frac{1}{2|\mathcal{B}|} \sum_{i=1}^{|\mathcal{B}|}\left(\underbrace{\log \frac{e^{{x}_i \cdot {y}_i / \tau}}{\sum_{j=1}^{|\mathcal{B}|} e^{{x}_i \cdot {y}_j / \tau}}}_{\text {image } \rightarrow \text { text softmax }}+\underbrace{\log \frac{e^{{x}_i \cdot {y}_i / \tau}}{\sum_{j=1}^{|\mathcal{B}|} e^{{x}_j \cdot {y}_i / \tau}}}_{\text {text } \rightarrow \text { image softmax }}\right)
+\label{eq:softmax_loss}
 \end{equation}
-where ${x}_i=\frac{f\left(I_i\right)}{\left\|f\left(I_i\right)\right\|_2}$ and ${y}_i=\frac{g\left(T_i\right)}{\left\|g\left(T_i\right)\right\|_2}$ are the normalized representations of the $i$-th image and text, respectively, and $\tau$ is a learnable temperature hyperparameter.
+where ${x}_i={f\left(I_i\right)}/{\left\|f\left(I_i\right)\right\|}$ and ${y}_i={g\left(T_i\right)}/{\left\|g\left(T_i\right)\right\|}$ are the normalized representations of the $i$-th image and text, respectively, and $\tau$ is a learnable temperature hyperparameter.
 % This bidirectional approach not only enhances the model's ability to create robust and versatile embeddings but also aligns with the intuitive understanding of how humans interact with and interpret multimodal information, further bridging the gap between artificial and natural intelligence in multimodal contexts.
@@
 Images are randomly cropped to the native resolution of the image encoder and randomly rotated at each training step.
-Given the relatively modest size of the fine-tuning set, a batch size $|\mathcal B| = 32$ is used throughout.
+Given the relatively modest size of the fine-tuning set, a batch size $|\mathcal B| = 32$ is used throughout; larger batch sizes were seen to susceptible to overfitting.
 We note that the positive and negative image-text association is noisy and imperfect, since multiple images can be associated with the same abstract.
@@
 We explore three different methods of training the model on our domain dataset: \emph{(1)} Fine-tuning the entire network, starting from the pre-trained base model; \emph{(2)} Freezing the base image and text encoders, and training a small projection head on top of these; and \emph{(3)} Training the entire model from scratch.
+For \emph{(2)}, we use a 2-layer MLP with 1024 units and a GELU activation layer, projecting to the 512-dimensional embedding space.
 
-Models were instantiated and trained using the \package{Jax} \citep{jax2018github} ecosystem.
-The model is trained over 20,000 steps with 5000 linear warmup steps and cosine decay using the AdamW optimizer \citep{DBLP:conf/iclr/LoshchilovH19,DBLP:journals/corr/KingmaB14} with peak learning rate of $10^{-5}$ and weight decay $10^{-3}$.
+Models were instantiated and trained using the \package{Jax} \citep{jax2018github} ecosystem. The model is trained over 20,000 steps with 2000 linear warmup steps 
+% and cosine decay 
+using the AdamW optimizer \citep{DBLP:conf/iclr/LoshchilovH19,DBLP:journals/corr/KingmaB14} with peak learning rate of $10^{-5}$ and weight decay $10^{-3}$.
 Training takes approximately 3 hours on 4 Nvidia A100 GPUs.
 
-% Describe evals
+The model is evaluated by tracking the loss in Eq.~\ref{eq:softmax_loss} as well as the top-$k\%$ retrieval accuracy on the held out validation set over the course of training.
+The retrieval accuracy is defined as the fraction of associated captions which fall within the top $k\%$ of captions by cosine similarity of the (normalized) embeddings $x_i \cdot y_j$, averaged over the images in the validation set.
 
@@
 
-% Effect of using summarization
-% Effect of using a base model vs fine-tuning from scratch
+\paragraph*{Validation metrics}
 
-\subsection{Fine-tuned Retrieval Accuracy}
-\label{sec:retrieval_acc}
+Fig.~\ref{fig:retrieval_acc} shows the contrastive loss (left) and the top-10\% retrieval accuracy on the held out validation set over the course of training, for different variations considered.
+The red lines show the metrics evaluated when training with batches where the image-text associations are randomly shuffled, serving as a baseline.
+This baseline is seen to do on par with random expectation, unlike the others, validating the presence of a positive image-text association signal in the dataset.
+%
+Interestingly, the base model performs better than random expectation, with a top-10\% retrieval accuracy of $\sim 15\%$.
+We will therefore compare the qualitative performance of the base model with the fine-tuned model on downstream retrieval tasks.
 
-\subsection{`Zero-shot' Hypothesis and Object Retrieval}
-\label{sec:zero_shot}
+The fiducial model with LLM-guided summarization (orange lines) is seen to perform significantly better than the model using raw abstracts as captions (purple line), validating the stronger association signal in the summarized dataset we curate.
+Fine-tuning a small MLP head over frozen vision and text backbones (green lines) and training from scratch (blue lines) show a non-trivial improvement in the retrieval metrics compared to the random baseline as well as base model, but with deteriorated performance compared to the fiducial set-up.
 
-\subsection{Text-to-Image Retrieval}
-\label{sec:tti}
+\begin{figure*}[!h]
+  \includegraphics[width=0.95\textwidth]{plots/val_metrics.pdf}
+  \caption{The CLIP-contrastive loss from Eq.~\ref{eq:softmax_loss} and the top-10\% retrieval accuracy computed on the validation set over the course of training.
+Shown for the fiducial set-up (orange), dataset using raw proposal abstracts (purple), only fine-tuning a small MLP head (green), training from scratch (blue), and a baseline trained with shuffled image-text pairs.}
+  \label{fig:retrieval_acc}
+  \end{figure*}
+
+\paragraph*{Image retrieval}
+
+Having aligned the image and text representations, we can query the validation set using natural language and show the most ``relevant'' images when ranked by cosine similarity.
+We show this in Figs.~\ref{fig:tti_base} and \ref{fig:tti} for the base model and fine-tuned model respectively using four simple curated queries: `globular clusters', `dwarf galaxy', `SN1987A', and `cluster lensing'.
+
+The base model shows signs of meaningful retrieval -- it returns observations that clearly visually resemble globular clusters (first row of Fig.~\ref{fig:tti_base}), for example.
+Beyond this, the `dwarf galaxy' query returns a mix of galaxies and globular clusters, and it is challenging to discern meaningful associations between the other retrieved images and corresponding query.
+
+The fine-tuned model (Fig.~\ref{fig:tti}) shows strikingly different behavior.
+For example, it is able to return images with processing and assembly artifacts particular to the telescope (the lines through the middle), which typically receive low similarity scores with the base model.
+The `dwarf galaxy' images correspond to proposals aiming to measure the kinematics of the stellar cores of dwarf galaxies.
+Supernova SN1987 itself can be seen as the most relevant image for the `SN1987A' query.
+Cluster-scale gravitational lenses are returned by the `cluster lensing' query, with lensing patterns visible in the images.
+
+\paragraph*{Text retrieval}
+
+Finally, we can use images from the validation set as queries and retrieve the most relevant text chunks (e.g., contained objects and use cases) from a curated list.
+We show the result of image-to-text retrieval in Fig.~\ref{fig:itt}, for the base as well as fine-tuned models, using the same genre of observations as for the text-to-image retrieval examples.
+We curate a list of possible text associations by querying the \textsc{Claude} large language model for such a list, which we show in App.~\ref{app:categories}.
+
+The top 3 text associations are shown for each image query.
+The `ground truth' summarized abstract is shown in the right column.
+The base model is seen to return a mix of relevant and less-relevant associations.
+While it can often return the nature of objects imaged, we observe it to seldom return scientific phenomena (e.g., `dark matter' as successfully done by the fine-tuned model in the second row).
+The third row (supernova 1987A) highlights interesting behavior -- the base model erroneously attributes the object at the center of the image to a gravitational lens or protoplanetary disk, while the fine-tuned model correctly identifies it as a supernova remnant (which play a crucial role for interstellar chemistry -- another returned snippet).
+
+\begin{figure*}[!h]
+\includegraphics[width=0.95\textwidth]{plots/tti_base.pdf}
+\caption{Image retrieval using the base CLIP model on four curated queries.}
+\label{fig:tti_base}
+\end{figure*}
+
+\begin{figure*}[!h]
+\includegraphics[width=0.95\textwidth]{plots/tti.pdf}
+\caption{Image retrieval using the fine-tuned CLIP model on four curated queries.}
+\label{fig:tti}
+\end{figure*}
+
+\begin{figure*}[!h]
+\includegraphics[width=0.95\textwidth]{plots/itt.pdf}
+\caption{Text associations from a curated list most closely matching a given image query, for both the fine-tuned and base models.
+The `ground truth' LLM-summarized abstract is shown in the right column.}
+\label{fig:itt}
+\end{figure*}
 
@@
 
-% \begin{figure*}[!h]
-% \includegraphics[width=0.95\textwidth]{example-image-a}
-% \caption{Overview of dataset}
-% \label{fig:dataset}
-% \end{figure*}
-
-\begin{figure*}[!h]
-\includegraphics[width=0.95\textwidth]{plots/val_metrics.pdf}
-\caption{Retrieval accuracy}
-\label{fig:retrieval_acc}
-\end{figure*}
-
-\begin{figure*}[!h]
-\includegraphics[width=0.95\textwidth]{plots/itt.pdf}
-\caption{Retrieval accuracy}
-\label{fig:itt}
-\end{figure*}
-
-\begin{figure*}[!h]
-\includegraphics[width=0.95\textwidth]{plots/tti_base.pdf}
-\caption{Retrieval accuracy}
-\label{fig:tti}
-\end{figure*}
-
-\begin{figure*}[!h]
-\includegraphics[width=0.95\textwidth]{plots/tti.pdf}
-\caption{Retrieval accuracy}
-\label{fig:tti_base}
-\end{figure*}
+We present \textsc{HubbleCLIP}, a multi-modal fountation model that associates observations imaged by the \hubble Space Telescope with natural language in a common embedding space.
+The model is fine-tuned from a pre-trained CLIP model on LLM-summarized versions of \hubble proposal abstracts, leveraging a noisy signal associating text and images.
+We show that \textsc{Hubble CLIP} significantly outperforms the base CLIP model in quantitative metrics, such as retrieval accuracy, as well as quality of text-to-image and image-to-text retrieval.
+The procedure demonstrates the efficacy on fine-tuning generalist pre-trained models on small amounts of domain-specific data, in particular astronomical datasets.
 
@@
 
-\subsubsection*{Code}
+\paragraph*{Code and data availability}
 
@@
 
-This work relied on the \package{Astroquery} \citep{2019AJ....157...98G}, \package{bitsandbytes} \citep{dettmers2022llmint8}, \package{Flax} \citep{flax2020github}, \package{Jax} \citep{jax2018github}, \package{Jupyter} \citep{Kluyver2016jupyter}, \package{Matplotlib} \citep{Hunter:2007}, \package{Numpy} \citep{harris2020array}, \package{Optax} \citep{deepmind2020jax}, \package{Outlines}, \package{Pandas}, \package{Pydantic}, \package{PyTorch} \citep{paszke2019pytorch}, \package{SciPy} \citep{2020SciPy-NMeth}, \package{Transformers} \citep{wolf2019huggingface}, and \package{wandb} \citep{wandb} software packages.
+\paragraph*{Software}
+
+This work relied on the \package{Astroquery} \citep{2019AJ....157...98G}, \package{BitsAndBytes} \citep{dettmers2022llmint8}, \package{Flax} \citep{flax2020github}, \package{Jax} \citep{jax2018github}, \package{Jupyter} \citep{Kluyver2016jupyter}, \package{Matplotlib} \citep{Hunter:2007}, \package{Numpy} \citep{harris2020array}, \package{Optax} \citep{deepmind2020jax}, \package{Outlines}, \package{Pandas} \citep{2020SciPy-NMeth}, \package{Pydantic}, \package{PyTorch} \citep{paszke2019pytorch}, \package{SciPy} \citep{2020SciPy-NMeth}, \package{Transformers} \citep{wolf2019huggingface}, and \package{Wandb} \citep{wandb} software packages.
 
@@
 
-\subsubsection*{Acknowledgments}
+\paragraph*{Acknowledgments}
+
 This work is supported by the National Science Foundation under Cooperative Agreement PHY-2019786 (The NSF AI Institute for Artificial Intelligence and Fundamental Interactions, \url{http://iaifi.org/}).
@@
 
-The following prompt is used to summarize the abstracts.
+The following prompt is used to summarize the abstracts using the \package{Outlines} package interfacing with \textsc{Mixtral-8x7B-Instruct}.
 
 \begin{lstlisting}[language=Python]
+import outlines
+
 @outlines.prompt
 def prompt_fn(abstract):
-
       """<s>[INST] You are an expert astrophysicist, with broad expertise across observational and theoretical astrophysics. You are able to extract core information from astrophysical texts.
@@
 
+\section{List of Categories}
+\label{app:categories}
+
+\begin{lstlisting}[language=Python]
+  ["star forming galaxies", "lyman alpha", "dust", "crowded stellar field", "core-collapse supernova", "cosmology", "gravitational lensing", "supernovae", "diffuse galaxies", "globular clusters", "stellar populations", "interstellar medium", "black holes", "dark matter", "galaxy clusters", "galaxy evolution", "galaxy formation", "quasars", "circumstellar disks", "exoplanets", "Kuiper Belt objects", "solar system objects", "cosmic web structure", "distant galaxies", "galaxy mergers", "galaxy interactions", "star formation", "stellar winds", "brown dwarfs", "white dwarfs", "nebulae", "star clusters", "galaxy archeology", "galactic structure", "active galactic nuclei", "gamma-ray bursts", "stellar nurseries", "intergalactic medium", "dark energy", "dwarf galaxies", "barred spiral galaxies", "irregular galaxies", "starburst galaxies", "low surface brightness galaxies", "ultra diffuse galaxies", "circumgalactic medium", "intracluster medium", "cosmic dust", "interstellar chemistry", "star formation histories", "initial mass function", "stellar proper motions", "binary star systems", "open clusters", "pre-main sequence stars", "protostars", "protoplanetary disks", "jets and outflows", "interstellar shocks", "planetary nebulae", "supernova remnants", "red giants", "Cepheid variables", "RR Lyrae variables", "stellar abundances", "stellar dynamics", "compact stellar remnants", "Einstein rings", "trans-Neptunian objects", "cosmic microwave background", "reionization epoch", "first stars", "first galaxies", "high-redshift quasars", "primordial black holes", "resolved binaries", "binary stars"]
+\end{lstlisting}
+
 \section{Additional Evaluation Metrics and Ablations}
@@
 \caption{Retrieval accuracy}
-\label{fig:retrieval_acc}
+\label{fig:retrieval_acc_supp}
 \end{figure*}

### 2024-01-03 summarize.py

@@ -1,4 +1,5 @@
 import sys, os
+
 sys.path.append("../")
 
 import outlines
@@ -16,6 +17,7 @@ import torch
 from utils.summarize_utils import prompt_fn, ConstrainedResponseHST
 from utils.abstract_utils import read_abstracts_file
 
+
 def summarize(
     data_dir: str = "./data/",
     observations_dir: str = "observations_v1",
@@ -28,35 +30,22 @@ def summarize(
     verbose: bool = False,
     n_max_tries: int = 5,
 ):
-    
     if model_name == "mistralai/Mixtral-8x7B-Instruct-v0.1":
-
         config = transformers.AutoConfig.from_pretrained(
-            model_name, trust_remote_code=True, asd=True,
+            model_name,
+            trust_remote_code=True,
+            asd=True,
         )
 
-        bnb_config = BitsAndBytesConfig(
-            load_in_4bit=True,
-            bnb_4bit_use_double_quant=True,
-            bnb_4bit_quant_type="nf4",
-            bnb_4bit_compute_dtype=torch.bfloat16
-        )
+        bnb_config = BitsAndBytesConfig(load_in_4bit=True, bnb_4bit_use_double_quant=True, bnb_4bit_quant_type="nf4", bnb_4bit_compute_dtype=torch.bfloat16)
 
         model = models.transformers(
-            
             model_name=model_name,
-            model_kwargs={
-                "config": config,
-                "quantization_config": bnb_config,
-                "trust_remote_code": True,
-                "device_map": "auto",
-                "load_in_4bit": True,
-                "cache_dir": "/n/holystore01/LABS/iaifi_lab/Users/smsharma/hf_cache/"
-            },
+            model_kwargs={"config": config, "quantization_config": bnb_config, "trust_remote_code": True, "device_map": "auto", "load_in_4bit": True, "cache_dir": "/n/holystore01/LABS/iaifi_lab/Users/smsharma/hf_cache/"},
         )
-        
+
     elif model_name == "TheBloke/OpenHermes-2.5-Mistral-7B-AWQ":
-        model = models.awq(model_name)  
+        model = models.awq(model_name)
     else:
         raise ValueError("Invalid model name.")
 
@@ -65,12 +54,12 @@ def summarize(
     abstracts_df = read_abstracts_file(abstract_file)
 
     # Drop rows with missing Cycle
-    abstracts_df = abstracts_df.dropna(subset=['Cycle'])
-    abstracts_df = abstracts_df[abstracts_df['Cycle'] != '']
[…]

### 2024-01-03 utils/summarize_utils.py

@@ -6,6 +6,7 @@ from enum import Enum
 
 import re
 
+
 class ConstrainedResponseHST(BaseModel):
     # objects_and_phenomena: conlist(str, min_length=1, max_length=5)
     # science_use_cases: conlist(str, min_length=1, max_length=5)
@@ -25,32 +26,31 @@ class ConstrainedResponseHST(BaseModel):
 
     #     return v
 
+
 @outlines.prompt
 def prompt_fn(abstract):
-        
     """<s>[INST] You are an expert astrophysicist, with broad expertise across observational and theoretical astrophysics. You are able to extract core information from astrophysical texts.
 
-Abstract: "{{abstract}}"
+    Abstract: "{{abstract}}"
 
-Based on the above observational proposal abstract, your task is to summarize the nature of the eventual observations. You will identify the astrophysical objects and phenomena, as well as the potential science use cases described in the abstract.
+    Based on the above observational proposal abstract, your task is to summarize the nature of the eventual observations. You will identify the astrophysical objects and phenomena, as well as the potential science use cases described in the abstract.
 
-Follow these instructions exactly:
-- Mention up to 5 items for both categories; do not mention more than 5 items in either category. 
-- Choose the most relevant ones if there are more than 5 items in a category.
-- Never mention the Hubble Space Telescope, HST, or the HST archive.
-- Mention the class (e.g., barred spiral galaxy) and not just the specific instance (e.g., Andromeda).
-- Name the objects in the science use cases, if appropriate.
-- Write out full names of objects in addition to acronyms.
-- Do not list irrelevant objects which do not describe the eventual observation, such as units or proposal Cycle numbers. List fewer but more relevant objects, if in doubt.
-- Each science case listed must be self-contained but succinct.
-- Only write in English.
-- Do not list items that are too generic (e.g., galaxy, faint object, kinematics)
-- The total length of text should not exceed 80 words.
-- Present your lists in a comma-separated format; no dashed or numbered lists.
+    Follow these instructions exactly:
+    - Mention up to 5 items for both categories; do not mention more than 5 items in either category.
+    - Choose the most relevant ones if there are more than 5 items in a category.
+    - Never mention the Hubble Space Telescope, HST, or the HST archive.
+    - Mention the class (e.g., barred spiral galaxy) and not just the specific instance (e.g., Andromeda).
+    - Name the objects in the science use cases, if appropriate.
[…]

### 2024-01-04 paper/main.tex
More text updates

@@
 
-Machine learning (ML) is starting to have a significant impact on the sciences, with astrophysics being no exception.
-Machine learning methods have demonstrated promise when applied to every part of the astrophysics pipeline, from instrument design, to data acquisition, to its analysis.
+Machine learning (ML) is beginning to have a significant impact in the sciences, with astrophysics being no exception.
+ML methods have demonstrated promise at every stage of the astrophysics pipeline, from instrument design, to data acquisition, to its analysis \citep{huertas2022dawes}.
 Until recently, most applications of ML within astrophysics have focused on replacing or augmenting traditional techniques with ML counterparts in order to improve performance on specific tasks.
+The \emph{Foundation Model} paradigm, in contrast, seeks to develop generalist models which can be deployed on a wide range of tasks.
+The paradigm has been highly successful in domains like computer vision and natural language processing, as demonstrated by the widespread adoption of tools like ChatGPT, Dall-E, Stable Diffusion, and CLIP.
+These models are typically pre-trained on massive amounts of unlabeled data using self-supervised learning techniques, enabling them to learn powerful representations which can be deployed on different tasks across domains.
+At the heart of the paradigm lies the triumph of scale -- scaling up model size, dataset size, and compute.
+However, foundation models can often benefit from \emph{fine tuning} using a small amounts of domain data, increasing their usefulness when applied to those target domains.
 
-The \emph{Foundation Model} paradigm, in contrast, seeks to develop generalist models which can be deployed on a wide range of tasks.
-The paradigm has been highly successful in domains like computer vision and natural language processing, as demonstrated by the widespread adoption of tools like CLIP, ChatGPT, Dall-E, and Stable Diffusion.
-These models are typically pre-trained on massive amounts of unlabeled data using self-supervised learning techniques, enabling them to learn powerful representations which can be optionally fine-tuned to address domain-specific tasks.
-At the heart of the paradigm lies the triumph of scale -- scaling up model size, dataset size, and compute.
-However, foundation models often benefit from fine tuning using a relatively small amounts of domain-specific data, increasing their usefulness when applied to those specific domains.
-
-There is considerable interest in developing custom foundation models for the sciences, with astrophysics being ripe for such an effort due to several reasons.
-The first is the availability of large amounts of publicly-available data as a contingency of publicy-funded data-taking efforts.
-The second is the multi-modality inherent astrophysical observations, with different types of data (e.g., images, spectra, light curves, textual descriptions) often available for each observation.
-This multi-modality was recently exploited to train \textsc{AstroCLIP}~\citep{lanusse2023astroclip} -- a joint representation between multi-band images and optical spectra from the Dark Energy Spectroscopic Instrument (DESI).
+There is considerable interest in developing custom foundation models for the sciences \citep{batatia2023foundation,subramanian2023towards}, with astrophysics being ripe for such an effort given the availability of large amounts of publicly-available data and diverse ways of interacting with it.
+The multi-modality inherent to the nature of astrophysical observations, with different types of data (e.g., images, spectra, light curves, textual descriptions) often available, presents a unique opportunity.
+This multi-modality was recently exploited in \textsc{AstroCLIP}~\citep{lanusse2023astroclip} -- method to obtain a joint embedding space between multi-band images and optical spectra from the Dark Energy Spectroscopic Instrument (DESI).
 \textsc{AstroLLaMA}~\citep{nguyen2023astrollama} is another recent effort to fine-tune a publicly-available model (\textsc{Llama-2}) on astrophysics-specific textual data from the arXiv.
 
-The CLIP~\citep{radford2021learning} family of models has shown strong performance on a variety of downstream tasks including zero-shot classification and image retrieval.
+The Contrastive Language-Image Pre-training~\citep[CLIP;][]{radford2021learning} family of foundation models, which in its original instantiation embed images and text-via-captions into a common embedding space via contrastive learning, has shown strong performance and generalization capabilities on a variety of downstream tasks including zero-shot classification and image retrieval.
+Besides \textsc{AstroCLIP}, the concept of contrastive pre-training has been used in a variety of domains, e.g \textsc{GeoCLIP}~\citep{cepeda2023geoclip} for geospatial location querying.
 
-\textsc{GeoCLIP}~\citep{cepeda2023geoclip}.
-
-The success of the foundation model paradigm partly relies on the ability to flexibly leverage text as a \emph{universal interface}.
+The success of the foundation model paradigm partly relies on the ability to flexibly leverage text as a \emph{universal interface} for interacting with models.
 In this work, we take this outlook and train a joint representation between observations taken by the \hubble Space Telescope (HST) and natural language.
-We do so by using the associations between observation proposals and corresponding downstream observations.
-We show that fine-tuning a CLIP (Contrastive Language-Image Pre-training) model on this data enables learning meaningful joint representations.
+We do so by leveraging associations between LLM-summarized abstracts of successful observing proposals and corresponding downstream image observations.
+We show that fine-tuning a pre-trained CLIP image-text model on this data enables learning semantically meaningful joint representations through quantitative and qualitative evaluation on retrieval tasks.
 
@@
 We curate a dataset of images of \hubble observations and corresponding text descriptions.
-We rely on summarized versions of proposal abstracts from the Proposal Abstracts Catalog\footnote{\url{https://archive.stsci.edu/hst/proposal_abstracts.html}} -- a catalog of all accepted \hubble proposals -- to derive captions for the observations.
+We rely on summarized versions of proposal abstracts from the Proposal Abstracts Catalog\footnote{\url{https://archive.stsci.edu/hst/proposal_abstracts.html}} -- a catalog of all accepted \hubble proposals to-date -- to derive captions for the observations.
 The HST has been operational for 33 years, having been launched on April 24, 1990.
@@
 Products of type \texttt{PREVIEW} are filtered in, corresponding to preview postcard images.
-Note that these are not science-grade observations, but rather lower-resolution images used for quick-look purposes; given the nature of associations we aim to learn, we deem this adequate for our current purposes.
-A maximum of 20 images are downloaded per proposal ID, selected at random, in order to avoid biasing the model towards proposals with a larger number of observations.
-Images are centered and resized to a resolution per side of 512 pixels before saving.
+We note that these are not science-grade observations, but rather lower-resolution images useful for quick-look purposes; given the nature of associations we aim to learn, this is adequate for our purposes.
+A maximum of 20 images are downloaded per proposal ID, selected at random, in order to avoid biasing the model towards proposals with a larger number of observations and survey-style campaigns.
+Images are centered and resized to a resolution-per-side of 512 pixels before.
 Color previews (i.e., observations taken with multiple wavelength filters assigned to individual RGB channels) are manually excluded via a filename filter in order to maintain consistency across the dataset; models trained on datasets with color images included were observed to show worse performance on generalization metrics.
-If no appropriate images corresponding to an abstract are found, the abstract is excluded from the dataset.
+If no appropriate images corresponding to an abstract are found, is is excluded from the dataset.
 
-In total 31,859 images corresponding to 4,438 abstracts are included in the fine-tuning dataset.
-3,194 images are held out for validation, with no abstract being common between training and validation sets in order to ensure an independent set of text-image pairs for testing.
+In total, 31,859 images corresponding to 4,438 abstracts are included in the fine-tuning dataset.
+3,194 images are held out for validation, with no abstract being common between training and validation sets in order to ensure an independent set of text-image pairs for evaluation.
 
@@
 These abstracts are written in a diversity of styles, formats, and lengths, being highly variable in the nature of content as well.
-Although the abstracts can be used as-is as image captions, we explore the use of summarization via guided LLM generation to standardize the captions used for fine-tuning the CLIP model.
-The goal is to summarize the objects and phenomena, as well as potential downstream science use cases corresponding to the HST observations in order to increase the signal between text and images.
+Although the abstracts can be used as-is as image captions, for our baseline model we summarize them via guided LLM generation to standardize the captions used for fine-tuning the CLIP model.
+The goal is to summarize the objects and phenomena, as well as potential downstream science use cases corresponding to the HST observations in order to increase the strength of the association signal between text and images.
 
 The method from \cite{willard2023efficient} is used to produce an LLM-generated summary of the abstract conforming to a particular schema, specified in JSON format.
-The schema is designed to represent a list of the objects (e.g., `Type Ia supernova') and phenomena (e.g., `gravitational lensing'), as well as potential downstream science uses cases (e.g., `set constraints on supernova explosion models') that could correspond to the eventual imaged observation given the abstract text.
+The schema is designed to represent a list of the objects (e.g., `Type Ia supernova') and phenomena (e.g., `gravitational lensing'), as well as potential downstream science uses cases (e.g., `set constraints on supernova explosion models') that could correspond to the eventual imaged observation given the abstract text, with a minimum of 1 and a maximum of 5 elements per list.
 
@@
 \end{lstlisting}
-which is then used to construct the summarized caption by combining the two key elements.
-Examples of raw abstracts and corresponding LLM-generated summaries are shown in Tab.~\ref{tab:datasetsumm}.
+which is then used to construct the summarized caption by combining the two key elements. Examples of raw abstracts and corresponding LLM-generated summaries are shown in Tab.~\ref{tab:datasetsumm}.
+
+We use the open-weights, instruction-tuned model \textsc{Mixtral-8x7B-Instruct}\footnote{\url{https://huggingface.co/mistralai/Mixtral-8x7B-Instruct-v0.1}} to generate the summaries, with guided generation performed using the \package{Outlines}\footnote{\url{https://github.com/outlines-dev/outlines}} package.
 Further details on the summarization procedure, including the prompts used and a more detailed description of guided generation, are provided in App.~\ref{app:summarization}.
 
-The open-weights, instruction-tuned model \textsc{Mixtral-8x7B-Instruct}\footnote{\url{https://huggingface.co/mistralai/Mixtral-8x7B-Instruct-v0.1}} is used to generate the summaries, with guided generation performed using the \package{Outlines}\footnote{\url{https://github.com/outlines-dev/outlines}} package.
-
-The goal if summarization-via-guided-generation is to increase the signal between text and images by standardizing the captions used for fine-tuning the CLIP model.
-
-% Note that the signal is very imperfect!
+We emphasize that the goal of summarization-via-guided-generation is to increase the signal between text and images by standardizing the captions used for fine-tuning the CLIP model, and compare the quantitative performance of the model vs using the raw abstracts in Sec.\ref{sec:results}.
+We also note that, even after summarization, the association signal is expected to be noisy, since the summarized caption may not always be usefully descriptive of the observed images.
 
@@
       \end{tabular}
-      \caption{Examples of images and corresponding captions, constructed using the LLM-extracted summaries. The CLIP model is fine-tuned on these text-image associations.}
+      \caption{Examples of images and corresponding captions, constructed using the LLM-extracted summaries.
+The CLIP model is fine-tuned on these text-image associations.
+\SM{Can add another row to fill the page}}
       \label{tab:dataset}
@@
 
-Our goal is to learn a semantically  meaningful joint representation of HST image observations and natural language, with the intention of using it for a variety of downstream tasks.
-We leverage the strong generalization capabilities demonstrated by CLIP (Contrastive Language-Image Pretraining) and adapt these to work with domain-specific \hubble data via fine-tuning; we describe these below.
+Our goal is to learn a semantically  meaningful joint representation between HST image observations and natural language.
+We leverage the strong generalization capabilities demonstrated by pre-trained CLIP models and adapt these to work with domain-specific \hubble data via fine-tuning.
 
-\subsection{Language-Image Pre-training}
-\label{sec:pretraining}
+\paragraph*{Pre-trained CLIP model}
 
@@
 Given a minibatch $\mathcal{B}$ of $|\mathcal{B}|$ image-text pairs $\{(I_i, T_i)\}$, the goal is to align the learned representations of corresponding (positive) pairs $(I_i, T_i)$ while repelling the representations of unaligned (negative) pairs $(I_i, T_{j\neq i})$.
-Image and text encoders $f(\cdot)$ and $g(\cdot)$ are used to map images and text to a common embedding space.
-The standard softmax-based bidirectional variation of the InfoNCE~\citep{oord2018representation} contrastive loss function, as used by CLIP, is particularly effective for multimodal learning.
-This bidirectionality is crucial as it ensures the model learns to map both images to text and text to images with equal importance.
-This symmetry in learning is essential for tasks that require a mutual understanding and interchangeability between visual and textual representations, such as image captioning, text-to-image synthesis, and cross-modal retrieval.
-The bidirectional loss is given by \citep{radford2021learning}
+Image and text encoders $f: I \rightarrow \mathbb R^{n_\text{emb}}$ and $g: T \rightarrow \mathbb R^{n_\text{emb}}$ are used to map images and text to a common embedding space of dimension $n_\text{emb}$.
+The standard softmax-based bidirectional variation of the InfoNCE~\citep{oord2018representation} contrastive loss function was introduce for training CLIP-style architectures\citep{radford2021learning}
 \begin{equation}
-\mathcal L(\mathcal B) = -\frac{1}{2|\mathcal{B}|} \sum_{i=1}^{|\mathcal{B}|}\left(\underbrace{\log \frac{e^{{x}_i \cdot {y}_i / \tau}}{\sum_{j=1}^{|\mathcal{B}|} e^{{x}_i \cdot {y}_j / \tau}}}_{\text {image } \rightarrow \text { text softmax }}+\underbrace{\log \frac{e^{{x}_i \cdot {y}_i / \tau}}{\sum_{j=1}^{|\mathcal{B}|} e^{{x}_j \cdot {y}_i / \tau}}}_{\text {text } \rightarrow \text { image softmax }}\right)
-\label{eq:softmax_loss}
+  \mathcal{L}(\mathcal{B})=-\frac{1}{2|\mathcal{B}|} \sum_{i=1}^{|\mathcal{B}|}\left(\log \frac{e^{x_i \cdot y_i / \tau}}{\sum_{j=1}^{|\mathcal{B}|} e^{x_i \cdot y_j / \tau}}+\log \frac{e^{x_i \cdot y_i / \tau}}{\sum_{j=1}^{|\mathcal{B}|} e^{x_j \cdot y_i / \tau}}\right)\label{eq:softmax_loss}
 \end{equation}
 where ${x}_i={f\left(I_i\right)}/{\left\|f\left(I_i\right)\right\|}$ and ${y}_i={g\left(T_i\right)}/{\left\|g\left(T_i\right)\right\|}$ are the normalized representations of the $i$-th image and text, respectively, and $\tau$ is a learnable temperature hyperparameter.
-% This bidirectional approach not only enhances the model's ability to create robust and versatile embeddings but also aligns with the intuitive understanding of how humans interact with and interpret multimodal information, further bridging the gap between artificial and natural intelligence in multimodal contexts.
+Note that this loss treats the image and text representations symmetrically, ensuring that the two modalities are on the same footing.
 
-As the base model, we use the CLIP-ViT-B/16 \citep{radford2021learning} variant trained by OpenAI.
-This model uses a (12-layer, 12-head, 768-embedding dimension) vision transformer as the image encoder and a (12-layer, 8-head, 512-embedding dimension) sequence transformer as the text backbone.
+We use the CLIP-ViT-B/16 \citep{radford2021learning} variant as the base pre-trained CLIP model.
+This model uses a 12-layer, 12-head, 768-embedding dimension vision transformer as the image encoder and a 12-layer, 8-head, 512-embedding dimension sequence transformer as the text backbone.
 The text encoder has a maximum length of 77 tokens and the image encoder a native resolution of $224\times224$.
@@
 In total, the model has 149,620,737 trainable parameters.
-The model was trained on 400 million image-text pairs from the internet.
+The model was originally trained on 400 million image-text pairs from internet data.
 
-\subsection{Fine-tuning Procedure and Evaluation}
-\label{sec:finetuning}
+\paragraph*{Fine-tuning procedure}
 
@@
 
-We explore three different methods of training the model on our domain dataset: \emph{(1)} Fine-tuning the entire network, starting from the pre-trained base model; \emph{(2)} Freezing the base image and text encoders, and training a small projection head on top of these; and \emph{(3)} Training the entire model from scratch.
-For \emph{(2)}, we use a 2-layer MLP with 1024 units and a GELU activation layer, projecting to the 512-dimensional embedding space.
+We explore three different methods of training the model on our domain dataset: \emph{(1)} Fine-tuning the entire network, starting from the pre-trained base model; \emph{(2)} Freezing the base image and text encoders, and training a small projection head; and \emph{(3)} Training the entire model from scratch.
+For \emph{(2)}, we use a 2-layer MLP with 1024 units and a GELU activation layer, projecting onto the 512-dimensional embedding space.
 
-Models were instantiated and trained using the \package{Jax} \citep{jax2018github} ecosystem. The model is trained over 20,000 steps with 2000 linear warmup steps 
+Models were instantiated and trained using the \package{Jax} \citep{jax2018github} ecosystem. Models were trained over 20,000 steps with 2000 linear warmup steps 
 % and cosine decay 
-using the AdamW optimizer \citep{DBLP:conf/iclr/LoshchilovH19,DBLP:journals/corr/KingmaB14} with peak learning rate of $10^{-5}$ and weight decay $10^{-3}$.
-Training takes approximately 3 hours on 4 Nvidia A100 GPUs.
+using the AdamW optimizer \citep{DBLP:conf/iclr/LoshchilovH19,DBLP:journals/corr/KingmaB14} with  %peak 
+learning rate of $10^{-5}$ and weight decay $10^{-3}$. Training takes approximately 3 hours on 4 Nvidia A100 GPUs.
+
+\paragraph*{Evaluation}
 
 The model is evaluated by tracking the loss in Eq.~\ref{eq:softmax_loss} as well as the top-$k\%$ retrieval accuracy on the held out validation set over the course of training.
-The retrieval accuracy is defined as the fraction of associated captions which fall within the top $k\%$ of captions by cosine similarity of the (normalized) embeddings $x_i \cdot y_j$, averaged over the images in the validation set.
+The retrieval accuracy is defined as the fraction of associated captions which fall within the top $k\%$ of captions by cosine similarity of the (normalized) embeddings $(x_i \cdot y_j)$, averaged over the images in the validation set.
+
+We also qualitatively evaluate the learned embeddings through image retrieval (i.e., retrieving the most relevant images from a set using natural language queries) and description retrieval (i.e., querying the astrophysical object classes and science use cases most relevant to a given observation) experiments.
 
@@
 
-Fig.~\ref{fig:retrieval_acc} shows the contrastive loss (left) and the top-10\% retrieval accuracy on the held out validation set over the course of training, for different variations considered.
+Figure~\ref{fig:retrieval_acc} shows the contrastive loss (left) and the top-10\% retrieval accuracy on the held out validation set over the course of training, for different variations considered.
 The red lines show the metrics evaluated when training with batches where the image-text associations are randomly shuffled, serving as a baseline.
-This baseline is seen to do on par with random expectation, unlike the others, validating the presence of a positive image-text association signal in the dataset.
+This baseline is seen to do on par with random expectation, unlike the others, validating the presence of a significant association signal in the dataset.
 %
 Interestingly, the base model performs better than random expectation, with a top-10\% retrieval accuracy of $\sim 15\%$.
-We will therefore compare the qualitative performance of the base model with the fine-tuned model on downstream retrieval tasks.
+We therefore compare the qualitative performance of the base model with the fine-tuned model on downstream retrieval tasks.
 
-The fiducial model with LLM-guided summarization (orange lines) is seen to perform significantly better than the model using raw abstracts as captions (purple line), validating the stronger association signal in the summarized dataset we curate.
-Fine-tuning a small MLP head over frozen vision and text backbones (green lines) and training from scratch (blue lines) show a non-trivial improvement in the retrieval metrics compared to the random baseline as well as base model, but with deteriorated performance compared to the fiducial set-up.
+The fiducial model with LLM-guided summarization (orange lines) is seen to perform significantly better than the model using raw abstracts as captions (purple line), validating the stronger association signal in the summarized dataset curated.
+Fine-tuning a small MLP head over frozen vision and text backbones (green lines) and training from scratch (blue lines) show a non-trivial improvement compared to the random baseline as well as base model, but with deteriorated performance compared to the fiducial set-up.
 
@@
 Having aligned the image and text representations, we can query the validation set using natural language and show the most ``relevant'' images when ranked by cosine similarity.
-We show this in Figs.~\ref{fig:tti_base} and \ref{fig:tti} for the base model and fine-tuned model respectively using four simple curated queries: `globular clusters', `dwarf galaxy', `SN1987A', and `cluster lensing'.
+We show these in Figs.~\ref{fig:tti_base} and \ref{fig:tti} for the base model and fine-tuned model respectively using four simple curated queries: `globular clusters', `dwarf galaxy', `SN1987A' (a specific, flagship supernova), and `cluster lensing'.
 
@@
 
-The fine-tuned model (Fig.~\ref{fig:tti}) shows strikingly different behavior.
-For example, it is able to return images with processing and assembly artifacts particular to the telescope (the lines through the middle), which typically receive low similarity scores with the base model.
+The fine-tuned model (Fig.~\ref{fig:tti}), meanwhile, shows strikingly different behavior.
+For example, it is able to return images with processing and assembly artifacts particular to HST (e.g., the lines through the middle in some images), which typically receive low similarity scores when using the base model.
 The `dwarf galaxy' images correspond to proposals aiming to measure the kinematics of the stellar cores of dwarf galaxies.
@@
 
-We present \textsc{HubbleCLIP}, a multi-modal fountation model that associates observations imaged by the \hubble Space Telescope with natural language in a common embedding space.
+We present \textsc{HubbleCLIP}, a multi-modal fountation model that associates observations imaged by the \hubble Space Telescope with natural language in a common, semantically-meaningful embedding space.
 The model is fine-tuned from a pre-trained CLIP model on LLM-summarized versions of \hubble proposal abstracts, leveraging a noisy signal associating text and images.
-We show that \textsc{Hubble CLIP} significantly outperforms the base CLIP model in quantitative metrics, such as retrieval accuracy, as well as quality of text-to-image and image-to-text retrieval.
+We show that \textsc{HubbleCLIP} significantly outperforms the base CLIP model in quantitative metrics, such as retrieval accuracy, as well as quality of text-to-image and image-to-text retrieval.
 The procedure demonstrates the efficacy on fine-tuning generalist pre-trained models on small amounts of domain-specific data, in particular astronomical datasets.
@@
 
+The following schema is used to guide the generation of the summaries.
+
+\begin{lstlisting}[language=Python]
+from pydantic import BaseModel, conlist
+
+class ConstrainedResponseHST(BaseModel):
+      objects_and_phenomena: conlist(str, min_length=1, max_length=5)
+      science_use_cases: conlist(str, min_length=1, max_length=5)
+\end{lstlisting}
+
 \section{List of Categories}

### 2024-01-11 configs/base.py
Added sum1 functionality

@@ -52,6 +52,12 @@ def get_config():
     training.ckpt_best_metric_best_mode = "min"  # "max" or "min" 
     training.ckpt_keep_top_n = 3  # Save the top `ckpt_keep_top_n` checkpoints based on `ckpt_best_metric`
 
+    # Sum1 options
+    config.sum1 = sum1 = ml_collections.ConfigDict()
+    sum1.use_sum1 = True
+    sum1.summaries_filename = "summary_v2"
+    sum1.sum1_filename = "summary_sum1_v2"
+
     # Optimizer (AdamW)
     config.optim = optim = ml_collections.ConfigDict()
     optim.schedule = "constant"

### 2024-01-11 train.py

@@ -11,6 +11,8 @@ from ml_collections import config_flags, ConfigDict
 from clu import metric_writers
 import wandb
 
+import pandas as pd
+
 import jax
 import jax.numpy as np
 import optax
@@ -73,6 +75,9 @@ def train(config: ConfigDict, workdir: str = "./logging/") -> train_state.TrainS
 
     batches = iter(train_ds)
 
+    # Convert to jnp type
+    dtype = getattr(np, config.clip.dtype)
+
     logging.info("Loaded the dataset")
 
     # # Model configs
@@ -83,7 +88,7 @@ def train(config: ConfigDict, workdir: str = "./logging/") -> train_state.TrainS
 
     # Use pre-trained model or train from scratch
     if config.clip.use_pretrained:
-        model = FlaxCLIPModel.from_pretrained(config.clip.pretrained_model_name, dtype=config.clip.dtype)
+        model = FlaxCLIPModel.from_pretrained(config.clip.pretrained_model_name, dtype=dtype)
         processor = AutoProcessor.from_pretrained(config.clip.pretrained_model_name)
     else:
         raise NotImplementedError
@@ -127,7 +132,7 @@ def train(config: ConfigDict, workdir: str = "./logging/") -> train_state.TrainS
 
     if config.clip.transfer_head:
 
-        model_transfer = FlaxCLIPModelTransfer(config=model.config, dtype=config.clip.dtype, d_head=config.clip.d_transfer_head)
+        model_transfer = FlaxCLIPModelTransfer(config=model.config, dtype=dtype, d_head=config.clip.d_transfer_head)
         
         # Transfer text and vision backbones
         model_transfer.params['text_model']['text_backbone'] = model.params['text_model']
@@ -193,6 +198,21 @@ def train(config: ConfigDict, workdir: str = "./logging/") -> train_state.TrainS
     if config.data.shuffle_within_batch:
         logging.info(f"Shuffling images within batch")
 
+    # If matching to sum1 summary
+    if config.sum1.use_sum1:
+        logging.info(f"Matching to summary {config.sum1.summaries_filename} with sum1 {config.sum1.sum1_filename}")
+        # Combine with data dir and add .csv extension
+        summaries_filename = os.path.join(config.data.data_dir, f"{config.sum1.summaries_filename}.csv")
+        sum1_filename = os.path.join(config.data.data_dir, f"{config.sum1.sum1_filename}.csv")
+
+        # Open dataframes
[…]

### 2024-01-11 utils/text_utils.py

@@ -1,13 +1,27 @@
 import nltk
 import jax
+import pandas as pd
 
-def process_truncate_captions(captions, key, max_length_words=None):
+
+def process_truncate_captions(captions, key, max_length_words=None, use_sum1=False, df_sum_merged=None):
     """ Process and truncate captions
     """
     captions = captions.numpy().tolist()
     captions = [c.decode("utf-8") for c in captions]
+
+    def get_objects_phenomena(caption):
+        first_part = caption.split(';')[0]
+        match = df_sum_merged[df_sum_merged['objects_phenomena_x'] == first_part]['objects_phenomena_y']
+        return match.values[0] if not (match.empty or pd.isna(match.values[0])) else "None"
+
+    # Get sum1 summaries by matching
+    if use_sum1 and df_sum_merged is not None:
+        results = pd.Series(captions).apply(get_objects_phenomena)
+        captions = list(results.values)
+
     if max_length_words is not None:
         captions = sample_and_pad(captions, key, max_length=max_length_words)
+
     return captions
 
 

### 2024-01-11 configs/base.py
Launch a run with sum1

@@ -54,7 +54,7 @@ def get_config():
 
     # Sum1 options
     config.sum1 = sum1 = ml_collections.ConfigDict()
-    sum1.use_sum1 = True
+    sum1.use_sum1 = False
     sum1.summaries_filename = "summary_v2"
     sum1.sum1_filename = "summary_sum1_v2"
 

### 2024-01-11 scripts/submit_train.sh

@@ -2,7 +2,7 @@
 
 #SBATCH --job-name=train
 #SBATCH --nodes=1
-#SBATCH --mem=200GB
+#SBATCH --mem=100GB
 #SBATCH --time=48:00:00
 #SBATCH --gres=gpu:4
 #SBATCH --account=iaifi_lab
@@ -27,7 +27,7 @@ alias jupyter=/n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/jupyter
 
 cd /n/holystore01/LABS/iaifi_lab/Users/smsharma/multimodal-data/
 
-/n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py  # Base config, with summary captions
+# /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py  # Base config, with summary captions
 # /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.optim.learning_rate=1e-6  # Lower LR
 
 # /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.clip.transfer_head=True  # Fine-tune just head
@@ -37,3 +37,5 @@ cd /n/holystore01/LABS/iaifi_lab/Users/smsharma/multimodal-data/
 # /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.training.loss_type="sigmoid"
 # /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.data.shuffle_within_batch=True  # Shuffle within batch
 # /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.training.batch_size=64  # Larger batch size
+
+/n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py  --config.sum1.use_sum1=True  # Base config, with summary captions

### 2024-01-11 scripts/submit_train.sh
Model train on Sum1 gives worse dot products

@@ -2,7 +2,7 @@
 
 #SBATCH --job-name=train
 #SBATCH --nodes=1
-#SBATCH --mem=100GB
+#SBATCH --mem=200GB
 #SBATCH --time=48:00:00
 #SBATCH --gres=gpu:4
 #SBATCH --account=iaifi_lab

### 2024-01-12 scripts/submit_train.sh
Update dot product plots with a smaller run; ran smaller number of steps

@@ -38,4 +38,5 @@ cd /n/holystore01/LABS/iaifi_lab/Users/smsharma/multimodal-data/
 # /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.data.shuffle_within_batch=True  # Shuffle within batch
 # /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.training.batch_size=64  # Larger batch size
 
-/n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py  --config.sum1.use_sum1=True  # Base config, with summary captions
+# /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py  --config.sum1.use_sum1=True  # Base config, with summary captions
+/n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py  --config.training.n_train_steps=5000  # Train for shorter

### 2024-01-12 train.py

@@ -211,7 +211,8 @@ def train(config: ConfigDict, workdir: str = "./logging/") -> train_state.TrainS
 
         # Merge on proposal ID
         df_sum_merged = pd.merge(df_summaries, df_sum1, on='proposal_id')
-
+    else:
+        df_sum_merged = None
 
     logging.info("Starting training...")
 

### 2024-01-14 paper/main.tex
Added dot-product plots to paper

@@
 The model is evaluated by tracking the loss in Eq.~\ref{eq:softmax_loss} as well as the top-$k\%$ retrieval accuracy on the held out validation set over the course of training.
-The retrieval accuracy is defined as the fraction of associated captions which fall within the top $k\%$ of captions by cosine similarity of the (normalized) embeddings $(x_i \cdot y_j)$, averaged over the images in the validation set.
+The retrieval accuracy is defined as the fraction of associated captions which fall within the top $k\%$ of captions by cosine similarity of the (normalized) embeddings $x_i \cdot y_j$, averaged over the images in the validation set.
 
@@
 
+\paragraph*{Cosine similarity distribution}
+
+The top row of Figure~\ref{fig:sim_valtrain} shows the distribution of cosine similarities between corresponding image and text embeddings, $x_i$ and $y_i$, for the LLM-summarized abstracts using the fine-tuned CLIP model (red line), with the ordering of text embeddings shuffled (dashed red line), and for the base CLIP model (blue line), shown for the set of validation pairs (top left) and same number of training set pairs (top right).
+The distribution for the base model is seen to be sharply peaked at a specific value, showing little diversity and being
+
 \begin{figure*}[!h]
@@
 \label{fig:itt}
+\end{figure*}
+
+\begin{figure*}[!h]
+\includegraphics[width=0.95\textwidth]{plots/sim_valtrain.pdf}
+\centering\includegraphics[width=0.45\textwidth]{plots/sim_summ1.pdf}
+\caption{Distribution of cosine similarities between corresponding image and text embeddings, $x_i$ and $y_i$.
+(Top row) For the LLM-summarized abstracts using the fine-tuned CLIP model (red line), with the ordering of text embeddings shuffled (dashed red line), and for the base CLIP model (blue line), shown for (top left) the set of validation pairs and (top right) same number of training set pairs.
+(Bottom) For the single-concept summaries, when evaluated using the CLIP model fine-tuned on LLM-summarized abstracts (red) and fine-tuned on the single-concept summaries themselves.}
+\label{fig:sim_valtrain}
 \end{figure*}

### 2024-01-14 paper/main.tex
Described results with single-concept summaries

--- 2024-01-04 paper/main.tex
@@
 While it can often return the nature of objects imaged, we observe it to seldom return scientific phenomena (e.g., `dark matter' as successfully done by the fine-tuned model in the second row).
-The third row (supernova 1987A) highlights interesting behavior -- the base model erroneously attributes the object at the center of the image to a gravitational lens or protoplanetary disk, while the fine-tuned model correctly identifies it as a supernova remnant (which play a crucial role for interstellar chemistry -- another returned snippet).
+The third row (supernova 1987A) highlights interesting behavior -- the base model erroneously attributes the object at the center of the image to a gravitational lens or protoplanetary disk, while the fine-tuned model correctly identifies it as a supernova remnant (which play a crucial role for interstellar chemistry -- another retrieved text snippet).
+
+\paragraph*{Cosine similarity distribution}
+
+The top row of Fig.~\ref{fig:sim_valtrain} shows the distribution of cosine similarities between corresponding image and text embeddings, $x_i$ and $y_i$, for the LLM-summarized abstracts using the fine-tuned CLIP model (red line), and for the base CLIP model (blue line), shown for the set of validation pairs (top left) and same number of training set pairs (top right).
+Distributions shuffling the order of text embeddings, randomizing the image-text correspondence, is shown as dashed lines.
+The distribution for the base model is seen to be sharply peaked at a specific value, showing little diversity and being very similar between the shuffled (dashed blue) and non-shuffled (solid blue) versions.
+Histograms for the fine-tuned model show a clear separation between the shuffled and corresponding text-image pair versions.
+
+Distributions for a set of training samples (top right) are visible more peaked towards higher values compared to those evaluated on validation samples (top left), indicating potential distribution shift and/or a degree of overfitting.
+
+Finally, in Fig.~\ref{fig:sim_valtrain} (bottom) we show the distribution of corresponding image-text pair cosine similarities for the single-concept summaries, evaluated using the CLIP model fine-tuned on summarized abstracts (red) and on the single-concept summaries themselves (purple).
+These distributions are peaked towards only marginally higher values compared to the baseline shuffled version, indicating that assigning $\mathcal O(100)$ labels to the images results in a relatively weak association signal compared to using noisy captions.
 
@@
 \label{fig:itt}
+\end{figure*}
+
+\begin{figure*}[!h]
+\includegraphics[width=0.95\textwidth]{plots/sim_valtrain.pdf}
+\centering\includegraphics[width=0.45\textwidth]{plots/sim_summ1.pdf}
+\caption{Distribution of cosine similarities between corresponding image and text embeddings, $x_i$ and $y_i$.
+(Top row) For the LLM-summarized abstracts using the fine-tuned CLIP model (red line), and for the base CLIP model (blue line), shown for (top left) the set of validation pairs and (top right) same number of training set pairs.
+Versions with the ordering of text embeddings shuffled are shown in dashed.
+(Bottom) For the single-concept summaries, when evaluated using the CLIP model fine-tuned on LLM-summarized abstracts (red) and fine-tuned on the single-concept summaries themselves (purple).}
+\label{fig:sim_valtrain}
 \end{figure*}

### 2024-01-14 paper/main.tex
Added prompts to app, more text updates

@@
 \end{lstlisting}
-which is then used to construct the summarized caption by combining the two key elements. Examples of raw abstracts and corresponding LLM-generated summaries are shown in Tab.~\ref{tab:datasetsumm}.
+which is then used to construct the summarized caption by combining the two key elements.
+Examples of raw abstract snippets and corresponding LLM-generated summaries are shown in Tab.~\ref{tab:datasetsumm}.
+See App.~\ref{app:guided-generation} for a more detailed description of this guidance generation process.
 
-We use the open-weights, instruction-tuned model \textsc{Mixtral-8x7B-Instruct}\footnote{\url{https://huggingface.co/mistralai/Mixtral-8x7B-Instruct-v0.1}} to generate the summaries, with guided generation performed using the \package{Outlines}\footnote{\url{https://github.com/outlines-dev/outlines}} package.
-Further details on the summarization procedure, including the prompts used and a more detailed description of guided generation, are provided in App.~\ref{app:summarization}.
+We use the open-weights, instruction-tuned model \textsc{Mixtral-8x7B-Instruct}~\citep{jiang2024mixtral} to generate the summaries, with guided generation performed using the \package{Outlines}\footnote{\url{https://github.com/outlines-dev/outlines}} package.
+Further details on the summarization procedure, including the prompts and schema used, are provided in App.~\ref{app:summarization}.
 
-We emphasize that the goal of summarization-via-guided-generation is to increase the signal between text and images by standardizing the captions used for fine-tuning the CLIP model, and compare the quantitative performance of the model vs using the raw abstracts in Sec.\ref{sec:results}.
+We emphasize that the goal of summarization-via-guided-generation is to increase the signal between text and images by standardizing the captions used for fine-tuning the CLIP model, and compare the quantitative performance of the model vs using the raw abstracts in Sec.~\ref{sec:results}.
 We also note that, even after summarization, the association signal is expected to be noisy, since the summarized caption may not always be usefully descriptive of the observed images.
+
+Finally, in order to test a further compression of the proposal abstracts, we use LLM-guided generation to also produce a list of single-concept summaries (e.g., `irregular galaxy', `quasars', `Galactic bulge', \ldots), and test whether this can lead to meaningful, generalizable learned associations with observations.
+The prompts and schemata for generating and assigning these are described in Apps.~\ref{app:singleconcept} and \ref{app:singleconceptassignments} respectively.
 
@@
 Finally, in Fig.~\ref{fig:sim_valtrain} (bottom) we show the distribution of corresponding image-text pair cosine similarities for the single-concept summaries, evaluated using the CLIP model fine-tuned on summarized abstracts (red) and on the single-concept summaries themselves (purple).
-These distributions are peaked towards only marginally higher values compared to the baseline shuffled version, indicating that assigning $\mathcal O(100)$ labels to the images results in a relatively weak association signal compared to using noisy captions.
+These distributions are peaked towards only marginally higher values compared to a baseline shuffled version, indicating that assigning $\mathcal O(100)$ labels to the images using an LLM results in a relatively weak association signal compared to using noisy captions.
 
@@
 
-The code, dataset, and models used in this work is available at \url{https://www.github.com/smsharma/HubbleCLIP}.
+The code, dataset, and fine-tuned models used in this work is available at \url{https://www.github.com/smsharma/HubbleCLIP}.
 
@@
 
+We thank Michael Brenner for helpful conversations.
 This work is supported by the National Science Foundation under Cooperative Agreement PHY-2019786 (The NSF AI Institute for Artificial Intelligence and Fundamental Interactions, \url{http://iaifi.org/}).
@@
 \section{Summarization via Regex-Guided Generation}
+\label{app:guided-generation}
+
+\SM{Technical description of regex-guided generation.}
+
+\section{Prompts and Schema Used for Summarization}
+\label{app:prompts}
+
+We list here the prompts and schemata (i.e.
+desired output formats) used at various stages for guided text generation via \package{Outlines} package interfacing with the \textsc{Mixtral-8x7B-Instruct} LLM.
+
+\subsection{Abstract summarization into objects, phenomena, and science use cases}
 \label{app:summarization}
 
-The following prompt is used to summarize the abstracts using the \package{Outlines} package interfacing with \textsc{Mixtral-8x7B-Instruct}.
+The following prompt is used to produce a list of possible objects and phenomena shown in HST observations downstream of a proposal abstract, as well as one to five possible science use cases.
 
@@
 
-The following schema is used to guide the generation of the summaries.
+The following schema is used to guide the generation of the summaries, intended to produce between one and five objects and hypotheses, as well as science use cases.
 
@@
 
-\section{List of Categories}
+\subsection{Generation of single-concept summaries}
+\label{app:singleconcept}
+
+The following prompt is used to generate a list of diverse single-concept summaries informed by a list of summarized abstracts:
+
+\begin{lstlisting}[language=Python]
+import outlines
+
+@outlines.prompt
+def prompt_fn(objects):
+    """<s>[INST] Please produce a list of around concepts characterizing prominent objects, phenomena, and science use cases of images observed by the Hubble Space Telescope.
+
+Here are some examples of objects:
+
+{{objects}}
+
+Follow these instructions exactly in your answer:
+- Do not output empty strings as elements.
+- Make sure that the list covers a diverse range of astronomical concepts, with items as different from each other as possible. 
+- Do not give specific names of objects, to make sure you span the widest possible range of concepts (e.g., "dwarf galaxy" is allowed, but NOT "Fornax", "Terzan 5", or  "NGC6440").
+- Do not return terms undescriptive of observations, e.g. "sloshing", "adiabatic", "interactions". Returning concrete physics objects, concepts, or phenomena.
+- Only output scientifically meaningful terms. E.g., NO "Cosmic Dance".
+- Do not duplicate entries. Do not reference any telescopes, observatories, or surveys.
+- Do not include units like "angular diameter distance", "parsec", or any other concepts that will not correlate with images of observations.
+- Use the above example list of objects only as inspiration to infer broad classes of objects.
+- Make sure each concept is succint, never more than 5 words.
+- Answer in JSON format.
+- The JSON should have the following keys {"galaxies", "stellar_physics", "exoplanets_planet_formation", "stellar_populations", "supermassive_black_holes", "solar_system", "integalactic_medium", "large_scale_structure"} reflecting rough observation categories.
+- Each category will have a list of objects and/or astronomical concepts.
+- Output up to 20 items and no more in each category
+[/INST]
+"""
+\end{lstlisting}
+
+The following schema guides generation, intended to produce 100 concepts reflecting the science categories of successful proposals\footnote{\url{https://www.stsci.edu/contents/newsletters/2023-volume-40-issue-02/hubble-cycle-31-proposal-selection}}.
+
+\begin{lstlisting}[language=Python]
+from pydantic import BaseModel, conlist
+
+class ScienceCategoriesHST(BaseModel):
+    galaxies: conlist(str, min_length=15, max_length=15)
+    stellar_physics: conlist(str, min_length=15, max_length=15)
+    exoplanets_planet_formation: conlist(str, min_length=15, max_length=15)
+    stellar_populations: conlist(str, min_length=10, max_length=10)
+    supermassive_black_holes: conlist(str, min_length=15, max_length=15)
+    solar_system: conlist(str, min_length=10, max_length=10)
+    integalactic_medium: conlist(str, min_length=10, max_length=10)
+    large_scale_structure: conlist(str, min_length=10, max_length=10)
+  \end{lstlisting}
+
+The output is then constrained to be one of the fixed number of concepts.
+
+\subsection{Assignment of abstracts to single-concept categories}
+\label{app:singleconceptassignments}
+
+Finally, the following prompt is used to assign concepts inferred in \ref{app:singleconcept} to each abstract:
+
+\begin{lstlisting}[language=Python]
+import outlines
+
+@outlines.prompt
+def prompt_fn(abs, cats):
+    """<s>[INST] The following is a successful proposal abstract for the Hubble Space Telescope: "{{abs}}"
+
+The following is a list of categories (astronomical concepts) that this abstract could correspond to.
+
+{{cats}}
+
+Please answer which of these listed concepts best describes this proposal, based on the objects and phenomena mentioned in the abstract.
+The concept should meaningfully be present in the abstract and the eventual observation.
+
+- For example, "The locations of supernovae {SNe} in the local stellar and gaseous environment in galaxies, as measured in high spatial resolution WFPC2 and ACS images, contain important clues to their progenitor stars." should return "supernova".
+- If the abstract centers calibration and/or instrumentation efforts, return calibration or instrumention".
+
+If no concept make sense, return "None". [/INST]
+"""
+\end{lstlisting}
+
+\section{List of Categories for Text Retrieval Task}
 \label{app:categories}
+
+The following curated categories are used in the text retrieval experiment in Sec.~\ref{sec:results}.
+These are intended to be broader than the single-concept summaries inferred in Sec.~\ref{app:singleconcept} and are derived by prompting \textsc{Claude}, without any external input e.g.
+proposal abstracts.
+\SM{Try to just use the single-concept summaries here to avoid the extra step}.
 
@@
 
-\section{Additional Evaluation Metrics and Ablations}
-\label{app:ablations}
+% \section{Additional Evaluation Metrics and Ablations}
+% \label{app:ablations}
 
-\begin{figure*}[!h]
-\includegraphics[width=0.95\textwidth]{plots/retrieval_acc.pdf}
-\caption{Retrieval accuracy}
-\label{fig:retrieval_acc_supp}
-\end{figure*}
+% \begin{figure*}[!h]
+% \includegraphics[width=0.95\textwidth]{plots/retrieval_acc.pdf}
+% \caption{Retrieval accuracy}
+% \label{fig:retrieval_acc_supp}
+% \end{figure*}
 

### 2024-01-14 paper/main.tex
More writing

@@
 We present \textsc{HubbleCLIP}, a multi-modal fountation model that associates observations imaged by the \hubble Space Telescope with natural language in a common, semantically-meaningful embedding space.
-The model is fine-tuned from a pre-trained CLIP model on LLM-summarized versions of \hubble proposal abstracts, leveraging a noisy signal associating text and images.
+The model is fine-tuned from a pre-trained CLIP model on summarized versions of \hubble proposal abstracts, leveraging a noisy association signal between text and images.
 We show that \textsc{HubbleCLIP} significantly outperforms the base CLIP model in quantitative metrics, such as retrieval accuracy, as well as quality of text-to-image and image-to-text retrieval.
-The procedure demonstrates the efficacy on fine-tuning generalist pre-trained models on small amounts of domain-specific data, in particular astronomical datasets.
+We also introduce a novel LLM summarization process which leverages constrained generation to significantly increase the association strength between images and captions while preserving salient information.
+Overall, the procedure demonstrates the efficacy on fine-tuning generalist pre-trained models on small amounts of domain-specific data, in particular astronomical datasets.
+
+Although the model explored here is fine-tuned using postage stamp images (i.e., preview-quality and not science-grade data), we highlight potential immediate as well as downstream use cases of this paradigm.
+In particular, a robust model trained using weakly-supervised image-text pairs can be used to query survey data e.g., PHANGS~\citep{lee2022phangs}, COSMOS~\citep{scoville2007cosmic} using natural language, as well as to quickly find distinguishing patterns in such data that may not be apparent using specialized models or manual inspection.
+The learned representations, having shown to correlate with physical characteristics of imaged objects, can also be fine-tuned via transfer learning to adapt to specific tasks~\citep[e.g., classification as in][]{wei2020deep}.
+Finally, while the CLIP model is restricted to retrieving nearest-neighbour associations within and across text/image modalities, the learned embeddings can be used as a starting point for training or fine-tuning multi-modal large-language models for interacting with survey data and receiving responses in natural language form, as well as grounding the responses based on known associations.
 

### 2024-01-15 configs/base.py
Produced gpt4-aided sum1

@@ -56,7 +56,7 @@ def get_config():
     config.sum1 = sum1 = ml_collections.ConfigDict()
     sum1.use_sum1 = False
     sum1.summaries_filename = "summary_v2"
-    sum1.sum1_filename = "summary_sum1_v2"
+    sum1.sum1_filename = "summary_sum1_v3"
 
     # Optimizer (AdamW)
     config.optim = optim = ml_collections.ConfigDict()

### 2024-01-15 scripts/submit_train.sh

@@ -38,5 +38,5 @@ cd /n/holystore01/LABS/iaifi_lab/Users/smsharma/multimodal-data/
 # /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.data.shuffle_within_batch=True  # Shuffle within batch
 # /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.training.batch_size=64  # Larger batch size
 
-# /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py  --config.sum1.use_sum1=True  # Base config, with summary captions
-/n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py  --config.training.n_train_steps=5000  # Train for shorter
+# /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py  --config.sum1.use_sum1=True --config.sum1.sum1_filename="summary_sum1_v3" # Base config, with summary captions
+/n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py  --config.training.n_train_steps=5000 --config.optim.learning_rate=1e-6  # Train for shorter

### 2024-01-15 paper/main.tex

@@
 Color previews (i.e., observations taken with multiple wavelength filters assigned to individual RGB channels) are manually excluded via a filename filter in order to maintain consistency across the dataset; models trained on datasets with color images included were observed to show worse performance on generalization metrics.
-If no appropriate images corresponding to an abstract are found, is is excluded from the dataset.
+If no appropriate images corresponding to an abstract are found, it is excluded from the dataset.
 
@@
 3,194 images are held out for validation, with no abstract being common between training and validation sets in order to ensure an independent set of text-image pairs for evaluation.
+
+We note that some fraction of the image-caption pairs in the constructed dataset will center instrumentation and/or calibration rather than primary scientific content.
+We choose to not filter out these pairs from our dataset, in order to have a larger sample of HST observations for the image encoder to leverage for adapting to the distinctive characteristics of Hubble images.
 

### 2024-01-16 scripts/submit_train.sh
Eval makes more sense with shuffled val set

@@ -32,11 +32,11 @@ cd /n/holystore01/LABS/iaifi_lab/Users/smsharma/multimodal-data/
 
 # /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.clip.transfer_head=True  # Fine-tune just head
 # /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.optim.schedule="cosine" # Cosine lr schedule
-# /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.data.caption_type="abstract"  --config.data.augment_subsample_text=True  # Full abstract
+/n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.data.caption_type="abstract"  --config.data.augment_subsample_text=True  # Full abstract
 # /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.clip.random_init_text=True --config.clip.random_init_vision=True  # From scratch
 # /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.training.loss_type="sigmoid"
 # /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.data.shuffle_within_batch=True  # Shuffle within batch
 # /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.training.batch_size=64  # Larger batch size
 
 # /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py  --config.sum1.use_sum1=True --config.sum1.sum1_filename="summary_sum1_v3" # Base config, with summary captions
-/n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py  --config.training.n_train_steps=5000 --config.optim.learning_rate=1e-6  # Train for shorter
[…]

### 2024-01-16 train.py

@@ -71,7 +71,7 @@ def train(config: ConfigDict, workdir: str = "./logging/") -> train_state.TrainS
     train_ds = make_dataloader(files_train, batch_size=config.training.batch_size, seed=config.seed, split='train', caption_type=config.data.caption_type, shuffle=True)
 
     files_val = tf.io.gfile.glob(f"{config.data.data_dir}/{config.data.tfrecords_dir}/*val*.tfrecord")
-    val_ds = make_dataloader(files_val, batch_size=config.training.batch_size_val, seed=config.seed, split='val', caption_type=config.data.caption_type, shuffle=False)
+    val_ds = make_dataloader(files_val, batch_size=config.training.batch_size_val, seed=config.seed, split='val', caption_type=config.data.caption_type, shuffle=True)
 
     batches = iter(train_ds)
 

### 2024-01-16 configs/base.py
Fix evals, new run order

@@ -47,7 +47,6 @@ def get_config():
     training.log_every_steps = 100
     training.eval_every_steps = 200
     training.loss_type = "softmax"  # "sigmoid" or "softmax"
-    training.n_eval_batches = 20  # How many batches to use for evaluation
     training.ckpt_best_metric = "loss"  # "loss" or "top_xx_accuracy"
     training.ckpt_best_metric_best_mode = "min"  # "max" or "min" 
     training.ckpt_keep_top_n = 3  # Save the top `ckpt_keep_top_n` checkpoints based on `ckpt_best_metric`

### 2024-01-16 scripts/submit_train.sh

@@ -27,16 +27,19 @@ alias jupyter=/n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/jupyter
 
 cd /n/holystore01/LABS/iaifi_lab/Users/smsharma/multimodal-data/
 
+# # Core runs
 # /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py  # Base config, with summary captions
-# /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.optim.learning_rate=1e-6  # Lower LR
-
 # /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.clip.transfer_head=True  # Fine-tune just head
-# /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.optim.schedule="cosine" # Cosine lr schedule
-/n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.data.caption_type="abstract"  --config.data.augment_subsample_text=True  # Full abstract
-# /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.clip.random_init_text=True --config.clip.random_init_vision=True  # From scratch
-# /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.training.loss_type="sigmoid"
-# /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.data.shuffle_within_batch=True  # Shuffle within batch
-# /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.training.batch_size=64  # Larger batch size
 
+# /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.data.caption_type="abstract"  --config.data.augment_subsample_text=True  # Full abstract
[…]

### 2024-01-16 train.py

@@ -234,9 +234,16 @@ def train(config: ConfigDict, workdir: str = "./logging/") -> train_state.TrainS
                 val_metrics = []
                 val_batches = iter(val_ds)
 
+                total_batches = sum(1 for _ in val_ds) - 1
+                current_batch = 0
+
                 # Validate on 10 batches
-                for _ in range(config.training.n_eval_batches):
-                    images, captions = next(val_batches)
+                for (images, captions) in val_batches:
+
+                    # Break if we've reached the end of the dataset (except for the last batch, which is likely partial)
+                    if current_batch == total_batches - 1:
+                        break
+
                     images = np.array(images)
 
                     # NOTE: Don't augment images for eval! NOTE: Changed my mind.
@@ -276,6 +283,9 @@ def train(config: ConfigDict, workdir: str = "./logging/") -> train_state.TrainS
 
                     metrics = eval_step(pstate, np.array(batch["input_ids"]), np.array(batch["pixel_values"]), np.array(batch["attention_mask"]), config.training.loss_type)
                     val_metrics.append(metrics)
+
+                    current_batch += 1  # Increment batch counter
+
 
                 def serialize_metrics(metrics):
                     """Convert all values in the metrics dict to Python standard types."""

### 2024-01-16 scripts/submit_train.sh
New runs

@@ -31,8 +31,8 @@ cd /n/holystore01/LABS/iaifi_lab/Users/smsharma/multimodal-data/
 # /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py  # Base config, with summary captions
 # /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.clip.transfer_head=True  # Fine-tune just head
 
-# /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.data.caption_type="abstract"  --config.data.augment_subsample_text=True  # Full abstract
-# /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.clip.random_init_text=True --config.clip.random_init_vision=True  # From scratch
+/n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.data.caption_type="abstract"  --config.data.augment_subsample_text=True  # Full abstract
+/n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.clip.random_init_text=True --config.clip.random_init_vision=True  # From scratch
 
 # /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.data.shuffle_within_batch=True  # Shuffle within batch
 # /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py  --config.sum1.use_sum1=True --config.sum1.sum1_filename="summary_sum1_v3" # Base config, with summary captions

### 2024-01-17 configs/base.py
Debugging, set up load_ckpt for eval

@@ -50,6 +50,9 @@ def get_config():
     training.ckpt_best_metric = "loss"  # "loss" or "top_xx_accuracy"
     training.ckpt_best_metric_best_mode = "min"  # "max" or "min" 
     training.ckpt_keep_top_n = 3  # Save the top `ckpt_keep_top_n` checkpoints based on `ckpt_best_metric`
+    training.load_ckpt = True
+    training.ckpt_run_name = "ancient-pine-88"
+
 
     # Sum1 options
     config.sum1 = sum1 = ml_collections.ConfigDict()

### 2024-01-17 train.py

@@ -49,7 +49,13 @@ def train(config: ConfigDict, workdir: str = "./logging/") -> train_state.TrainS
             config=wandb_config,
         )
         wandb.define_metric("*", step_metric="train/step")  # Set default x-axis as 'train/step'
-        workdir = os.path.join(workdir, run.group, run.name)
+        
+        # If load model from checkpoint, set from config file
+        if config.training.load_ckpt:
+            workdir = os.path.join(workdir, run.group, config.training.ckpt_run_name)
+            logging.info(f"Loading checkpoint from run {config.training.ckpt_run_name}")
+        else:
+            workdir = os.path.join(workdir, run.group, run.name)
 
         # Recursively create workdir
         os.makedirs(workdir, exist_ok=True)
@@ -188,6 +194,18 @@ def train(config: ConfigDict, workdir: str = "./logging/") -> train_state.TrainS
 
     # State
     state = train_state.TrainState.create(apply_fn=model.__call__ if (config.clip.use_pretrained) else model.apply, params=params, tx=tx)
+
+    # Load checkpoint
+    if config.training.load_ckpt:
+        restore_args = flax.training.orbax_utils.restore_args_from_target(state, mesh=None)
+        restored_state = ckpt_mgr.restore(ckpt_mgr.latest_step(), items=state, restore_kwargs={'restore_args': restore_args})
+
+        if state is restored_state:
+            raise FileNotFoundError(f"Did not load checkpoint correctly")
+        else:
+            state = restored_state
+            logging.info(f"Loaded checkpoint from step {ckpt_mgr.latest_step()}")
+
     pstate = replicate(state)
 
     # Log info about augmentations
@@ -221,6 +239,7 @@ def train(config: ConfigDict, workdir: str = "./logging/") -> train_state.TrainS
         for step in steps:
 
             # Use same rng for eval every time
+            # NOTE: Should move this down inside the eval if
             rng_eval = jax.random.PRNGKey(config.seed)
 
             # Eval portion

### 2024-02-19 paper/main.tex
update paper with new results

--- 2024-01-14 paper/main.tex
@@
 \begin{abstract}
-We present a multi-modal model which associates astronomical observations imaged by the \hubble Space Telescope (HST) with natural language.
-The model is fine-tuned from a base Contrastive Language–Image Pre-training (CLIP) model using successful proposal abstracts corresponding to HST observations, summarized via guided large language model (LLM) generation.
-We show that the model embodies a meaningful joint representation between observations and text through experiments targeting observation retrieval (i.e., retrieving the most relevant observations from a set using natural language queries) and description retrieval (i.e., querying the astrophysical object classes and science use cases most relevant to a given observation).
-The model demonstrates the potential for using generalist rather than task-specific models for astrophysics research, in particular by leveraging text as an interface.
+We train a multi-modal model which associates astronomical observations imaged by the \hubble Space Telescope (HST) with natural language.
+The model is fine-tuned from a pre-trained Contrastive Language–Image Pre-training (CLIP) model using successful proposal abstracts corresponding to HST observations, optionally summarized via guided large language model (LLM) generation.
+We show that the model embodies a meaningful joint representation between observations and text through experiments targeting image retrieval (i.e., retrieving the most relevant observations from a set using natural language queries) and description retrieval (i.e., querying the astrophysical object classes and science use cases most relevant to a given observation).
+Our study demonstrates the potential for using generalist rather than task-specific models for astrophysics research, in particular by leveraging text as an interface.
 \end{abstract}
@@
 Machine learning (ML) is beginning to have a significant impact in the sciences, with astrophysics being no exception.
-ML methods have demonstrated promise at every stage of the astrophysics pipeline, from instrument design, to data acquisition, to its analysis \citep{huertas2022dawes}.
-Until recently, most applications of ML within astrophysics have focused on replacing or augmenting traditional techniques with ML counterparts in order to improve performance on specific tasks.
-The \emph{Foundation Model} paradigm, in contrast, seeks to develop generalist models which can be deployed on a wide range of tasks.
+ML methods have demonstrated promise at every stage of the research pipeline, from instrument design, to data acquisition, to its analysis \citep{huertas2022dawes}.
+Until recently, most applications of ML within astrophysics have focused on replacing or augmenting traditional techniques with ML counterparts in order to improve performance on specific, individual tasks.
+The \emph{Foundation Model} paradigm, in contrast, seeks to develop generalist models which can be deployed to tackle a wide range of tasks.
 The paradigm has been highly successful in domains like computer vision and natural language processing, as demonstrated by the widespread adoption of tools like ChatGPT, Dall-E, Stable Diffusion, and CLIP.
@@
 At the heart of the paradigm lies the triumph of scale -- scaling up model size, dataset size, and compute.
-However, foundation models can often benefit from \emph{fine tuning} using a small amounts of domain data, increasing their usefulness when applied to those target domains.
+However, foundation models can often benefit from additional training ({fine tuning}) using a relatively small amounts of domain-specific data in order to increase their usefulness when applied to specialized domains.
 
-There is considerable interest in developing custom foundation models for the sciences \citep{batatia2023foundation,subramanian2023towards}, with astrophysics being ripe for such an effort given the availability of large amounts of publicly-available data and diverse ways of interacting with it.
-The multi-modality inherent to the nature of astrophysical observations, with different types of data (e.g., images, spectra, light curves, textual descriptions) often available, presents a unique opportunity.
-This multi-modality was recently exploited in \textsc{AstroCLIP}~\citep{lanusse2023astroclip} -- method to obtain a joint embedding space between multi-band images and optical spectra from the Dark Energy Spectroscopic Instrument (DESI).
-\textsc{AstroLLaMA}~\citep{nguyen2023astrollama} is another recent effort to fine-tune a publicly-available model (\textsc{Llama-2}) on astrophysics-specific textual data from the arXiv.
+There is considerable interest in developing custom foundation models for the sciences \citep[e.g., ][]{batatia2023foundation,subramanian2023towards}, with astrophysics being ripe for such an effort given the availability of large amounts of publicly-available data and diverse ways of interacting with it.
+The multi-modality inherent to the nature of astrophysical observations, with different types of data (e.g., images, spectra, light curves, textual descriptions) often available for a given object, presents a unique opportunity.
+This multi-modality was recently exploited in \textsc{AstroCLIP}~\citep{lanusse2023astroclip} to obtain a joint embedding space between multi-band images and optical spectra from the Dark Energy Spectroscopic Instrument (DESI).
+\textsc{AstroLLaMA}~\citep{nguyen2023astrollama,perkowski2024astrollama} is another recent effort to fine-tune a publicly-available model (\textsc{Llama-2}) on astrophysics-specific textual data from the arXiv.
 
-The Contrastive Language-Image Pre-training~\citep[CLIP;][]{radford2021learning} family of foundation models, which in its original instantiation embed images and text-via-captions into a common embedding space via contrastive learning, has shown strong performance and generalization capabilities on a variety of downstream tasks including zero-shot classification and image retrieval.
-Besides \textsc{AstroCLIP}, the concept of contrastive pre-training has been used in a variety of domains, e.g \textsc{GeoCLIP}~\citep{cepeda2023geoclip} for geospatial location querying.
+The Contrastive Language-Image Pre-training~\citep[CLIP;][]{radford2021learning} family of foundation models, which in their original instantiation embed images and associated captions into a common embedding space via contrastive learning, have shown strong performance and generalization capabilities on a variety of downstream tasks including zero-shot classification and image retrieval.
+The concept of associating diverse modalities via contrastive training has been used on other scientific domains~\citep[e.g.,][]{liu2023text,Sanchez-Fernandez2022.11.17.516915,lanusse2023astroclip,cepeda2023geoclip}, and has been shown to be effective in learning semantically meaningful joint representations.
 
-The success of the foundation model paradigm partly relies on the ability to flexibly leverage text as a \emph{universal interface} for interacting with models.
+The success of the foundation model paradigm relies in part on the ability to flexibly leverage text as a \emph{universal interface} for interacting with models.
 In this work, we take this outlook and train a joint representation between observations taken by the \hubble Space Telescope (HST) and natural language.
 We do so by leveraging associations between LLM-summarized abstracts of successful observing proposals and corresponding downstream image observations.
-We show that fine-tuning a pre-trained CLIP image-text model on this data enables learning semantically meaningful joint representations through quantitative and qualitative evaluation on retrieval tasks.
+We show that fine-tuning a pre-trained CLIP image-text model on this data results in meaningful joint representations through quantitative and qualitative evaluation.
 
@@
 We curate a dataset of images of \hubble observations and corresponding text descriptions.
-We rely on summarized versions of proposal abstracts from the Proposal Abstracts Catalog\footnote{\url{https://archive.stsci.edu/hst/proposal_abstracts.html}} -- a catalog of all accepted \hubble proposals to-date -- to derive captions for the observations.
-The HST has been operational for 33 years, having been launched on April 24, 1990.
-We use available proposals and observations up to Cycle 30, which commenced data-taking in 2022.
+We rely on proposal abstracts from the Proposal Abstracts Catalog\footnote{\url{https://archive.stsci.edu/hst/proposal_abstracts.html}} -- a catalog of all accepted \hubble proposals to-date -- to derive captions for the observations, optionally summarizing the abstract via guided LLM generation (described in Sec.~\ref{sec:summarization} below).
+The HST has been operational since its launch on April 24, 1990, and we use available proposals and observations up to Cycle 30, which commenced data-taking in 2022.
 
-Examples of images and corresponding captions are shown in Tab.~\ref{tab:dataset}.
-It can be seen that these images have specific characteristics as well as artifacts particular to the nature of data-taking which distinguish them from the distribution of natural images typically used for large-scale pre-training of foundation models.
+Examples of images and corresponding (summarized) captions are shown in Tab.~\ref{tab:dataset}.
+It can be seen that these images have specific characteristics as well as artifacts particular to the nature of data-taking and processing which distinguish them from the distribution of natural images typically used for large-scale pre-training of foundation models.
 This further motivates the need for fine-tuning on domain-specific data.
@@
 
-Observations corresponding to individual proposal IDs are queried through the Mikulski Archive for Space Telescopes (MAST) via the \package{Astroquery} interface.
+Observations corresponding to individual proposal IDs are queried through the Mikulski Archive for Space Telescopes (MAST)\footnote{\url{https://mast.stsci.edu/}} via the \package{Astroquery} \citep{2019AJ....157...98G} interface.
 Products of type \texttt{PREVIEW} are filtered in, corresponding to preview postcard images.
-We note that these are not science-grade observations, but rather lower-resolution images useful for quick-look purposes; given the nature of associations we aim to learn, this is adequate for our purposes.
+We note that these are not science-grade observations, but rather lower-resolution images useful for diagnostic purposes.
 A maximum of 20 images are downloaded per proposal ID, selected at random, in order to avoid biasing the model towards proposals with a larger number of observations and survey-style campaigns.
@@
 
+We note that some fraction of the image-caption pairs in the constructed dataset will center instrumentation and/or calibration rather than primarily scientific content.
+We choose to not filter out these pairs from our dataset, in order to have a larger sample of HST observations that the image encoder can leverage to adapting to the distinctive characteristics of \hubble images.
+
 % Mention that the signal is noisy!
 
-\subsection{Summarization via Guided Generation}
+\subsection{Abstract Summarization via Guided Generation}
 \label{sec:summarization}
 
-Raw abstracts summarize the corresponding successful HST observing proposals, which intend to make the case for allocating \hubble telescope time towards a particular set of observations.
+Raw proposal abstracts summarize the corresponding successful HST observing proposals, which intend to make the case for allocating \hubble telescope time towards a particular set of observations.
 These abstracts are written in a diversity of styles, formats, and lengths, being highly variable in the nature of content as well.
-Although the abstracts can be used as-is as image captions, for our baseline model we summarize them via guided LLM generation to standardize the captions used for fine-tuning the CLIP model.
+Although the abstracts can be used as-is as image captions, we experiment with summarizing them via guided large language model (LLM) generation to standardize the captions used for fine-tuning the CLIP model.
 The goal is to summarize the objects and phenomena, as well as potential downstream science use cases corresponding to the HST observations in order to increase the strength of the association signal between text and images.
@@
 
-The procedure guides the generation of LLM outputs while ensuring that the schema is respected at every point in the generation by masking out tokens that would violate the intended format.
+The procedure guides the generation of LLM outputs while ensuring that the schema is respected at every step in the autoregressive generation process by masking out tokens that would violate the intended format.
 By framing the problem in terms of transitions between a set of finite-state machines, \cite{willard2023efficient} showed that guided generation can be performed with negligible overhead compared to unconstrained generation.
@@
 Examples of raw abstract snippets and corresponding LLM-generated summaries are shown in Tab.~\ref{tab:datasetsumm}.
-See App.~\ref{app:guided-generation} for a more detailed description of this guidance generation process.
+See App.~\ref{app:guided-generation} for a more detailed description of the guidance generation process.
 
@@
 We emphasize that the goal of summarization-via-guided-generation is to increase the signal between text and images by standardizing the captions used for fine-tuning the CLIP model, and compare the quantitative performance of the model vs using the raw abstracts in Sec.~\ref{sec:results}.
-We also note that, even after summarization, the association signal is expected to be noisy, since the summarized caption may not always be usefully descriptive of the observed images.
+We also note that, even after summarization, the association signal is expected to be noisy, since aspects of the summarized caption may not be directly descriptive of the observed images, as can be seen from Tab.~\ref{tab:dataset}.
 
 Finally, in order to test a further compression of the proposal abstracts, we use LLM-guided generation to also produce a list of single-concept summaries (e.g., `irregular galaxy', `quasars', `Galactic bulge', \ldots), and test whether this can lead to meaningful, generalizable learned associations with observations.
-The prompts and schemata for generating and assigning these are described in Apps.~\ref{app:singleconcept} and \ref{app:singleconceptassignments} respectively.
+The prompts and schemata for generating these discrete categories and assigning observations to them are described in Apps.~\ref{app:singleconcept} and \ref{app:singleconceptassignments} respectively.
 
@@
       \end{tabular}
-      \caption{Examples of images and corresponding captions, constructed using the LLM-extracted summaries.
-The CLIP model is fine-tuned on these text-image associations.
-\SM{Can add another row to fill the page}}
+      \caption{Examples of images and corresponding captions, constructed using the LLM-extracted summaries. The CLIP model is fine-tuned using these text-image associations.}
       \label{tab:dataset}
@@
           \end{tabular}
-          \caption{Examples of the initial parts of raw proposal abstracts (second column) and LLM (\textsc{Mixtral-8x7B})-extracted summaries (rightmost two columns), separately extracting objects and phenomena as well as potential downstream science use cases.
-The LLM-extracted summaries are used for associating text with observations.}
+          \caption{Examples of the initial parts of raw proposal abstracts (second column) and LLM (\textsc{Mixtral-8x7B})-extracted summaries (rightmost two columns), separately extracting objects and phenomena as well as potential downstream science use cases.}
           \label{tab:datasetsumm}
@@
 
-Our goal is to learn a semantically  meaningful joint representation between HST image observations and natural language.
-We leverage the strong generalization capabilities demonstrated by pre-trained CLIP models and adapt these to work with domain-specific \hubble data via fine-tuning.
+Our goal is to learn a semantically  meaningful joint representation between HST observation images and natural language.
+We leverage the strong generalization capabilities demonstrated by pre-trained CLIP models and adapt these to work with domain-specific \hubble data via fine tuning.
 
@@
 Image and text encoders $f: I \rightarrow \mathbb R^{n_\text{emb}}$ and $g: T \rightarrow \mathbb R^{n_\text{emb}}$ are used to map images and text to a common embedding space of dimension $n_\text{emb}$.
-The standard softmax-based bidirectional variation of the InfoNCE~\citep{oord2018representation} contrastive loss function was introduce for training CLIP-style architectures\citep{radford2021learning}
+The standard softmax-based bidirectional variation of the InfoNCE~\citep{oord2018representation} contrastive loss function was introduce for training CLIP-style architectures \citep{radford2021learning}
 \begin{equation}
@@
 The text encoder has a maximum length of 77 tokens and the image encoder a native resolution of $224\times224$.
-Linear projection layers map the outputs of the image and text encoders to a common embedding space of dimension 512.
+Linear projection layers map the outputs of the image and text encoders to a common embedding space of dimension $n_\text{emb}=512$.
 In total, the model has 149,620,737 trainable parameters.
@@
 
-The base CLIP model is fine-tuned using the dataset described in Sec.~\ref{sec:dataset}, using either the LLM-summarized data or the raw proposal abstracts.
+The base CLIP model is fine-tuned using the dataset described in Sec.~\ref{sec:dataset}, using either the LLM-summarized data, the raw proposal abstracts, or single-concept summaries.
 When using raw proposal abstracts, random chunks of the text delimited by periods are selected on the fly to fit within the maximum token length of the text encoder.
 Images are randomly cropped to the native resolution of the image encoder and randomly rotated at each training step.
-Given the relatively modest size of the fine-tuning set, a batch size $|\mathcal B| = 32$ is used throughout; larger batch sizes were seen to susceptible to overfitting.
+Given the relatively modest size of the fine-tuning set, a batch size $|\mathcal B| = 32$ is used throughout; larger batch sizes were seen to be susceptible to overfitting.
 We note that the positive and negative image-text association is noisy and imperfect, since multiple images can be associated with the same abstract.
@@
 
-Models were instantiated and trained using the \package{Jax} \citep{jax2018github} ecosystem. Models were trained over 20,000 steps with 2000 linear warmup steps 
+Models were trained over 20,000 steps with 2000 linear warmup steps 
 % and cosine decay 
@@
 learning rate of $10^{-5}$ and weight decay $10^{-3}$. Training takes approximately 3 hours on 4 Nvidia A100 GPUs.
+Models were instantiated and trained using the \package{Jax} \citep{jax2018github} ecosystem.
 
@@
 The model is evaluated by tracking the loss in Eq.~\ref{eq:softmax_loss} as well as the top-$k\%$ retrieval accuracy on the held out validation set over the course of training.
-The retrieval accuracy is defined as the fraction of associated captions which fall within the top $k\%$ of captions by cosine similarity of the (normalized) embeddings $x_i \cdot y_j$, averaged over the images in the validation set.
+The retrieval accuracy is defined as the fraction of associated captions which fall within the top $k\%$ of captions by cosine similarity of the normalized embeddings $x_i \cdot y_j$, averaged over the images in the validation set.
 
@@
 
-Figure~\ref{fig:retrieval_acc} shows the contrastive loss (left) and the top-10\% retrieval accuracy on the held out validation set over the course of training, for different variations considered.
-The red lines show the metrics evaluated when training with batches where the image-text associations are randomly shuffled, serving as a baseline.
-This baseline is seen to do on par with random expectation, unlike the others, validating the presence of a significant association signal in the dataset.
+Figure~\ref{fig:retrieval_acc} shows the contrastive loss (left) and the top-10\% retrieval accuracy on the held out validation set over the course of training, for different training configurations considered.
+The purple lines show the metrics evaluated when training with batches where the image-text associations are randomly shuffled, serving as a baseline.
+This baseline is seen to do on par with random expectation, unlike the others, validating the presence of a significant association signal between images and text in the dataset.
 %
-Interestingly, the base model performs better than random expectation, with a top-10\% retrieval accuracy of $\sim 15\%$.
+Interestingly, the base pre-trained model performs better than random expectation, with a top-10\% retrieval accuracy of $\sim 15\%$.
 We therefore compare the qualitative performance of the base model with the fine-tuned model on downstream retrieval tasks.
 
-The fiducial model with LLM-guided summarization (orange lines) is seen to perform significantly better than the model using raw abstracts as captions (purple line), validating the stronger association signal in the summarized dataset curated.
-Fine-tuning a small MLP head over frozen vision and text backbones (green lines) and training from scratch (blue lines) show a non-trivial improvement compared to the random baseline as well as base model, but with deteriorated performance compared to the fiducial set-up.
+The model with LLM-guided summarization (yellow lines) is seen to perform on-par with the model using raw abstracts as captions (orange line), despite the stronger association signal in the summarized dataset curated.
+Fine-tuning a small MLP head over frozen vision and text backbones (green lines), training from scratch (blue lines), and using the single-concept summaries (red lines) show a non-trivial improvement compared to the random baseline as well as base model, but with deteriorated performance compared to fine-tuning with either summarized or raw abstracts.
 
@@
   \includegraphics[width=0.95\textwidth]{plots/val_metrics.pdf}
-  \caption{The CLIP-contrastive loss from Eq.~\ref{eq:softmax_loss} and the top-10\% retrieval accuracy computed on the validation set over the course of training.
-Shown for the fiducial set-up (orange), dataset using raw proposal abstracts (purple), only fine-tuning a small MLP head (green), training from scratch (blue), and a baseline trained with shuffled image-text pairs.}
+  \caption{The CLIP contrastive loss from Eq.~\ref{eq:softmax_loss} (left) and the top-10\% retrieval accuracy (right) computed on the validation set over the course of training.
+Shown for the dataset with summarized abstracts (orange), dataset using raw proposal abstracts (yellow), dataset with single-concept summaries (red), only fine-tuning a small MLP head (green), training from scratch (blue), and trained with shuffled image-text pairs (purple).}
   \label{fig:retrieval_acc}
@@
 
-Having aligned the image and text representations, we can query the validation set using natural language and show the most ``relevant'' images when ranked by cosine similarity.
-We show these in Figs.~\ref{fig:tti_base} and \ref{fig:tti} for the base model and fine-tuned model respectively using four simple curated queries: `globular clusters', `dwarf galaxy', `SN1987A' (a specific, flagship supernova), and `cluster lensing'.
+Having aligned the image and text representations, we can query the validation set using natural language and show the most `relevant' images when ranked by cosine similarity.
+We show these in Figs.~\ref{fig:tti_base} and \ref{fig:tti} for the base model and fine-tuned model respectively using four simple curated queries: `jupiter', `dwarf galaxy', `SN1987A' (a specific, prominent supernova), and `strong lensing'.
 
-The base model shows signs of meaningful retrieval -- it returns observations that clearly visually resemble globular clusters (first row of Fig.~\ref{fig:tti_base}), for example.
-Beyond this, the `dwarf galaxy' query returns a mix of galaxies and globular clusters, and it is challenging to discern meaningful associations between the other retrieved images and corresponding query.
+While the base model shows some signs of meaningful retrieval (e.g., the image of jupiter in the first row of Fig.~\ref{fig:tti_base}, and images of galaxies in the second row), it is challenging to discern meaningful strong associations between the retrieved images and corresponding query.
 
 The fine-tuned model (Fig.~\ref{fig:tti}), meanwhile, shows strikingly different behavior. 
-For example, it is able to return images with processing and assembly artifacts particular to HST (e.g., the lines through the middle in some images), which typically receive low similarity scores when using the base model.
+% For example, it is able to return images with processing and assembly artifacts particular to HST (e.g., the lines through the middle in some images), which typically receive low similarity scores when using the base model.
+Images of Jupiter are returned for the `jupiter' query, with the planet visible in all four images.
 The `dwarf galaxy' images correspond to proposals aiming to measure the kinematics of the stellar cores of dwarf galaxies.
-Supernova SN1987 itself can be seen as the most relevant image for the `SN1987A' query.
-Cluster-scale gravitational lenses are returned by the `cluster lensing' query, with lensing patterns visible in the images.
+Supernova SN1987 itself can be seen in the most relevant images for the `SN1987A' query, with the fourth image corresponding to a supernova remnant.
+Cluster-scale as well as galaxy-scale gravitational lenses are returned by the `strong lensing' query, with lensing patterns visible in the images.
 
@@
 
-Finally, we can use images from the validation set as queries and retrieve the most relevant text chunks (e.g., contained objects and use cases) from a curated list.
-We show the result of image-to-text retrieval in Fig.~\ref{fig:itt}, for the base as well as fine-tuned models, using the same genre of observations as for the text-to-image retrieval examples.
-We curate a list of possible text associations by querying the \textsc{Claude} large language model for such a list, which we show in App.~\ref{app:categories}.
+We can use images from the validation set as queries and retrieve the most relevant text chunks (e.g., contained objects and use cases) from a curated list.
+We show the result of image-to-text retrieval in Fig.~\ref{fig:itt}, for the base as well as fine-tuned models, using four observations from the validation set.
+We curate a list of possible text associations by querying the \textsc{Claude}\footnote{\url{https://claude.ai/}} large language model for such a list, which we show in App.~\ref{app:categories}.
 
-The top 3 text associations are shown for each image query.
+The top four text associations are shown for each image query.
 The `ground truth' summarized abstract is shown in the right column.
-The base model is seen to return a mix of relevant and less-relevant associations.
-While it can often return the nature of objects imaged, we observe it to seldom return scientific phenomena (e.g., `dark matter' as successfully done by the fine-tuned model in the second row).
-The third row (supernova 1987A) highlights interesting behavior -- the base model erroneously attributes the object at the center of the image to a gravitational lens or protoplanetary disk, while the fine-tuned model correctly identifies it as a supernova remnant (which play a crucial role for interstellar chemistry -- another retrieved text snippet).
+The base as well as fine-tuned models are seen to return a mix of relevant and less-relevant associations, although showing different qualitative behavior.
+% While it can often return the nature of objects imaged, we observe it to seldom return scientific phenomena (e.g., `dark matter' as successfully done by the fine-tuned model in the second row). 
+The second row (supernova 1987A) highlights an interesting pattern -- the base model erroneously attributes the object at the center of the image to a gravitational lens, while the fine-tuned model correctly identifies it as a supernova remnant.
 
-\paragraph*{Cosine similarity distribution}
+\paragraph*{Statistical distributions on the validation set}
 
-The top row of Fig.~\ref{fig:sim_valtrain} shows the distribution of cosine similarities between corresponding image and text embeddings, $x_i$ and $y_i$, for the LLM-summarized abstracts using the fine-tuned CLIP model (red line), and for the base CLIP model (blue line), shown for the set of validation pairs (top left) and same number of training set pairs (top right).
-Distributions shuffling the order of text embeddings, randomizing the image-text correspondence, is shown as dashed lines.
-The distribution for the base model is seen to be sharply peaked at a specific value, showing little diversity and being very similar between the shuffled (dashed blue) and non-shuffled (solid blue) versions.
-Histograms for the fine-tuned model show a clear separation between the shuffled and corresponding text-image pair versions.
+Figure~\ref{fig:sim_valtrain} (top left) shows the distribution of cosine similarities between corresponding image and text embeddings, $x_i$ and $y_i$, for the LLM-summarized abstracts using the fine-tuned CLIP model (red line), and for the base CLIP model (blue line).
+Distributions after shuffling the order of text embeddings -- therefore randomizing the image-text correspondence -- are shown as dashed lines.
+The distributions for the base model is seen to be sharply peaked at a specific value, showing little diversity and being very similar between the shuffled (dashed blue) and non-shuffled (solid blue) versions.
+Histograms for the fine-tuned model, on the other hand, show a clear separation between the shuffled and corresponding text-image pair versions.
 
-Distributions for a set of training samples (top right) are visible more peaked towards higher values compared to those evaluated on validation samples (top left), indicating potential distribution shift and/or a degree of overfitting.
+Similarly, Fig.~\ref{fig:sim_valtrain} (top right) shows the distribution of cosine similarities between corresponding image-text pairs for the single-concept captions (solid red) and the shuffled versions (dashed red).
+The distribution for corresponding pairs is only modestly shifted compared to the shuffled version, indicating that assigning $\mathcal O(400)$ labels to the images using an LLM results in a weaker association signal compared to using noisy captions.
+The purple line shows the distribution of cosine similarities between corresponding pairs with single-concept summaries, but evaluated using a model trained on the LLM-summarized abstracts.
+This distribution is more skewed towards higher similarity values, indicating that the model trained on diverse caption can still better capture the association signal in the single-concept summaries compared to one trained specifically using the single-concept summaries.
 
-Finally, in Fig.~\ref{fig:sim_valtrain} (bottom) we show the distribution of corresponding image-text pair cosine similarities for the single-concept summaries, evaluated using the CLIP model fine-tuned on summarized abstracts (red) and on the single-concept summaries themselves (purple).
-These distributions are peaked towards only marginally higher values compared to a baseline shuffled version, indicating that assigning $\mathcal O(100)$ labels to the images using an LLM results in a relatively weak association signal compared to using noisy captions.
+Finally, Fig.~\ref{fig:sim_valtrain} (bottom) shows the image-to-text retrieval accuracy -- i.e., the fraction of correct abstracts which fall within the top $k$-fraction of captions by cosine similarity, averaged over images in the validation set -- as a function of the retrieval fraction $k$.
+In this case, we evaluate three models -- fine tuned on raw abstracts (red), fine-tuned on LLM-summarized abstracts (blue), and fine-tuned on single-concept summaries (green), and the base model (purple) -- on the summarized abstracts.
+Remarkably, the model trained on raw abstracts shows very similar performance on the summarized abstracts compared to that trained on the summarized abstracts themselves, indicating that \emph{(1)} the association signal is preserved in the summarization process, and \emph{(2)} the model is able to effectively leverage meaningful concepts in the noisy raw abstracts through weak supervision.
 
@@
 \begin{figure*}[!h]
-\includegraphics[width=0.95\textwidth]{plots/sim_valtrain.pdf}
-\centering\includegraphics[width=0.45\textwidth]{plots/sim_summ1.pdf}
+\includegraphics[width=0.45\textwidth]{plots/sim_val.pdf}
+\includegraphics[width=0.45\textwidth]{plots/sim_summ1.pdf}
+\centering\includegraphics[width=0.45\textwidth]{plots/retrieval.pdf}
 \caption{Distribution of cosine similarities between corresponding image and text embeddings, $x_i$ and $y_i$.
@@
 We show that \textsc{HubbleCLIP} significantly outperforms the base CLIP model in quantitative metrics, such as retrieval accuracy, as well as quality of text-to-image and image-to-text retrieval.
-We also introduce a novel LLM summarization process which leverages constrained generation to significantly increase the association strength between images and captions while preserving salient information.
+We also introduce a novel LLM summarization process which leverages constrained generation to increase the association strength between images and captions while preserving salient information.
 Overall, the procedure demonstrates the efficacy on fine-tuning generalist pre-trained models on small amounts of domain-specific data, in particular astronomical datasets.
@@
 In particular, a robust model trained using weakly-supervised image-text pairs can be used to query survey data e.g., PHANGS~\citep{lee2022phangs}, COSMOS~\citep{scoville2007cosmic} using natural language, as well as to quickly find distinguishing patterns in such data that may not be apparent using specialized models or manual inspection.
-The learned representations, having shown to correlate with physical characteristics of imaged objects, can also be fine-tuned via transfer learning to adapt to specific tasks~\citep[e.g., classification as in][]{wei2020deep}.
-Finally, while the CLIP model is restricted to retrieving nearest-neighbour associations within and across text/image modalities, the learned embeddings can be used as a starting point for training or fine-tuning multi-modal large-language models for interacting with survey data and receiving responses in natural language form, as well as grounding the responses based on known associations.
+The learned representations, having shown to correlate with physical characteristics of imaged objects, can also be fine-tuned via transfer learning to adapt to specific tasks e.g., classification~\citep{wei2020deep} or segmentation~\citep{hausen2020morpheus}.
+Finally, while the CLIP model is restricted to retrieving nearest-neighbour associations within and across text/image modalities, the learned embeddings can be used as a starting point for training or fine-tuning multi-modal large-language models for interacting with survey data and receiving responses in natural language form, as well as grounding the responses based on an existing set of observations.
 
@@
 These are intended to be broader than the single-concept summaries inferred in Sec.~\ref{app:singleconcept} and are derived by prompting \textsc{Claude}, without any external input e.g.
-proposal abstracts.
-\SM{Try to just use the single-concept summaries here to avoid the extra step}.
+proposal abstracts, to produce a list of categories corresponding to HST observations.
+% \SM{Try to just use the single-concept summaries here to avoid the extra step}.
 

### 2024-02-27 paper/main.tex
Added percents to tex source
Text comments and updates

@@
 
-\title{\textsc{HubbleCLIP}: Associating Astronomical Observations and Natural Language with Multi-Modal Models}
-
-% Authors must not appear in the submitted version. They should be hidden
-% as long as the tmlr package is used without the [accepted] or [preprint] options.
-% Non-anonymous submissions will be rejected without review.
+\title{\textsc{PAPERCLIP}: Associating Astronomical Observations and Natural Language with Multi-Modal Models}
 
@@
 }
-% The \author macro works with any number of authors. Use \AND 
-% to separate the names and addresses of multiple authors.
 
@@
 \begin{abstract}
-We train a multi-modal model which associates astronomical observations imaged by the \hubble Space Telescope (HST) with natural language.
-The model is fine-tuned from a pre-trained Contrastive Language–Image Pre-training (CLIP) model using successful proposal abstracts corresponding to HST observations, optionally summarized via guided large language model (LLM) generation.
+We present PAPERCLIP (Proposal Abstracts Provide an Effective Representation for Contrastive Language-Image Pretraining), a method which associates astronomical observations imaged by astronomical surveys like the \hubble Space Telescope (HST) with natural language.
+The model is fine-tuned from a pre-trained Contrastive Language–Image Pre-training (CLIP) model using successful proposal abstracts corresponding to HST observations, optionally summarized via guided generation using large language models (LLMs).
 We show that the model embodies a meaningful joint representation between observations and text through experiments targeting image retrieval (i.e., retrieving the most relevant observations from a set using natural language queries) and description retrieval (i.e., querying the astrophysical object classes and science use cases most relevant to a given observation).
@@
 
+\SM{Add TOC.}
+\SM{Change PAPERCLIP everywhere.}
+\SM{Add some qualitative evals with the abstract model.}
+
 Machine learning (ML) is beginning to have a significant impact in the sciences, with astrophysics being no exception.
+%
 ML methods have demonstrated promise at every stage of the research pipeline, from instrument design, to data acquisition, to its analysis \citep{huertas2022dawes}.
+%
 Until recently, most applications of ML within astrophysics have focused on replacing or augmenting traditional techniques with ML counterparts in order to improve performance on specific, individual tasks.
+%
 The \emph{Foundation Model} paradigm, in contrast, seeks to develop generalist models which can be deployed to tackle a wide range of tasks.
+%
 The paradigm has been highly successful in domains like computer vision and natural language processing, as demonstrated by the widespread adoption of tools like ChatGPT, Dall-E, Stable Diffusion, and CLIP.
+%
 These models are typically pre-trained on massive amounts of unlabeled data using self-supervised learning techniques, enabling them to learn powerful representations which can be deployed on different tasks across domains.
+%
 At the heart of the paradigm lies the triumph of scale -- scaling up model size, dataset size, and compute.
+%
 However, foundation models can often benefit from additional training ({fine tuning}) using a relatively small amounts of domain-specific data in order to increase their usefulness when applied to specialized domains.
@@
 There is considerable interest in developing custom foundation models for the sciences \citep[e.g., ][]{batatia2023foundation,subramanian2023towards}, with astrophysics being ripe for such an effort given the availability of large amounts of publicly-available data and diverse ways of interacting with it.
+%
 The multi-modality inherent to the nature of astrophysical observations, with different types of data (e.g., images, spectra, light curves, textual descriptions) often available for a given object, presents a unique opportunity.
+%
 This multi-modality was recently exploited in \textsc{AstroCLIP}~\citep{lanusse2023astroclip} to obtain a joint embedding space between multi-band images and optical spectra from the Dark Energy Spectroscopic Instrument (DESI).
+%
 \textsc{AstroLLaMA}~\citep{nguyen2023astrollama,perkowski2024astrollama} is another recent effort to fine-tune a publicly-available model (\textsc{Llama-2}) on astrophysics-specific textual data from the arXiv.
@@
 The Contrastive Language-Image Pre-training~\citep[CLIP;][]{radford2021learning} family of foundation models, which in their original instantiation embed images and associated captions into a common embedding space via contrastive learning, have shown strong performance and generalization capabilities on a variety of downstream tasks including zero-shot classification and image retrieval.
+%
 The concept of associating diverse modalities via contrastive training has been used on other scientific domains~\citep[e.g.,][]{liu2023text,Sanchez-Fernandez2022.11.17.516915,lanusse2023astroclip,cepeda2023geoclip}, and has been shown to be effective in learning semantically meaningful joint representations.
 
+\SM{Put in this work we at the start of the paragraph. Also put the paragraph right after AstroLlama, reworking some of the logic.} \SM{Define PAPERCLIP, funny footnote about PAPERCLIFT.}
+
 The success of the foundation model paradigm relies in part on the ability to flexibly leverage text as a \emph{universal interface} for interacting with models.
+%
 In this work, we take this outlook and train a joint representation between observations taken by the \hubble Space Telescope (HST) and natural language.
+%
 We do so by leveraging associations between LLM-summarized abstracts of successful observing proposals and corresponding downstream image observations.
+%
 We show that fine-tuning a pre-trained CLIP image-text model on this data results in meaningful joint representations through quantitative and qualitative evaluation.
@@
 The paper is organized as follows.
+%
 In Sec.~\ref{sec:dataset}, we describe the dataset used in this work, including its curation and processing.
+%
 In Sec.~\ref{sec:methodology}, we describe the methodology used to train and evaluate the model.
+%
 In Sec.~\ref{sec:results}, we present the results of our experiments on image and text retrieval tasks.
+%
 We discuss future prospects and conclude in Sec.~\ref{sec:conclusion}.
@@
 We curate a dataset of images of \hubble observations and corresponding text descriptions.
+%
 We rely on proposal abstracts from the Proposal Abstracts Catalog\footnote{\url{https://archive.stsci.edu/hst/proposal_abstracts.html}} -- a catalog of all accepted \hubble proposals to-date -- to derive captions for the observations, optionally summarizing the abstract via guided LLM generation (described in Sec.~\ref{sec:summarization} below).
+%
 The HST has been operational since its launch on April 24, 1990, and we use available proposals and observations up to Cycle 30, which commenced data-taking in 2022.
+%
 
-Examples of images and corresponding (summarized) captions are shown in Tab.~\ref{tab:dataset}.
+Examples of images and corresponding (summarized) captions are shown in Tab.~\ref{tab:dataset}. \SM{Push table earlier.}
+%
 It can be seen that these images have specific characteristics as well as artifacts particular to the nature of data-taking and processing which distinguish them from the distribution of natural images typically used for large-scale pre-training of foundation models.
+%
 This further motivates the need for fine-tuning on domain-specific data.
@@
 Observations corresponding to individual proposal IDs are queried through the Mikulski Archive for Space Telescopes (MAST)\footnote{\url{https://mast.stsci.edu/}} via the \package{Astroquery} \citep{2019AJ....157...98G} interface.
+%
 Products of type \texttt{PREVIEW} are filtered in, corresponding to preview postcard images.
+%
 We note that these are not science-grade observations, but rather lower-resolution images useful for diagnostic purposes.
+%
 A maximum of 20 images are downloaded per proposal ID, selected at random, in order to avoid biasing the model towards proposals with a larger number of observations and survey-style campaigns.
-Images are centered and resized to a resolution-per-side of 512 pixels before.
+%
+Images are centered and resized to a resolution-per-side of 512 pixels.
+%
 Color previews (i.e., observations taken with multiple wavelength filters assigned to individual RGB channels) are manually excluded via a filename filter in order to maintain consistency across the dataset; models trained on datasets with color images included were observed to show worse performance on generalization metrics.
+%
 If no appropriate images corresponding to an abstract are found, it is excluded from the dataset.
@@
 In total, 31,859 images corresponding to 4,438 abstracts are included in the fine-tuning dataset.
+%
 3,194 images are held out for validation, with no abstract being common between training and validation sets in order to ensure an independent set of text-image pairs for evaluation.
+\SM{Put number of held out abstracts here}
 
-We note that some fraction of the image-caption pairs in the constructed dataset will center instrumentation and/or calibration rather than primarily scientific content.
+We note that some fraction of the image-caption pairs in the constructed dataset will primarily concern instrumentation and/or calibration rather than scientific content.
+%
 We choose to not filter out these pairs from our dataset, in order to have a larger sample of HST observations that the image encoder can leverage to adapting to the distinctive characteristics of \hubble images.
@@
 Raw proposal abstracts summarize the corresponding successful HST observing proposals, which intend to make the case for allocating \hubble telescope time towards a particular set of observations.
+%
 These abstracts are written in a diversity of styles, formats, and lengths, being highly variable in the nature of content as well.
+%
 Although the abstracts can be used as-is as image captions, we experiment with summarizing them via guided large language model (LLM) generation to standardize the captions used for fine-tuning the CLIP model.
+%
 The goal is to summarize the objects and phenomena, as well as potential downstream science use cases corresponding to the HST observations in order to increase the strength of the association signal between text and images.
@@
 The method from \cite{willard2023efficient} is used to produce an LLM-generated summary of the abstract conforming to a particular schema, specified in JSON format.
+%
 The schema is designed to represent a list of the objects (e.g., `Type Ia supernova') and phenomena (e.g., `gravitational lensing'), as well as potential downstream science uses cases (e.g., `set constraints on supernova explosion models') that could correspond to the eventual imaged observation given the abstract text, with a minimum of 1 and a maximum of 5 elements per list.
+\SM{Emphasize that the schema ensures the format, but the prompt is necessary for it to make sense.}
 
 The procedure guides the generation of LLM outputs while ensuring that the schema is respected at every step in the autoregressive generation process by masking out tokens that would violate the intended format.
+%
 By framing the problem in terms of transitions between a set of finite-state machines, \cite{willard2023efficient} showed that guided generation can be performed with negligible overhead compared to unconstrained generation.
+%
 This ensures that the output of the LLM strictly conforms to the format of the following example:
@@
 which is then used to construct the summarized caption by combining the two key elements.
+%
 Examples of raw abstract snippets and corresponding LLM-generated summaries are shown in Tab.~\ref{tab:datasetsumm}.
+%
 See App.~\ref{app:guided-generation} for a more detailed description of the guidance generation process.
@@
 We use the open-weights, instruction-tuned model \textsc{Mixtral-8x7B-Instruct}~\citep{jiang2024mixtral} to generate the summaries, with guided generation performed using the \package{Outlines}\footnote{\url{https://github.com/outlines-dev/outlines}} package.
+%
 Further details on the summarization procedure, including the prompts and schema used, are provided in App.~\ref{app:summarization}.
@@
 We emphasize that the goal of summarization-via-guided-generation is to increase the signal between text and images by standardizing the captions used for fine-tuning the CLIP model, and compare the quantitative performance of the model vs using the raw abstracts in Sec.~\ref{sec:results}.
+%
 We also note that, even after summarization, the association signal is expected to be noisy, since aspects of the summarized caption may not be directly descriptive of the observed images, as can be seen from Tab.~\ref{tab:dataset}.
@@
 Finally, in order to test a further compression of the proposal abstracts, we use LLM-guided generation to also produce a list of single-concept summaries (e.g., `irregular galaxy', `quasars', `Galactic bulge', \ldots), and test whether this can lead to meaningful, generalizable learned associations with observations.
+%
 The prompts and schemata for generating these discrete categories and assigning observations to them are described in Apps.~\ref{app:singleconcept} and \ref{app:singleconceptassignments} respectively.
+\SM{Each abstract is assigned one summary.
+The generation of the single concepts is different }
 
@@
       \end{tabular}
-      \caption{Examples of images and corresponding captions, constructed using the LLM-extracted summaries. The CLIP model is fine-tuned using these text-image associations.}
+      \caption{Examples of images and corresponding captions, constructed using the LLM-extracted summaries.
+%
+The CLIP model is fine-tuned using these text-image associations. \SM{Semicolon separates, but the model doesn't know about it.}}
       \label{tab:dataset}
@@
           \end{tabular}
-          \caption{Examples of the initial parts of raw proposal abstracts (second column) and LLM (\textsc{Mixtral-8x7B})-extracted summaries (rightmost two columns), separately extracting objects and phenomena as well as potential downstream science use cases.}
+          \caption{Examples of the initial parts of raw proposal abstracts (second column) and LLM (\textsc{Mixtral-8x7B})-extracted summaries (rightmost two columns), separately extracting objects and phenomena as well as potential downstream science use cases.
+\SM{Cite all proposal.} \SM{Add a year with the obs.
+cycle.}.}
           \label{tab:datasetsumm}
@@
 Our goal is to learn a semantically  meaningful joint representation between HST observation images and natural language.
+%
 We leverage the strong generalization capabilities demonstrated by pre-trained CLIP models and adapt these to work with domain-specific \hubble data via fine tuning.
@@
 CLIP \citep[Contrastive Language-Image Pretraining;][]{radford2021learning} is a multi-modal model pre-trained on a large corpus of image-text pairs via weak supervision using a contrastive loss.
+%
 Given a minibatch $\mathcal{B}$ of $|\mathcal{B}|$ image-text pairs $\{(I_i, T_i)\}$, the goal is to align the learned representations of corresponding (positive) pairs $(I_i, T_i)$ while repelling the representations of unaligned (negative) pairs $(I_i, T_{j\neq i})$.
+%
 Image and text encoders $f: I \rightarrow \mathbb R^{n_\text{emb}}$ and $g: T \rightarrow \mathbb R^{n_\text{emb}}$ are used to map images and text to a common embedding space of dimension $n_\text{emb}$.
+%
 The standard softmax-based bidirectional variation of the InfoNCE~\citep{oord2018representation} contrastive loss function was introduce for training CLIP-style architectures \citep{radford2021learning}
@@
 where ${x}_i={f\left(I_i\right)}/{\left\|f\left(I_i\right)\right\|}$ and ${y}_i={g\left(T_i\right)}/{\left\|g\left(T_i\right)\right\|}$ are the normalized representations of the $i$-th image and text, respectively, and $\tau$ is a learnable temperature hyperparameter.
+%
 Note that this loss treats the image and text representations symmetrically, ensuring that the two modalities are on the same footing.
@@
 We use the CLIP-ViT-B/16 \citep{radford2021learning} variant as the base pre-trained CLIP model.
+%
 This model uses a 12-layer, 12-head, 768-embedding dimension vision transformer as the image encoder and a 12-layer, 8-head, 512-embedding dimension sequence transformer as the text backbone.
+%
 The text encoder has a maximum length of 77 tokens and the image encoder a native resolution of $224\times224$.
+%
 Linear projection layers map the outputs of the image and text encoders to a common embedding space of dimension $n_\text{emb}=512$.
+%
 In total, the model has 149,620,737 trainable parameters.
+%
 The model was originally trained on 400 million image-text pairs from internet data.
+\SM{$\tau$ was initialized to the pre-trained value.} \SM{HuggingFace link?} \SM{What area is maintained.} \SM{It's only 90 deg rotations.}
+%
 
@@
 The base CLIP model is fine-tuned using the dataset described in Sec.~\ref{sec:dataset}, using either the LLM-summarized data, the raw proposal abstracts, or single-concept summaries.
+%
 When using raw proposal abstracts, random chunks of the text delimited by periods are selected on the fly to fit within the maximum token length of the text encoder.
+%
 Images are randomly cropped to the native resolution of the image encoder and randomly rotated at each training step.
+%
 Given the relatively modest size of the fine-tuning set, a batch size $|\mathcal B| = 32$ is used throughout; larger batch sizes were seen to be susceptible to overfitting.
+%
 We note that the positive and negative image-text association is noisy and imperfect, since multiple images can be associated with the same abstract.
@@
 We explore three different methods of training the model on our domain dataset: \emph{(1)} Fine-tuning the entire network, starting from the pre-trained base model; \emph{(2)} Freezing the base image and text encoders, and training a small projection head; and \emph{(3)} Training the entire model from scratch.
+%
 For \emph{(2)}, we use a 2-layer MLP with 1024 units and a GELU activation layer, projecting onto the 512-dimensional embedding space.
@@
 using the AdamW optimizer \citep{DBLP:conf/iclr/LoshchilovH19,DBLP:journals/corr/KingmaB14} with  %peak 
-learning rate of $10^{-5}$ and weight decay $10^{-3}$. Training takes approximately 3 hours on 4 Nvidia A100 GPUs.
-Models were instantiated and trained using the \package{Jax} \citep{jax2018github} ecosystem.
+learning rate of $10^{-5}$ and weight decay $10^{-3}$.
+%
+Training takes approximately 3 hours on 4 Nvidia A100 GPUs.
+Models were instantiated and trained using packages from the \package{Jax} \citep{jax2018github} ecosystem.
+%
 
-\paragraph*{Evaluation}
+\paragraph*{Evaluation metrics}
 
 The model is evaluated by tracking the loss in Eq.~\ref{eq:softmax_loss} as well as the top-$k\%$ retrieval accuracy on the held out validation set over the course of training.
+%
 The retrieval accuracy is defined as the fraction of associated captions which fall within the top $k\%$ of captions by cosine similarity of the normalized embeddings $x_i \cdot y_j$, averaged over the images in the validation set.
@@
 We also qualitatively evaluate the learned embeddings through image retrieval (i.e., retrieving the most relevant images from a set using natural language queries) and description retrieval (i.e., querying the astrophysical object classes and science use cases most relevant to a given observation) experiments.
+\SM{Add some expressions.}
+
+\SM{Put in the Claude thing here also.}
+
+\SM{Numbered subsections.}
 
@@
 
-Figure~\ref{fig:retrieval_acc} shows the contrastive loss (left) and the top-10\% retrieval accuracy on the held out validation set over the course of training, for different training configurations considered.
+Figure~\ref{fig:retrieval_acc} shows the contrastive loss (left) and the top-10\% retrieval accuracy (right) on the held out validation set over the course of training, for different training configurations considered.
+%
 The purple lines show the metrics evaluated when training with batches where the image-text associations are randomly shuffled, serving as a baseline.
+%
 This baseline is seen to do on par with random expectation, unlike the others, validating the presence of a significant association signal between images and text in the dataset.
@@
 Interestingly, the base pre-trained model performs better than random expectation, with a top-10\% retrieval accuracy of $\sim 15\%$.
+%
 We therefore compare the qualitative performance of the base model with the fine-tuned model on downstream retrieval tasks.
@@
 The model with LLM-guided summarization (yellow lines) is seen to perform on-par with the model using raw abstracts as captions (orange line), despite the stronger association signal in the summarized dataset curated.
+%
 Fine-tuning a small MLP head over frozen vision and text backbones (green lines), training from scratch (blue lines), and using the single-concept summaries (red lines) show a non-trivial improvement compared to the random baseline as well as base model, but with deteriorated performance compared to fine-tuning with either summarized or raw abstracts.
@@
   \caption{The CLIP contrastive loss from Eq.~\ref{eq:softmax_loss} (left) and the top-10\% retrieval accuracy (right) computed on the validation set over the course of training.
-Shown for the dataset with summarized abstracts (orange), dataset using raw proposal abstracts (yellow), dataset with single-concept summaries (red), only fine-tuning a small MLP head (green), training from scratch (blue), and trained with shuffled image-text pairs (purple).}
+Shown for the dataset with summarized abstracts (orange), dataset using raw proposal abstracts (yellow), dataset with single-concept summaries (red), only fine-tuning a small MLP head (green), training from scratch (blue), and trained with shuffled image-text pairs (purple).
+\SM{Try not splitting the legend.} \SM{Would dashing be useful at all as a way to signify relationships.}}
   \label{fig:retrieval_acc}
@@
 Having aligned the image and text representations, we can query the validation set using natural language and show the most `relevant' images when ranked by cosine similarity.
+%
 We show these in Figs.~\ref{fig:tti_base} and \ref{fig:tti} for the base model and fine-tuned model respectively using four simple curated queries: `jupiter', `dwarf galaxy', `SN1987A' (a specific, prominent supernova), and `strong lensing'.
@@
 The fine-tuned model (Fig.~\ref{fig:tti}), meanwhile, shows strikingly different behavior.
+%
 % For example, it is able to return images with processing and assembly artifacts particular to HST (e.g., the lines through the middle in some images), which typically receive low similarity scores when using the base model.
+%
 Images of Jupiter are returned for the `jupiter' query, with the planet visible in all four images.
+%
 The `dwarf galaxy' images correspond to proposals aiming to measure the kinematics of the stellar cores of dwarf galaxies.
+%
 Supernova SN1987 itself can be seen in the most relevant images for the `SN1987A' query, with the fourth image corresponding to a supernova remnant.
+%
 Cluster-scale as well as galaxy-scale gravitational lenses are returned by the `strong lensing' query, with lensing patterns visible in the images.
@@
 We can use images from the validation set as queries and retrieve the most relevant text chunks (e.g., contained objects and use cases) from a curated list.
+%
 We show the result of image-to-text retrieval in Fig.~\ref{fig:itt}, for the base as well as fine-tuned models, using four observations from the validation set.
+%
 We curate a list of possible text associations by querying the \textsc{Claude}\footnote{\url{https://claude.ai/}} large language model for such a list, which we show in App.~\ref{app:categories}.
+\SM{Introduce this earlier!}
 
 The top four text associations are shown for each image query.
+%
 The `ground truth' summarized abstract is shown in the right column.
+%
 The base as well as fine-tuned models are seen to return a mix of relevant and less-relevant associations, although showing different qualitative behavior.
-% While it can often return the nature of objects imaged, we observe it to seldom return scientific phenomena (e.g., `dark matter' as successfully done by the fine-tuned model in the second row). 
+%
+
 The second row (supernova 1987A) highlights an interesting pattern -- the base model erroneously attributes the object at the center of the image to a gravitational lens, while the fine-tuned model correctly identifies it as a supernova remnant.
@@
 Figure~\ref{fig:sim_valtrain} (top left) shows the distribution of cosine similarities between corresponding image and text embeddings, $x_i$ and $y_i$, for the LLM-summarized abstracts using the fine-tuned CLIP model (red line), and for the base CLIP model (blue line).
+%
 Distributions after shuffling the order of text embeddings -- therefore randomizing the image-text correspondence -- are shown as dashed lines.
+%
 The distributions for the base model is seen to be sharply peaked at a specific value, showing little diversity and being very similar between the shuffled (dashed blue) and non-shuffled (solid blue) versions.
+%
 Histograms for the fine-tuned model, on the other hand, show a clear separation between the shuffled and corresponding text-image pair versions.
@@
 Similarly, Fig.~\ref{fig:sim_valtrain} (top right) shows the distribution of cosine similarities between corresponding image-text pairs for the single-concept captions (solid red) and the shuffled versions (dashed red).
+%
 The distribution for corresponding pairs is only modestly shifted compared to the shuffled version, indicating that assigning $\mathcal O(400)$ labels to the images using an LLM results in a weaker association signal compared to using noisy captions.
+%
 The purple line shows the distribution of cosine similarities between corresponding pairs with single-concept summaries, but evaluated using a model trained on the LLM-summarized abstracts.
+%
 This distribution is more skewed towards higher similarity values, indicating that the model trained on diverse caption can still better capture the association signal in the single-concept summaries compared to one trained specifically using the single-concept summaries.
@@
 Finally, Fig.~\ref{fig:sim_valtrain} (bottom) shows the image-to-text retrieval accuracy -- i.e., the fraction of correct abstracts which fall within the top $k$-fraction of captions by cosine similarity, averaged over images in the validation set -- as a function of the retrieval fraction $k$.
+%
 In this case, we evaluate three models -- fine tuned on raw abstracts (red), fine-tuned on LLM-summarized abstracts (blue), and fine-tuned on single-concept summaries (green), and the base model (purple) -- on the summarized abstracts.
+%
 Remarkably, the model trained on raw abstracts shows very similar performance on the summarized abstracts compared to that trained on the summarized abstracts themselves, indicating that \emph{(1)} the association signal is preserved in the summarization process, and \emph{(2)} the model is able to effectively leverage meaningful concepts in the noisy raw abstracts through weak supervision.
@@
 \includegraphics[width=0.95\textwidth]{plots/tti_base.pdf}
-\caption{Image retrieval using the base CLIP model on four curated queries.}
+\caption{Image retrieval using the base CLIP model on four curated queries.
+\SM{Horizontal lines} \SM{Query not rotated} \SM{Same for next Figure.} \SM{Put text ``Base Model'' in consistent colour to legend.}}
 \label{fig:tti_base}
@@
 \caption{Text associations from a curated list most closely matching a given image query, for both the fine-tuned and base models.
-The `ground truth' LLM-summarized abstract is shown in the right column.}
+The `ground truth' LLM-summarized abstract is shown in the right column.
+\SM{Swap order of cols 2 and 3.}}
 \label{fig:itt}
@@
 Versions with the ordering of text embeddings shuffled are shown in dashed.
-(Bottom) For the single-concept summaries, when evaluated using the CLIP model fine-tuned on LLM-summarized abstracts (red) and fine-tuned on the single-concept summaries themselves (purple).}
+(Bottom) For the single-concept summaries, when evaluated using the CLIP model fine-tuned on LLM-summarized abstracts (red) and fine-tuned on the single-concept summaries themselves (purple).
+\SM{Make the colours here consistent with previous Figure.}}
 \label{fig:sim_valtrain}
@@
 We present \textsc{HubbleCLIP}, a multi-modal fountation model that associates observations imaged by the \hubble Space Telescope with natural language in a common, semantically-meaningful embedding space.
+%
 The model is fine-tuned from a pre-trained CLIP model on summarized versions of \hubble proposal abstracts, leveraging a noisy association signal between text and images.
+%
 We show that \textsc{HubbleCLIP} significantly outperforms the base CLIP model in quantitative metrics, such as retrieval accuracy, as well as quality of text-to-image and image-to-text retrieval.
+%
 We also introduce a novel LLM summarization process which leverages constrained generation to increase the association strength between images and captions while preserving salient information.
+%
 Overall, the procedure demonstrates the efficacy on fine-tuning generalist pre-trained models on small amounts of domain-specific data, in particular astronomical datasets.
@@
 Although the model explored here is fine-tuned using postage stamp images (i.e., preview-quality and not science-grade data), we highlight potential immediate as well as downstream use cases of this paradigm.
+%
 In particular, a robust model trained using weakly-supervised image-text pairs can be used to query survey data e.g., PHANGS~\citep{lee2022phangs}, COSMOS~\citep{scoville2007cosmic} using natural language, as well as to quickly find distinguishing patterns in such data that may not be apparent using specialized models or manual inspection.
+%
 The learned representations, having shown to correlate with physical characteristics of imaged objects, can also be fine-tuned via transfer learning to adapt to specific tasks e.g., classification~\citep{wei2020deep} or segmentation~\citep{hausen2020morpheus}.
+%
 Finally, while the CLIP model is restricted to retrieving nearest-neighbour associations within and across text/image modalities, the learned embeddings can be used as a starting point for training or fine-tuning multi-modal large-language models for interacting with survey data and receiving responses in natural language form, as well as grounding the responses based on an existing set of observations.
-
-% \subsubsection*{Broader Impact Statement}
-% In this optional section, TMLR encourages authors to discuss possible repercussions of their work,
-% notably any potential negative impact that a user of this research should be aware of. 
-% Authors should consult the TMLR Ethics Guidelines available on the TMLR website
-% for guidance on how to approach this subject.
-
-% Should write BIS!
-
-% \subsubsection*{Author Contributions}
-% If you'd like to, you may include a section for author contributions as is done
-% in many journals. This is optional and at the discretion of the authors. Only add
-% this information once your submission is accepted and deanonymized.
 
@@
 We thank Michael Brenner for helpful conversations.
+%
 This work is supported by the National Science Foundation under Cooperative Agreement PHY-2019786 (The NSF AI Institute for Artificial Intelligence and Fundamental Interactions, \url{http://iaifi.org/}).
-This material is based upon work supported by the U.S.
-Department of Energy, Office of Science, Office of High Energy Physics of U.S.
-Department of Energy under grant Contract Number  DE-SC0012567.
+%
+This material is based upon work supported by the U.S. Department of Energy, Office of Science, Office of High Energy Physics of U.S. Department of Energy under grant Contract Number  DE-SC0012567. 
+%
 YS was supported by the Research Science Institute (RSI) program at MIT.
+%
 The computations in this paper were run on the FASRC Cannon cluster supported by the FAS Division of Science Research Computing Group at Harvard University.
@@
 We list here the prompts and schemata (i.e.
+%
 desired output formats) used at various stages for guided text generation via \package{Outlines} package interfacing with the \textsc{Mixtral-8x7B-Instruct} LLM.
@@
 The following schema guides generation, intended to produce 100 concepts reflecting the science categories of successful proposals\footnote{\url{https://www.stsci.edu/contents/newsletters/2023-volume-40-issue-02/hubble-cycle-31-proposal-selection}}.
+%
 
@@
 The following curated categories are used in the text retrieval experiment in Sec.~\ref{sec:results}.
+%
 These are intended to be broader than the single-concept summaries inferred in Sec.~\ref{app:singleconcept} and are derived by prompting \textsc{Claude}, without any external input e.g.
+%
 proposal abstracts, to produce a list of categories corresponding to HST observations.
-% \SM{Try to just use the single-concept summaries here to avoid the extra step}.
 
@@
 
-% \begin{figure*}[!h]
-% \includegraphics[width=0.95\textwidth]{plots/retrieval_acc.pdf}
-% \caption{Retrieval accuracy}
-% \label{fig:retrieval_acc_supp}
-% \end{figure*}
-
 \end{document}

### 2024-02-29 paper/main.tex
Start edits, changed paragraph to subsection

@@
 
+\tableofcontents
+
 \section{Introduction}
@@
 
-\SM{Add TOC.}
-\SM{Change PAPERCLIP everywhere.}
 \SM{Add some qualitative evals with the abstract model.}
@@
 %
-The paradigm has been highly successful in domains like computer vision and natural language processing, as demonstrated by the widespread adoption of tools like ChatGPT, Dall-E, Stable Diffusion, and CLIP.
+The paradigm has been highly successful in domains like computer vision and natural language processing, as demonstrated by the widespread use and adoption of tools like ChatGPT, Dall-E, Stable Diffusion, and CLIP.
 %
@@
 
-The Contrastive Language-Image Pre-training~\citep[CLIP;][]{radford2021learning} family of foundation models, which in their original instantiation embed images and associated captions into a common embedding space via contrastive learning, have shown strong performance and generalization capabilities on a variety of downstream tasks including zero-shot classification and image retrieval.
+In this paper we describe \text{PAPERCLIP} (Proposal Abstracts Provide an Effective Representation for Contrastive Language-Image Pretraining\footnote{Technically we fine tune rather than pretrain, but ``PAPERCLIFT'' was rejected by the senior author of this paper.}), a method that connects astronomical image observations with natural language by leveraging the association between abstracts of successful observing proposals and corresponding downstream image observations.
+%
+Concretely, here we showcase the method using observations imaged by the \hubble Space Telescope (HST).
+%
+We show that fine-tuning a pre-trained CLIP (Contrastive Language-Image Pre-training)~\cite{radford2021learning} image-text model on observation-abstract associations results in meaningful joint representations through quantitative and qualitative evaluation.
+%
+This opens up the possibility of interacting with astronomical survey data using free-form natural language as an interface, which is a cornerstone of the success of the modern foundation model paradigm.
+%
+
+The CLIP family of foundation models, which in their original instantiation embed images and associated captions into a common embedding space via contrastive learning, have shown strong performance and generalization capabilities on a variety of downstream tasks including zero-shot classification and image retrieval.
 %
@@
 
-\SM{Put in this work we at the start of the paragraph. Also put the paragraph right after AstroLlama, reworking some of the logic.} \SM{Define PAPERCLIP, funny footnote about PAPERCLIFT.}
-
-The success of the foundation model paradigm relies in part on the ability to flexibly leverage text as a \emph{universal interface} for interacting with models.
-%
-In this work, we take this outlook and train a joint representation between observations taken by the \hubble Space Telescope (HST) and natural language.
-%
-We do so by leveraging associations between LLM-summarized abstracts of successful observing proposals and corresponding downstream image observations.
-%
-We show that fine-tuning a pre-trained CLIP image-text model on this data results in meaningful joint representations through quantitative and qualitative evaluation.
+% \SM{Put in this work we at the start of the paragraph. Also put the paragraph right after AstroLlama, reworking some of the logic.} \SM{Define PAPERCLIP, funny footnote about PAPERCLIFT.}
 
@@
 
-Examples of images and corresponding (summarized) captions are shown in Tab.~\ref{tab:dataset}. \SM{Push table earlier.}
+Examples of images and corresponding (summarized) captions are shown in Tab.~\ref{tab:dataset}. % \SM{Push table earlier.}
 %
@@
 This further motivates the need for fine-tuning on domain-specific data.
+
+\datafolder{./plots/data/}
+
+\begin{table}[h!]
+  \centering
+  \begin{tabular}{m{0.21\textwidth} p{1.9cm} p{1.9cm} m{8cm}}
+      \toprule
+      \centering \bfseries Image & \centering \bfseries Obs. cycle \\ (Year) & \centering \bfseries Prop. ID & \centering \bfseries LLM-extracted summary \tabularnewline
+      \midrule
+      \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_0.pdf} & \centering \input{\thedatafolder/cycle_0.txt} \\ (2019) & \centering \input{\thedatafolder/id_0.txt} &  {\scriptsize \input{\thedatafolder/sum_0.txt}} \tabularnewline
+      \midrule
+      \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_1.pdf} & \centering \input{\thedatafolder/cycle_1.txt} \\ (2013) & \centering \input{\thedatafolder/id_1.txt} &  {\scriptsize \input{\thedatafolder/sum_1.txt}} \tabularnewline
+      \midrule
+      \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_2.pdf} & \centering \input{\thedatafolder/cycle_2.txt} \\ (1999) & \centering \input{\thedatafolder/id_2.txt} &  {\scriptsize \input{\thedatafolder/sum_2.txt}} \tabularnewline
+      \midrule
+      \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_3.pdf} & \centering \input{\thedatafolder/cycle_3.txt} \\ (2016) & \centering \input{\thedatafolder/id_3.txt} &  {\scriptsize \input{\thedatafolder/sum_3.txt}} \tabularnewline
+      \bottomrule
+  \end{tabular}
+  \caption{Examples of images and corresponding captions, constructed using the LLM-extracted summaries.
+%
+The CLIP model is fine-tuned using these text-image associations.
+%
+Although the summarized abstract is extracted as two parts (objects/phenomena and science use cases), the model is trained using the entire summary as a single caption.}
+  \label{tab:dataset}
+\end{table}
 
@@
 3,194 images are held out for validation, with no abstract being common between training and validation sets in order to ensure an independent set of text-image pairs for evaluation.
-\SM{Put number of held out abstracts here}
+The held out images correspond to 438 unique abstracts.
+% Double check numbers in final version.
 
@@
 The schema is designed to represent a list of the objects (e.g., `Type Ia supernova') and phenomena (e.g., `gravitational lensing'), as well as potential downstream science uses cases (e.g., `set constraints on supernova explosion models') that could correspond to the eventual imaged observation given the abstract text, with a minimum of 1 and a maximum of 5 elements per list.
-\SM{Emphasize that the schema ensures the format, but the prompt is necessary for it to make sense.}
 
@@
 %
+While the schema-guided generation ensures the format of the output, the prompt and choice of LLM will influence the content of the generated summaries.
+%
 Further details on the summarization procedure, including the prompts and schema used, are provided in App.~\ref{app:summarization}.
@@
 The generation of the single concepts is different }
-
-\datafolder{./plots/data/}
-
-\begin{table}[h!]
-      \centering
-      \begin{tabular}{m{0.21\textwidth} p{1.9cm} p{1.9cm} m{8cm}}
-          \toprule
-          \centering \bfseries Image & \centering \bfseries Obs. cycle & \centering \bfseries Prop. ID & \centering \bfseries LLM-extracted summary \tabularnewline
-          \midrule
-          \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_0.pdf} & \centering \input{\thedatafolder/cycle_0.txt} & \centering \input{\thedatafolder/id_0.txt} &  {\scriptsize \input{\thedatafolder/sum_0.txt}} \tabularnewline
-          \midrule
-          \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_1.pdf} & \centering \input{\thedatafolder/cycle_1.txt} & \centering \input{\thedatafolder/id_1.txt} &  {\scriptsize \input{\thedatafolder/sum_1.txt}} \tabularnewline
-          \midrule
-          \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_2.pdf} & \centering \input{\thedatafolder/cycle_2.txt} & \centering \input{\thedatafolder/id_2.txt} &  {\scriptsize \input{\thedatafolder/sum_2.txt}} \tabularnewline
-          \midrule
-          \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_3.pdf} & \centering \input{\thedatafolder/cycle_3.txt} & \centering \input{\thedatafolder/id_3.txt} &  {\scriptsize \input{\thedatafolder/sum_3.txt}} \tabularnewline
-          \bottomrule
-      \end{tabular}
-      \caption{Examples of images and corresponding captions, constructed using the LLM-extracted summaries.
-%
-The CLIP model is fine-tuned using these text-image associations. \SM{Semicolon separates, but the model doesn't know about it.}}
-      \label{tab:dataset}
-  \end{table}
 
@@
           \end{tabular}
-          \caption{Examples of the initial parts of raw proposal abstracts (second column) and LLM (\textsc{Mixtral-8x7B})-extracted summaries (rightmost two columns), separately extracting objects and phenomena as well as potential downstream science use cases.
-\SM{Cite all proposal.} \SM{Add a year with the obs.
-cycle.}.}
+          \caption{Examples of the initial parts of raw proposal abstracts (second column) and LLM (\textsc{Mixtral-8x7B})-extracted summaries (rightmost two columns), separately extracting objects and phenomena as well as potential downstream science use cases.}
           \label{tab:datasetsumm}
@@
 
-\paragraph*{Pre-trained CLIP model}
+\subsection{Contrastive Language-Image Pretraining (CLIP)}
 
-CLIP \citep[Contrastive Language-Image Pretraining;][]{radford2021learning} is a multi-modal model pre-trained on a large corpus of image-text pairs via weak supervision using a contrastive loss.
+Contrastive Language-Image Pretraining \citep[CLIP;][]{radford2021learning} is a multi-modal model pre-trained on a large corpus of image-text pairs via weak supervision using a contrastive loss.
 %
@@
 
-\paragraph*{Fine-tuning procedure}
+\subsection{Fine-tuning procedure}
 
@@
 
-\paragraph*{Evaluation metrics}
+\subsection{Evaluation metrics}
 
@@
 
-\paragraph*{Validation metrics}
+\subsection{Validation metrics over training}
 
@@
 
-\paragraph*{Image retrieval}
+\subsection{Image retrieval task}
 
@@
 
-\paragraph*{Text retrieval}
+\subsection{Text retrieval task}
 
@@
 
-\paragraph*{Statistical distributions on the validation set}
+\subsection{Distribution of learned representations}
 
@@
 
-We present \textsc{HubbleCLIP}, a multi-modal fountation model that associates observations imaged by the \hubble Space Telescope with natural language in a common, semantically-meaningful embedding space.
+We present \textsc{PAPERCLIP}, a multi-modal fountation model that associates observations imaged by the \hubble Space Telescope with natural language in a common, semantically-meaningful embedding space.
 %
@@
 %
-We show that \textsc{HubbleCLIP} significantly outperforms the base CLIP model in quantitative metrics, such as retrieval accuracy, as well as quality of text-to-image and image-to-text retrieval.
+We show that \textsc{PAPERCLIP} significantly outperforms the base CLIP model in quantitative metrics, such as retrieval accuracy, as well as quality of text-to-image and image-to-text retrieval.
 %

### 2024-03-03 utils/text_utils.py
Plot updates

@@ -3,16 +3,21 @@ import jax
 import pandas as pd
 
 
-def process_truncate_captions(captions, key, max_length_words=None, use_sum1=False, df_sum_merged=None):
-    """ Process and truncate captions
-    """
+def process_truncate_captions(
+    captions, key, max_length_words=None, use_sum1=False, df_sum_merged=None
+):
+    """Process and truncate captions"""
     captions = captions.numpy().tolist()
     captions = [c.decode("utf-8") for c in captions]
 
     def get_objects_phenomena(caption):
-        first_part = caption.split(';')[0]
-        match = df_sum_merged[df_sum_merged['objects_phenomena_x'] == first_part]['objects_phenomena_y']
-        return match.values[0] if not (match.empty or pd.isna(match.values[0])) else "None"
+        first_part = caption.split(";")[0]
+        match = df_sum_merged[df_sum_merged["objects_phenomena_x"] == first_part][
+            "objects_phenomena_y"
+        ]
+        return (
+            match.values[0] if not (match.empty or pd.isna(match.values[0])) else "None"
+        )
 
     # Get sum1 summaries by matching
     if use_sum1 and df_sum_merged is not None:
@@ -25,22 +30,30 @@ def process_truncate_captions(captions, key, max_length_words=None, use_sum1=Fal
     return captions
 
 
-def tokenize_captions(captions, tokenizer, max_length_tokens, max_length_words=None, key=None):
-    """ Tokenizer if using custom CLIP model
-    """
-    captions = process_truncate_captions(captions, key, max_length_words=max_length_words)
-    captions = tokenizer(captions, padding="max_length", truncation=True, max_length=max_length_tokens, return_tensors="np")
+def tokenize_captions(
+    captions, tokenizer, max_length_tokens, max_length_words=None, key=None
+):
+    """Tokenizer if using custom CLIP model"""
+    captions = process_truncate_captions(
+        captions, key, max_length_words=max_length_words
+    )
+    captions = tokenizer(
+        captions,
+        padding="max_length",
+        truncation=True,
+        max_length=max_length_tokens,
+        return_tensors="np",
+    )
     return captions["input_ids"], captions["attention_mask"]
 
 
 def sample_and_pad(captions, key, max_length=77):
-    """ Sample a random sequence of sentences up to max_length in length
-    """
+    """Sample a random sequence of sentences up to max_length in length"""
     sampled_seqs = []
 
     for caption in captions:
         sentences = nltk.sent_tokenize(caption)
-        
+
[…]

### 2024-03-03 paper/main.tex

@@
 %
-Although the summarized abstract is extracted as two parts (objects/phenomena and science use cases), the model is trained using the entire summary as a single caption.}
+Although the summarized abstract is extracted as two parts (objects/phenomena and science use cases), the model is trained using the entire summary as a single caption.
+\SM{\hubble image should be mentioned (maybe in column title)} \SM{Just have the raw abstract, not summarized here.}}
   \label{tab:dataset}
@@
 
-\subsection{Data Selection and Pre-Processing}
+\subsection{\hubble Data Selection and Pre-Processing}
 
@@
           \end{tabular}
-          \caption{Examples of the initial parts of raw proposal abstracts (second column) and LLM (\textsc{Mixtral-8x7B})-extracted summaries (rightmost two columns), separately extracting objects and phenomena as well as potential downstream science use cases.}
+          \caption{Examples of the initial parts of raw proposal abstracts (second column) and LLM (\textsc{Mixtral-8x7B})-extracted summaries (rightmost two columns), separately extracting objects and phenomena as well as potential downstream science use cases.
+\SM{No need to have the raw abstracts here now, just the summaries.}}
           \label{tab:datasetsumm}
@@
 The model was originally trained on 400 million image-text pairs from internet data.
-\SM{$\tau$ was initialized to the pre-trained value.} \SM{HuggingFace link?} \SM{What area is maintained.} \SM{It's only 90 deg rotations.}
 %
@@
 %
-Images are randomly cropped to the native resolution of the image encoder and randomly rotated at each training step.
+Images are randomly cropped to the native resolution of the image encoder (maintaining $\sim 20\%$ of the area of the original image) and randomly rotated by discrete increments of $90^\circ$ at each training step.
 %
@@
 We note that the positive and negative image-text association is noisy and imperfect, since multiple images can be associated with the same abstract.
+The temperature hyperparameter $\tau$ was initialized to its pre-trained value.
 
@@
 
-We also qualitatively evaluate the learned embeddings through image retrieval (i.e., retrieving the most relevant images from a set using natural language queries) and description retrieval (i.e., querying the astrophysical object classes and science use cases most relevant to a given observation) experiments.
+We also qualitatively evaluate the learned embeddings through image retrieval (i.e., retrieving the most relevant images from the validation set using natural language queries) and description retrieval (i.e., querying the astrophysical object classes and science use cases most relevant to a given observation, akin to zero-shot classification) experiments.
+%
+For the description/text retrieval evaluation, we curate a list of possible text associations (i.e., classes) by querying the \textsc{Claude}\footnote{\url{https://claude.ai/}} large language model for such a list, which we show in App.~\ref{app:categories}.
+
 \SM{Add some expressions.}
@@
 
-\SM{Numbered subsections.}
-
 \section{Results and Discussion}
@@
 
-\subsection{Validation metrics over training}
+\subsection{Validation metrics during model training}
 
@@
 %
-The purple lines show the metrics evaluated when training with batches where the image-text associations are randomly shuffled, serving as a baseline.
+The purple lines show the metrics evaluated when training with batches where the image-text associations are randomly shuffled, serving as a baseline. \SM{Avoid the word baseline.}
 %
@@
 The model with LLM-guided summarization (yellow lines) is seen to perform on-par with the model using raw abstracts as captions (orange line), despite the stronger association signal in the summarized dataset curated.
+\SM{Version baseline}
 %
-Fine-tuning a small MLP head over frozen vision and text backbones (green lines), training from scratch (blue lines), and using the single-concept summaries (red lines) show a non-trivial improvement compared to the random baseline as well as base model, but with deteriorated performance compared to fine-tuning with either summarized or raw abstracts.
+Fine-tuning a small MLP head over frozen vision and text backbones (green lines), training from scratch (blue lines), and using the single-concept summaries (red lines) show a non-trivial improvement compared to the random baseline \SM{Maybe random baseline is okay} as well as base model, but with deteriorated performance compared to fine-tuning with either summarized or raw abstracts.
 
@@
 Shown for the dataset with summarized abstracts (orange), dataset using raw proposal abstracts (yellow), dataset with single-concept summaries (red), only fine-tuning a small MLP head (green), training from scratch (blue), and trained with shuffled image-text pairs (purple).
-\SM{Try not splitting the legend.} \SM{Would dashing be useful at all as a way to signify relationships.}}
+\SM{Try not splitting the legend.} \SM{Would dashing be useful at all as a way to signify relationships.} \SM{Put an arrow for what is better.} \SM{Mention that fine tuning head isn't that good.}}
   \label{fig:retrieval_acc}
@@
 Cluster-scale as well as galaxy-scale gravitational lenses are returned by the `strong lensing' query, with lensing patterns visible in the images.
+
+\SM{Describe science usefulness, see appendix.}
 
@@
 The distributions for the base model is seen to be sharply peaked at a specific value, showing little diversity and being very similar between the shuffled (dashed blue) and non-shuffled (solid blue) versions.
+\SM{Make very clear that it is the evals that are shuffled, not batches during training.}
 %
@@
 \caption{Image retrieval using the base CLIP model on four curated queries.
-\SM{Horizontal lines} \SM{Query not rotated} \SM{Same for next Figure.} \SM{Put text ``Base Model'' in consistent colour to legend.}}
+\SM{Horizontal lines} \SM{Query not rotated} \SM{Same for next Figure.} \SM{Put text ``Base Model'' in consistent colour to legend.} \SM{Say top 4} \SM{Say that for science we care about the distributions of things.}}
 \label{fig:tti_base}
@@
 The `ground truth' LLM-summarized abstract is shown in the right column.
-\SM{Swap order of cols 2 and 3.}}
+\SM{Say ground truth summarized abstract in the column}\SM{Swap order of cols 2 and 3.}}
 \label{fig:itt}
@@
 (Bottom) For the single-concept summaries, when evaluated using the CLIP model fine-tuned on LLM-summarized abstracts (red) and fine-tuned on the single-concept summaries themselves (purple).
-\SM{Make the colours here consistent with previous Figure.}}
+\SM{Make the colours here consistent with previous Figure.} \SM{Language should be correct pairs and incorrect pairs in top left.} \SM{Totally axe single concept.}}
 \label{fig:sim_valtrain}
@@
 We also introduce a novel LLM summarization process which leverages constrained generation to increase the association strength between images and captions while preserving salient information.
+\SM{We don't actually show this.}
 %

### 2024-03-03 paper/main.tex
add table

@@
 \usepackage{multirow} 
+\usepackage{multicol}
 \usepackage{listings}
@@
       \toprule
-      \centering \bfseries Image & \centering \bfseries Obs. cycle \\ (Year) & \centering \bfseries Prop. ID & \centering \bfseries LLM-extracted summary \tabularnewline
+      \centering \bfseries \hubble Image & \centering \bfseries Obs. cycle \\ (Year) & \centering \bfseries Prop. ID & \centering \bfseries Proposal abstract (clipped) \tabularnewline
       \midrule
-      \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_0.pdf} & \centering \input{\thedatafolder/cycle_0.txt} \\ (2019) & \centering \input{\thedatafolder/id_0.txt} &  {\scriptsize \input{\thedatafolder/sum_0.txt}} \tabularnewline
+      \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_0.pdf} & \centering \input{\thedatafolder/cycle_0.txt} \\ (2019) & \centering \input{\thedatafolder/id_0.txt} &  {\scriptsize \input{\thedatafolder/abs1_0.txt}} \tabularnewline
       \midrule
-      \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_1.pdf} & \centering \input{\thedatafolder/cycle_1.txt} \\ (2013) & \centering \input{\thedatafolder/id_1.txt} &  {\scriptsize \input{\thedatafolder/sum_1.txt}} \tabularnewline
+      \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_1.pdf} & \centering \input{\thedatafolder/cycle_1.txt} \\ (2013) & \centering \input{\thedatafolder/id_1.txt} &  {\scriptsize \input{\thedatafolder/abs1_1.txt}} \tabularnewline
       \midrule
-      \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_2.pdf} & \centering \input{\thedatafolder/cycle_2.txt} \\ (1999) & \centering \input{\thedatafolder/id_2.txt} &  {\scriptsize \input{\thedatafolder/sum_2.txt}} \tabularnewline
+      \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_2.pdf} & \centering \input{\thedatafolder/cycle_2.txt} \\ (1999) & \centering \input{\thedatafolder/id_2.txt} &  {\scriptsize \input{\thedatafolder/abs1_2.txt}} \tabularnewline
       \midrule
-      \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_3.pdf} & \centering \input{\thedatafolder/cycle_3.txt} \\ (2016) & \centering \input{\thedatafolder/id_3.txt} &  {\scriptsize \input{\thedatafolder/sum_3.txt}} \tabularnewline
+      \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_3.pdf} & \centering \input{\thedatafolder/cycle_3.txt} \\ (2016) & \centering \input{\thedatafolder/id_3.txt} &  {\scriptsize \input{\thedatafolder/abs1_3.txt}} \tabularnewline
       \bottomrule
@@
 %
-Although the summarized abstract is extracted as two parts (objects/phenomena and science use cases), the model is trained using the entire summary as a single caption.
-\SM{\hubble image should be mentioned (maybe in column title)} \SM{Just have the raw abstract, not summarized here.}}
+Although the summarized abstract is extracted as two parts (objects/phenomena and science use cases), the model is trained using the entire summary as a single caption.}
   \label{tab:dataset}
@@
 This ensures that the output of the LLM strictly conforms to the format of the following example:
+
 \begin{lstlisting}[language=Python]
@@
 \end{lstlisting}
+
 which is then used to construct the summarized caption by combining the two key elements.
@@
 
-Finally, in order to test a further compression of the proposal abstracts, we use LLM-guided generation to also produce a list of single-concept summaries (e.g., `irregular galaxy', `quasars', `Galactic bulge', \ldots), and test whether this can lead to meaningful, generalizable learned associations with observations.
-%
-The prompts and schemata for generating these discrete categories and assigning observations to them are described in Apps.~\ref{app:singleconcept} and \ref{app:singleconceptassignments} respectively.
+% Finally, in order to test a further compression of the proposal abstracts, we use LLM-guided generation to also produce a list of single-concept summaries (e.g., `irregular galaxy', `quasars', `Galactic bulge', \ldots), and test whether this can lead to meaningful, generalizable learned associations with observations.
+% %
+% The prompts and schemata for generating these discrete categories and assigning observations to them are described in Apps.~\ref{app:singleconcept} and \ref{app:singleconceptassignments} respectively.
 \SM{Each abstract is assigned one summary.
@@
       \begin{table}[h!]
+        \renewcommand{\arraystretch}{2}
           \centering
@@
           \end{tabular}
-          \caption{Examples of the initial parts of raw proposal abstracts (second column) and LLM (\textsc{Mixtral-8x7B})-extracted summaries (rightmost two columns), separately extracting objects and phenomena as well as potential downstream science use cases.
-\SM{No need to have the raw abstracts here now, just the summaries.}}
+          \caption{Examples of the initial parts of raw proposal abstracts (second column) and LLM (\textsc{Mixtral-8x7B})-extracted summaries (rightmost two columns), separately extracting objects and phenomena as well as potential downstream science use cases.}
           \label{tab:datasetsumm}
@@
 
-\begin{figure*}[!h]
-\includegraphics[width=0.95\textwidth]{plots/tti.pdf}
-\caption{Image retrieval using the fine-tuned CLIP model on four curated queries.}
-\label{fig:tti}
-\end{figure*}
+% \begin{figure*}[!h]
+% \includegraphics[width=0.95\textwidth]{plots/tti.pdf}
+% \caption{Image retrieval using the fine-tuned CLIP model on four curated queries.}
+% \label{fig:tti}
+% \end{figure*}
+
+\begin{table}[h!]
+  \centering
+  \begin{tabular}{m{3cm} p{3cm} p{3cm} p{3cm} p{3cm}}
+      \toprule
+      \centering \bfseries Query & \multicolumn{4}{c}{\bfseries{Top-4 most similar images}} \tabularnewline
+      \midrule
+       \texttt{\input{\thedatafolder/query_tti_0.txt}} \vspace{20mm} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_0_0.pdf} \\ \input{\thedatafolder/propid_tti_0_0.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_0_1.pdf} \\ \input{\thedatafolder/propid_tti_0_1.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_0_2.pdf} \\ \input{\thedatafolder/propid_tti_0_2.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_0_3.pdf} \\ \input{\thedatafolder/propid_tti_0_3.txt}  \tabularnewline
+      \midrule
+      \texttt{\input{\thedatafolder/query_tti_1.txt}} \vspace{20mm} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_1_0.pdf} \\ \input{\thedatafolder/propid_tti_1_0.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_1_1.pdf} \\ \input{\thedatafolder/propid_tti_1_1.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_1_2.pdf} \\ \input{\thedatafolder/propid_tti_1_2.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_1_3.pdf} \\ \input{\thedatafolder/propid_tti_1_3.txt}  \tabularnewline
+      \midrule
+      \texttt{\input{\thedatafolder/query_tti_2.txt}} \vspace{20mm} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_2_0.pdf} \\ \input{\thedatafolder/propid_tti_2_0.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_2_1.pdf} \\ \input{\thedatafolder/propid_tti_2_1.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_2_2.pdf} \\ \input{\thedatafolder/propid_tti_2_2.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_2_3.pdf} \\ \input{\thedatafolder/propid_tti_2_3.txt}  \tabularnewline
+      \midrule
+      \texttt{\input{\thedatafolder/query_tti_3.txt}} \vspace{20mm} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_3_0.pdf} \\ \input{\thedatafolder/propid_tti_3_0.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_3_1.pdf} \\ \input{\thedatafolder/propid_tti_3_1.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_3_2.pdf} \\ \input{\thedatafolder/propid_tti_3_2.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_3_3.pdf} \\ \input{\thedatafolder/propid_tti_3_3.txt}  \tabularnewline
+      \bottomrule
+  \end{tabular}
+  \caption{Image retrieval using the fine-tuned CLIP model on four curated queries}
+  \label{tab:tti}
+\end{table}
 
@@
 \includegraphics[width=0.45\textwidth]{plots/sim_val.pdf}
-\includegraphics[width=0.45\textwidth]{plots/sim_summ1.pdf}
-\centering\includegraphics[width=0.45\textwidth]{plots/retrieval.pdf}
+% \includegraphics[width=0.45\textwidth]{plots/sim_summ1.pdf}
+\includegraphics[width=0.45\textwidth]{plots/retrieval.pdf}
 \caption{Distribution of cosine similarities between corresponding image and text embeddings, $x_i$ and $y_i$.
@@
 Versions with the ordering of text embeddings shuffled are shown in dashed.
-(Bottom) For the single-concept summaries, when evaluated using the CLIP model fine-tuned on LLM-summarized abstracts (red) and fine-tuned on the single-concept summaries themselves (purple).
-\SM{Make the colours here consistent with previous Figure.} \SM{Language should be correct pairs and incorrect pairs in top left.} \SM{Totally axe single concept.}}
+\SM{Make the colours here consistent with previous Figure.} \SM{Language should be correct pairs and incorrect pairs in top left.}}
 \label{fig:sim_valtrain}

### 2024-03-03 paper/main.tex
merge

@@
 \usepackage{listings}
+\usepackage{verbatim}
 
@@
 
-\begin{figure*}[!h]
-\includegraphics[width=0.95\textwidth]{plots/tti_base.pdf}
+\begin{table}[h!]
+  \centering
+  \begin{tabular}{m{3cm} p{3cm} p{3cm} p{3cm} p{3cm}}
+      \toprule
+      \centering \bfseries Query & \multicolumn{4}{c}{\bfseries{Top-4 most similar images}} \tabularnewline
+      \midrule
+       \texttt{\input{\thedatafolder/query_tti_base_0.txt}} \vspace{20mm} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_0_0.pdf} \\ \input{\thedatafolder/propid_tti_base_0_0.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_0_1.pdf} \\ \input{\thedatafolder/propid_tti_base_0_1.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_0_2.pdf} \\ \input{\thedatafolder/propid_tti_base_0_2.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_0_3.pdf} \\ \input{\thedatafolder/propid_tti_base_0_3.txt}  \tabularnewline
+      \midrule
+      \texttt{\input{\thedatafolder/query_tti_base_1.txt}} \vspace{20mm} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_1_0.pdf} \\ \input{\thedatafolder/propid_tti_base_1_0.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_1_1.pdf} \\ \input{\thedatafolder/propid_tti_base_1_1.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_1_2.pdf} \\ \input{\thedatafolder/propid_tti_base_1_2.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_1_3.pdf} \\ \input{\thedatafolder/propid_tti_base_1_3.txt}  \tabularnewline
+      \midrule
+      \texttt{\input{\thedatafolder/query_tti_base_2.txt}} \vspace{20mm} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_2_0.pdf} \\ \input{\thedatafolder/propid_tti_base_2_0.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_2_1.pdf} \\ \input{\thedatafolder/propid_tti_base_2_1.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_2_2.pdf} \\ \input{\thedatafolder/propid_tti_base_2_2.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_2_3.pdf} \\ \input{\thedatafolder/propid_tti_base_2_3.txt}  \tabularnewline
+      \midrule
+      \texttt{\input{\thedatafolder/query_tti_base_3.txt}} \vspace{20mm} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_3_0.pdf} \\ \input{\thedatafolder/propid_tti_base_3_0.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_3_1.pdf} \\ \input{\thedatafolder/propid_tti_base_3_1.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_3_2.pdf} \\ \input{\thedatafolder/propid_tti_base_3_2.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_3_3.pdf} \\ \input{\thedatafolder/propid_tti_base_3_3.txt}  \tabularnewline
+      \bottomrule
+  \end{tabular}
   \caption{Image retrieval using the base CLIP model on four curated queries.
 \SM{Horizontal lines} \SM{Query not rotated} \SM{Same for next Figure.} \SM{Put text ``Base Model'' in consistent colour to legend.} \SM{Say top 4} \SM{Say that for science we care about the distributions of things.}}
-\label{fig:tti_base}
-\end{figure*}
+  \label{tab:tti_base}
+\end{table}
+
+% \begin{figure*}[!h]
+% \includegraphics[width=0.95\textwidth]{plots/tti_base.pdf}
+% \caption{Image retrieval using the base CLIP model on four curated queries.
+\SM{Horizontal lines} \SM{Query not rotated} \SM{Same for next Figure.} \SM{Put text ``Base Model'' in consistent colour to legend.} \SM{Say top 4} \SM{Say that for science we care about the distributions of things.}}
+% \label{fig:tti_base}
+% \end{figure*}
 
@@
 
-\begin{figure*}[!h]
-\includegraphics[width=0.95\textwidth]{plots/itt.pdf}
+% \begin{figure*}[!h]
+% \includegraphics[width=0.95\textwidth]{plots/itt.pdf}
+% \caption{Text associations from a curated list most closely matching a given image query, for both the fine-tuned and base models.
+The `ground truth' LLM-summarized abstract is shown in the right column.
+\SM{Say ground truth summarized abstract in the column}\SM{Swap order of cols 2 and 3.}}
+% \label{fig:itt}
+% \end{figure*}
+
+% Set line spacing to 0.5
+
+\begin{table}[h!]
+  \centering
+  \renewcommand{\arraystretch}{0.1}
+  \begin{tabular}{m{3cm} m{3cm} m{3cm} m{6cm}}
+      \toprule
+      \centering \bfseries \hubble Image & \centering \bfseries Obs. cycle \\ (Year) & \centering \bfseries Prop. ID & \centering \bfseries Proposal abstract (clipped) \tabularnewline
+      \midrule
+      \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_0.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_0.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_0.txt} &  {\scriptsize \input{\thedatafolder/abs_itt_0.txt}} \tabularnewline
+      \midrule
+      \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_1.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_1.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_1.txt} &  {\scriptsize \input{\thedatafolder/abs_itt_1.txt}} \tabularnewline
+      \midrule
+      \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_2.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_2.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_2.txt} &  {\scriptsize \input{\thedatafolder/abs_itt_2.txt}} \tabularnewline
+      \midrule
+      \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_3.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_3.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_3.txt} &  {\scriptsize \input{\thedatafolder/abs_itt_3.txt}} \tabularnewline
+      \bottomrule
+  \end{tabular}
   \caption{Text associations from a curated list most closely matching a given image query, for both the fine-tuned and base models.
@@
 \SM{Say ground truth summarized abstract in the column}\SM{Swap order of cols 2 and 3.}}
-\label{fig:itt}
-\end{figure*}
+  \label{tab:itt}
+\end{table}
 
@@
 
-\subsection{Generation of single-concept summaries}
-\label{app:singleconcept}
+% \subsection{Generation of single-concept summaries}
+% \label{app:singleconcept}
 
-The following prompt is used to generate a list of diverse single-concept summaries informed by a list of summarized abstracts:
+% The following prompt is used to generate a list of diverse single-concept summaries informed by a list of summarized abstracts:
 
-\begin{lstlisting}[language=Python]
-import outlines
+% \begin{lstlisting}[language=Python]
+% import outlines
 
-@outlines.prompt
-def prompt_fn(objects):
-    """<s>[INST] Please produce a list of around concepts characterizing prominent objects, phenomena, and science use cases of images observed by the Hubble Space Telescope.
+% @outlines.prompt
+% def prompt_fn(objects):
+%     """<s>[INST] Please produce a list of around concepts characterizing prominent objects, phenomena, and science use cases of images observed by the Hubble Space Telescope.
 
-Here are some examples of objects:
+% Here are some examples of objects:
 
-{{objects}}
+% {{objects}}
 
-Follow these instructions exactly in your answer:
-- Do not output empty strings as elements.
-- Make sure that the list covers a diverse range of astronomical concepts, with items as different from each other as possible. 
-- Do not give specific names of objects, to make sure you span the widest possible range of concepts (e.g., "dwarf galaxy" is allowed, but NOT "Fornax", "Terzan 5", or  "NGC6440").
-- Do not return terms undescriptive of observations, e.g. "sloshing", "adiabatic", "interactions". Returning concrete physics objects, concepts, or phenomena.
-- Only output scientifically meaningful terms. E.g., NO "Cosmic Dance".
-- Do not duplicate entries. Do not reference any telescopes, observatories, or surveys.
-- Do not include units like "angular diameter distance", "parsec", or any other concepts that will not correlate with images of observations.
-- Use the above example list of objects only as inspiration to infer broad classes of objects.
-- Make sure each concept is succint, never more than 5 words.
-- Answer in JSON format.
-- The JSON should have the following keys {"galaxies", "stellar_physics", "exoplanets_planet_formation", "stellar_populations", "supermassive_black_holes", "solar_system", "integalactic_medium", "large_scale_structure"} reflecting rough observation categories.
-- Each category will have a list of objects and/or astronomical concepts.
-- Output up to 20 items and no more in each category
-[/INST]
-"""
-\end{lstlisting}
+% Follow these instructions exactly in your answer:
+% - Do not output empty strings as elements.
+% - Make sure that the list covers a diverse range of astronomical concepts, with items as different from each other as possible. 
+% - Do not give specific names of objects, to make sure you span the widest possible range of concepts (e.g., "dwarf galaxy" is allowed, but NOT "Fornax", "Terzan 5", or  "NGC6440").
+% - Do not return terms undescriptive of observations, e.g. "sloshing", "adiabatic", "interactions". Returning concrete physics objects, concepts, or phenomena.
+% - Only output scientifically meaningful terms. E.g., NO "Cosmic Dance".
+% - Do not duplicate entries. Do not reference any telescopes, observatories, or surveys.
+% - Do not include units like "angular diameter distance", "parsec", or any other concepts that will not correlate with images of observations.
+% - Use the above example list of objects only as inspiration to infer broad classes of objects.
+% - Make sure each concept is succint, never more than 5 words.
+% - Answer in JSON format.
+% - The JSON should have the following keys {"galaxies", "stellar_physics", "exoplanets_planet_formation", "stellar_populations", "supermassive_black_holes", "solar_system", "integalactic_medium", "large_scale_structure"} reflecting rough observation categories.
+% - Each category will have a list of objects and/or astronomical concepts.
+% - Output up to 20 items and no more in each category
+% [/INST]
+% """
+% \end{lstlisting}
 
-The following schema guides generation, intended to produce 100 concepts reflecting the science categories of successful proposals\footnote{\url{https://www.stsci.edu/contents/newsletters/2023-volume-40-issue-02/hubble-cycle-31-proposal-selection}}.
-%
+% The following schema guides generation, intended to produce 100 concepts reflecting the science categories of successful proposals\footnote{\url{https://www.stsci.edu/contents/newsletters/2023-volume-40-issue-02/hubble-cycle-31-proposal-selection}}.
+% %
 
-\begin{lstlisting}[language=Python]
-from pydantic import BaseModel, conlist
+% \begin{lstlisting}[language=Python]
+% from pydantic import BaseModel, conlist
 
-class ScienceCategoriesHST(BaseModel):
-    galaxies: conlist(str, min_length=15, max_length=15)
-    stellar_physics: conlist(str, min_length=15, max_length=15)
-    exoplanets_planet_formation: conlist(str, min_length=15, max_length=15)
-    stellar_populations: conlist(str, min_length=10, max_length=10)
-    supermassive_black_holes: conlist(str, min_length=15, max_length=15)
-    solar_system: conlist(str, min_length=10, max_length=10)
-    integalactic_medium: conlist(str, min_length=10, max_length=10)
-    large_scale_structure: conlist(str, min_length=10, max_length=10)
-  \end{lstlisting}
+% class ScienceCategoriesHST(BaseModel):
+%     galaxies: conlist(str, min_length=15, max_length=15)
+%     stellar_physics: conlist(str, min_length=15, max_length=15)
+%     exoplanets_planet_formation: conlist(str, min_length=15, max_length=15)
+%     stellar_populations: conlist(str, min_length=10, max_length=10)
+%     supermassive_black_holes: conlist(str, min_length=15, max_length=15)
+%     solar_system: conlist(str, min_length=10, max_length=10)
+%     integalactic_medium: conlist(str, min_length=10, max_length=10)
+%     large_scale_structure: conlist(str, min_length=10, max_length=10)
+%   \end{lstlisting}
 
-The output is then constrained to be one of the fixed number of concepts.
+% The output is then constrained to be one of the fixed number of concepts.
 
-\subsection{Assignment of abstracts to single-concept categories}
-\label{app:singleconceptassignments}
+% \subsection{Assignment of abstracts to single-concept categories}
+% \label{app:singleconceptassignments}
 
-Finally, the following prompt is used to assign concepts inferred in \ref{app:singleconcept} to each abstract:
+% Finally, the following prompt is used to assign concepts inferred in \ref{app:singleconcept} to each abstract:
 
-\begin{lstlisting}[language=Python]
-import outlines
+% \begin{lstlisting}[language=Python]
+% import outlines
 
-@outlines.prompt
-def prompt_fn(abs, cats):
-    """<s>[INST] The following is a successful proposal abstract for the Hubble Space Telescope: "{{abs}}"
+% @outlines.prompt
+% def prompt_fn(abs, cats):
+%     """<s>[INST] The following is a successful proposal abstract for the Hubble Space Telescope: "{{abs}}"
 
-The following is a list of categories (astronomical concepts) that this abstract could correspond to.
+% The following is a list of categories (astronomical concepts) that this abstract could correspond to.
 
-{{cats}}
+% {{cats}}
 
-Please answer which of these listed concepts best describes this proposal, based on the objects and phenomena mentioned in the abstract.
-The concept should meaningfully be present in the abstract and the eventual observation.
+% Please answer which of these listed concepts best describes this proposal, based on the objects and phenomena mentioned in the abstract.
+% The concept should meaningfully be present in the abstract and the eventual observation.
 
-- For example, "The locations of supernovae {SNe} in the local stellar and gaseous environment in galaxies, as measured in high spatial resolution WFPC2 and ACS images, contain important clues to their progenitor stars." should return "supernova".
-- If the abstract centers calibration and/or instrumentation efforts, return calibration or instrumention".
+% - For example, "The locations of supernovae {SNe} in the local stellar and gaseous environment in galaxies, as measured in high spatial resolution WFPC2 and ACS images, contain important clues to their progenitor stars." should return "supernova".
+% - If the abstract centers calibration and/or instrumentation efforts, return calibration or instrumention".
 
-If no concept make sense, return "None". [/INST]
-"""
-\end{lstlisting}
+% If no concept make sense, return "None". [/INST]
+% """
+% \end{lstlisting}
 
@@
 %
-These are intended to be broader than the single-concept summaries inferred in Sec.~\ref{app:singleconcept} and are derived by prompting \textsc{Claude}, without any external input e.g.
+These are derived by prompting \textsc{Claude}, without any external input e.g.
 %

### 2024-03-04 paper/main.tex
Pass through the text

@@
 \usepackage{verbatim}
+\usepackage{caption}
+\usepackage{bbm}
 
@@
 \usepackage{color}
+
 \definecolor{deepblue}{rgb}{0,0,0.5}
@@
 \definecolor{deepgreen}{rgb}{0,0.5,0}
+\definecolor{deeppurple}{rgb}{0.60,0.42,0.63}
 
 \definecolor{xlinkcolor}{rgb}{0.7752941176470588, 0.22078431372549023, 0.2262745098039215}
+
 \definecolor{codegreen}{rgb}{0,0.4,0}
@@
 \begin{abstract}
-We present PAPERCLIP (Proposal Abstracts Provide an Effective Representation for Contrastive Language-Image Pretraining), a method which associates astronomical observations imaged by astronomical surveys like the \hubble Space Telescope (HST) with natural language.
-The model is fine-tuned from a pre-trained Contrastive Language–Image Pre-training (CLIP) model using successful proposal abstracts corresponding to HST observations, optionally summarized via guided generation using large language models (LLMs).
-We show that the model embodies a meaningful joint representation between observations and text through experiments targeting image retrieval (i.e., retrieving the most relevant observations from a set using natural language queries) and description retrieval (i.e., querying the astrophysical object classes and science use cases most relevant to a given observation).
-Our study demonstrates the potential for using generalist rather than task-specific models for astrophysics research, in particular by leveraging text as an interface.
+We present PAPERCLIP (Proposal Abstracts Provide an Effective Representation for Contrastive Language-Image Pre-training), a method which associates astronomical observations imaged by surveys and telescopes with natural language.
+The model is fine-tuned from a pre-trained Contrastive Language–Image Pre-training (CLIP) model using successful proposal abstracts, optionally summarized via guided generation using large language models (LLMs), and corresponding downstream observations.
+Using observations from the \hubble Space Telescope (HST) as an example, we show that the model embodies a meaningful joint representation between observations and text through experiments targeting image retrieval (i.e., retrieving the most relevant observations from a set using natural language queries) and description retrieval (i.e., querying the astrophysical object classes and science use cases most relevant to a given observation).
+Our study demonstrates the potential for using generalist rather than task-specific models for finding patterns in astronomical data, in particular by leveraging text as an interface.
 \end{abstract}
@@
 
-\SM{Add some qualitative evals with the abstract model.}
+% \SM{Add some qualitative evals with the abstract model.}
 
-Machine learning (ML) is beginning to have a significant impact in the sciences, with astrophysics being no exception.
+Machine learning (ML) is starting to have a significant impact in the sciences, with astrophysics being no exception.
 %
@@
 %
-Until recently, most applications of ML within astrophysics have focused on replacing or augmenting traditional techniques with ML counterparts in order to improve performance on specific, individual tasks.
+Until recently, most applications of ML within astrophysics have focused on augmenting traditional techniques in order to improve performance on specific tasks.
 %
-The \emph{Foundation Model} paradigm, in contrast, seeks to develop generalist models which can be deployed to tackle a wide range of tasks.
+The {Foundation Model} paradigm, in contrast, seeks to develop generalist models which can be deployed to simultaneously tackle a wide range of tasks \citep{bommasani2021opportunities}.
 %
-The paradigm has been highly successful in domains like computer vision and natural language processing, as demonstrated by the widespread use and adoption of tools like ChatGPT, Dall-E, Stable Diffusion, and CLIP.
+% The paradigm has been highly successful in domains like computer vision and natural language processing, as demonstrated by the widespread use and adoption of tools like ChatGPT, Dall-E, Stable Diffusion, and CLIP.
 %
-These models are typically pre-trained on massive amounts of unlabeled data using self-supervised learning techniques, enabling them to learn powerful representations which can be deployed on different tasks across domains.
-%
-At the heart of the paradigm lies the triumph of scale -- scaling up model size, dataset size, and compute.
+% At the heart of the paradigm lies the triumph of scale -- scaling up model size, dataset size, and compute.
+% %
+These models are typically pre-trained on massive amounts of unlabeled data using self-supervised or weakly-supervised learning techniques, enabling them to learn powerful representations which can be deployed on different tasks across domains.
 %
@@
 %
-The multi-modality inherent to the nature of astrophysical observations, with different types of data (e.g., images, spectra, light curves, textual descriptions) often available for a given object, presents a unique opportunity.
+The multi-modality inherent to astrophysical observations, with different types of data (e.g., images, spectra, light curves, textual descriptions) often available for a given target object, presents a unique opportunity.
 %
@@
 
-In this paper we describe \text{PAPERCLIP} (Proposal Abstracts Provide an Effective Representation for Contrastive Language-Image Pretraining\footnote{Technically we fine tune rather than pretrain, but ``PAPERCLIFT'' was rejected by the senior author of this paper.}), a method that connects astronomical image observations with natural language by leveraging the association between abstracts of successful observing proposals and corresponding downstream image observations.
+In this paper we describe \text{PAPERCLIP} (Proposal Abstracts Provide an Effective Representation for Contrastive Language-Image Pre-training\footnote{Technically we fine tune rather than pre train, but ``PAPERCLIFT'' was rejected by the senior author of this paper.}), a method that connects astronomical image observations with natural language by leveraging the association between abstracts of successful observing proposals and images corresponding to downstream observations.
 %
@@
 %
-We show that fine-tuning a pre-trained CLIP (Contrastive Language-Image Pre-training)~\cite{radford2021learning} image-text model on observation-abstract associations results in meaningful joint representations through quantitative and qualitative evaluation.
+We show that fine-tuning a pre-trained CLIP ~\citep[Contrastive Language-Image Pre-training; ][]{radford2021learning} image-text model on observation-abstract associations results in meaningful joint representations through quantitative and qualitative evaluation.
 %
@@
 
-The CLIP family of foundation models, which in their original instantiation embed images and associated captions into a common embedding space via contrastive learning, have shown strong performance and generalization capabilities on a variety of downstream tasks including zero-shot classification and image retrieval.
+The CLIP family of foundation models, which in their original form embed images and associated captions into a common representation space via contrastive learning, have shown strong performance and generalization capabilities on a variety of downstream tasks including zero-shot classification and image retrieval.
 %
-The concept of associating diverse modalities via contrastive training has been used on other scientific domains~\citep[e.g.,][]{liu2023text,Sanchez-Fernandez2022.11.17.516915,lanusse2023astroclip,cepeda2023geoclip}, and has been shown to be effective in learning semantically meaningful joint representations.
+The concept of associating diverse modalities via contrastive training has been employed in other scientific domains~\citep[e.g.,][]{liu2023text,Sanchez-Fernandez2022.11.17.516915,lanusse2023astroclip,cepeda2023geoclip}, and has been shown to be effective in learning semantically meaningful joint representations.
+Here we present for the first an application associating astronomical data with the text modality.
 
@@
 %
-In Sec.~\ref{sec:dataset}, we describe the dataset used in this work, including its curation and processing.
+In Sec.~\ref{sec:dataset}, we describe the \hubble dataset used in this work, including the curation and processing of observations as well as text captions.
 %
@@
 %
-In Sec.~\ref{sec:results}, we present the results of our experiments on image and text retrieval tasks.
+In Sec.~\ref{sec:results}, we present quantitative and qualitative results of our experiments on image and text retrieval tasks.
 %
@@
 %
-We rely on proposal abstracts from the Proposal Abstracts Catalog\footnote{\url{https://archive.stsci.edu/hst/proposal_abstracts.html}} -- a catalog of all accepted \hubble proposals to-date -- to derive captions for the observations, optionally summarizing the abstract via guided LLM generation (described in Sec.~\ref{sec:summarization} below).
+We rely on proposal abstracts from the Proposal Abstracts Catalog\footnote{\url{https://archive.stsci.edu/hst/proposal_abstracts.html}} -- a catalog of all accepted \hubble proposals to-date -- to derive captions for the observations, optionally summarizing the abstract via guided generation using LLMs (described in Sec.~\ref{sec:summarization} below).
 %
@@
 
-Examples of images and corresponding (summarized) captions are shown in Tab.~\ref{tab:dataset}. % \SM{Push table earlier.}
+Examples of images and corresponding (clipped) proposal abstracts are shown in Tab.~\ref{tab:dataset}. % \SM{Push table earlier.}
 %
-It can be seen that these images have specific characteristics as well as artifacts particular to the nature of data-taking and processing which distinguish them from the distribution of natural images typically used for large-scale pre-training of foundation models.
+It can be seen that these images have specific characteristics as well as artifacts particular to the nature of data-taking and processing which distinguish them from the distribution of natural images typically used for large-scale pre-training of foundation models \citep{deng2009imagenet}.
 %
@@
       \toprule
-      \centering \bfseries \hubble Image & \centering \bfseries Obs. cycle \\ (Year) & \centering \bfseries Prop. ID & \centering \bfseries Proposal abstract (clipped) \tabularnewline
+      \centering \bfseries \hubble image & \centering \bfseries Obs. cycle \\ (Year) & \centering \bfseries Prop. ID & \centering \bfseries Proposal abstract (clipped) \tabularnewline
       \midrule
-      \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_0.pdf} & \centering \input{\thedatafolder/cycle_0.txt} \\ (2019) & \centering \input{\thedatafolder/id_0.txt} &  {\scriptsize \input{\thedatafolder/abs1_0.txt}} \tabularnewline
+      \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_0.pdf} & \centering \input{\thedatafolder/cycle_0.txt} \\ (2019) & \centering \input{\thedatafolder/id_0.txt} &  {\scriptsize \input{\thedatafolder/abs1_0.txt}} \tabularnewline
       \midrule
-      \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_1.pdf} & \centering \input{\thedatafolder/cycle_1.txt} \\ (2013) & \centering \input{\thedatafolder/id_1.txt} &  {\scriptsize \input{\thedatafolder/abs1_1.txt}} \tabularnewline
+      \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_1.pdf} & \centering \input{\thedatafolder/cycle_1.txt} \\ (2013) & \centering \input{\thedatafolder/id_1.txt} &  {\scriptsize \input{\thedatafolder/abs1_1.txt}} \tabularnewline
       \midrule
-      \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_2.pdf} & \centering \input{\thedatafolder/cycle_2.txt} \\ (1999) & \centering \input{\thedatafolder/id_2.txt} &  {\scriptsize \input{\thedatafolder/abs1_2.txt}} \tabularnewline
+      \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_2.pdf} & \centering \input{\thedatafolder/cycle_2.txt} \\ (1999) & \centering \input{\thedatafolder/id_2.txt} &  {\scriptsize \input{\thedatafolder/abs1_2.txt}} \tabularnewline
       \midrule
-      \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_3.pdf} & \centering \input{\thedatafolder/cycle_3.txt} \\ (2016) & \centering \input{\thedatafolder/id_3.txt} &  {\scriptsize \input{\thedatafolder/abs1_3.txt}} \tabularnewline
+      \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_3.pdf} & \centering \input{\thedatafolder/cycle_3.txt} \\ (2016) & \centering \input{\thedatafolder/id_3.txt} &  {\scriptsize \input{\thedatafolder/abs1_3.txt}} \tabularnewline
       \bottomrule
   \end{tabular}
-  \caption{Examples of images and corresponding captions, constructed using the LLM-extracted summaries.
-%
-The CLIP model is fine-tuned using these text-image associations.
-%
-Although the summarized abstract is extracted as two parts (objects/phenomena and science use cases), the model is trained using the entire summary as a single caption.}
+  \caption{Examples of images (left-most column) and corresponding clipped proposal abstracts (right-most column).
+The observation cycle and corresponding year, as well as proposal ID are shown in the second and third columns, respectively.
+The proposal ID links to the Mikulski Archive for Space Telescopes (MAST) page corresponding to the proposal.}
   \label{tab:dataset}
@@
 %
-We choose to not filter out these pairs from our dataset, in order to have a larger sample of HST observations that the image encoder can leverage to adapting to the distinctive characteristics of \hubble images.
+We choose to not filter out these pairs from our dataset, in order to have a larger sample of HST observations that the image encoder can leverage to adapt to the distinctive characteristics of \hubble images.
 
@@
 %
-These abstracts are written in a diversity of styles, formats, and lengths, being highly variable in the nature of content as well.
+These abstracts are written in a diversity of styles, formats, and lengths while being highly variable in the nature of content.
 %
@@
 %
-The goal is to summarize the objects and phenomena, as well as potential downstream science use cases corresponding to the HST observations in order to increase the strength of the association signal between text and images.
+The intended goal is to summarize the objects and phenomena, as well as potential downstream science use cases corresponding to the HST observations in order to increase the strength of the association signal between text and images.
 
@@
 %
-By framing the problem in terms of transitions between a set of finite-state machines, \cite{willard2023efficient} showed that guided generation can be performed with negligible overhead compared to unconstrained generation.
+By framing the problem in terms of transitions between a set of finite states (a finite-state machine), \cite{willard2023efficient} showed that guided generation can be performed with negligible overhead compared to unconstrained generation.
 %
-This ensures that the output of the LLM strictly conforms to the format of the following example:
+This ensures that, in this case, the output of the LLM strictly conforms to the format of the following example:
 
@@
 %
-See App.~\ref{app:guided-generation} for a more detailed description of the guidance generation process.
+See App.~\ref{app:guided-generation} for a more detailed description of the guidance generation method used here.
 
+While the schema-guided generation ensures the \emph{format} of the output, the prompt and choice of LLM will dictate the \emph{content} of the generated summaries.
+%
 We use the open-weights, instruction-tuned model \textsc{Mixtral-8x7B-Instruct}~\citep{jiang2024mixtral} to generate the summaries, with guided generation performed using the \package{Outlines}\footnote{\url{https://github.com/outlines-dev/outlines}} package.
-%
-While the schema-guided generation ensures the format of the output, the prompt and choice of LLM will influence the content of the generated summaries.
 %
@@
 
-We emphasize that the goal of summarization-via-guided-generation is to increase the signal between text and images by standardizing the captions used for fine-tuning the CLIP model, and compare the quantitative performance of the model vs using the raw abstracts in Sec.~\ref{sec:results}.
+We emphasize that the intended goal of summarization-via-guided-generation is to increase the signal between text and images by standardizing the captions used for fine-tuning the CLIP model, and compare the quantitative performance of the model vs using the raw abstracts in Sec.~\ref{sec:results}.
 %
-We also note that, even after summarization, the association signal is expected to be noisy, since aspects of the summarized caption may not be directly descriptive of the observed images, as can be seen from Tab.~\ref{tab:dataset}.
+We also note that, even after summarization, the association signal is expected to be noisy, since aspects of the summarized caption may not be directly descriptive of the observed images, as can be seen from Tab.~\ref{tab:datasetsumm}.
 
@@
           \centering
-          \begin{tabular}{m{1.8cm} m{8cm} m{5cm} m{6cm}}
+          \begin{tabular}{m{1.8cm} m{8cm} m{5cm} m{6.5cm}}
               \toprule
@@
               \input{\thedatafolder/id1_2.txt} & {\scriptsize \input{\thedatafolder/abs1_2.txt}} & {\scriptsize \input{\thedatafolder/obj1_2.txt}} & {\scriptsize \input{\thedatafolder/sci1_2.txt}} \tabularnewline
-              \midrule
-              \input{\thedatafolder/id1_3.txt} & {\scriptsize \input{\thedatafolder/abs1_3.txt}} & {\scriptsize \input{\thedatafolder/obj1_3.txt}} & {\scriptsize \input{\thedatafolder/sci1_3.txt}} \tabularnewline
+              % \midrule
+              % \input{\thedatafolder/id1_3.txt} & {\scriptsize \input{\thedatafolder/abs1_3.txt}} & {\scriptsize \input{\thedatafolder/obj1_3.txt}} & {\scriptsize \input{\thedatafolder/sci1_3.txt}} \tabularnewline
               \bottomrule
           \end{tabular}
-          \caption{Examples of the initial parts of raw proposal abstracts (second column) and LLM (\textsc{Mixtral-8x7B})-extracted summaries (rightmost two columns), separately extracting objects and phenomena as well as potential downstream science use cases.}
+          \caption{Examples of the clipped raw proposal abstracts (second column) and LLM (\textsc{Mixtral-8x7B})-extracted summaries (right-most two columns), separately extracting objects and phenomena as well as potential downstream science use cases.}
           \label{tab:datasetsumm}
@@
 
-Our goal is to learn a semantically  meaningful joint representation between HST observation images and natural language.
+Our goal is to learn a semantically meaningful joint representation between images corresponding to HST observation and natural (English) language.
 %
@@
 
-\subsection{Contrastive Language-Image Pretraining (CLIP)}
+\subsection{Contrastive Language-Image Pre-training (CLIP)}
 
-Contrastive Language-Image Pretraining \citep[CLIP;][]{radford2021learning} is a multi-modal model pre-trained on a large corpus of image-text pairs via weak supervision using a contrastive loss.
+CLIP \citep[Contrastive Language-Image Pre-training;][]{radford2021learning} is a multi-modal neural network model pre-trained on a large corpus of image-text pairs via weak supervision using a contrastive loss.
 %
@@
 %
-The standard softmax-based bidirectional variation of the InfoNCE~\citep{oord2018representation} contrastive loss function was introduce for training CLIP-style architectures \citep{radford2021learning}
+The standard softmax-based bidirectional variant of the InfoNCE~\citep{oord2018representation} contrastive loss function was introduced for training CLIP-style architectures \citep{radford2021learning}
+%
 \begin{equation}
@@
 \end{equation}
+%
 where ${x}_i={f\left(I_i\right)}/{\left\|f\left(I_i\right)\right\|}$ and ${y}_i={g\left(T_i\right)}/{\left\|g\left(T_i\right)\right\|}$ are the normalized representations of the $i$-th image and text, respectively, and $\tau$ is a learnable temperature hyperparameter.
@@
 
-We use the CLIP-ViT-B/16 \citep{radford2021learning} variant as the base pre-trained CLIP model.
+We use the CLIP-ViT-B/16\footnote{\url{https://huggingface.co/openai/clip-vit-base-patch16}} \citep{radford2021learning} variant as the base pre-trained CLIP model.
 %
@@
 %
-The text encoder has a maximum length of 77 tokens and the image encoder a native resolution of $224\times224$.
+The text encoder has a maximum length of 77 tokens and the image encoder a native resolution of $224\times224$ pixels.
 %
@@
 %
-The model was originally trained on 400 million image-text pairs from internet data.
+The model was originally trained on 400 million image-text pairs from internet data \citep{radford2021learning}.
 %
 
-\subsection{Fine-tuning procedure}
+\subsection{Fine-tuning Procedure}
 
-The base CLIP model is fine-tuned using the dataset described in Sec.~\ref{sec:dataset}, using either the LLM-summarized data, the raw proposal abstracts, or single-concept summaries.
+The base CLIP model is fine-tuned using the dataset described in Sec.~\ref{sec:dataset}, using either the LLM-summarized abstracts or raw proposal abstracts.
 %
@@
 %
-Images are randomly cropped to the native resolution of the image encoder (maintaining $\sim 20\%$ of the area of the original image) and randomly rotated by discrete increments of $90^\circ$ at each training step.
+Images are augmented via random four-fold rotations (increments of $90^\circ$) and randomly cropped to the native resolution of the image encoder, maintaining $\sim 20\%$ of the area of the original image, at each training step.
 %
-Given the relatively modest size of the fine-tuning set, a batch size $|\mathcal B| = 32$ is used throughout; larger batch sizes were seen to be susceptible to overfitting.
+Given the relatively modest size of the fine-tuning dataset, a batch size $|\mathcal B| = 32$ is used throughout; larger batch sizes were observed to be susceptible to overfitting.
+%
+The temperature hyperparameter $\tau$ was initialized to its pre-trained value.
 %
 We note that the positive and negative image-text association is noisy and imperfect, since multiple images can be associated with the same abstract.
-The temperature hyperparameter $\tau$ was initialized to its pre-trained value.
 
@@
 %
-For \emph{(2)}, we use a 2-layer MLP with 1024 units and a GELU activation layer, projecting onto the 512-dimensional embedding space.
+For \emph{(2)}, we use a 2-layer MLP with 1024 hidden units and a GELU activation layer, projecting onto the 512-dimensional embedding space.
 
-Models were trained over 20,000 steps with 2000 linear warmup steps 
+All models were trained over 20,000 steps with 2000 linear warmup steps 
 % and cosine decay 
 using the AdamW optimizer \citep{DBLP:conf/iclr/LoshchilovH19,DBLP:journals/corr/KingmaB14} with  %peak 
-learning rate of $10^{-5}$ and weight decay $10^{-3}$.
+learning rate $10^{-5}$ and weight decay $10^{-3}$.
 %
 Training takes approximately 3 hours on 4 Nvidia A100 GPUs.
-Models were instantiated and trained using packages from the \package{Jax} \citep{jax2018github} ecosystem.
+Models were instantiated using \package{Transformers} \citep{wolf2019huggingface} and trained using packages from the \package{Jax} \citep{jax2018github} ecosystem.
 %
 
-\subsection{Evaluation metrics}
+\subsection{Evaluation Metrics}
+\label{sec:eval}
 
-The model is evaluated by tracking the loss in Eq.~\ref{eq:softmax_loss} as well as the top-$k\%$ retrieval accuracy on the held out validation set over the course of training.
+The model is evaluated by tracking the contrastive loss in Eq.~\ref{eq:softmax_loss} as well as the top-$k\%$ retrieval accuracy on the held out validation set over the course of training.
 %
-The retrieval accuracy is defined as the fraction of associated captions which fall within the top $k\%$ of captions by cosine similarity of the normalized embeddings $x_i \cdot y_j$, averaged over the images in the validation set.
+The retrieval accuracy is defined as the fraction of associated captions which fall within the top $k\%$ of captions by cosine similarity of the normalized image and caption embeddings, averaged over the images in the validation set:
+\begin{equation}
+\text{Retrieval accuracy} = \frac{1}{|\mathcal V|} \sum_{i=1}^{|\mathcal V|} \mathbbm{1}\left[\operatorname{rank}\left({x}_i \cdot {y}_{i}; \{{x}_i \cdot {y}_{j}\}_{j=1}^{|\mathcal V|}\right) \leq \left\lfloor\frac{k}{100}|\mathcal V|\right\rfloor\right]
+\label{eq:retrieval_accuracy}
+\end{equation}
+where $|\mathcal V|$ is the total number of images in the validation set, $\mathbbm{1}(\cdot)$ is the indicator function that returns 1 if the condition inside the brackets is true, and 0 otherwise, $\operatorname{rank}\left({x}_i \cdot {y}_{i}; \{{x}_i \cdot {y}_{j}\}_{j=1}^{|\mathcal V|}\right)$ is a function that returns the rank of the cosine similarity between ${x}_i$ and ${y}_{i}$ among the cosine similarities between ${x}_i$ and all captions ${y}_j$ in the validation set, and $k$ is the percentage of top captions considered for the retrieval accuracy.
 
@@
 %
-For the description/text retrieval evaluation, we curate a list of possible text associations (i.e., classes) by querying the \textsc{Claude}\footnote{\url{https://claude.ai/}} large language model for such a list, which we show in App.~\ref{app:categories}.
-
-\SM{Add some expressions.}
-
-\SM{Put in the Claude thing here also.}
+For the description/text retrieval evaluation, we curate a list of possible text associations (i.e., classes) by querying the \textsc{Claude 2}\footnote{\url{https://claude.ai/}} large language model for such a list, which we show in App.~\ref{app:categories}.
 
@@
 
-\subsection{Validation metrics during model training}
+\subsection{Validation Metrics}
 
-Figure~\ref{fig:retrieval_acc} shows the contrastive loss (left) and the top-10\% retrieval accuracy (right) on the held out validation set over the course of training, for different training configurations considered.
+Figure~\ref{fig:retrieval_acc} shows the contrastive loss (left) and the top-10\% retrieval accuracy (right) evaluated on the held out validation set over the course of training, for different training configurations considered.
 %
-The purple lines show the metrics evaluated when training with batches where the image-text associations are randomly shuffled, serving as a baseline. \SM{Avoid the word baseline.}
+The dashed-orange lines show the metrics evaluated when training with batches where the image-text associations are randomly shuffled.
 %
-This baseline is seen to do on par with random expectation, unlike the others, validating the presence of a significant association signal between images and text in the dataset.
+This randomized baseline is seen to do on par with random expectation, unlike the others, validating the presence of a significant association signal between images and text in the dataset.
 %
@@
 
-The model with LLM-guided summarization (yellow lines) is seen to perform on-par with the model using raw abstracts as captions (orange line), despite the stronger association signal in the summarized dataset curated.
-\SM{Version baseline}
+The model with using LLM-summarized abstracts (red lines) is seen to perform slightly worse than the model using raw abstracts as captions (blue lines), despite the curation of the summarized-abstract dataset intended to provide a stronger image-text association signal.
 %
-Fine-tuning a small MLP head over frozen vision and text backbones (green lines), training from scratch (blue lines), and using the single-concept summaries (red lines) show a non-trivial improvement compared to the random baseline \SM{Maybe random baseline is okay} as well as base model, but with deteriorated performance compared to fine-tuning with either summarized or raw abstracts.
+Fine-tuning a small MLP head over frozen vision and text backbones (dotted green lines) and training from scratch (yellow lines) show a non-trivial improvement compared to the base model, but with deteriorated performance compared to fine-tuning with either summarized or raw abstracts.
 
 \begin{figure*}[!h]
-  \includegraphics[width=0.95\textwidth]{plots/val_metrics.pdf}
+\includegraphics[width=0.99\textwidth]{plots/val_metrics.pdf}
 \caption{The CLIP contrastive loss from Eq.~\ref{eq:softmax_loss} (left) and the top-10\% retrieval accuracy (right) computed on the validation set over the course of training.
 Shown for the dataset with summarized abstracts (orange), dataset using raw proposal abstracts (yellow), dataset with single-concept summaries (red), only fine-tuning a small MLP head (green), training from scratch (blue), and trained with shuffled image-text pairs (purple).
-\SM{Try not splitting the legend.} \SM{Would dashing be useful at all as a way to signify relationships.} \SM{Put an arrow for what is better.} \SM{Mention that fine tuning head isn't that good.}}
+\SM{Put an arrow for what is better.}}
 \label{fig:retrieval_acc}
@@
 
-\subsection{Image retrieval task}
+\subsection{Image Retrieval}
 
@@
 %
-We show these in Figs.~\ref{fig:tti_base} and \ref{fig:tti} for the base model and fine-tuned model respectively using four simple curated queries: `jupiter', `dwarf galaxy', `SN1987A' (a specific, prominent supernova), and `strong lensing'.
+We show these in Tabs.~\ref{tab:tti_base} and \ref{tab:tti} for the base model and fine-tuned model respectively using four simple curated queries: \texttt{Jupiter}, \texttt{dwarf galaxy} (small galaxies that typically orbit larger galaxies like the Milky Way), \texttt{SN1987A} (a specific, prominent supernova), and \texttt{strong lensing} (i.e., the bending of background light due to the gravitational effects of a foreground distribution of matter).
 
-While the base model shows some signs of meaningful retrieval (e.g., the image of jupiter in the first row of Fig.~\ref{fig:tti_base}, and images of galaxies in the second row), it is challenging to discern meaningful strong associations between the retrieved images and corresponding query.
+While the base model shows some signs of meaningful retrieval (e.g., the image of jupiter in the first row of Tab.~\ref{tab:tti_base}, and images of galaxies in the second row), it is challenging to discern meaningful strong associations between the retrieved images and corresponding query.
 
-The fine-tuned model (Fig.~\ref{fig:tti}), meanwhile, shows strikingly different behavior.
+The model fine-tuned with summarized abstracts (Tab.~\ref{tab:tti}), meanwhile, shows strikingly different behavior.
 %
@@
 %
-Images of Jupiter are returned for the `jupiter' query, with the planet visible in all four images.
+Images of Jupiter are returned for the \texttt{Jupiter} query, with the planet clearly visible in all four images.
 %
-The `dwarf galaxy' images correspond to proposals aiming to measure the kinematics of the stellar cores of dwarf galaxies.
+The \texttt{dwarf galaxy}-queried images correspond to proposals aiming to measure the kinematics of the stellar cores of dwarf galaxies.
 %
-Supernova SN1987 itself can be seen in the most relevant images for the `SN1987A' query, with the fourth image corresponding to a supernova remnant.
+Supernova SN1987 itself can be seen in the most relevant images for the \texttt{SN1987A} query, with the fourth image corresponding to a supernova remnant.
 %
-Cluster-scale as well as galaxy-scale gravitational lenses are returned by the `strong lensing' query, with lensing patterns visible in the images.
+Cluster-scale as well as galaxy-scale gravitational lenses are returned by the \texttt{strong lensing} query, with lensing patterns visible in the images.
 
@@
 
-\subsection{Text retrieval task}
+\subsection{Text Retrieval}
 
-We can use images from the validation set as queries and retrieve the most relevant text chunks (e.g., contained objects and use cases) from a curated list.
+We can use images from the validation set as queries and retrieve the most relevant text chunks (e.g., contained objects and use cases) from a curated list as described in Sec.~\ref{sec:eval}.
 %
-We show the result of image-to-text retrieval in Fig.~\ref{fig:itt}, for the base as well as fine-tuned models, using four observations from the validation set.
-%
-We curate a list of possible text associations by querying the \textsc{Claude}\footnote{\url{https://claude.ai/}} large language model for such a list, which we show in App.~\ref{app:categories}.
-\SM{Introduce this earlier!}
+We show the result of image-to-text retrieval in Tab.~\ref{tab:itt}, for the base (second column) as well as summary fine-tuned (third column) models, using four observations (left-most column) from the validation set.
+% %
+% We curate a list of possible text associations by querying the \textsc{Claude}\footnote{\url{https://claude.ai/}} large language model for such a list, which we show in App.~\ref{app:categories}.
 
@@
 The base as well as fine-tuned models are seen to return a mix of relevant and less-relevant associations, although showing different qualitative behavior.
+Purely qualitatively, the fine-tuned model is seen to consistently return more relevant associations compared to the base model.
 %
 
-The second row (supernova 1987A) highlights an interesting pattern -- the base model erroneously attributes the object at the center of the image to a gravitational lens, while the fine-tuned model correctly identifies it as a supernova remnant.
+The second row (an image of supernova 1987A) highlights an interesting pattern -- the base model erroneously attributes the object at the center of the image to a gravitational lens, while the fine-tuned model correctly identifies it as a supernova remnant.
 
-\subsection{Distribution of learned representations}
+\subsection{Distribution of Learned Representations}
 
-Figure~\ref{fig:sim_valtrain} (top left) shows the distribution of cosine similarities between corresponding image and text embeddings, $x_i$ and $y_i$, for the LLM-summarized abstracts using the fine-tuned CLIP model (red line), and for the base CLIP model (blue line).
+Figure~\ref{fig:sim_valtrain} (left) shows the distribution of cosine similarities between corresponding image and text embeddings, $x_i$ and $y_i$, for the LLM-summarized abstracts using the fine-tuned CLIP model (red line), and for the base CLIP model (purple line).
 %
-Distributions after shuffling the order of text embeddings -- therefore randomizing the image-text correspondence -- are shown as dashed lines.
+Distributions evaluated for a shuffled order of text embeddings -- therefore randomizing the image-text correspondence during evaluation -- are shown as dashed lines.
+We note that the shuffling here is performed at the evaluation stage, and not the training stage.
 %
-The distributions for the base model is seen to be sharply peaked at a specific value, showing little diversity and being very similar between the shuffled (dashed blue) and non-shuffled (solid blue) versions.
-\SM{Make very clear that it is the evals that are shuffled, not batches during training.}
+The distributions for the base model is seen to be sharply peaked at a specific value, showing little diversity and being very similar between the shuffled (dashed purple) and non-shuffled (solid purple) versions.
 %
-Histograms for the fine-tuned model, on the other hand, show a clear separation between the shuffled and corresponding text-image pair versions.
+Histograms for the fine-tuned model, on the other hand, show a clear separation when evaluated on shuffled (dashed red) and corresponding (solid red) text-image pairs.
 
-Similarly, Fig.~\ref{fig:sim_valtrain} (top right) shows the distribution of cosine similarities between corresponding image-text pairs for the single-concept captions (solid red) and the shuffled versions (dashed red).
+% Similarly, Fig.~\ref{fig:sim_valtrain} (top right) shows the distribution of cosine similarities between corresponding image-text pairs for the single-concept captions (solid red) and the shuffled versions (dashed red).
+% %
+% The distribution for corresponding pairs is only modestly shifted compared to the shuffled version, indicating that assigning $\mathcal O(400)$ labels to the images using an LLM results in a weaker association signal compared to using noisy captions.
+% %
+% The purple line shows the distribution of cosine similarities between corresponding pairs with single-concept summaries, but evaluated using a model trained on the LLM-summarized abstracts.
+% %
+% This distribution is more skewed towards higher similarity values, indicating that the model trained on diverse caption can still better capture the association signal in the single-concept summaries compared to one trained specifically using the single-concept summaries.
+
+Finally, Fig.~\ref{fig:sim_valtrain} (right) shows the retrieval accuracy as defined in Eq.~\ref{eq:retrieval_accuracy} as a function of the retrieval fraction $k$.
 %
-The distribution for corresponding pairs is only modestly shifted compared to the shuffled version, indicating that assigning $\mathcal O(400)$ labels to the images using an LLM results in a weaker association signal compared to using noisy captions.
+In this case, we evaluate four models -- fine tuned on raw abstracts (blue), fine-tuned on LLM-summarized abstracts (red), trained on LLM-summarized abstracts from scratch (yellow), and the base model (purple) -- all on the summarized abstracts.
 %
-The purple line shows the distribution of cosine similarities between corresponding pairs with single-concept summaries, but evaluated using a model trained on the LLM-summarized abstracts.
-%
-This distribution is more skewed towards higher similarity values, indicating that the model trained on diverse caption can still better capture the association signal in the single-concept summaries compared to one trained specifically using the single-concept summaries.
-
-Finally, Fig.~\ref{fig:sim_valtrain} (bottom) shows the image-to-text retrieval accuracy -- i.e., the fraction of correct abstracts which fall within the top $k$-fraction of captions by cosine similarity, averaged over images in the validation set -- as a function of the retrieval fraction $k$.
-%
-In this case, we evaluate three models -- fine tuned on raw abstracts (red), fine-tuned on LLM-summarized abstracts (blue), and fine-tuned on single-concept summaries (green), and the base model (purple) -- on the summarized abstracts.
-%
-Remarkably, the model trained on raw abstracts shows very similar performance on the summarized abstracts compared to that trained on the summarized abstracts themselves, indicating that \emph{(1)} the association signal is preserved in the summarization process, and \emph{(2)} the model is able to effectively leverage meaningful concepts in the noisy raw abstracts through weak supervision.
+Remarkably, the model trained on raw abstracts shows very similar performance when evaluated on the summarized abstracts compared to that trained on the summarized abstracts themselves, indicating that \emph{(1)} the image-text association signal is preserved in the summarization process, and \emph{(2)} the model is able to effectively leverage meaningful concepts in the noisy raw abstracts through weak supervision.
 
@@
       \toprule
-      \centering \bfseries Query & \multicolumn{4}{c}{\bfseries{Top-4 most similar images}} \tabularnewline
+      \centering \bfseries Query & \multicolumn{4}{c}{\bfseries{Top-4 most similar images using \textcolor{deeppurple}{base CLIP model}}} \tabularnewline
       \midrule
-       \texttt{\input{\thedatafolder/query_tti_base_0.txt}} \vspace{20mm} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_0_0.pdf} \\ \input{\thedatafolder/propid_tti_base_0_0.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_0_1.pdf} \\ \input{\thedatafolder/propid_tti_base_0_1.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_0_2.pdf} \\ \input{\thedatafolder/propid_tti_base_0_2.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_0_3.pdf} \\ \input{\thedatafolder/propid_tti_base_0_3.txt}  \tabularnewline
+       \texttt{\input{\thedatafolder/query_tti_base_0.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_0_0.pdf} \\ \input{\thedatafolder/propid_tti_base_0_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_0_1.pdf} \\ \input{\thedatafolder/propid_tti_base_0_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_0_2.pdf} \\ \input{\thedatafolder/propid_tti_base_0_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_0_3.pdf} \\ \input{\thedatafolder/propid_tti_base_0_3.txt}  \tabularnewline
       \midrule
-      \texttt{\input{\thedatafolder/query_tti_base_1.txt}} \vspace{20mm} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_1_0.pdf} \\ \input{\thedatafolder/propid_tti_base_1_0.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_1_1.pdf} \\ \input{\thedatafolder/propid_tti_base_1_1.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_1_2.pdf} \\ \input{\thedatafolder/propid_tti_base_1_2.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_1_3.pdf} \\ \input{\thedatafolder/propid_tti_base_1_3.txt}  \tabularnewline
+      \texttt{\input{\thedatafolder/query_tti_base_1.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_1_0.pdf} \\ \input{\thedatafolder/propid_tti_base_1_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_1_1.pdf} \\ \input{\thedatafolder/propid_tti_base_1_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_1_2.pdf} \\ \input{\thedatafolder/propid_tti_base_1_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_1_3.pdf} \\ \input{\thedatafolder/propid_tti_base_1_3.txt}  \tabularnewline
       \midrule
-      \texttt{\input{\thedatafolder/query_tti_base_2.txt}} \vspace{20mm} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_2_0.pdf} \\ \input{\thedatafolder/propid_tti_base_2_0.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_2_1.pdf} \\ \input{\thedatafolder/propid_tti_base_2_1.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_2_2.pdf} \\ \input{\thedatafolder/propid_tti_base_2_2.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_2_3.pdf} \\ \input{\thedatafolder/propid_tti_base_2_3.txt}  \tabularnewline
+      \texttt{\input{\thedatafolder/query_tti_base_2.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_2_0.pdf} \\ \input{\thedatafolder/propid_tti_base_2_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_2_1.pdf} \\ \input{\thedatafolder/propid_tti_base_2_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_2_2.pdf} \\ \input{\thedatafolder/propid_tti_base_2_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_2_3.pdf} \\ \input{\thedatafolder/propid_tti_base_2_3.txt}  \tabularnewline
       \midrule
-      \texttt{\input{\thedatafolder/query_tti_base_3.txt}} \vspace{20mm} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_3_0.pdf} \\ \input{\thedatafolder/propid_tti_base_3_0.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_3_1.pdf} \\ \input{\thedatafolder/propid_tti_base_3_1.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_3_2.pdf} \\ \input{\thedatafolder/propid_tti_base_3_2.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_3_3.pdf} \\ \input{\thedatafolder/propid_tti_base_3_3.txt}  \tabularnewline
+      \texttt{\input{\thedatafolder/query_tti_base_3.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_3_0.pdf} \\ \input{\thedatafolder/propid_tti_base_3_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_3_1.pdf} \\ \input{\thedatafolder/propid_tti_base_3_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_3_2.pdf} \\ \input{\thedatafolder/propid_tti_base_3_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_3_3.pdf} \\ \input{\thedatafolder/propid_tti_base_3_3.txt}  \tabularnewline
       \bottomrule
   \end{tabular}
-  \caption{Image retrieval using the base CLIP model on four curated queries.
-\SM{Horizontal lines} \SM{Query not rotated} \SM{Same for next Figure.} \SM{Put text ``Base Model'' in consistent colour to legend.} \SM{Say top 4} \SM{Say that for science we care about the distributions of things.}}
+  \caption{Image retrieval using the \textbf{\textcolor{deeppurple}{base CLIP model}} on four curated queries. \SM{Say that for science we care about the distributions of things.}}
   \label{tab:tti_base}
@@
       \toprule
-      \centering \bfseries Query & \multicolumn{4}{c}{\bfseries{Top-4 most similar images}} \tabularnewline
+      \centering \bfseries Query & \multicolumn{4}{c}{\bfseries{Top-4 most similar images using \textcolor{deepred}{fine-tuned CLIP model}}} \tabularnewline
       \midrule
-       \texttt{\input{\thedatafolder/query_tti_0.txt}} \vspace{20mm} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_0_0.pdf} \\ \input{\thedatafolder/propid_tti_0_0.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_0_1.pdf} \\ \input{\thedatafolder/propid_tti_0_1.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_0_2.pdf} \\ \input{\thedatafolder/propid_tti_0_2.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_0_3.pdf} \\ \input{\thedatafolder/propid_tti_0_3.txt}  \tabularnewline
+       \texttt{\input{\thedatafolder/query_tti_0.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_0_0.pdf} \\ \input{\thedatafolder/propid_tti_0_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_0_1.pdf} \\ \input{\thedatafolder/propid_tti_0_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_0_2.pdf} \\ \input{\thedatafolder/propid_tti_0_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_0_3.pdf} \\ \input{\thedatafolder/propid_tti_0_3.txt}  \tabularnewline
       \midrule
-      \texttt{\input{\thedatafolder/query_tti_1.txt}} \vspace{20mm} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_1_0.pdf} \\ \input{\thedatafolder/propid_tti_1_0.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_1_1.pdf} \\ \input{\thedatafolder/propid_tti_1_1.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_1_2.pdf} \\ \input{\thedatafolder/propid_tti_1_2.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_1_3.pdf} \\ \input{\thedatafolder/propid_tti_1_3.txt}  \tabularnewline
+      \texttt{\input{\thedatafolder/query_tti_1.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_1_0.pdf} \\ \input{\thedatafolder/propid_tti_1_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_1_1.pdf} \\ \input{\thedatafolder/propid_tti_1_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_1_2.pdf} \\ \input{\thedatafolder/propid_tti_1_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_1_3.pdf} \\ \input{\thedatafolder/propid_tti_1_3.txt}  \tabularnewline
       \midrule
-      \texttt{\input{\thedatafolder/query_tti_2.txt}} \vspace{20mm} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_2_0.pdf} \\ \input{\thedatafolder/propid_tti_2_0.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_2_1.pdf} \\ \input{\thedatafolder/propid_tti_2_1.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_2_2.pdf} \\ \input{\thedatafolder/propid_tti_2_2.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_2_3.pdf} \\ \input{\thedatafolder/propid_tti_2_3.txt}  \tabularnewline
+      \texttt{\input{\thedatafolder/query_tti_2.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_2_0.pdf} \\ \input{\thedatafolder/propid_tti_2_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_2_1.pdf} \\ \input{\thedatafolder/propid_tti_2_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_2_2.pdf} \\ \input{\thedatafolder/propid_tti_2_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_2_3.pdf} \\ \input{\thedatafolder/propid_tti_2_3.txt}  \tabularnewline
       \midrule
-      \texttt{\input{\thedatafolder/query_tti_3.txt}} \vspace{20mm} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_3_0.pdf} \\ \input{\thedatafolder/propid_tti_3_0.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_3_1.pdf} \\ \input{\thedatafolder/propid_tti_3_1.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_3_2.pdf} \\ \input{\thedatafolder/propid_tti_3_2.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_3_3.pdf} \\ \input{\thedatafolder/propid_tti_3_3.txt}  \tabularnewline
+      \texttt{\input{\thedatafolder/query_tti_3.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_3_0.pdf} \\ \input{\thedatafolder/propid_tti_3_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_3_1.pdf} \\ \input{\thedatafolder/propid_tti_3_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_3_2.pdf} \\ \input{\thedatafolder/propid_tti_3_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_3_3.pdf} \\ \input{\thedatafolder/propid_tti_3_3.txt}  \tabularnewline
       \bottomrule
   \end{tabular}
-  \caption{Image retrieval using the fine-tuned CLIP model on four curated queries}
+  \caption{Image retrieval using the \textbf{\textcolor{deepred}{fine-tuned CLIP model}} on four curated queries}
   \label{tab:tti}
@@
   \renewcommand{\arraystretch}{0.1}
-  \begin{tabular}{m{3cm} m{3cm} m{3cm} m{6cm}}
+  \begin{tabular}{m{3cm} m{3.2cm} m{3.2cm} m{5cm}}
       \toprule
-      \centering \bfseries \hubble Image & \centering \bfseries Obs. cycle \\ (Year) & \centering \bfseries Prop. ID & \centering \bfseries Proposal abstract (clipped) \tabularnewline
+      \centering \bfseries \hubble image & \centering \textbf{Top-4 text} \\ \textbf{\textcolor{deeppurple}{(base)}} & \centering  \textbf{Top-4 text} \\ \textbf{\textcolor{deepred}{(fine-tuned)}} & \centering \textbf{Summarized abstract} \\ (objects; `ground truth') \tabularnewline
       \midrule
@@
   \end{tabular}
-  \caption{Text associations from a curated list most closely matching a given image query, for both the fine-tuned and base models.
-The `ground truth' LLM-summarized abstract is shown in the right column.
-\SM{Say ground truth summarized abstract in the column}\SM{Swap order of cols 2 and 3.}}
+  \caption{Text associations from a curated list most closely matching a given image query, for both the \textcolor{deepred}{fine-tuned} and \textcolor{deeppurple}{base} models.
+The `ground truth' LLM-summarized abstract is shown in the right column.}
   \label{tab:itt}
@@
 (Top row) For the LLM-summarized abstracts using the fine-tuned CLIP model (red line), and for the base CLIP model (blue line), shown for (top left) the set of validation pairs and (top right) same number of training set pairs.
-Versions with the ordering of text embeddings shuffled are shown in dashed.
-\SM{Make the colours here consistent with previous Figure.} \SM{Language should be correct pairs and incorrect pairs in top left.}}
+Versions with the ordering of text embeddings shuffled are shown in dashed.}
 \label{fig:sim_valtrain}
@@
 
-We present \textsc{PAPERCLIP}, a multi-modal fountation model that associates observations imaged by the \hubble Space Telescope with natural language in a common, semantically-meaningful embedding space.
+We present \textsc{PAPERCLIP}, a method for training domain-specific multi-modal models for astrophysics that associates observations imaged by the telescopes with natural language in a common, semantically-meaningful embedding space.
 %
-The model is fine-tuned from a pre-trained CLIP model on summarized versions of \hubble proposal abstracts, leveraging a noisy association signal between text and images.
+We showcase an application to \hubble Space Telescope (HST) observations, where the model is fine-tuned from a pre-trained CLIP model using (optionally summarized) abstracts of successful \hubble proposals, leveraging a noisy association signal between text and images.
 %
@@
 %
-We also introduce a novel LLM summarization process which leverages constrained generation to increase the association strength between images and captions while preserving salient information.
-\SM{We don't actually show this.}
+We also introduce a novel LLM summarization process which leverages guided generation to distill the semantic content of the proposal abstracts while preserving salient information. 
 %
@@
 \section{Prompts and Schema Used for Summarization}
-\label{app:prompts}
+\label{app:summarization}
 
@@
 desired output formats) used at various stages for guided text generation via \package{Outlines} package interfacing with the \textsc{Mixtral-8x7B-Instruct} LLM.
-
-\subsection{Abstract summarization into objects, phenomena, and science use cases}
-\label{app:summarization}
 
@@
 %
-These are derived by prompting \textsc{Claude}, without any external input e.g.
+These are derived by prompting \textsc{Claude 2}, without any external input e.g.
 %

### 2024-03-05 paper/main.tex
Tweaks in app

@@
 \usepackage{bbm}
+\usepackage{fontawesome}
 
@@
 \definecolor{backcolour}{rgb}{0.95,0.95,0.92}
+
+\newcommand{\githubmaster}{\href{https://github.com/smsharma/HubbleCLIP}{\faGithub}\xspace}
 
@@
 Our study demonstrates the potential for using generalist rather than task-specific models for finding patterns in astronomical data, in particular by leveraging text as an interface.
+\githubmaster
 \end{abstract}
@@
 
-\subsection{\hubble Data Selection and Pre-Processing}
+\subsection{\hubble Data Selection and Pre-processing}
 
@@
 \caption{The CLIP contrastive loss from Eq.~\ref{eq:softmax_loss} (left) and the top-10\% retrieval accuracy (right) computed on the validation set over the course of training.
-Shown for the dataset with summarized abstracts (orange), dataset using raw proposal abstracts (yellow), dataset with single-concept summaries (red), only fine-tuning a small MLP head (green), training from scratch (blue), and trained with shuffled image-text pairs (purple).
-\SM{Put an arrow for what is better.}}
+Shown for the dataset with summarized abstracts (orange), dataset using raw proposal abstracts (yellow), dataset with single-concept summaries (red), only fine-tuning a small MLP head (green), training from scratch (blue), and trained with shuffled image-text pairs (purple).}
 \label{fig:retrieval_acc}
@@
   \end{tabular}
-  \caption{Image retrieval using the \textbf{\textcolor{deeppurple}{base CLIP model}} on four curated queries. \SM{Say that for science we care about the distributions of things.}}
+  \caption{Image retrieval using the \textbf{\textcolor{deeppurple}{base CLIP model}} (CLIP-ViT-B/16) on four curated queries.}
   \label{tab:tti_base}
@@
       \toprule
-      \centering \bfseries Query & \multicolumn{4}{c}{\bfseries{Top-4 most similar images using \textcolor{deepred}{fine-tuned CLIP model}}} \tabularnewline
+      \centering \bfseries Query & \multicolumn{4}{c}{\bfseries{Top-4 most similar images using \textcolor{deepred}{summary fine-tuned CLIP model}}} \tabularnewline
       \midrule
@@
   \end{tabular}
-  \caption{Image retrieval using the \textbf{\textcolor{deepred}{fine-tuned CLIP model}} on four curated queries}
+  \caption{Image retrieval using the \textbf{\textcolor{deepred}{summary fine-tuned CLIP model}} on four curated queries}
   \label{tab:tti}
@@
       \toprule
-      \centering \bfseries \hubble image & \centering \textbf{Top-4 text} \\ \textbf{\textcolor{deeppurple}{(base)}} & \centering  \textbf{Top-4 text} \\ \textbf{\textcolor{deepred}{(fine-tuned)}} & \centering \textbf{Summarized abstract} \\ (objects; `ground truth') \tabularnewline
+      \centering \bfseries \hubble image & \centering \textbf{Top-4 text} \\ {\textcolor{deeppurple}{(base)}} & \centering  \textbf{Top-4 text} \\ {\textcolor{deepred}{(summary fine-tuned)}} & \centering \textbf{Summarized abstract} \\ (objects; `ground truth') \tabularnewline
       \midrule
@@
   \end{tabular}
-  \caption{Text associations from a curated list most closely matching a given image query, for both the \textcolor{deepred}{fine-tuned} and \textcolor{deeppurple}{base} models.
+  \caption{Text associations from a curated list most closely matching a given image query, for both the \textcolor{deepred}{summary fine-tuned} and \textcolor{deeppurple}{base} (CLIP-ViT-B/16) models.
 The `ground truth' LLM-summarized abstract is shown in the right column.}
@@
 \includegraphics[width=0.45\textwidth]{plots/retrieval.pdf}
-\caption{Distribution of cosine similarities between corresponding image and text embeddings, $x_i$ and $y_i$.
-(Top row) For the LLM-summarized abstracts using the fine-tuned CLIP model (red line), and for the base CLIP model (blue line), shown for (top left) the set of validation pairs and (top right) same number of training set pairs.
-Versions with the ordering of text embeddings shuffled are shown in dashed.}
+\caption{(Left) Distribution of cosine similarities between corresponding image and text embeddings, $x_i$ and $y_i$.
+(left) For the LLM-summarized abstracts using the fine-tuned CLIP model (red line), and for the base CLIP model (purple line).
+Versions with the ordering of text embeddings shuffled during evaluation are shown in dashed.
+(Right) Retrieval accuracy as a function of the retrieval fraction $k$ for the fine-tuned model on the summarized abstracts (red), fine-tuned on raw abstracts (blue), trained on LLM-summarized abstracts from scratch (yellow), and the base model (purple).}
 \label{fig:sim_valtrain}
@@
 \appendix
-\section{Summarization via Regex-Guided Generation}
+
+\section{Details on the Abstract Summarization Procedure}
+
+\subsection{Guided LLM Generation with \package{Outlines}}
 \label{app:guided-generation}
 
-\SM{Technical description of regex-guided generation.}
+We employ the guided generation method introduced by \citet{willard2023efficient} and implemented in \package{Outlines} to ensure that the LLM summarization of the raw proposal abstracts adheres to specific pattern, specified in JSON format (Sec.~\ref{app:summarization} below), which we briefly describe here.
+This approach represents the desired output format as a finite-state machine (FSM) that encodes the JSON schema as a regular expression.
+The JSON schema constraint is therefore first converted into a regular expression.
 
-\section{Prompts and Schema Used for Summarization}
+The key idea then is to pre-compute an index that maps each state of the FSM to the subset of tokens from the LLM's vocabulary that can be generated from that state while still allowing for a valid completion of the pattern.
+By doing so, we can efficiently determine the valid next tokens at each step of the generation process without having to check the entire vocabulary.
+
+Formally, let $\mathcal{M} = (Q, \Sigma, \delta, q_0, F)$ be the FSM representing the regular expression, where $Q$ is the set of states, $\Sigma$ is the alphabet of the regular expression, $\delta: Q \times \Sigma \rightarrow Q$ is the transition function between states, $q_0$ is the start state, and $F$ is the set of accept states which terminate the generation.
+An index $\sigma: Q \rightarrow \mathcal{P}(V)$ is first constructed, where $V$ is the LLM's token vocabulary and $\mathcal{P}(V)$ denotes the power set of $V$.
+For each state $q \in Q$, $\sigma(q)$ contains the allowed tokens that can be generated from state $q$ while maintaining the possibility of reaching an accept state.
+The construction of $\sigma$ involves finding all token sequences that, when processed by the FSM starting from each state $q$, lead to an accept state.
+
+During the token-by-token generation process, we keep track of the current FSM state $q_t$ after sampling each token $v_t$.
+At each step $t$, we mask the LLM's output logits based on the valid next tokens $\sigma(q_t)$, setting the logits of invalid tokens to $-\infty$.
+The next token is then sampled from the categorical distribution defined by the unmasked logits, and the FSM transitions to the next state $q_{t+1} = \delta(q_t, v_{t+1})$, where $v_{t+1} \in \Sigma$ is the token in the regular expression alphabet corresponding to the sampled token.
+This process continues until an accept state with no outgoing transitions is reached, indicating a valid completion of the pattern.
+
+% The regex-guided generation method ensures that the LLM's output always adheres to the specified regular expression pattern, without incurring the computational overhead of checking the validity of each possible next token at every generation step.
+This approach can be extended to more complex patterns, such as context-free grammars, by using augmented pushdown automata to track the generation state and valid next tokens.
+
+% In our application, we use the regex-guided generation method to constrain LLM outputs to follow specific JSON schemas, which are automatically converted into regular expressions.
+This ensures that the generated summaries of astronomy observation proposals adhere to the desired format, enabling more reliable extraction of key information.
+% The pre-computed index $\sigma$ allows for efficient determination of valid next tokens at each generation step, making the regex-guided method computationally feasible for large-scale applications.
+
+\subsection{Prompts and Schema Used for Summarization}
 \label{app:summarization}
 
-We list here the prompts and schemata (i.e.
+We list here the prompts and schema (i.e.
 %
-desired output formats) used at various stages for guided text generation via \package{Outlines} package interfacing with the \textsc{Mixtral-8x7B-Instruct} LLM.
+desired output formats) used for guided text generation via \package{Outlines} package interfacing with the \textsc{Mixtral-8x7B-Instruct} open-weights LLM.
+
+The following schema is used to guide the generation of the summaries, intended to produce between one and five objects and hypotheses, as well as science use cases.
+
+\begin{lstlisting}[language=Python]
+from pydantic import BaseModel, conlist
+
+class ConstrainedResponseHST(BaseModel):
+      objects_and_phenomena: conlist(str, min_length=1, max_length=5)
+      science_use_cases: conlist(str, min_length=1, max_length=5)
+\end{lstlisting}
 
@@
 """
-\end{lstlisting}
-
-The following schema is used to guide the generation of the summaries, intended to produce between one and five objects and hypotheses, as well as science use cases.
-
-\begin{lstlisting}[language=Python]
-from pydantic import BaseModel, conlist
-
-class ConstrainedResponseHST(BaseModel):
-      objects_and_phenomena: conlist(str, min_length=1, max_length=5)
-      science_use_cases: conlist(str, min_length=1, max_length=5)
 \end{lstlisting}
@@
 %
-These are derived by prompting \textsc{Claude 2}, without any external input e.g.
-%
-proposal abstracts, to produce a list of categories corresponding to HST observations.
+These are derived by prompting \textsc{Claude 2}, without any external input (e.g. proposal abstracts) to produce a list of categories corresponding to typical HST observations.
 
@@
 
+\section{Evaluation of Model Trained on Raw Abstracts}
+
+\begin{table}[h!]
+  \centering
+  \begin{tabular}{m{3cm} p{3cm} p{3cm} p{3cm} p{3cm}}
+      \toprule
+      \centering \bfseries Query & \multicolumn{4}{c}{\bfseries{Top-4 most similar images using \textcolor{deepblue}{abstract fine-tuned CLIP model}}} \tabularnewline
+      \midrule
+       \texttt{\input{\thedatafolder/query_tti_abs_0.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_0_0.pdf} \\ \input{\thedatafolder/propid_tti_abs_0_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_0_1.pdf} \\ \input{\thedatafolder/propid_tti_abs_0_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_0_2.pdf} \\ \input{\thedatafolder/propid_tti_abs_0_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_0_3.pdf} \\ \input{\thedatafolder/propid_tti_abs_0_3.txt}  \tabularnewline
+      \midrule
+      \texttt{\input{\thedatafolder/query_tti_abs_1.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_1_0.pdf} \\ \input{\thedatafolder/propid_tti_abs_1_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_1_1.pdf} \\ \input{\thedatafolder/propid_tti_abs_1_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_1_2.pdf} \\ \input{\thedatafolder/propid_tti_abs_1_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_1_3.pdf} \\ \input{\thedatafolder/propid_tti_abs_1_3.txt}  \tabularnewline
+      \midrule
+      \texttt{\input{\thedatafolder/query_tti_abs_2.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_2_0.pdf} \\ \input{\thedatafolder/propid_tti_abs_2_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_2_1.pdf} \\ \input{\thedatafolder/propid_tti_abs_2_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_2_2.pdf} \\ \input{\thedatafolder/propid_tti_abs_2_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_2_3.pdf} \\ \input{\thedatafolder/propid_tti_abs_2_3.txt}  \tabularnewline
+      \midrule
+      \texttt{\input{\thedatafolder/query_tti_abs_3.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_3_0.pdf} \\ \input{\thedatafolder/propid_tti_abs_3_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_3_1.pdf} \\ \input{\thedatafolder/propid_tti_abs_3_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_3_2.pdf} \\ \input{\thedatafolder/propid_tti_abs_3_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_3_3.pdf} \\ \input{\thedatafolder/propid_tti_abs_3_3.txt}  \tabularnewline
+      \bottomrule
+  \end{tabular}
+  \caption{Image retrieval using the \textbf{\textcolor{deepblue}{abstract fine-tuned CLIP model}} on four curated queries}
+  \label{tab:tti_abs}
+\end{table}
+
+\begin{table}[h!]
+  \centering
+  \renewcommand{\arraystretch}{0.1}
+  \begin{tabular}{m{3cm} m{3.2cm} m{3.2cm} m{5cm}}
+      \toprule
+      \centering \bfseries \hubble image & \centering \textbf{Top-4 text} \\ {\textcolor{deeppurple}{(base)}} & \centering  \textbf{Top-4 text} \\ {\textcolor{deepblue}{(abstract fine-tuned)}} & \centering \textbf{Summarized abstract} \\ (objects; `ground truth') \tabularnewline
+      \midrule
+      \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_abs_0.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_0.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_abs_0.txt} &  {\scriptsize \input{\thedatafolder/abs_itt_abs_0.txt}} \tabularnewline
+      \midrule
+      \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_abs_1.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_1.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_abs_1.txt} &  {\scriptsize \input{\thedatafolder/abs_itt_abs_1.txt}} \tabularnewline
+      \midrule
+      \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_abs_2.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_2.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_abs_2.txt} &  {\scriptsize \input{\thedatafolder/abs_itt_abs_2.txt}} \tabularnewline
+      \midrule
+      \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_abs_3.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_3.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_abs_3.txt} &  {\scriptsize \input{\thedatafolder/abs_itt_abs_3.txt}} \tabularnewline
+      \bottomrule
+  \end{tabular}
+  \caption{Text associations from a curated list most closely matching a given image query, for both the \textcolor{deepblue}{abstract fine-tuned} and \textcolor{deeppurple}{base} (CLIP-ViT-B/16) models.
+The `ground truth' LLM-summarized abstract is shown in the right column.}
+  \label{tab:itt_abs}
+\end{table}
+
 \end{document}

### 2024-03-05 paper/main.tex
Add blurbs about raw abs trained model in app

@@
 
-\subsection{Validation Metrics}
+\subsection{Quantitative Evaluation}
+
+\paragraph*{Validation metrics during training}
 
@@
 
+\paragraph*{Distribution of text-image cosine similarities}
+
+Figure~\ref{fig:sim_valtrain} (left) shows the distribution of cosine similarities between corresponding image and text embeddings, $x_i$ and $y_i$, for the LLM-summarized abstracts using the fine-tuned CLIP model (red line), and for the base CLIP model (purple line).
+%
+Distributions evaluated for a shuffled order of text embeddings -- therefore randomizing the image-text correspondence during evaluation -- are shown as dashed lines.
+We note that the shuffling here is performed at the evaluation stage, and not the training stage.
+%
+The distributions for the base model is seen to be sharply peaked at a specific value, showing little diversity and being very similar between the shuffled (dashed purple) and non-shuffled (solid purple) versions.
+%
+Histograms for the fine-tuned model, on the other hand, show a clear separation when evaluated on shuffled (dashed red) and corresponding (solid red) text-image pairs.
+
+% Similarly, Fig.~\ref{fig:sim_valtrain} (top right) shows the distribution of cosine similarities between corresponding image-text pairs for the single-concept captions (solid red) and the shuffled versions (dashed red).
+% %
+% The distribution for corresponding pairs is only modestly shifted compared to the shuffled version, indicating that assigning $\mathcal O(400)$ labels to the images using an LLM results in a weaker association signal compared to using noisy captions.
+% %
+% The purple line shows the distribution of cosine similarities between corresponding pairs with single-concept summaries, but evaluated using a model trained on the LLM-summarized abstracts.
+% %
+% This distribution is more skewed towards higher similarity values, indicating that the model trained on diverse caption can still better capture the association signal in the single-concept summaries compared to one trained specifically using the single-concept summaries.
+
+\paragraph*{Retrieval accuracy}
+
+Finally, Fig.~\ref{fig:sim_valtrain} (right) shows the retrieval accuracy as defined in Eq.~\ref{eq:retrieval_accuracy} as a function of the retrieval fraction $k$.
+%
+In this case, we evaluate four models -- fine tuned on raw abstracts (blue), fine-tuned on LLM-summarized abstracts (red), trained on LLM-summarized abstracts from scratch (yellow), and the base model (purple) -- all on the summarized abstracts.
+%
+Remarkably, the model trained on raw abstracts shows very similar performance when evaluated on the summarized abstracts compared to that trained on the summarized abstracts themselves, indicating that \emph{(1)} the image-text association signal is preserved in the summarization process, and \emph{(2)} the model is able to effectively leverage meaningful concepts in the noisy raw abstracts through weak supervision.
+
+\begin{figure*}[!h]
+  \includegraphics[width=0.45\textwidth]{plots/sim_val.pdf}
+  % \includegraphics[width=0.45\textwidth]{plots/sim_summ1.pdf}
+  \includegraphics[width=0.45\textwidth]{plots/retrieval.pdf}
+  \caption{(Left) Distribution of cosine similarities between corresponding image and text embeddings, $x_i$ and $y_i$.
+(left) For the LLM-summarized abstracts using the fine-tuned CLIP model (red line), and for the base CLIP model (purple line).
+Versions with the ordering of text embeddings shuffled during evaluation are shown in dashed.
+(Right) Retrieval accuracy as a function of the retrieval fraction $k$ for the fine-tuned model on the summarized abstracts (red), fine-tuned on raw abstracts (blue), trained on LLM-summarized abstracts from scratch (yellow), and the base model (purple).}
+  \label{fig:sim_valtrain}
+  \end{figure*}
+
 \subsection{Image Retrieval}
@@
 
-\SM{Describe science usefulness, see appendix.}
-
 \subsection{Text Retrieval}
@@
 
-\subsection{Distribution of Learned Representations}
-
-Figure~\ref{fig:sim_valtrain} (left) shows the distribution of cosine similarities between corresponding image and text embeddings, $x_i$ and $y_i$, for the LLM-summarized abstracts using the fine-tuned CLIP model (red line), and for the base CLIP model (purple line).
-%
-Distributions evaluated for a shuffled order of text embeddings -- therefore randomizing the image-text correspondence during evaluation -- are shown as dashed lines.
-We note that the shuffling here is performed at the evaluation stage, and not the training stage.
-%
-The distributions for the base model is seen to be sharply peaked at a specific value, showing little diversity and being very similar between the shuffled (dashed purple) and non-shuffled (solid purple) versions.
-%
-Histograms for the fine-tuned model, on the other hand, show a clear separation when evaluated on shuffled (dashed red) and corresponding (solid red) text-image pairs.
-
-% Similarly, Fig.~\ref{fig:sim_valtrain} (top right) shows the distribution of cosine similarities between corresponding image-text pairs for the single-concept captions (solid red) and the shuffled versions (dashed red).
-% %
-% The distribution for corresponding pairs is only modestly shifted compared to the shuffled version, indicating that assigning $\mathcal O(400)$ labels to the images using an LLM results in a weaker association signal compared to using noisy captions.
-% %
-% The purple line shows the distribution of cosine similarities between corresponding pairs with single-concept summaries, but evaluated using a model trained on the LLM-summarized abstracts.
-% %
-% This distribution is more skewed towards higher similarity values, indicating that the model trained on diverse caption can still better capture the association signal in the single-concept summaries compared to one trained specifically using the single-concept summaries.
-
-Finally, Fig.~\ref{fig:sim_valtrain} (right) shows the retrieval accuracy as defined in Eq.~\ref{eq:retrieval_accuracy} as a function of the retrieval fraction $k$.
-%
-In this case, we evaluate four models -- fine tuned on raw abstracts (blue), fine-tuned on LLM-summarized abstracts (red), trained on LLM-summarized abstracts from scratch (yellow), and the base model (purple) -- all on the summarized abstracts.
-%
-Remarkably, the model trained on raw abstracts shows very similar performance when evaluated on the summarized abstracts compared to that trained on the summarized abstracts themselves, indicating that \emph{(1)} the image-text association signal is preserved in the summarization process, and \emph{(2)} the model is able to effectively leverage meaningful concepts in the noisy raw abstracts through weak supervision.
+Note that we choose to illustrate qualitative performance on text and image retrieval using the model fine-tuned on summarized abstracts, rather than raw abstracts.
+We show analogous results for the model fine-tuned on raw abstracts in App.~\ref{app:eval_raw}.
+Although the two models show very similar quantitative performance on retrieval metrics (as shown in Fig.~\ref{fig:sim_valtrain}), they exhibit characteristic behavior in terms of objects (images/text) retrieved.
+We emphasize that for scientific usefulness, retrieving the `most similar' objects would not be the goal, but rather identifying interesting candidates to follow up on; both models perform sensibly in this regard.
 
@@
 \end{table}
-
-\begin{figure*}[!h]
-\includegraphics[width=0.45\textwidth]{plots/sim_val.pdf}
-% \includegraphics[width=0.45\textwidth]{plots/sim_summ1.pdf}
-\includegraphics[width=0.45\textwidth]{plots/retrieval.pdf}
-\caption{(Left) Distribution of cosine similarities between corresponding image and text embeddings, $x_i$ and $y_i$.
-(left) For the LLM-summarized abstracts using the fine-tuned CLIP model (red line), and for the base CLIP model (purple line).
-Versions with the ordering of text embeddings shuffled during evaluation are shown in dashed.
-(Right) Retrieval accuracy as a function of the retrieval fraction $k$ for the fine-tuned model on the summarized abstracts (red), fine-tuned on raw abstracts (blue), trained on LLM-summarized abstracts from scratch (yellow), and the base model (purple).}
-\label{fig:sim_valtrain}
-\end{figure*}
 
@@
 \section{Evaluation of Model Trained on Raw Abstracts}
+\label{app:eval_raw}
 
@@
   \renewcommand{\arraystretch}{0.1}
-  \begin{tabular}{m{3cm} m{3.2cm} m{3.2cm} m{5cm}}
+  \begin{tabular}{m{3cm} m{3.2cm} m{3.2cm} m{3.2cm}}
       \toprule
-      \centering \bfseries \hubble image & \centering \textbf{Top-4 text} \\ {\textcolor{deeppurple}{(base)}} & \centering  \textbf{Top-4 text} \\ {\textcolor{deepblue}{(abstract fine-tuned)}} & \centering \textbf{Summarized abstract} \\ (objects; `ground truth') \tabularnewline
+      \centering \bfseries \hubble image & \centering \textbf{Top-4 text} \\ {\textcolor{deeppurple}{(base)}} & \centering  \textbf{Top-4 text} \\ {\textcolor{deepblue}{(abstract fine-tuned)}} & \centering  \textbf{Top-4 text} \\ {\textcolor{deepred}{(summary fine-tuned)}} \tabularnewline
       \midrule
-      \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_abs_0.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_0.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_abs_0.txt} &  {\scriptsize \input{\thedatafolder/abs_itt_abs_0.txt}} \tabularnewline
+      \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_abs_0.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_0.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_abs_0.txt} &  {\scriptsize \verbatiminput{\thedatafolder/sci_itt_0.txt}} \tabularnewline
       \midrule
-      \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_abs_1.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_1.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_abs_1.txt} &  {\scriptsize \input{\thedatafolder/abs_itt_abs_1.txt}} \tabularnewline
+      \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_abs_1.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_1.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_abs_1.txt} &  {\scriptsize \verbatiminput{\thedatafolder/sci_itt_1.txt}} \tabularnewline
       \midrule
-      \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_abs_2.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_2.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_abs_2.txt} &  {\scriptsize \input{\thedatafolder/abs_itt_abs_2.txt}} \tabularnewline
+      \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_abs_2.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_2.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_abs_2.txt} &  {\scriptsize \verbatiminput{\thedatafolder/sci_itt_2.txt}} \tabularnewline
       \midrule
-      \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_abs_3.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_3.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_abs_3.txt} &  {\scriptsize \input{\thedatafolder/abs_itt_abs_3.txt}} \tabularnewline
+      \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_abs_3.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_3.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_abs_3.txt} &  {\scriptsize \verbatiminput{\thedatafolder/sci_itt_3.txt}} \tabularnewline
       \bottomrule
   \end{tabular}
-  \caption{Text associations from a curated list most closely matching a given image query, for both the \textcolor{deepblue}{abstract fine-tuned} and \textcolor{deeppurple}{base} (CLIP-ViT-B/16) models.
-The `ground truth' LLM-summarized abstract is shown in the right column.}
+  \caption{Text associations from a curated list most closely matching a given image query (first column), for the \textcolor{deeppurple}{base} (CLIP-ViT-B/16) model, \textcolor{deepblue}{abstract fine-tuned} model, and \textcolor{deepred}{summary fine-tuned} model.}
   \label{tab:itt_abs}

### 2024-03-06 paper/main.tex
Moar text changes

@@
 \usepackage{fontawesome}
+\usepackage{fancyvrb}
+
+% For theorems and such
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{mathtools}
+\usepackage{amsthm}
 
@@
 
-\definecolor{deepblue}{rgb}{0,0,0.5}
-\definecolor{deepred}{rgb}{0.6,0,0}
+\definecolor{deepblue}{rgb}{0.29411765 0.45882353 0.61960784}
+\definecolor{deepred}{rgb}{0.74509804 0.21176471 0.23921569}
 \definecolor{deepgreen}{rgb}{0,0.5,0}
-\definecolor{deeppurple}{rgb}{0.60,0.42,0.63}
+\definecolor{deeppurple}{rgb}{0.52941176 0.32941176 0.56470588}
 
@@
 
+% \definecolor{codegray}{RGB}{240,240,240}
+\definecolor{codeblue}{RGB}{0,0,255}
+% \definecolor{codegreen}{RGB}{0,150,0}
+% \definecolor{codepurple}{RGB}{150,0,150}
+
 \newcommand{\githubmaster}{\href{https://github.com/smsharma/HubbleCLIP}{\faGithub}\xspace}
@@
 \newcommand{\hubble}{\emph{Hubble}\xspace}
+
+\newcommand{\eqrefb}[1]{(\ref{#1})}
 
@@
   commentstyle=\color{codegreen},
-  keywordstyle=\color{magenta},
+   keywordstyle=\color{codegreen}, 
    numberstyle=\tiny\color{codegray}, 
    stringstyle=\color{codegreen},
-  basicstyle=\ttfamily\footnotesize,
+   basicstyle=\ttfamily\footnotesize\linespread{1.1},
    breakatwhitespace=false, 
-  breaklines=true,
-  captionpos=b,
-  keepspaces=true,
-  numbers=left,
-  numbersep=5pt,
-  showspaces=false,
-  showstringspaces=false,
-  showtabs=false,
-  frame=single, % adds a frame around the code
-  rulecolor=\color{black}, % if you want to change the frame color
-  tabsize=2
+   breaklines=true, captionpos=b, 
+   keepspaces=true, numbers=left, 
+   numbersep=8pt, showspaces=false, 
+   showstringspaces=false, showtabs=false, 
+   frame=single, 
+   framerule=0.2pt, 
+   rulecolor=\color{codegray}, 
+   tabsize=2, 
+   aboveskip=1.5ex, 
+   belowskip=1.5ex,
+   xleftmargin=15pt, 
+   xrightmargin=15pt, 
+   extendedchars=true, 
+   columns=flexible, 
+   linewidth=\textwidth 
   }
@@
 
+% Define Verbatim environment with custom style
+\DefineVerbatimEnvironment{jsoncode}{Verbatim}{
+  commandchars=\\\{\},
+  % frame=single,
+  % framerule=0.5pt,
+  rulecolor=\color{codegray},
+  fillcolor=\color{codegray},
+  % label=JSON,
+  labelposition=topline,
+  fontsize=\small,
+  baselinestretch=1.1,
+  formatcom=\color{codeblue},
+  xleftmargin=15pt,
+  xrightmargin=15pt,
+  tabsize=2
+}
+
 \begin{document}
@@
 \begin{abstract}
-We present PAPERCLIP (Proposal Abstracts Provide an Effective Representation for Contrastive Language-Image Pre-training), a method which associates astronomical observations imaged by surveys and telescopes with natural language.
-The model is fine-tuned from a pre-trained Contrastive Language–Image Pre-training (CLIP) model using successful proposal abstracts, optionally summarized via guided generation using large language models (LLMs), and corresponding downstream observations.
-Using observations from the \hubble Space Telescope (HST) as an example, we show that the model embodies a meaningful joint representation between observations and text through experiments targeting image retrieval (i.e., retrieving the most relevant observations from a set using natural language queries) and description retrieval (i.e., querying the astrophysical object classes and science use cases most relevant to a given observation).
+We present PAPERCLIP (Proposal Abstracts Provide an Effective Representation for Contrastive Language-Image Pre-training), a method which associates astronomical observations imaged by surveys and telescopes with natural language  using a neural network model.
+The model is fine-tuned from a pre-trained Contrastive Language–Image Pre-training (CLIP) model using successful observing proposal abstracts, optionally summarized via guided generation using large language models (LLMs), and corresponding downstream observations.
+Using observations from the \hubble Space Telescope (HST) as an example, we show that the fine-tuned model embodies a meaningful joint representation between observations and text through tests targeting image retrieval (i.e., querying the closest observations using natural language queries) and description retrieval (i.e., querying the astrophysical object classes and science use cases closest to a given observation).
 Our study demonstrates the potential for using generalist rather than task-specific models for finding patterns in astronomical data, in particular by leveraging text as an interface.
@@
 % %
-These models are typically pre-trained on massive amounts of unlabeled data using self-supervised or weakly-supervised learning techniques, enabling them to learn powerful representations which can be deployed on different tasks across domains.
+These models are typically pre-trained on massive amounts of unlabeled data using self-supervised or weakly-supervised learning techniques, enabling them to learn powerful representations which can be used downstream.
 %
-However, foundation models can often benefit from additional training ({fine tuning}) using a relatively small amounts of domain-specific data in order to increase their usefulness when applied to specialized domains.
+Foundation models can often benefit from additional training ({fine tuning}) using a relatively small amounts of domain-specific data in order to increase their usefulness when applied to specialized domains.
 
-There is considerable interest in developing custom foundation models for the sciences \citep[e.g., ][]{batatia2023foundation,subramanian2023towards}, with astrophysics being ripe for such an effort given the availability of large amounts of publicly-available data and diverse ways of interacting with it.
+There is considerable interest in developing custom foundation models for the sciences \citep[e.g., ][]{batatia2023foundation,subramanian2023towards}, with astrophysics being ripe for such an effort given the availability of large amounts of publicly-available data and diverse modes of interacting with it.
 %
@@
 %
-We show that fine-tuning a pre-trained CLIP ~\citep[Contrastive Language-Image Pre-training; ][]{radford2021learning} image-text model on observation-abstract associations results in meaningful joint representations through quantitative and qualitative evaluation.
+We show that fine-tuning a pre-trained CLIP ~\citep[Contrastive Language-Image Pre-training; ][]{radford2021learning} image-text model on observation-abstract associations results in meaningful joint representations through quantitative and qualitative evaluation tests.
 %
-This opens up the possibility of interacting with astronomical survey data using free-form natural language as an interface, which is a cornerstone of the success of the modern foundation model paradigm.
+The method opens up the possibility of interacting with astronomical survey data using free-form natural language as an interface, which is a cornerstone of the success of the modern foundation model paradigm.
 %
@@
 The concept of associating diverse modalities via contrastive training has been employed in other scientific domains~\citep[e.g.,][]{liu2023text,Sanchez-Fernandez2022.11.17.516915,lanusse2023astroclip,cepeda2023geoclip}, and has been shown to be effective in learning semantically meaningful joint representations.
-Here we present for the first an application associating astronomical data with the text modality.
+Here we present for the first time an application associating astronomical data with the text modality.
 
@@
 %
-In Sec.~\ref{sec:results}, we present quantitative and qualitative results of our experiments on image and text retrieval tasks.
+In Sec.~\ref{sec:results}, we present quantitative and qualitative results of our experiments on retrieval tasks.
 %
@@
 
-We curate a dataset of images of \hubble observations and corresponding text descriptions.
+We curate a dataset of images of \hubble Space Telescope (HST) observations and corresponding text descriptions from publicly-available sources.
 %
-We rely on proposal abstracts from the Proposal Abstracts Catalog\footnote{\url{https://archive.stsci.edu/hst/proposal_abstracts.html}} -- a catalog of all accepted \hubble proposals to-date -- to derive captions for the observations, optionally summarizing the abstract via guided generation using LLMs (described in Sec.~\ref{sec:summarization} below).
+We rely on proposal abstracts from the Proposal Abstracts Catalog\footnote{\url{https://archive.stsci.edu/hst/proposal_abstracts.html}} -- a catalog of successful HST proposals -- to generate captions for the observations, optionally summarizing the abstracts via guided generation using LLMs (described in Sec.~\ref{sec:summarization} below).
 %
-The HST has been operational since its launch on April 24, 1990, and we use available proposals and observations up to Cycle 30, which commenced data-taking in 2022.
+The HST has been operational since its launch on April 24, 1990, and we use available proposals and observations up to the Cycle 30 science program, which commenced data-taking in 2022.
 %
 
-Examples of images and corresponding (clipped) proposal abstracts are shown in Tab.~\ref{tab:dataset}. % \SM{Push table earlier.}
+Table \ref{tab:dataset} shows examples of images and their corresponding (clipped) proposal abstracts.
 %
-It can be seen that these images have specific characteristics as well as artifacts particular to the nature of data-taking and processing which distinguish them from the distribution of natural images typically used for large-scale pre-training of foundation models \citep{deng2009imagenet}.
+It can be seen that the images in this dataset exhibit specific characteristics as well as artifacts particular to HST data-taking and processing which distinguishes them from the distribution of natural images typically used for large-scale pre-training of foundation models \citep{deng2009imagenet}.
 %
-This further motivates the need for fine-tuning on domain-specific data.
+This further motivates the need for fine tuning on domain-specific data.
+
+% \section{Dataset Construction}
+% \label{sec:dataset}
+
+% We curate a dataset of \hubble Space Telescope (HST) observations and corresponding text descriptions from publicly available sources.
+The dataset consists of images from the HST and associated proposal abstracts from the Hubble Space Telescope Proposal Abstracts Catalog\footnote{\url{https://archive.stsci.edu/hst/proposal_abstracts.html}}, which contains all accepted HST proposals since the telescope's launch on April 24, 1990.
+Our dataset covers observations up to Cycle 30, which began collecting data in 2022.
+
+% To generate captions for the observations, we use the proposal abstracts directly or optionally summarize them using guided language model generation (described in Sec.\ref{sec:summarization}).
+
+% The images in this dataset exhibit specific characteristics and artifacts inherent to the HST's data acquisition and processing pipeline, setting them apart from typical natural image datasets used for large-scale pre-training of foundation models \citep{deng2009imagenet}.
+This difference highlights the necessity of fine-tuning models on domain-specific data.
+
+% We emphasize that all data used in this work is publicly available, ensuring transparency and enabling reproducibility.
+The HST images can be accessed through the Hubble Legacy Archive\footnote{\url{https://hla.stsci.edu/}}, while the proposal abstracts are available in the Hubble Space Telescope Proposal Abstracts Catalog.
 
@@
       \midrule
-      \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_0.pdf} & \centering \input{\thedatafolder/cycle_0.txt} \\ (2019) & \centering \input{\thedatafolder/id_0.txt} &  {\scriptsize \input{\thedatafolder/abs1_0.txt}} \tabularnewline
+      \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_2.pdf} & \centering \input{\thedatafolder/cycle_2.txt} \\ (1999) & \centering \input{\thedatafolder/id_2.txt} &  {\scriptsize \input{\thedatafolder/abs1_2.txt}} \tabularnewline
       \midrule
@@
       \midrule
-      \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_2.pdf} & \centering \input{\thedatafolder/cycle_2.txt} \\ (1999) & \centering \input{\thedatafolder/id_2.txt} &  {\scriptsize \input{\thedatafolder/abs1_2.txt}} \tabularnewline
+      \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_3.pdf} & \centering \input{\thedatafolder/cycle_3.txt} \\ (2016) & \centering \input{\thedatafolder/id_3.txt} &  {\scriptsize \input{\thedatafolder/abs1_3.txt}} \tabularnewline
       \midrule
-      \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_3.pdf} & \centering \input{\thedatafolder/cycle_3.txt} \\ (2016) & \centering \input{\thedatafolder/id_3.txt} &  {\scriptsize \input{\thedatafolder/abs1_3.txt}} \tabularnewline
+      \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_0.pdf} & \centering \input{\thedatafolder/cycle_0.txt} \\ (2019) & \centering \input{\thedatafolder/id_0.txt} &  {\scriptsize \input{\thedatafolder/abs1_0.txt}} \tabularnewline
       \bottomrule
   \end{tabular}
-  \caption{Examples of images (left-most column) and corresponding clipped proposal abstracts (right-most column).
+  \caption{Examples of \hubble images (left-most column) and corresponding clipped proposal abstracts (right-most column).
 The observation cycle and corresponding year, as well as proposal ID are shown in the second and third columns, respectively.
@@
 
-Observations corresponding to individual proposal IDs are queried through the Mikulski Archive for Space Telescopes (MAST)\footnote{\url{https://mast.stsci.edu/}} via the \package{Astroquery} \citep{2019AJ....157...98G} interface.
+Observations corresponding to individual proposal IDs are queried through the Mikulski Archive for Space Telescopes (MAST)\footnote{\url{https://mast.stsci.edu/}} via the \package{Astroquery} \citep{2019AJ....157...98G} API.
 %
@@
 %
-Color previews (i.e., observations taken with multiple wavelength filters assigned to individual RGB channels) are manually excluded via a filename filter in order to maintain consistency across the dataset; models trained on datasets with color images included were observed to show worse performance on generalization metrics.
+Color previews (i.e., observations taken with multiple wavelength filters assigned to individual RGB channels) are manually excluded via a filename filter in order to maintain consistency across the dataset; models trained on datasets with color images included were observed to show worse performance on evaluation metrics.
 %
@@
 
-In total, 31,859 images corresponding to 4,438 abstracts are included in the fine-tuning dataset.
+In total 31,859 images corresponding to 4,438 abstracts are included in the fine-tuning dataset.
 %
-3,194 images are held out for validation, with no abstract being common between training and validation sets in order to ensure an independent set of text-image pairs for evaluation.
+3,194 images are held out for validation, with no abstract being common between training and validation sets in order to ensure an independent set of image-text pairs for evaluation.
 The held out images correspond to 438 unique abstracts.
@@
 %
-We choose to not filter out these pairs from our dataset, in order to have a larger sample of HST observations that the image encoder can leverage to adapt to the distinctive characteristics of \hubble images.
+We choose to not filter out these pairs from our dataset, in order to have a larger sample of HST observations that the model can leverage to adapt to the distinctive characteristics of \hubble images.
 
@@
 %
-These abstracts are written in a diversity of styles, formats, and lengths while being highly variable in the nature of content.
+These abstracts are written in a diversity of styles, formats, and lengths while also being highly variable in their content.
 %
@@
 %
-The intended goal is to summarize the objects and phenomena, as well as potential downstream science use cases corresponding to the HST observations in order to increase the strength of the association signal between text and images.
+Captions are summarized by extracting a list of objects and phenomena, as well as potential downstream science use cases, corresponding to the eventual imaged observation.
+The intended goal of the summarization process is to increase the strength of the association signal between text and images.
 
@@
 
-The procedure guides the generation of LLM outputs while ensuring that the schema is respected at every step in the autoregressive generation process by masking out tokens that would violate the intended format.
+The procedure guides the generation of LLM outputs while ensuring that the schema is respected at every step in the generation process by masking out tokens that would violate the intended format.
 %
 By framing the problem in terms of transitions between a set of finite states (a finite-state machine), \cite{willard2023efficient} showed that guided generation can be performed with negligible overhead compared to unconstrained generation.
-%
-This ensures that, in this case, the output of the LLM strictly conforms to the format of the following example:
-
-\begin{lstlisting}[language=Python]
-{
-  'objects_and_phenomena': ['star forming galaxy', 'lensed galaxy', ...], 
-  'science_use_cases': ['measure lensing magnification', 'probe spectral energy distributions', ...]
-}
-\end{lstlisting}
-
-which is then used to construct the summarized caption by combining the two key elements.
-%
-Examples of raw abstract snippets and corresponding LLM-generated summaries are shown in Tab.~\ref{tab:datasetsumm}.
 %
@@
 
-We emphasize that the intended goal of summarization-via-guided-generation is to increase the signal between text and images by standardizing the captions used for fine-tuning the CLIP model, and compare the quantitative performance of the model vs using the raw abstracts in Sec.~\ref{sec:results}.
+The guided generation process ensures that, in this case, the output of the LLM strictly conforms to the format of the following example:
+% \SM{Have it be centered}
+% \begin{lstlisting}[language=Python]
+% {
+%   'objects_and_phenomena': ['star forming galaxy', 'lensed galaxy', ...], 
+%   'science_use_cases': ['measure lensing magnification', 'probe spectral energy distributions', ...]
+% }
+% \end{lstlisting}
+\begin{center}
+  \begin{jsoncode}
+    \centering
+          \color{black}\{
+            \color{codegreen}'objects_and_phenomena'\color{black}: [\color{codegreen}'star forming galaxy', 'lensed galaxy'\color{black}], 
+            \color{codegreen}'science_use_cases'\color{black}: [\color{codegreen}'measure lensing magnification'\color{black}]
+          \color{black}\}
+  \end{jsoncode}
+  \end{center}
+which is then used to construct the summarized caption by combining the two key elements.
 %
-We also note that, even after summarization, the association signal is expected to be noisy, since aspects of the summarized caption may not be directly descriptive of the observed images, as can be seen from Tab.~\ref{tab:datasetsumm}.
+Examples of raw abstract snippets and corresponding LLM-generated summaries are shown in Tab.~\ref{tab:datasetsumm}.
+%
+% We emphasize that the intended goal of summarization-via-guided-generation is to increase the signal between text and images by standardizing the captions used for fine-tuning the CLIP model, and compare the quantitative performance of the model vs using the raw abstracts in Sec.~\ref{sec:results}.
+%
+We train separate models using the raw abstracts and the LLM-generated summaries, and compare their performance on downstream tasks in Sec.~\ref{sec:results}.
+%
+We note that, even after summarization, the association signal is expected to be noisy, since aspects of the summarized caption may not be directly descriptive of the observed images.
 
@@
           \midrule
-              \input{\thedatafolder/id1_0.txt} & {\scriptsize \input{\thedatafolder/abs1_0.txt}} & {\scriptsize \input{\thedatafolder/obj1_0.txt}} & {\scriptsize \input{\thedatafolder/sci1_0.txt}} \tabularnewline
+          \input{\thedatafolder/id1_2.txt} & {\scriptsize \input{\thedatafolder/abs1_2.txt}} & {\scriptsize \input{\thedatafolder/obj1_2.txt}} & {\scriptsize \input{\thedatafolder/sci1_2.txt}} \tabularnewline
           \midrule
@@
           \midrule
-              \input{\thedatafolder/id1_2.txt} & {\scriptsize \input{\thedatafolder/abs1_2.txt}} & {\scriptsize \input{\thedatafolder/obj1_2.txt}} & {\scriptsize \input{\thedatafolder/sci1_2.txt}} \tabularnewline
-              % \midrule
-              % \input{\thedatafolder/id1_3.txt} & {\scriptsize \input{\thedatafolder/abs1_3.txt}} & {\scriptsize \input{\thedatafolder/obj1_3.txt}} & {\scriptsize \input{\thedatafolder/sci1_3.txt}} \tabularnewline
+          \input{\thedatafolder/id1_3.txt} & {\scriptsize \input{\thedatafolder/abs1_3.txt}} & {\scriptsize \input{\thedatafolder/obj1_3.txt}} & {\scriptsize \input{\thedatafolder/sci1_3.txt}} \tabularnewline
+          \midrule
+          \input{\thedatafolder/id1_0.txt} & {\scriptsize \input{\thedatafolder/abs1_0.txt}} & {\scriptsize \input{\thedatafolder/obj1_0.txt}} & {\scriptsize \input{\thedatafolder/sci1_0.txt}} \tabularnewline
           \bottomrule
       \end{tabular}
-          \caption{Examples of the clipped raw proposal abstracts (second column) and LLM (\textsc{Mixtral-8x7B})-extracted summaries (right-most two columns), separately extracting objects and phenomena as well as potential downstream science use cases.}
+      \caption{Examples of the clipped \hubble proposal abstracts (second column) and LLM (\textsc{Mixtral-8x7B})-extracted summaries (right-most two columns), separately extracting objects and phenomena as well as potential downstream science use cases.
+\SM{Don't rotate?}}
       \label{tab:datasetsumm}
@@
 %
-We leverage the strong generalization capabilities demonstrated by pre-trained CLIP models and adapt these to work with domain-specific \hubble data via fine tuning.
+With PAPERCLIP, we leverage the strong generalization capabilities demonstrated by pre-trained CLIP models and adapt these to work with domain-specific \hubble data via fine tuning.
 
@@
 %
-The standard softmax-based bidirectional variant of the InfoNCE~\citep{oord2018representation} contrastive loss function was introduced for training CLIP-style architectures \citep{radford2021learning}
+We use the standard softmax-based bidirectional variant of the InfoNCE~\citep{oord2018representation} contrastive loss function introduced for training CLIP-style architectures \citep{radford2021learning}
 %
 \begin{equation}
-  \mathcal{L}(\mathcal{B})=-\frac{1}{2|\mathcal{B}|} \sum_{i=1}^{|\mathcal{B}|}\left(\log \frac{e^{x_i \cdot y_i / \tau}}{\sum_{j=1}^{|\mathcal{B}|} e^{x_i \cdot y_j / \tau}}+\log \frac{e^{x_i \cdot y_i / \tau}}{\sum_{j=1}^{|\mathcal{B}|} e^{x_j \cdot y_i / \tau}}\right)\label{eq:softmax_loss}
+  \label{eq:softmax_loss}
+  \mathcal{L}(\mathcal{B})=-\frac{1}{2|\mathcal{B}|} \sum_{i=1}^{|\mathcal{B}|}\left(\log \frac{e^{x_i \cdot y_i / \tau}}{\sum_{j=1}^{|\mathcal{B}|} e^{x_i \cdot y_j / \tau}}+\log \frac{e^{x_i \cdot y_i / \tau}}{\sum_{j=1}^{|\mathcal{B}|} e^{x_j \cdot y_i / \tau}}\right)
 \end{equation}
@@
 %
-The model was originally trained on 400 million image-text pairs from internet data \citep{radford2021learning}.
+This model was originally pre-trained on 400 million image-text pairs from internet data \citep{radford2021learning}.
 %
@@
 %
-We note that the positive and negative image-text association is noisy and imperfect, since multiple images can be associated with the same abstract.
+We emphasize that the positive and negative image-text association is noisy and imperfect, since multiple images can be associated with the same abstract, and the goal of the fine-tuning process is to leverage the signal contained in this noisy association.
 
-We explore three different methods of training the model on our domain dataset: \emph{(1)} Fine-tuning the entire network, starting from the pre-trained base model; \emph{(2)} Freezing the base image and text encoders, and training a small projection head; and \emph{(3)} Training the entire model from scratch.
+We explore three different methods of training the model on our domain dataset: \emph{(1)} Fine-tuning the entire network starting from the pre-trained base model; \emph{(2)} Freezing the base image/text encoders and training a small projection head; and \emph{(3)} Training the entire model from scratch.
 %
-For \emph{(2)}, we use a 2-layer MLP with 1024 hidden units and a GELU activation layer, projecting onto the 512-dimensional embedding space.
+For \emph{(2)}, we use a 2-layer MLP with 1024 hidden units and a GELU activation layer, projecting onto the 512-dimensional common embedding space.
 
@@
 Training takes approximately 3 hours on 4 Nvidia A100 GPUs.
-Models were instantiated using \package{Transformers} \citep{wolf2019huggingface} and trained using packages from the \package{Jax} \citep{jax2018github} ecosystem.
+Models were instantiated using the \package{Transformers} \citep{wolf2019huggingface} library and trained using packages from the \package{Jax} \citep{jax2018github} ecosystem.
 %
@@
 
-The model is evaluated by tracking the contrastive loss in Eq.~\ref{eq:softmax_loss} as well as the top-$k\%$ retrieval accuracy on the held out validation set over the course of training.
+The model is evaluated by tracking the contrastive loss in Eq.~\eqrefb{eq:softmax_loss} as well as the top-$k\%$ retrieval accuracy on the held out validation set over the course of training.
 %
@@
 \begin{equation}
-\text{Retrieval accuracy} = \frac{1}{|\mathcal V|} \sum_{i=1}^{|\mathcal V|} \mathbbm{1}\left[\operatorname{rank}\left({x}_i \cdot {y}_{i}; \{{x}_i \cdot {y}_{j}\}_{j=1}^{|\mathcal V|}\right) \leq \left\lfloor\frac{k}{100}|\mathcal V|\right\rfloor\right]
+\text{Retrieval accuracy}_k = \frac{1}{|\mathcal V|} \sum_{i=1}^{|\mathcal V|} \mathbbm{1}\left[\operatorname{rank}\left({x}_i \cdot {y}_{i}; \{{x}_i \cdot {y}_{j}\}_{j=1}^{|\mathcal V|}\right) \leq \left\lfloor\frac{k}{100}|\mathcal V|\right\rfloor\right]
 \label{eq:retrieval_accuracy}
 \end{equation}
-where $|\mathcal V|$ is the total number of images in the validation set, $\mathbbm{1}(\cdot)$ is the indicator function that returns 1 if the condition inside the brackets is true, and 0 otherwise, $\operatorname{rank}\left({x}_i \cdot {y}_{i}; \{{x}_i \cdot {y}_{j}\}_{j=1}^{|\mathcal V|}\right)$ is a function that returns the rank of the cosine similarity between ${x}_i$ and ${y}_{i}$ among the cosine similarities between ${x}_i$ and all captions ${y}_j$ in the validation set, and $k$ is the percentage of top captions considered for the retrieval accuracy.
+where $|\mathcal V|$ is the total number of images in the validation set, $\mathbbm{1}(\cdot)$ is the indicator function that returns 1 if the condition inside the brackets is true and 0 otherwise, $\operatorname{rank}\left({x}_i \cdot {y}_{i}; \{{x}_i \cdot {y}_{j}\}_{j=1}^{|\mathcal V|}\right)$ is a function that returns the rank of the cosine similarity between ${x}_i$ and ${y}_{i}$ among the cosine similarities between ${x}_i$ and all captions ${y}_j$ in the validation set, and $k$ is the percentage of top captions considered for the retrieval accuracy.
+Note that this metric is symmetric in the image and text modalities.
 
@@
 %
-For the description/text retrieval evaluation, we curate a list of possible text associations (i.e., classes) by querying the \textsc{Claude 2}\footnote{\url{https://claude.ai/}} large language model for such a list, which we show in App.~\ref{app:categories}.
+For the description/text retrieval evaluation, we curate a list of possible text associations (i.e., classes) by querying the \textsc{Claude 2}\footnote{\url{https://claude.ai/}} large language model, which we show in App.~\ref{app:categories}.
 
@@
 %
-The dashed-orange lines show the metrics evaluated when training with batches where the image-text associations are randomly shuffled.
+The dashed orange lines show the metrics evaluated when training with batches where the image-text associations are randomly shuffled.
 %
-This randomized baseline is seen to do on par with random expectation, unlike the others, validating the presence of a significant association signal between images and text in the dataset.
+This randomized baseline is seen to do on par with random expectation (i.e., a 10\% retrieval accuracy), unlike the others, validating the presence of a significant association signal between images and text in the dataset.
 %
@@
 %
-We therefore compare the qualitative performance of the base model with the fine-tuned model on downstream retrieval tasks.
+We therefore also compare the qualitative performance of the base model with the fine-tuned models on downstream retrieval tasks.
 
-The model with using LLM-summarized abstracts (red lines) is seen to perform slightly worse than the model using raw abstracts as captions (blue lines), despite the curation of the summarized-abstract dataset intended to provide a stronger image-text association signal.
+The model trained using LLM-summarized abstracts (red lines) is seen to perform slightly worse than the model using raw abstracts as captions (blue lines), despite the curation of the summarized-abstract dataset intended to provide a stronger image-text association signal.
 %
-Fine-tuning a small MLP head over frozen vision and text backbones (dotted green lines) and training from scratch (yellow lines) show a non-trivial improvement compared to the base model, but with deteriorated performance compared to fine-tuning with either summarized or raw abstracts.
+Fine-tuning a small MLP head over frozen vision and text backbones (dotted green lines) and training from scratch (yellow lines) show a non-trivial improvement compared to the base model, although with deteriorated performance compared to fine-tuning with either summarized or raw abstracts.
 
@@
 \includegraphics[width=0.99\textwidth]{plots/val_metrics.pdf}
-\caption{The CLIP contrastive loss from Eq.~\ref{eq:softmax_loss} (left) and the top-10\% retrieval accuracy (right) computed on the validation set over the course of training.
-Shown for the dataset with summarized abstracts (orange), dataset using raw proposal abstracts (yellow), dataset with single-concept summaries (red), only fine-tuning a small MLP head (green), training from scratch (blue), and trained with shuffled image-text pairs (purple).}
+\caption{The CLIP contrastive loss from Eq.~\eqrefb{eq:softmax_loss} (left) and the top-10\% retrieval accuracy from Eq.~\eqrefb{eq:retrieval_accuracy} (right) computed on the validation set over the course of training.
+Shown for the dataset with summarized abstracts as captions (red), dataset using raw proposal abstracts as captions (blue), only fine-tuning a small MLP head (dotted green), training from scratch (purple), and trained with shuffled image-text pairs (dashed orange).}
 \label{fig:retrieval_acc}
@@
 
-Figure~\ref{fig:sim_valtrain} (left) shows the distribution of cosine similarities between corresponding image and text embeddings, $x_i$ and $y_i$, for the LLM-summarized abstracts using the fine-tuned CLIP model (red line), and for the base CLIP model (purple line).
+Figure~\ref{fig:sim_valtrain} (left) shows the distribution of cosine similarities between corresponding image and text embeddings, $x_i$ and $y_i$, for the base CLIP model (purple line), and for the LLM-summarized abstracts using the fine-tuned CLIP model (red line).
 %
@@
 %
-Histograms for the fine-tuned model, on the other hand, show a clear separation when evaluated on shuffled (dashed red) and corresponding (solid red) text-image pairs.
+Distributions for the fine-tuned model, on the other hand, show a clear separation when evaluated on shuffled (dashed red) and corresponding (solid red) text-image pairs.
 
@@
 
-Finally, Fig.~\ref{fig:sim_valtrain} (right) shows the retrieval accuracy as defined in Eq.~\ref{eq:retrieval_accuracy} as a function of the retrieval fraction $k$.
+Figure~\ref{fig:sim_valtrain} (right) shows the retrieval accuracy, as defined in Eq.~\eqrefb{eq:retrieval_accuracy}, as a function of the retrieval fraction $k$.
 %
-In this case, we evaluate four models -- fine tuned on raw abstracts (blue), fine-tuned on LLM-summarized abstracts (red), trained on LLM-summarized abstracts from scratch (yellow), and the base model (purple) -- all on the summarized abstracts.
+In this case, we evaluate all four models (fine tuned on raw abstracts (blue), fine-tuned on LLM-summarized abstracts (red), trained on LLM-summarized abstracts from scratch (yellow), and the base model (purple)) on the same captions dataset, the summarized abstracts, for a direct comparison.
 %
@@
 \begin{figure*}[!h]
-  \includegraphics[width=0.45\textwidth]{plots/sim_val.pdf}
+  \includegraphics[width=0.49\textwidth]{plots/sim_val.pdf}
   % \includegraphics[width=0.45\textwidth]{plots/sim_summ1.pdf}
-  \includegraphics[width=0.45\textwidth]{plots/retrieval.pdf}
-  \caption{(Left) Distribution of cosine similarities between corresponding image and text embeddings, $x_i$ and $y_i$.
-(left) For the LLM-summarized abstracts using the fine-tuned CLIP model (red line), and for the base CLIP model (purple line).
-Versions with the ordering of text embeddings shuffled during evaluation are shown in dashed.
+  \includegraphics[width=0.49\textwidth]{plots/retrieval.pdf}
+  \caption{(Left) Distribution of cosine similarities between corresponding image and text embeddings, $x_i$ and $y_i$, shown when using the base CLIP model (purple lines), and the summary fine-tuned CLIP model (red line).
+Dashed lines correspond to models evaluated on image-text pairs with associations shuffled.
 (Right) Retrieval accuracy as a function of the retrieval fraction $k$ for the fine-tuned model on the summarized abstracts (red), fine-tuned on raw abstracts (blue), trained on LLM-summarized abstracts from scratch (yellow), and the base model (purple).}
@@
 
-Having aligned the image and text representations, we can query the validation set using natural language and show the most `relevant' images when ranked by cosine similarity.
+Having aligned the image and text representations, we can embed a natural language query using the model and show the closest images by embedding when ranked by cosine similarity.
 %
-We show these in Tabs.~\ref{tab:tti_base} and \ref{tab:tti} for the base model and fine-tuned model respectively using four simple curated queries: \texttt{Jupiter}, \texttt{dwarf galaxy} (small galaxies that typically orbit larger galaxies like the Milky Way), \texttt{SN1987A} (a specific, prominent supernova), and \texttt{strong lensing} (i.e., the bending of background light due to the gravitational effects of a foreground distribution of matter).
+We show these in Tabs.~\ref{tab:tti_base} and \ref{tab:tti} for the base and fine-tuned models respectively using four simple curated queries: \texttt{dwarf galaxy} (small galaxies that typically orbit larger galaxies like the Milky Way), \texttt{Jupiter},  \texttt{SN1987A} (a specific, prominent supernova), and \texttt{strong lensing} (the phenomenon of bending of light due to the gravitational influence of a foreground distribution of matter).
+The proposal ID corresponding to the retrieved images is shown below each image, which contains a hyperlink to the MAST page corresponding to the proposal.
 
-While the base model shows some signs of meaningful retrieval (e.g., the image of jupiter in the first row of Tab.~\ref{tab:tti_base}, and images of galaxies in the second row), it is challenging to discern meaningful strong associations between the retrieved images and corresponding query.
+While the base model shows some signs of meaningful retrieval (e.g., the image of Jupiter in the second row of Tab.~\ref{tab:tti_base}, and images of galaxies in first row), it is challenging to discern meaningful, strong associations between the retrieved images and corresponding query.
 
@@
 %
-Supernova SN1987 itself can be seen in the most relevant images for the \texttt{SN1987A} query, with the fourth image corresponding to a supernova remnant.
+Supernova SN1987 itself can be seen in the three closest images for the \texttt{SN1987A} query with the fourth image being a supernova remnant.
 %
@@
 
-We can use images from the validation set as queries and retrieve the most relevant text chunks (e.g., contained objects and use cases) from a curated list as described in Sec.~\ref{sec:eval}.
+We can use images from the validation set as queries and retrieve the most relevant text chunks (e.g., objects and use cases) from a curated list as described in Sec.~\ref{sec:eval}.
 %
@@
 
-Note that we choose to illustrate qualitative performance on text and image retrieval using the model fine-tuned on summarized abstracts, rather than raw abstracts.
+Note that we chose to illustrate qualitative performance on text and image retrieval using the model fine-tuned on summarized abstracts, rather than raw abstracts.
 We show analogous results for the model fine-tuned on raw abstracts in App.~\ref{app:eval_raw}.
 Although the two models show very similar quantitative performance on retrieval metrics (as shown in Fig.~\ref{fig:sim_valtrain}), they exhibit characteristic behavior in terms of objects (images/text) retrieved.
-We emphasize that for scientific usefulness, retrieving the `most similar' objects would not be the goal, but rather identifying interesting candidates to follow up on; both models perform sensibly in this regard.
+We emphasize that for scientific usefulness, the most useful goal is not necessarily to retrieve the objects with the highest similarity scores, but rather to identify a diverse set of interesting candidates for manual follow-up and further analysis; both models are seen to perform sensibly, even if differently, in this regard.
 
@@
       \midrule
+      \texttt{\input{\thedatafolder/query_tti_base_1.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_1_0.pdf} \\ \input{\thedatafolder/propid_tti_base_1_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_1_1.pdf} \\ \input{\thedatafolder/propid_tti_base_1_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_1_2.pdf} \\ \input{\thedatafolder/propid_tti_base_1_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_1_3.pdf} \\ \input{\thedatafolder/propid_tti_base_1_3.txt}  \tabularnewline
+      \midrule
        \texttt{\input{\thedatafolder/query_tti_base_0.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_0_0.pdf} \\ \input{\thedatafolder/propid_tti_base_0_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_0_1.pdf} \\ \input{\thedatafolder/propid_tti_base_0_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_0_2.pdf} \\ \input{\thedatafolder/propid_tti_base_0_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_0_3.pdf} \\ \input{\thedatafolder/propid_tti_base_0_3.txt}  \tabularnewline
-      \midrule
-      \texttt{\input{\thedatafolder/query_tti_base_1.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_1_0.pdf} \\ \input{\thedatafolder/propid_tti_base_1_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_1_1.pdf} \\ \input{\thedatafolder/propid_tti_base_1_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_1_2.pdf} \\ \input{\thedatafolder/propid_tti_base_1_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_1_3.pdf} \\ \input{\thedatafolder/propid_tti_base_1_3.txt}  \tabularnewline
       \midrule
@@
   \end{tabular}
-  \caption{Image retrieval using the \textbf{\textcolor{deeppurple}{base CLIP model}} (CLIP-ViT-B/16) on four curated queries.}
+  \caption{For four text queries (left-most column), the four most similar images from the validation dataset by cosine similarity when using the \textbf{\textcolor{deeppurple}{base CLIP model}} (CLIP-ViT-B/16).
+The proposal ID associated with each image is given below the image and contains a hyperlink to the MAST page corresponding to the proposal.}
   \label{tab:tti_base}
@@
       \midrule
+      \texttt{\input{\thedatafolder/query_tti_1.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_1_0.pdf} \\ \input{\thedatafolder/propid_tti_1_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_1_1.pdf} \\ \input{\thedatafolder/propid_tti_1_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_1_2.pdf} \\ \input{\thedatafolder/propid_tti_1_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_1_3.pdf} \\ \input{\thedatafolder/propid_tti_1_3.txt}  \tabularnewline
+      \midrule
        \texttt{\input{\thedatafolder/query_tti_0.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_0_0.pdf} \\ \input{\thedatafolder/propid_tti_0_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_0_1.pdf} \\ \input{\thedatafolder/propid_tti_0_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_0_2.pdf} \\ \input{\thedatafolder/propid_tti_0_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_0_3.pdf} \\ \input{\thedatafolder/propid_tti_0_3.txt}  \tabularnewline
-      \midrule
-      \texttt{\input{\thedatafolder/query_tti_1.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_1_0.pdf} \\ \input{\thedatafolder/propid_tti_1_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_1_1.pdf} \\ \input{\thedatafolder/propid_tti_1_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_1_2.pdf} \\ \input{\thedatafolder/propid_tti_1_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_1_3.pdf} \\ \input{\thedatafolder/propid_tti_1_3.txt}  \tabularnewline
       \midrule
@@
   \end{tabular}
-  \caption{Image retrieval using the \textbf{\textcolor{deepred}{summary fine-tuned CLIP model}} on four curated queries}
+  \caption{Same as Tab.~\ref{tab:tti_base}, but using the \textbf{\textcolor{deepred}{summary fine-tuned CLIP model}}.}
   \label{tab:tti}
@@
   \renewcommand{\arraystretch}{0.1}
-  \begin{tabular}{m{3cm} m{3.2cm} m{3.2cm} m{5cm}}
+  \begin{tabular}{m{3cm} m{3.6cm} m{3.6cm} m{4cm}}
       \toprule
-      \centering \bfseries \hubble image & \centering \textbf{Top-4 text} \\ {\textcolor{deeppurple}{(base)}} & \centering  \textbf{Top-4 text} \\ {\textcolor{deepred}{(summary fine-tuned)}} & \centering \textbf{Summarized abstract} \\ (objects; `ground truth') \tabularnewline
+      \centering \bfseries \hubble image & \centering \textbf{Top-4 text} \\ \small\textbf{\textcolor{deeppurple}{(base)}} & \centering  \textbf{Top-4 text} \\ \small\textbf{\textcolor{deepred}{(summary fine-tuned)}} & \centering \textbf{Summarized abstract} \\ (objects; `ground truth') \tabularnewline
       \midrule
@@
   \end{tabular}
-  \caption{Text associations from a curated list most closely matching a given image query, for both the \textcolor{deepred}{summary fine-tuned} and \textcolor{deeppurple}{base} (CLIP-ViT-B/16) models.
-The `ground truth' LLM-summarized abstract is shown in the right column.}
+  \caption{Text snippets from a curated list most closely matching a given image query (left-most column) by cosine similarity, shown for the \textcolor{deeppurple}{base} (CLIP-ViT-B/16) and \textcolor{deepred}{summary fine-tuned} models.
+The `ground truth' LLM-summarized abstract is shown in the right-most column.}
   \label{tab:itt}
@@
 
-We present \textsc{PAPERCLIP}, a method for training domain-specific multi-modal models for astrophysics that associates observations imaged by the telescopes with natural language in a common, semantically-meaningful embedding space.
+In this paper we present \textsc{PAPERCLIP}, a method for training domain-specific multi-modal models for astrophysics that associates observations imaged by telescopes with natural language in a common, semantically-meaningful embedding space.
 %
@@
 %
-We also introduce a novel LLM summarization process which leverages guided generation to distill the semantic content of the proposal abstracts while preserving salient information. 
+We also introduce a novel LLM summarization process which leverages guided generation to distill the content of proposal abstracts while preserving salient information. 
 %
@@
 
-Although the model explored here is fine-tuned using postage stamp images (i.e., preview-quality and not science-grade data), we highlight potential immediate as well as downstream use cases of this paradigm.
+Although the model explored here is fine-tuned using postage stamp images (i.e., preview-quality and not science-grade data), we highlight potential immediate as well as downstream use cases of the PAPERCLIP paradigm.
 %
-In particular, a robust model trained using weakly-supervised image-text pairs can be used to query survey data e.g., PHANGS~\citep{lee2022phangs}, COSMOS~\citep{scoville2007cosmic} using natural language, as well as to quickly find distinguishing patterns in such data that may not be apparent using specialized models or manual inspection.
+A model trained using weakly-supervised image-text pairs can be used to query survey data e.g., PHANGS~\citep{lee2022phangs}, COSMOS~\citep{scoville2007cosmic} using natural language, as well as to quickly find distinguishing patterns in such data that may not be apparent using specialized models or manual inspection.
 %
-The learned representations, having shown to correlate with physical characteristics of imaged objects, can also be fine-tuned via transfer learning to adapt to specific tasks e.g., classification~\citep{wei2020deep} or segmentation~\citep{hausen2020morpheus}.
+The learned representations, having shown to correlate with physical characteristics of imaged objects, can also be fine-tuned via transfer learning to adapt to either specific tasks e.g., classification~\citep{wei2020deep} or segmentation~\citep{hausen2020morpheus}, or observations imaged by other telescopes.
 %
+
 Finally, while the CLIP model is restricted to retrieving nearest-neighbour associations within and across text/image modalities, the learned embeddings can be used as a starting point for training or fine-tuning multi-modal large-language models for interacting with survey data and receiving responses in natural language form, as well as grounding the responses based on an existing set of observations.
 
-\paragraph*{Code and data availability}
+\subsubsection*{Code and Data Availability}
 
-The code, dataset, and fine-tuned models used in this work is available at \url{https://www.github.com/smsharma/HubbleCLIP}.
+The code, dataset, and fine-tuned models used in this work are available at \url{https://www.github.com/smsharma/HubbleCLIP}.
 
-\paragraph*{Software}
+\subsubsection*{Sofware}
 
@@
 
-\paragraph*{Acknowledgments}
+\subsubsection*{Broader Impact}
+
+This work relies on using abstracts from successful \hubble Space Telescope observing proposals to fine-tune machine learning models.
+While these abstracts are publicly accessible, the authors likely did not anticipate their text being used to train machine learning models, raising questions around consent and appropriate use of data.
+Since this research intends to develop methods to aid astronomical research and does not use sensitive personal information or target commercial gain, we believe that the scientific benefits outweigh the potential concerns in this case while acknowledging good-faith arguments to the contrary.
+As the use of foundation models in the sciences increases, it will be important for the astronomy community to consider norms and guidelines around the appropriate use of various data sources for model training to ensure transparency and maintain trust in the community.
+
+\subsubsection*{Acknowledgments}
 
@@
 YS was supported by the Research Science Institute (RSI) program at MIT.
+%
+This research was supported by an award from Google,  ``Interpretation of Multimodal Images from Astronomy''.
+%
+This research was supported by the Munich Institute for Astro-, Particle and BioPhysics (MIAPbP), which is funded by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) under Germany's Excellence Strategy – EXC-2094 – 390783311.
 %
@@
 
+In the main text, we illustrated qualitative evaluation (image and text retrieval) for the model fine-tuned on summarized abstracts.
+Here, we show the same for the model fine-tuned on raw proposal abstracts.
+Table~\ref{tab:tti_abs} shows the top-4 most similar images for the abstract fine-tuned CLIP model on the same curated queries as in Tab.~\ref{tab:tti} for the summary fine-tuned model.
+Table~\ref{tab:itt_abs} shows text associations from the curated list most closely matching the image queries, for the base and abstract fine-tuned models, as well as the summary fine-tuned model, for comparison.
+Although qualitatively different behavior is observed for both tasks, the objects retrieved are seen to correspond well the given image/text queries.
+
 \begin{table}[h!]
@@
       \midrule
+      \texttt{\input{\thedatafolder/query_tti_abs_1.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_1_0.pdf} \\ \input{\thedatafolder/propid_tti_abs_1_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_1_1.pdf} \\ \input{\thedatafolder/propid_tti_abs_1_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_1_2.pdf} \\ \input{\thedatafolder/propid_tti_abs_1_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_1_3.pdf} \\ \input{\thedatafolder/propid_tti_abs_1_3.txt}  \tabularnewline
+      \midrule
       \texttt{\input{\thedatafolder/query_tti_abs_0.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_0_0.pdf} \\ \input{\thedatafolder/propid_tti_abs_0_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_0_1.pdf} \\ \input{\thedatafolder/propid_tti_abs_0_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_0_2.pdf} \\ \input{\thedatafolder/propid_tti_abs_0_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_0_3.pdf} \\ \input{\thedatafolder/propid_tti_abs_0_3.txt}  \tabularnewline
-      \midrule
-      \texttt{\input{\thedatafolder/query_tti_abs_1.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_1_0.pdf} \\ \input{\thedatafolder/propid_tti_abs_1_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_1_1.pdf} \\ \input{\thedatafolder/propid_tti_abs_1_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_1_2.pdf} \\ \input{\thedatafolder/propid_tti_abs_1_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_1_3.pdf} \\ \input{\thedatafolder/propid_tti_abs_1_3.txt}  \tabularnewline
       \midrule
@@
   \end{tabular}
-  \caption{Image retrieval using the \textbf{\textcolor{deepblue}{abstract fine-tuned CLIP model}} on four curated queries}
+  \caption{Same as Tabs.~\ref{tab:tti_base} and \ref{tab:tti}, but using the \textbf{\textcolor{deepblue}{abstract fine-tuned CLIP model}}.}
   \label{tab:tti_abs}
@@
 
-\begin{table}[h!]
+\begin{table}[t!]
   \centering
   \renewcommand{\arraystretch}{0.1}
-  \begin{tabular}{m{3cm} m{3.2cm} m{3.2cm} m{3.2cm}}
+  \begin{tabular}{m{3cm} m{3.9cm} m{3.9cm} m{3.9cm}}
       \toprule
-      \centering \bfseries \hubble image & \centering \textbf{Top-4 text} \\ {\textcolor{deeppurple}{(base)}} & \centering  \textbf{Top-4 text} \\ {\textcolor{deepblue}{(abstract fine-tuned)}} & \centering  \textbf{Top-4 text} \\ {\textcolor{deepred}{(summary fine-tuned)}} \tabularnewline
+      \centering \bfseries \hubble image & \centering \textbf{Top-4 text} \\ \textbf{\textcolor{deeppurple}{(base)}} & \centering  \textbf{Top-4 text} \\ \textbf{\textcolor{deepblue}{(abstract fine-tuned)}} & \centering  \textbf{Top-4 text} \\ \textbf{\textcolor{deepred}{(summary fine-tuned)}} \tabularnewline
       \midrule
@@
   \end{tabular}
-  \caption{Text associations from a curated list most closely matching a given image query (first column), for the \textcolor{deeppurple}{base} (CLIP-ViT-B/16) model, \textcolor{deepblue}{abstract fine-tuned} model, and \textcolor{deepred}{summary fine-tuned} model.}
+  \caption{Text associations from a curated list most closely matching four image queries (first column, the same as in Tab.~\ref{tab:itt}), for the \textcolor{deeppurple}{base} (CLIP-ViT-B/16), \textcolor{deepblue}{abstract fine-tuned}, and \textcolor{deepred}{summary fine-tuned} models.}
   \label{tab:itt_abs}

### 2024-03-07 paper/main.tex

@@
 The model is fine-tuned from a pre-trained Contrastive Language–Image Pre-training (CLIP) model using successful observing proposal abstracts, optionally summarized via guided generation using large language models (LLMs), and corresponding downstream observations.
-Using observations from the \hubble Space Telescope (HST) as an example, we show that the fine-tuned model embodies a meaningful joint representation between observations and text through tests targeting image retrieval (i.e., querying the closest observations using natural language queries) and description retrieval (i.e., querying the astrophysical object classes and science use cases closest to a given observation).
+Using observations from the \hubble Space Telescope (HST) as an example, we show that the fine-tuned model embodies a meaningful joint representation between observations and text through tests targeting image retrieval (i.e., finding the most relevant observations using natural language queries) and description retrieval (i.e., querying for astrophysical object classes and science use cases most relevant to a given observation).
 Our study demonstrates the potential for using generalist rather than task-specific models for finding patterns in astronomical data, in particular by leveraging text as an interface.
@@
 % %
-These models are typically pre-trained on massive amounts of unlabeled data using self-supervised or weakly-supervised learning techniques, enabling them to learn powerful representations which can be used downstream.
+These models are typically pre-trained on massive amounts of unlabeled data using self-supervised or weakly-supervised learning techniques, enabling them to learn powerful representations which can then be used downstream in different ways.
 %
@@
 %
-This multi-modality was recently exploited in \textsc{AstroCLIP}~\citep{lanusse2023astroclip} to obtain a joint embedding space between multi-band images and optical spectra from the Dark Energy Spectroscopic Instrument (DESI).
+This multi-modality was recently exploited in \textsc{AstroCLIP}~\citep{lanusse2023astroclip} to construct a joint embedding space between multi-band images and optical spectra from the Dark Energy Spectroscopic Instrument (DESI).
 %
@@
 %
-We show that fine-tuning a pre-trained CLIP ~\citep[Contrastive Language-Image Pre-training; ][]{radford2021learning} image-text model on observation-abstract associations results in meaningful joint representations through quantitative and qualitative evaluation tests.
+We show that fine-tuning a pre-trained CLIP ~\citep[Contrastive Language-Image Pre-training; ][]{radford2021learning} image-text model on observation-abstract pairs results in meaningful joint representations through quantitative and qualitative evaluation tests.
 %
@@
 
-We curate a dataset of images of \hubble Space Telescope (HST) observations and corresponding text descriptions from publicly-available sources.
+We curate a dataset of \hubble Space Telescope (HST) image observations and corresponding text descriptions from publicly available sources.
 %
-We rely on proposal abstracts from the Proposal Abstracts Catalog\footnote{\url{https://archive.stsci.edu/hst/proposal_abstracts.html}} -- a catalog of successful HST proposals -- to generate captions for the observations, optionally summarizing the abstracts via guided generation using LLMs (described in Sec.~\ref{sec:summarization} below).
+We rely on proposal abstracts from the Proposal Abstracts Catalog\footnote{\url{https://archive.stsci.edu/hst/proposal_abstracts.html}} -- a catalog of successful HST proposals -- to generate captions for the observations, optionally summarizing them via guided generation using LLMs (described in Sec.~\ref{sec:summarization} below).
 %
@@
 The observation cycle and corresponding year, as well as proposal ID are shown in the second and third columns, respectively.
-The proposal ID links to the Mikulski Archive for Space Telescopes (MAST) page corresponding to the proposal.}
+The proposal IDs link to the Mikulski Archive for Space Telescopes (MAST) page corresponding to the proposal.}
   \label{tab:dataset}
@@
 %
-We note that, even after summarization, the association signal is expected to be noisy, since aspects of the summarized caption may not be directly descriptive of the observed images.
+We note that, even after summarization, the association signal is expected to be noisy, since parts of the summarized caption may not be directly descriptive of the observed images.
+The goal of the fine-tuning process is to leverage the signal contained in this noisy association.
 
@@
 %
-Note that this loss treats the image and text representations symmetrically, ensuring that the two modalities are on the same footing.
+Note that this loss treats the image and text representations symmetrically, ensuring that the two modalities are considered on the same footing.
 
@@
 %
-This model uses a 12-layer, 12-head, 768-embedding dimension vision transformer as the image encoder and a 12-layer, 8-head, 512-embedding dimension sequence transformer as the text backbone.
+This model uses a 12-layer, 12-head, 768-embedding dimension vision transformer as the image encoder and a 12-layer, 8-head, 512-embedding dimension text sequence transformer as the text backbone.
 %
@@
 
-The base CLIP model is fine-tuned using the dataset described in Sec.~\ref{sec:dataset}, using either the LLM-summarized abstracts or raw proposal abstracts.
+The base CLIP model is fine-tuned using the dataset described in Sec.~\ref{sec:dataset}, using either the LLM-summarized abstracts or raw proposal abstracts paired with observations.
 %
@@
 %
-In this case, we evaluate all four models (fine tuned on raw abstracts (blue), fine-tuned on LLM-summarized abstracts (red), trained on LLM-summarized abstracts from scratch (yellow), and the base model (purple)) on the same captions dataset, the summarized abstracts, for a direct comparison.
+In this case, we evaluate all four models (fine tuned on raw abstracts (blue), fine-tuned on LLM-summarized abstracts (red), trained on LLM-summarized abstracts from scratch (yellow), and the base model (purple)) on the same captions dataset -- the summarized abstracts -- for a direct comparison.
 %
@@
 
-Having aligned the image and text representations, we can embed a natural language query using the model and show the closest images by embedding when ranked by cosine similarity.
+Having aligned the image and text representations, we can embed a natural language query using the model and show the closest images by embedding from the validation set when ranked by cosine similarity.
 %
 We show these in Tabs.~\ref{tab:tti_base} and \ref{tab:tti} for the base and fine-tuned models respectively using four simple curated queries: \texttt{dwarf galaxy} (small galaxies that typically orbit larger galaxies like the Milky Way), \texttt{Jupiter},  \texttt{SN1987A} (a specific, prominent supernova), and \texttt{strong lensing} (the phenomenon of bending of light due to the gravitational influence of a foreground distribution of matter).
-The proposal ID corresponding to the retrieved images is shown below each image, which contains a hyperlink to the MAST page corresponding to the proposal.
+The proposal ID corresponding to the retrieved images is shown below each image, and contains a hyperlink to the MAST page corresponding to the proposal for further details.
 
@@
 
-The model fine-tuned with summarized abstracts (Tab.~\ref{tab:tti}), meanwhile, shows strikingly different behavior.
+The model fine-tuned with summarized abstracts, meanwhile, shows strikingly different behavior (Tab.~\ref{tab:tti}).
 %
@@
 %
+The \texttt{dwarf galaxy}-queried images correspond to proposals aiming to measure the kinematics of the stellar cores of dwarf galaxies.
+%
 Images of Jupiter are returned for the \texttt{Jupiter} query, with the planet clearly visible in all four images.
-%
-The \texttt{dwarf galaxy}-queried images correspond to proposals aiming to measure the kinematics of the stellar cores of dwarf galaxies.
 %
@@
 The second row (an image of supernova 1987A) highlights an interesting pattern -- the base model erroneously attributes the object at the center of the image to a gravitational lens, while the fine-tuned model correctly identifies it as a supernova remnant.
+The kind of reasonable misattribution is common when querying the base model, and largely absent in the fine-tuned model.
 
@@
 We show analogous results for the model fine-tuned on raw abstracts in App.~\ref{app:eval_raw}.
-Although the two models show very similar quantitative performance on retrieval metrics (as shown in Fig.~\ref{fig:sim_valtrain}), they exhibit characteristic behavior in terms of objects (images/text) retrieved.
-We emphasize that for scientific usefulness, the most useful goal is not necessarily to retrieve the objects with the highest similarity scores, but rather to identify a diverse set of interesting candidates for manual follow-up and further analysis; both models are seen to perform sensibly, even if differently, in this regard.
+Although the two models show very similar quantitative performance on retrieval metrics (as shown in Fig.~\ref{fig:sim_valtrain}), they exhibit characteristically different behaviors in terms of objects (images/text) retrieved.
+We emphasize that for scientific usefulness, the goal is not necessarily to correctly retrieve the most ``relevant'' objects, but rather to identify a diverse set of interesting candidates for manual follow-up and further analysis; both models are seen to perform sensibly, even if differently, in this regard.
 
@@
   \end{tabular}
-  \caption{Text snippets from a curated list most closely matching a given image query (left-most column) by cosine similarity, shown for the \textcolor{deeppurple}{base} (CLIP-ViT-B/16) and \textcolor{deepred}{summary fine-tuned} models.
-The `ground truth' LLM-summarized abstract is shown in the right-most column.}
+  \caption{Text snippets from a curated list most closely matching a given image query (left-most column) by cosine similarity of respective embeddings, shown for the \textcolor{deeppurple}{base} (CLIP-ViT-B/16) and \textcolor{deepred}{summary fine-tuned} models.
+The `ground truth' LLM-summarized abstract (only objects/phenomena) is shown in the right-most column.}
   \label{tab:itt}
@@
 %
-We showcase an application to \hubble Space Telescope (HST) observations, where the model is fine-tuned from a pre-trained CLIP model using (optionally summarized) abstracts of successful \hubble proposals, leveraging a noisy association signal between text and images.
+We showcase an application to \hubble Space Telescope (HST) observations, where the model is fine-tuned from a pre-trained CLIP model using abstracts of successful \hubble proposals, optionally summarized, leveraging a noisy association signal between text and images.
 %
@@
 %
-Overall, the procedure demonstrates the efficacy on fine-tuning generalist pre-trained models on small amounts of domain-specific data, in particular astronomical datasets.
+Overall, the procedure demonstrates the efficacy of fine-tuning generalist pre-trained models on small amounts of domain-specific data, in particular astronomical datasets, and leveraging text as an interface.
 
-Although the model explored here is fine-tuned using postage stamp images (i.e., preview-quality and not science-grade data), we highlight potential immediate as well as downstream use cases of the PAPERCLIP paradigm.
+Although the model explored here is fine-tuned using postage stamp images (i.e., preview-quality and not science-grade data), we highlight potential immediate as well as downstream use cases.
 %
-A model trained using weakly-supervised image-text pairs can be used to query survey data e.g., PHANGS~\citep{lee2022phangs}, COSMOS~\citep{scoville2007cosmic} using natural language, as well as to quickly find distinguishing patterns in such data that may not be apparent using specialized models or manual inspection.
+A model trained using weakly-supervised image-text pairs can be used to query survey data e.g., PHANGS~\citep{lee2022phangs}, COSMOS~\citep{scoville2007cosmic} using natural language, as well as to efficiently find patterns in such data that may not be apparent using specialized models or manual inspection.
 %
@@
 
-This work relies on using abstracts from successful \hubble Space Telescope observing proposals to fine-tune machine learning models.
-While these abstracts are publicly accessible, the authors likely did not anticipate their text being used to train machine learning models, raising questions around consent and appropriate use of data.
+This work relies on using abstracts from successful \hubble Space Telescope observing proposals as part of a dataset for training and evaluating machine learning models.
+While these abstracts are publicly accessible, the authors likely did not anticipate their text being used in this manner, raising questions around consent and appropriate use of data.
 Since this research intends to develop methods to aid astronomical research and does not use sensitive personal information or target commercial gain, we believe that the scientific benefits outweigh the potential concerns in this case while acknowledging good-faith arguments to the contrary.
-As the use of foundation models in the sciences increases, it will be important for the astronomy community to consider norms and guidelines around the appropriate use of various data sources for model training to ensure transparency and maintain trust in the community.
+As the use of foundation models in the sciences increases, it will be important for the astronomy community to consider norms and guidelines around the appropriate use of various data sources for model training and evaluation to ensure transparency and maintain trust in the community.
 
@@
 Table~\ref{tab:itt_abs} shows text associations from the curated list most closely matching the image queries, for the base and abstract fine-tuned models, as well as the summary fine-tuned model, for comparison.
-Although qualitatively different behavior is observed for both tasks, the objects retrieved are seen to correspond well the given image/text queries.
+Although qualitatively different behavior is observed for both tasks, the objects retrieved are seen to, in most cases, meaningfully correspond to the given image/text queries.
 

### 2024-03-07 paper/main.tex
Address jdt comments

@@
 
-\def\preprintno{XXXX} % Insert correct preprint number
+\def\preprintno{XXXX \SM{Requested from Charles}} % Insert correct preprint number
 
@@
 
-There is considerable interest in developing custom foundation models for the sciences \citep[e.g., ][]{batatia2023foundation,subramanian2023towards}, with astrophysics being ripe for such an effort given the availability of large amounts of publicly-available data and diverse modes of interacting with it.
+There is considerable interest in developing custom foundation models for the sciences \citep[e.g., ][]{batatia2023foundation,subramanian2023towards}, with astrophysics being ripe for such an effort given the large amounts of publicly-available data and diverse modes of interacting with it.
 %
@@
 
-In this paper we describe \text{PAPERCLIP} (Proposal Abstracts Provide an Effective Representation for Contrastive Language-Image Pre-training\footnote{Technically we fine tune rather than pre train, but ``PAPERCLIFT'' was rejected by the senior author of this paper.}), a method that connects astronomical image observations with natural language by leveraging the association between abstracts of successful observing proposals and images corresponding to downstream observations.
+In this paper, we describe \text{PAPERCLIP} (Proposal Abstracts Provide an Effective Representation \SM{I like `Representation' rather than `Refinement', I think the former emphasizes the central abstracts-as-captions aspect a bit more} for Contrastive Language-Image Pre-training\footnote{Technically, we fine tune rather than pre train, but ``PAPERCLIFT'' was rejected by the senior author of this paper.}), a method that connects astronomical image observations with natural language by leveraging the association between abstracts of successful observing proposals and images corresponding to downstream observations.
 %
@@
 %
-The method opens up the possibility of interacting with astronomical survey data using free-form natural language as an interface, which is a cornerstone of the success of the modern foundation model paradigm.
+Our method opens up the possibility of interacting with astronomical survey data using free-form natural language as an interface, which is a cornerstone of the success of the modern foundation model paradigm.
 %
@@
 The concept of associating diverse modalities via contrastive training has been employed in other scientific domains~\citep[e.g.,][]{liu2023text,Sanchez-Fernandez2022.11.17.516915,lanusse2023astroclip,cepeda2023geoclip}, and has been shown to be effective in learning semantically meaningful joint representations.
-Here we present for the first time an application associating astronomical data with the text modality.
+Here, we present for the first time an application associating astronomical data with the text modality.
 
@@
 
-The paper is organized as follows.
+The rest of this paper is organized as follows.
 %
@@
   \centering
-  \begin{tabular}{m{0.21\textwidth} p{1.9cm} p{1.9cm} m{8cm}}
+  \begin{tabular}{m{0.20\textwidth} p{1.9cm} p{1.9cm} m{7.5cm}}
       \toprule
@@
   \caption{Examples of \hubble images (left-most column) and corresponding clipped proposal abstracts (right-most column).
-The observation cycle and corresponding year, as well as proposal ID are shown in the second and third columns, respectively.
+The observation cycle and corresponding year, as well as proposal ID, are shown in the second and third columns, respectively.
 The proposal IDs link to the Mikulski Archive for Space Telescopes (MAST) page corresponding to the proposal.}
@@
 
-In total 31,859 images corresponding to 4,438 abstracts are included in the fine-tuning dataset.
+In total, 31,859 images corresponding to 4,438 abstracts are included in the fine-tuning dataset.
 %
@@
 \end{equation}
-where $|\mathcal V|$ is the total number of images in the validation set, $\mathbbm{1}(\cdot)$ is the indicator function that returns 1 if the condition inside the brackets is true and 0 otherwise, $\operatorname{rank}\left({x}_i \cdot {y}_{i}; \{{x}_i \cdot {y}_{j}\}_{j=1}^{|\mathcal V|}\right)$ is a function that returns the rank of the cosine similarity between ${x}_i$ and ${y}_{i}$ among the cosine similarities between ${x}_i$ and all captions ${y}_j$ in the validation set, and $k$ is the percentage of top captions considered for the retrieval accuracy.
+where $|\mathcal V|$ is the total number of images in the validation set, $\mathbbm{1}[\cdot]$ is the indicator function that returns 1 if the condition inside the brackets is true and 0 otherwise, $\operatorname{rank}\left({x}_i \cdot {y}_{i}; \{{x}_i \cdot {y}_{j}\}_{j=1}^{|\mathcal V|}\right)$ is a function that returns the rank of the cosine similarity between ${x}_i$ and ${y}_{i}$ among the cosine similarities between ${x}_i$ and all captions ${y}_j$ in the validation set, and $k$ is the percentage of top captions considered for the retrieval accuracy.
 Note that this metric is symmetric in the image and text modalities.
@@
 %
-For the description/text retrieval evaluation, we curate a list of possible text associations (i.e., classes) by querying the \textsc{Claude 2}\footnote{\url{https://claude.ai/}} large language model, which we show in App.~\ref{app:categories}.
+For the description/text retrieval evaluation, we define a list of possible text associations (i.e., classes), which we show in App.~\ref{app:categories}, by querying the \textsc{Claude 2}\footnote{\url{https://claude.ai/}} large language model along with manual curation.
 
@@
 %
-Fine-tuning a small MLP head over frozen vision and text backbones (dotted green lines) and training from scratch (yellow lines) show a non-trivial improvement compared to the base model, although with deteriorated performance compared to fine-tuning with either summarized or raw abstracts.
+Fine-tuning a small MLP head over frozen vision and text backbones (dotted green lines) and training from scratch with summarized abstracts as captions (yellow lines) show a non-trivial improvement compared to the base model, although with deteriorated performance compared to fine-tuning with either summarized or raw abstracts.
 
@@
 \caption{The CLIP contrastive loss from Eq.~\eqrefb{eq:softmax_loss} (left) and the top-10\% retrieval accuracy from Eq.~\eqrefb{eq:retrieval_accuracy} (right) computed on the validation set over the course of training.
-Shown for the dataset with summarized abstracts as captions (red), dataset using raw proposal abstracts as captions (blue), only fine-tuning a small MLP head (dotted green), training from scratch (purple), and trained with shuffled image-text pairs (dashed orange).}
+Shown for the dataset with summarized abstracts as captions (red), dataset using raw proposal abstracts as captions (blue), only fine-tuning a small MLP head (dotted green), training from scratch with summarized abstracts as captions (yellow), and trained with shuffled image-text pairs (dashed orange).}
 \label{fig:retrieval_acc}
@@
 
-Figure~\ref{fig:sim_valtrain} (right) shows the retrieval accuracy, as defined in Eq.~\eqrefb{eq:retrieval_accuracy}, as a function of the retrieval fraction $k$.
+Figure~\ref{fig:sim_valtrain} (right) shows the retrieval accuracy, as defined in Eq.~\eqrefb{eq:retrieval_accuracy}, as a function of the retrieval fraction $k\%$.
 %
@@
 Dashed lines correspond to models evaluated on image-text pairs with associations shuffled.
-(Right) Retrieval accuracy as a function of the retrieval fraction $k$ for the fine-tuned model on the summarized abstracts (red), fine-tuned on raw abstracts (blue), trained on LLM-summarized abstracts from scratch (yellow), and the base model (purple).}
+(Right) Retrieval accuracy as a function of the retrieval fraction $k$ for the fine-tuned model on the summarized abstracts (red), fine-tuned on raw abstracts (blue), trained on summarized abstracts from scratch (yellow), and the base model (purple).}
   \label{fig:sim_valtrain}
@@
 The second row (an image of supernova 1987A) highlights an interesting pattern -- the base model erroneously attributes the object at the center of the image to a gravitational lens, while the fine-tuned model correctly identifies it as a supernova remnant.
-The kind of reasonable misattribution is common when querying the base model, and largely absent in the fine-tuned model.
+This kind of reasonable misattribution is common when querying the base model, and largely absent in the fine-tuned model.
 
@@
   \centering
-  \begin{tabular}{m{3cm} p{3cm} p{3cm} p{3cm} p{3cm}}
+  \begin{tabular}{m{2.7cm} p{2.9cm} p{2.9cm} p{2.9cm} p{2.9cm}}
       \toprule
@@
 
-In this paper we present \textsc{PAPERCLIP}, a method for training domain-specific multi-modal models for astrophysics that associates observations imaged by telescopes with natural language in a common, semantically-meaningful embedding space.
+In this paper, we present \textsc{PAPERCLIP}, a method for training domain-specific multi-modal models for astrophysics that associates observations imaged by telescopes with natural language in a common, semantically-meaningful embedding space.
 %
@@
 
-\subsubsection*{Sofware}
+\subsubsection*{Software}
 
@@
 Since this research intends to develop methods to aid astronomical research and does not use sensitive personal information or target commercial gain, we believe that the scientific benefits outweigh the potential concerns in this case while acknowledging good-faith arguments to the contrary.
-As the use of foundation models in the sciences increases, it will be important for the astronomy community to consider norms and guidelines around the appropriate use of various data sources for model training and evaluation to ensure transparency and maintain trust in the community.
+As the use of foundation models in the sciences increases, it will be important for the astronomy community to consider norms and guidelines around the appropriate use and attribution of various data sources for model training and evaluation, including qualitative textual data, to ensure transparency and maintain trust in the community.
 
@@
 
-We employ the guided generation method introduced by \citet{willard2023efficient} and implemented in \package{Outlines} to ensure that the LLM summarization of the raw proposal abstracts adheres to specific pattern, specified in JSON format (Sec.~\ref{app:summarization} below), which we briefly describe here.
+As mention in Sec.~\ref{sec:summarization}, we employ the guided generation method introduced by \citet{willard2023efficient} and implemented in \package{Outlines} to ensure that the LLM summarization of the raw proposal abstracts adheres to specific pattern, specified in JSON format (Sec.~\ref{app:summarization} below), which we briefly describe here.
 This approach represents the desired output format as a finite-state machine (FSM) that encodes the JSON schema as a regular expression.
@@
 
-We list here the prompts and schema (i.e.
-%
-desired output formats) used for guided text generation via \package{Outlines} package interfacing with the \textsc{Mixtral-8x7B-Instruct} open-weights LLM.
+We list here the prompts and schema (i.e., desired output formats) used for guided text generation via \package{Outlines} package interfacing with the \textsc{Mixtral-8x7B-Instruct} open-weights LLM.
 
@@
 
-The following prompt is used to produce a list of possible objects and phenomena shown in HST observations downstream of a proposal abstract, as well as one to five possible science use cases.
+The following prompt function is used to produce a list of possible objects and phenomena shown in HST observations downstream of a proposal abstract, as well as one to five possible science use cases.
 
@@
 %
-These are derived by prompting \textsc{Claude 2}, without any external input (e.g. proposal abstracts) to produce a list of categories corresponding to typical HST observations.
+These are derived by initially prompting \textsc{Claude 2}, having attached a subsample of 30 proposal abstracts in the online interface to be used as context, to produce a list of categories corresponding to typical HST observations.
+The list is then manually curated to remove similar entries and ensure a representative sample of categories.
+\SM{I realized when looking up the prompt that I did actually attach a sample of abstracts as context}
 
@@
 \end{lstlisting}
+
+The following prompt is used to generate the initial list before manual curation: \emph{``Here is a list of Hubble proposals.
+Base on this, please provide a list of about 100 strings, each describing a science target or use case for observations imaged by the Hubble Space Telescope.
+You may use these proposals and also rely on your general knowledge.
+For example, ["gravitational lensing", "supernovae", "diffuse galaxies", ...]''}
 
@@
   \centering
-  \begin{tabular}{m{3cm} p{3cm} p{3cm} p{3cm} p{3cm}}
+  \begin{tabular}{m{2.7cm} p{2.9cm} p{2.9cm} p{2.9cm} p{2.9cm}}
     \toprule

### 2024-03-07 paper/hubble_paperclip.tex
Rename files to hubble_paperclip; version sent out for comments

--- 2024-03-07 paper/main.tex
@@
 
-\def\preprintno{XXXX \SM{Requested from Charles}} % Insert correct preprint number
+\def\preprintno{XXXX} % Insert correct preprint number
 
@@
 \fancypagestyle{firstpage}{
-    \rhead{MIT-CTP/\preprintno}
-    % Define other header and footer elements if necessary
+    % \rhead{MIT-CTP/\preprintno}
 }
@@
 
-In this paper, we describe \text{PAPERCLIP} (Proposal Abstracts Provide an Effective Representation \SM{I like `Representation' rather than `Refinement', I think the former emphasizes the central abstracts-as-captions aspect a bit more} for Contrastive Language-Image Pre-training\footnote{Technically, we fine tune rather than pre train, but ``PAPERCLIFT'' was rejected by the senior author of this paper.}), a method that connects astronomical image observations with natural language by leveraging the association between abstracts of successful observing proposals and images corresponding to downstream observations.
+In this paper, we describe \text{PAPERCLIP} (Proposal Abstracts Provide an Effective Representation for Contrastive Language-Image Pre-training\footnote{Technically, we fine tune rather than pre train, but ``PAPERCLIFT'' was rejected by the senior author of this paper.}), a method that connects astronomical image observations with natural language by leveraging the association between abstracts of successful observing proposals and images corresponding to downstream observations.
 %
@@
       \end{tabular}
-      \caption{Examples of the clipped \hubble proposal abstracts (second column) and LLM (\textsc{Mixtral-8x7B})-extracted summaries (right-most two columns), separately extracting objects and phenomena as well as potential downstream science use cases.
-\SM{Don't rotate?}}
+      \caption{Examples of the clipped \hubble proposal abstracts (second column) and LLM (\textsc{Mixtral-8x7B})-extracted summaries (right-most two columns), separately extracting objects and phenomena as well as potential downstream science use cases.}
       \label{tab:datasetsumm}
@@
 This work relies on using abstracts from successful \hubble Space Telescope observing proposals as part of a dataset for training and evaluating machine learning models.
-While these abstracts are publicly accessible, the authors likely did not anticipate their text being used in this manner, raising questions around consent and appropriate use of data.
+While these abstracts are publicly available, the authors likely did not anticipate their text being used in this manner, raising questions around consent, attribution, and appropriate use of data.
 Since this research intends to develop methods to aid astronomical research and does not use sensitive personal information or target commercial gain, we believe that the scientific benefits outweigh the potential concerns in this case while acknowledging good-faith arguments to the contrary.
-As the use of foundation models in the sciences increases, it will be important for the astronomy community to consider norms and guidelines around the appropriate use and attribution of various data sources for model training and evaluation, including qualitative textual data, to ensure transparency and maintain trust in the community.
+As the use of foundation models in the sciences increases, it will be important for the community to consider norms and guidelines around the appropriate use and attribution of various data sources for model training and evaluation, including qualitative textual data, to ensure transparency and maintain trust.
 
@@
 
-\bibliography{main}
+\bibliography{hubble_paperclip}
 \bibliographystyle{tmlr}
@@
 
-The following schema is used to guide the generation of the summaries, intended to produce between one and five objects and hypotheses, as well as science use cases.
+The following schema is used to guide the generation of the summaries, intended to produce between one and five objects and hypotheses, as well as science use cases. \\
 
@@
 The following prompt function is used to produce a list of possible objects and phenomena shown in HST observations downstream of a proposal abstract, as well as one to five possible science use cases.
+\\
 
@@
 The list is then manually curated to remove similar entries and ensure a representative sample of categories.
-\SM{I realized when looking up the prompt that I did actually attach a sample of abstracts as context}
+\\
 

### 2024-03-09 scripts/submit_train.sh
Add ablation tests

@@ -6,7 +6,7 @@
 #SBATCH --time=48:00:00
 #SBATCH --gres=gpu:4
 #SBATCH --account=iaifi_lab
-#SBATCH -p iaifi_gpu_requeue
+#SBATCH -p iaifi_gpu_priority
 
 export TF_CPP_MIN_LOG_LEVEL="2"
 
@@ -29,8 +29,8 @@ cd /n/holystore01/LABS/iaifi_lab/Users/smsharma/multimodal-data/
 
 # Core runs
 
-/n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.data.caption_type="summary"  # Base config, with summary captions
-/n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.clip.transfer_head=True  # Fine-tune just head
+# /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.data.caption_type="summary"  # Base config, with summary captions
+# /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.clip.transfer_head=True  # Fine-tune just head
 
 # /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.data.caption_type="abstract"  --config.data.augment_subsample_text=True  # Full abstract
 # /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.clip.random_init_text=True --config.clip.random_init_vision=True  # From scratch
@@ -45,3 +45,4 @@ cd /n/holystore01/LABS/iaifi_lab/Users/smsharma/multimodal-data/
 # /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.training.loss_type="sigmoid"
 # /n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python -u train.py --config ./configs/base.py --config.training.batch_size=64  # Larger batch size
[…]

### 2024-03-10 paper/hubble_paperclip.tex
Add ablations app

@@
 \usepackage{booktabs}
-\usepackage{lipsum}
 \usepackage{xspace}
@@
 \usepackage{amssymb}
+\usepackage{amsthm}
 \usepackage{mathtools}
-\usepackage{amsthm}
 
@@
 
+\definecolor{xlinkcolor}{rgb}{0.7752941176470588, 0.22078431372549023, 0.2262745098039215}
+
+% For model colours
 \definecolor{deepblue}{rgb}{0.29411765 0.45882353 0.61960784}
@@
 
-\definecolor{xlinkcolor}{rgb}{0.7752941176470588, 0.22078431372549023, 0.2262745098039215}
-
+% For code listings
 \definecolor{codegreen}{rgb}{0,0.4,0}
@@
 \definecolor{backcolour}{rgb}{0.95,0.95,0.92}
-
-% \definecolor{codegray}{RGB}{240,240,240}
-\definecolor{codeblue}{RGB}{0,0,255}
-% \definecolor{codegreen}{RGB}{0,150,0}
-% \definecolor{codepurple}{RGB}{150,0,150}
 
@@
 \newcommand{\ie}{{i.\,e.}\xspace}
+\newcommand{\hubble}{\emph{Hubble}\xspace}
+\newcommand{\eqrefb}[1]{(\ref{#1})}
+
 \newcommand{\SM}[1]{\textcolor{blue}{[SM: #1]}}
-\newcommand{\hubble}{\emph{Hubble}\xspace}
-
-\newcommand{\eqrefb}[1]{(\ref{#1})}
+% Changes
+\newcommand{\changes}[1]{\textcolor{blue}{#1}}
 
@@
 \fancypagestyle{firstpage}{
-    % \rhead{MIT-CTP/\preprintno}
+    \rhead{MIT-CTP/\preprintno}
 }
@@
 
+\newcommand{\datafolder}[1]{\def\thedatafolder{#1}}
+
+% Define Verbatim environment with custom style
+\DefineVerbatimEnvironment{jsoncode}{Verbatim}{
+  commandchars=\\\{\},
+  rulecolor=\color{codegray},
+  fillcolor=\color{codegray},
+  labelposition=topline,
+  fontsize=\small,
+  baselinestretch=1.1,
+  formatcom=\color{codegreen},
+  xleftmargin=15pt,
+  xrightmargin=15pt,
+  tabsize=2
+}
+
 \title{\textsc{PAPERCLIP}: Associating Astronomical Observations and Natural Language with Multi-Modal Models}
@@
 
-\newcommand{\datafolder}[1]{\def\thedatafolder{#1}}
-
-% Define Verbatim environment with custom style
-\DefineVerbatimEnvironment{jsoncode}{Verbatim}{
-  commandchars=\\\{\},
-  % frame=single,
-  % framerule=0.5pt,
-  rulecolor=\color{codegray},
-  fillcolor=\color{codegray},
-  % label=JSON,
-  labelposition=topline,
-  fontsize=\small,
-  baselinestretch=1.1,
-  formatcom=\color{codeblue},
-  xleftmargin=15pt,
-  xrightmargin=15pt,
-  tabsize=2
-}
-
 \begin{document}
@@
 Using observations from the \hubble Space Telescope (HST) as an example, we show that the fine-tuned model embodies a meaningful joint representation between observations and text through tests targeting image retrieval (i.e., finding the most relevant observations using natural language queries) and description retrieval (i.e., querying for astrophysical object classes and science use cases most relevant to a given observation).
-Our study demonstrates the potential for using generalist rather than task-specific models for finding patterns in astronomical data, in particular by leveraging text as an interface.
+Our study demonstrates the potential for using generalist \changes{foundation models} rather than task-specific models for \changes{interacting with} astronomical data by leveraging text as an interface.
 \githubmaster
@@
 
+\begin{table}[h!]
+  \renewcommand{\arraystretch}{2}
+    \centering
+    \begin{tabular}{m{1.8cm} m{6.6cm} m{6.6cm}}
+        \toprule
+        \bfseries Prop. ID & \multicolumn{2}{c}{\bfseries LLM-extracted summary} \tabularnewline
+        \cmidrule(r){2-3}
+        & \centering\arraybackslash \bfseries Objects and phenomena & \centering\arraybackslash \bfseries Science use cases \tabularnewline
+        \midrule
+        \input{\thedatafolder/id1_2.txt} & {\scriptsize \input{\thedatafolder/obj1_2.txt}} & {\scriptsize \input{\thedatafolder/sci1_2.txt}} \tabularnewline
+        \midrule
+        \input{\thedatafolder/id1_1.txt} & {\scriptsize \input{\thedatafolder/obj1_1.txt}} & {\scriptsize \input{\thedatafolder/sci1_1.txt}} \tabularnewline
+        \midrule
+        \input{\thedatafolder/id1_3.txt} & {\scriptsize \input{\thedatafolder/obj1_3.txt}} & {\scriptsize \input{\thedatafolder/sci1_3.txt}} \tabularnewline
+        \midrule
+        \input{\thedatafolder/id1_0.txt} & {\scriptsize \input{\thedatafolder/obj1_0.txt}} & {\scriptsize \input{\thedatafolder/sci1_0.txt}} \tabularnewline
+        \bottomrule
+    \end{tabular}
+    \caption{\changes{For the \hubble proposal abstracts shown in Tab.~\ref{tab:dataset},} the LLM (\textsc{Mixtral-8x7B})-extracted summaries showing objects and phenomena (middle column) as well as potential downstream science use cases (last column) separately.}
+    \label{tab:datasetsumm}
+\end{table}
+
 \subsection{\hubble Data Selection and Pre-processing}
@@
 
-\begin{landscape}
-  \begin{table}[h!]
-    \renewcommand{\arraystretch}{2}
-      \centering
-      \begin{tabular}{m{1.8cm} m{8cm} m{5cm} m{6.5cm}}
-          \toprule
-          \bfseries Prop. ID & \centering\arraybackslash \bfseries Proposal abstract & \multicolumn{2}{c}{\bfseries LLM-extracted summary} \tabularnewline
-          \cmidrule(r){3-4}
-          & & \centering\arraybackslash \bfseries Objects and phenomena & \centering\arraybackslash \bfseries Science use cases \tabularnewline
-          \midrule
-          \input{\thedatafolder/id1_2.txt} & {\scriptsize \input{\thedatafolder/abs1_2.txt}} & {\scriptsize \input{\thedatafolder/obj1_2.txt}} & {\scriptsize \input{\thedatafolder/sci1_2.txt}} \tabularnewline
-          \midrule
-          \input{\thedatafolder/id1_1.txt} & {\scriptsize \input{\thedatafolder/abs1_1.txt}} & {\scriptsize \input{\thedatafolder/obj1_1.txt}} & {\scriptsize \input{\thedatafolder/sci1_1.txt}} \tabularnewline
-          \midrule
-          \input{\thedatafolder/id1_3.txt} & {\scriptsize \input{\thedatafolder/abs1_3.txt}} & {\scriptsize \input{\thedatafolder/obj1_3.txt}} & {\scriptsize \input{\thedatafolder/sci1_3.txt}} \tabularnewline
-          \midrule
-          \input{\thedatafolder/id1_0.txt} & {\scriptsize \input{\thedatafolder/abs1_0.txt}} & {\scriptsize \input{\thedatafolder/obj1_0.txt}} & {\scriptsize \input{\thedatafolder/sci1_0.txt}} \tabularnewline
-          \bottomrule
-      \end{tabular}
-      \caption{Examples of the clipped \hubble proposal abstracts (second column) and LLM (\textsc{Mixtral-8x7B})-extracted summaries (right-most two columns), separately extracting objects and phenomena as well as potential downstream science use cases.}
-      \label{tab:datasetsumm}
-  \end{table}
-  \end{landscape}
+% \begin{landscape}
+%   \begin{table}[h!]
+%     \renewcommand{\arraystretch}{2}
+%       \centering
+%       \begin{tabular}{m{1.8cm} m{8cm} m{5cm} m{6.5cm}}
+%           \toprule
+%           \bfseries Prop. ID & \centering\arraybackslash \bfseries Proposal abstract & \multicolumn{2}{c}{\bfseries LLM-extracted summary} \tabularnewline
+%           \cmidrule(r){3-4}
+%           & & \centering\arraybackslash \bfseries Objects and phenomena & \centering\arraybackslash \bfseries Science use cases \tabularnewline
+%           \midrule
+%           \input{\thedatafolder/id1_2.txt} & {\scriptsize \input{\thedatafolder/abs1_2.txt}} & {\scriptsize \input{\thedatafolder/obj1_2.txt}} & {\scriptsize \input{\thedatafolder/sci1_2.txt}} \tabularnewline
+%           \midrule
+%           \input{\thedatafolder/id1_1.txt} & {\scriptsize \input{\thedatafolder/abs1_1.txt}} & {\scriptsize \input{\thedatafolder/obj1_1.txt}} & {\scriptsize \input{\thedatafolder/sci1_1.txt}} \tabularnewline
+%           \midrule
+%           \input{\thedatafolder/id1_3.txt} & {\scriptsize \input{\thedatafolder/abs1_3.txt}} & {\scriptsize \input{\thedatafolder/obj1_3.txt}} & {\scriptsize \input{\thedatafolder/sci1_3.txt}} \tabularnewline
+%           \midrule
+%           \input{\thedatafolder/id1_0.txt} & {\scriptsize \input{\thedatafolder/abs1_0.txt}} & {\scriptsize \input{\thedatafolder/obj1_0.txt}} & {\scriptsize \input{\thedatafolder/sci1_0.txt}} \tabularnewline
+%           \bottomrule
+%       \end{tabular}
+%       \caption{Examples of the clipped \hubble proposal abstracts (second column) and LLM (\textsc{Mixtral-8x7B})-extracted summaries (right-most two columns), separately extracting objects and phenomena as well as potential downstream science use cases.}
+%       \label{tab:datasetsumm}
+%   \end{table}
+%   \end{landscape}
 
@@
 
-We use the CLIP-ViT-B/16\footnote{\url{https://huggingface.co/openai/clip-vit-base-patch16}} \citep{radford2021learning} variant as the base pre-trained CLIP model.
+We use the \texttt{CLIP-ViT-B/16} \citep{radford2021learning} variant as the base pre-trained CLIP model.
 %
@@
 %
-In total, the model has 149,620,737 trainable parameters.
+In total, the model contains $\sim 149$ million trainable parameters.
 %
@@
 Remarkably, the model trained on raw abstracts shows very similar performance when evaluated on the summarized abstracts compared to that trained on the summarized abstracts themselves, indicating that \emph{(1)} the image-text association signal is preserved in the summarization process, and \emph{(2)} the model is able to effectively leverage meaningful concepts in the noisy raw abstracts through weak supervision.
+\changes{The significantly worse performance of the model trained from scratch, compared to the fine-tuned models, highlights the crucial role of the inductive bias inherent in the base pretrained model, which effectively captures the rich associations between images and language.}
 
@@
   \includegraphics[width=0.49\textwidth]{plots/sim_val.pdf}
-  % \includegraphics[width=0.45\textwidth]{plots/sim_summ1.pdf}
   \includegraphics[width=0.49\textwidth]{plots/retrieval.pdf}
@@
   \end{figure*}
+
+\changes{We show retrieval accuracy performance for additional variations on the model and training configuration in App.~\ref{app:ablations}.}
 
@@
 
+\section{\changes{Variations on Model and Training}}
+\label{app:ablations}
+
+Figure~\ref{fig:sim_app} shows the retrieval accuracy as defined in Eq.~\eqrefb{eq:retrieval_accuracy} as a function of the retrieval fraction for further variations of the model or training, evaluated and trained on summarized abstracts.
+The red line corresponds to the model trained on summarized abstract described in the main text (fine-tuned on \texttt{CLIP-ViT-B/16} with constant learning rate $\mathrm{LR}=10^{-5}$ after linear warmup).
+The purple line corresponds to the base \texttt{CLIP-ViT-B/16} model.
+
+Curves for the model fine-tuned on the larger base CLIP model \texttt{CLIP-ViT-L/14} (dotted red), with a smaller learning rate $\mathrm{LR}=10^{-6}$ (dashed green), and with a cosine learning rate schedule (green) are also shown.
+All these models are seen to perform similarly, with the exception of the model trained with smaller learning rate showing degraded performance.
+Given the similar performance between \texttt{CLIP-ViT-L/14} ($\sim 428$ million parameters) and \texttt{CLIP-ViT-B/16} ($\sim 149$ million parameters), we chose the latter as the base model in the main text for computational efficiency.
+
+\begin{figure*}[!h]
+  \centering
+  \includegraphics[width=0.62\textwidth]{plots/retrieval_app.pdf}
+  \caption{Same as Fig.~\ref{fig:sim_valtrain} (right) -- retrieval accuracy as a function of the retrieval fraction -- for further variations on the model or training.
+The red and purple lines correspond to the model trained on summarized abstract, described in the main text, and the base \texttt{CLIP-ViT-B/16} model, respectively.
+Curves for the model fine-tuned on the larger base CLIP model \texttt{CLIP-ViT-L/14} (dotted red), with a smaller learning rate $\mathrm{LR}=10^{-6}$ (dashed green), and with a cosine learning rate schedule (green) are also shown.}
+  \label{fig:sim_app}
+  \end{figure*}
+
 \end{document}

### 2024-03-11 paper/hubble_paperclip.tex
Add overview fig

@@
 
-\newcommand{\githubmaster}{\href{https://github.com/smsharma/HubbleCLIP}{\faGithub}\xspace}
+\newcommand{\githubmaster}{\href{https://www.github.com/smsharma/PAPERCLIP-Hubble}{\faGithub}\xspace}
 
@@
   backgroundcolor=\color{backcolour}, 
-  commentstyle=\color{codegreen},
-   keywordstyle=\color{codegreen}, 
+  commentstyle=\color{deepgreen},
+  keywordstyle=\color{deepred}, 
   numberstyle=\tiny\color{codegray}, 
-   stringstyle=\color{codegreen},
+  stringstyle=\color{deepgreen},
   basicstyle=\ttfamily\footnotesize\linespread{1.1},
@@
 %
-% The paradigm has been highly successful in domains like computer vision and natural language processing, as demonstrated by the widespread use and adoption of tools like ChatGPT, Dall-E, Stable Diffusion, and CLIP.
-%
-% At the heart of the paradigm lies the triumph of scale -- scaling up model size, dataset size, and compute.
-% %
 These models are typically pre-trained on massive amounts of unlabeled data using self-supervised or weakly-supervised learning techniques, enabling them to learn powerful representations which can then be used downstream in different ways.
@@
 %
-Concretely, here we showcase the method using observations imaged by the \hubble Space Telescope (HST).
+Concretely, we showcase the method using observations imaged by the \hubble Space Telescope (HST).
 %
@@
 
-% \SM{Put in this work we at the start of the paragraph. Also put the paragraph right after AstroLlama, reworking some of the logic.} \SM{Define PAPERCLIP, funny footnote about PAPERCLIFT.}
-
 The rest of this paper is organized as follows.
@@
 We discuss future prospects and conclude in Sec.~\ref{sec:conclusion}.
+
+\begin{figure*}[!t]
+\centering
+\includegraphics[width=0.99\textwidth]{plots/figure.pdf}
+\caption{\changes{Overview of the PAPERCLIP method.
+(Left) A pre-trained CLIP model is fine-tuned using a dataset of \hubble observations and corresponding proposal abstracts.
+(Right) The fine-tuned model can then be used for downstream tasks such as observation retrieval (i.e., finding the observations most relevant to a given text query).
+The proposal abstract snippet here corresponds to proposal ID \href{https://archive.stsci.edu/proposal_search.php?id=16914&mission=hst}{16914}}}
+\label{fig:overview}
+\end{figure*}
 
@@
 This further motivates the need for fine tuning on domain-specific data.
-
-% \section{Dataset Construction}
-% \label{sec:dataset}
-
-% We curate a dataset of \hubble Space Telescope (HST) observations and corresponding text descriptions from publicly available sources.
-The dataset consists of images from the HST and associated proposal abstracts from the Hubble Space Telescope Proposal Abstracts Catalog\footnote{\url{https://archive.stsci.edu/hst/proposal_abstracts.html}}, which contains all accepted HST proposals since the telescope's launch on April 24, 1990.
-Our dataset covers observations up to Cycle 30, which began collecting data in 2022.
-
-% To generate captions for the observations, we use the proposal abstracts directly or optionally summarize them using guided language model generation (described in Sec.\ref{sec:summarization}).
-
-% The images in this dataset exhibit specific characteristics and artifacts inherent to the HST's data acquisition and processing pipeline, setting them apart from typical natural image datasets used for large-scale pre-training of foundation models \citep{deng2009imagenet}.
-This difference highlights the necessity of fine-tuning models on domain-specific data.
-
-% We emphasize that all data used in this work is publicly available, ensuring transparency and enabling reproducibility.
-The HST images can be accessed through the Hubble Legacy Archive\footnote{\url{https://hla.stsci.edu/}}, while the proposal abstracts are available in the Hubble Space Telescope Proposal Abstracts Catalog.
 
@@
 The guided generation process ensures that, in this case, the output of the LLM strictly conforms to the format of the following example:
-% \SM{Have it be centered}
-% \begin{lstlisting}[language=Python]
-% {
-%   'objects_and_phenomena': ['star forming galaxy', 'lensed galaxy', ...], 
-%   'science_use_cases': ['measure lensing magnification', 'probe spectral energy distributions', ...]
-% }
-% \end{lstlisting}
 \begin{center}
@@
 %
-% We emphasize that the intended goal of summarization-via-guided-generation is to increase the signal between text and images by standardizing the captions used for fine-tuning the CLIP model, and compare the quantitative performance of the model vs using the raw abstracts in Sec.~\ref{sec:results}.
-%
 We train separate models using the raw abstracts and the LLM-generated summaries, and compare their performance on downstream tasks in Sec.~\ref{sec:results}.
@@
 The goal of the fine-tuning process is to leverage the signal contained in this noisy association.
-
-% Finally, in order to test a further compression of the proposal abstracts, we use LLM-guided generation to also produce a list of single-concept summaries (e.g., `irregular galaxy', `quasars', `Galactic bulge', \ldots), and test whether this can lead to meaningful, generalizable learned associations with observations.
-% %
-% The prompts and schemata for generating these discrete categories and assigning observations to them are described in Apps.~\ref{app:singleconcept} and \ref{app:singleconceptassignments} respectively.
-\SM{Each abstract is assigned one summary.
-The generation of the single concepts is different }
-
-% \begin{landscape}
-%   \begin{table}[h!]
-%     \renewcommand{\arraystretch}{2}
-%       \centering
-%       \begin{tabular}{m{1.8cm} m{8cm} m{5cm} m{6.5cm}}
-%           \toprule
-%           \bfseries Prop. ID & \centering\arraybackslash \bfseries Proposal abstract & \multicolumn{2}{c}{\bfseries LLM-extracted summary} \tabularnewline
-%           \cmidrule(r){3-4}
-%           & & \centering\arraybackslash \bfseries Objects and phenomena & \centering\arraybackslash \bfseries Science use cases \tabularnewline
-%           \midrule
-%           \input{\thedatafolder/id1_2.txt} & {\scriptsize \input{\thedatafolder/abs1_2.txt}} & {\scriptsize \input{\thedatafolder/obj1_2.txt}} & {\scriptsize \input{\thedatafolder/sci1_2.txt}} \tabularnewline
-%           \midrule
-%           \input{\thedatafolder/id1_1.txt} & {\scriptsize \input{\thedatafolder/abs1_1.txt}} & {\scriptsize \input{\thedatafolder/obj1_1.txt}} & {\scriptsize \input{\thedatafolder/sci1_1.txt}} \tabularnewline
-%           \midrule
-%           \input{\thedatafolder/id1_3.txt} & {\scriptsize \input{\thedatafolder/abs1_3.txt}} & {\scriptsize \input{\thedatafolder/obj1_3.txt}} & {\scriptsize \input{\thedatafolder/sci1_3.txt}} \tabularnewline
-%           \midrule
-%           \input{\thedatafolder/id1_0.txt} & {\scriptsize \input{\thedatafolder/abs1_0.txt}} & {\scriptsize \input{\thedatafolder/obj1_0.txt}} & {\scriptsize \input{\thedatafolder/sci1_0.txt}} \tabularnewline
-%           \bottomrule
-%       \end{tabular}
-%       \caption{Examples of the clipped \hubble proposal abstracts (second column) and LLM (\textsc{Mixtral-8x7B})-extracted summaries (right-most two columns), separately extracting objects and phenomena as well as potential downstream science use cases.}
-%       \label{tab:datasetsumm}
-%   \end{table}
-%   \end{landscape}
 
@@
 
-% Similarly, Fig.~\ref{fig:sim_valtrain} (top right) shows the distribution of cosine similarities between corresponding image-text pairs for the single-concept captions (solid red) and the shuffled versions (dashed red).
-% %
-% The distribution for corresponding pairs is only modestly shifted compared to the shuffled version, indicating that assigning $\mathcal O(400)$ labels to the images using an LLM results in a weaker association signal compared to using noisy captions.
-% %
-% The purple line shows the distribution of cosine similarities between corresponding pairs with single-concept summaries, but evaluated using a model trained on the LLM-summarized abstracts.
-% %
-% This distribution is more skewed towards higher similarity values, indicating that the model trained on diverse caption can still better capture the association signal in the single-concept summaries compared to one trained specifically using the single-concept summaries.
-
 \paragraph*{Retrieval accuracy}
@@
 %
-% For example, it is able to return images with processing and assembly artifacts particular to HST (e.g., the lines through the middle in some images), which typically receive low similarity scores when using the base model.
-%
 The \texttt{dwarf galaxy}-queried images correspond to proposals aiming to measure the kinematics of the stellar cores of dwarf galaxies.
@@
 We show the result of image-to-text retrieval in Tab.~\ref{tab:itt}, for the base (second column) as well as summary fine-tuned (third column) models, using four observations (left-most column) from the validation set.
-% %
-% We curate a list of possible text associations by querying the \textsc{Claude}\footnote{\url{https://claude.ai/}} large language model for such a list, which we show in App.~\ref{app:categories}.
+%
 
@@
   \end{tabular}
-  \caption{For four text queries (left-most column), the four most similar images from the validation dataset by cosine similarity when using the \textbf{\textcolor{deeppurple}{base CLIP model}} (CLIP-ViT-B/16).
+  \caption{For four text queries (left-most column), the four most similar images from the validation dataset by cosine similarity when using the \textbf{\textcolor{deeppurple}{base CLIP model}} (\texttt{CLIP-ViT-B/16}).
 The proposal ID associated with each image is given below the image and contains a hyperlink to the MAST page corresponding to the proposal.}
@@
 \end{table}
-
-% \begin{figure*}[!h]
-% \includegraphics[width=0.95\textwidth]{plots/tti_base.pdf}
-% \caption{Image retrieval using the base CLIP model on four curated queries.
-\SM{Horizontal lines} \SM{Query not rotated} \SM{Same for next Figure.} \SM{Put text ``Base Model'' in consistent colour to legend.} \SM{Say top 4} \SM{Say that for science we care about the distributions of things.}}
-% \label{fig:tti_base}
-% \end{figure*}
-
-% \begin{figure*}[!h]
-% \includegraphics[width=0.95\textwidth]{plots/tti.pdf}
-% \caption{Image retrieval using the fine-tuned CLIP model on four curated queries.}
-% \label{fig:tti}
-% \end{figure*}
 
@@
 
-% \begin{figure*}[!h]
-% \includegraphics[width=0.95\textwidth]{plots/itt.pdf}
-% \caption{Text associations from a curated list most closely matching a given image query, for both the fine-tuned and base models.
-The `ground truth' LLM-summarized abstract is shown in the right column.
-\SM{Say ground truth summarized abstract in the column}\SM{Swap order of cols 2 and 3.}}
-% \label{fig:itt}
-% \end{figure*}
-
-% Set line spacing to 0.5
-
 \begin{table}[h!]
@@
   \end{tabular}
-  \caption{Text snippets from a curated list most closely matching a given image query (left-most column) by cosine similarity of respective embeddings, shown for the \textcolor{deeppurple}{base} (CLIP-ViT-B/16) and \textcolor{deepred}{summary fine-tuned} models.
+  \caption{Text snippets from a curated list most closely matching a given image query (left-most column) by cosine similarity of respective embeddings, shown for the \textcolor{deeppurple}{base} (\texttt{CLIP-ViT-B/16}) and \textcolor{deepred}{summary fine-tuned} models.
 The `ground truth' LLM-summarized abstract (only objects/phenomena) is shown in the right-most column.}
@@
 
-The code, dataset, and fine-tuned models used in this work are available at \url{https://www.github.com/smsharma/HubbleCLIP}.
+The code, dataset, and fine-tuned models used in this work are available at \url{https://www.github.com/smsharma/PAPERCLIP-Hubble}.
 
@@
 
-% The regex-guided generation method ensures that the LLM's output always adheres to the specified regular expression pattern, without incurring the computational overhead of checking the validity of each possible next token at every generation step.
-This approach can be extended to more complex patterns, such as context-free grammars, by using augmented pushdown automata to track the generation state and valid next tokens.
-
-% In our application, we use the regex-guided generation method to constrain LLM outputs to follow specific JSON schemas, which are automatically converted into regular expressions.
-This ensures that the generated summaries of astronomy observation proposals adhere to the desired format, enabling more reliable extraction of key information.
-% The pre-computed index $\sigma$ allows for efficient determination of valid next tokens at each generation step, making the regex-guided method computationally feasible for large-scale applications.
-
 \subsection{Prompts and Schema Used for Summarization}
@@
 \end{lstlisting}
-
-% \subsection{Generation of single-concept summaries}
-% \label{app:singleconcept}
-
-% The following prompt is used to generate a list of diverse single-concept summaries informed by a list of summarized abstracts:
-
-% \begin{lstlisting}[language=Python]
-% import outlines
-
-% @outlines.prompt
-% def prompt_fn(objects):
-%     """<s>[INST] Please produce a list of around concepts characterizing prominent objects, phenomena, and science use cases of images observed by the Hubble Space Telescope.
-
-% Here are some examples of objects:
-
-% {{objects}}
-
-% Follow these instructions exactly in your answer:
-% - Do not output empty strings as elements.
-% - Make sure that the list covers a diverse range of astronomical concepts, with items as different from each other as possible. 
-% - Do not give specific names of objects, to make sure you span the widest possible range of concepts (e.g., "dwarf galaxy" is allowed, but NOT "Fornax", "Terzan 5", or  "NGC6440").
-% - Do not return terms undescriptive of observations, e.g. "sloshing", "adiabatic", "interactions". Returning concrete physics objects, concepts, or phenomena.
-% - Only output scientifically meaningful terms. E.g., NO "Cosmic Dance".
-% - Do not duplicate entries. Do not reference any telescopes, observatories, or surveys.
-% - Do not include units like "angular diameter distance", "parsec", or any other concepts that will not correlate with images of observations.
-% - Use the above example list of objects only as inspiration to infer broad classes of objects.
-% - Make sure each concept is succint, never more than 5 words.
-% - Answer in JSON format.
-% - The JSON should have the following keys {"galaxies", "stellar_physics", "exoplanets_planet_formation", "stellar_populations", "supermassive_black_holes", "solar_system", "integalactic_medium", "large_scale_structure"} reflecting rough observation categories.
-% - Each category will have a list of objects and/or astronomical concepts.
-% - Output up to 20 items and no more in each category
-% [/INST]
-% """
-% \end{lstlisting}
-
-% The following schema guides generation, intended to produce 100 concepts reflecting the science categories of successful proposals\footnote{\url{https://www.stsci.edu/contents/newsletters/2023-volume-40-issue-02/hubble-cycle-31-proposal-selection}}.
-% %
-
-% \begin{lstlisting}[language=Python]
-% from pydantic import BaseModel, conlist
-
-% class ScienceCategoriesHST(BaseModel):
-%     galaxies: conlist(str, min_length=15, max_length=15)
-%     stellar_physics: conlist(str, min_length=15, max_length=15)
-%     exoplanets_planet_formation: conlist(str, min_length=15, max_length=15)
-%     stellar_populations: conlist(str, min_length=10, max_length=10)
-%     supermassive_black_holes: conlist(str, min_length=15, max_length=15)
-%     solar_system: conlist(str, min_length=10, max_length=10)
-%     integalactic_medium: conlist(str, min_length=10, max_length=10)
-%     large_scale_structure: conlist(str, min_length=10, max_length=10)
-%   \end{lstlisting}
-
-% The output is then constrained to be one of the fixed number of concepts.
-
-% \subsection{Assignment of abstracts to single-concept categories}
-% \label{app:singleconceptassignments}
-
-% Finally, the following prompt is used to assign concepts inferred in \ref{app:singleconcept} to each abstract:
-
-% \begin{lstlisting}[language=Python]
-% import outlines
-
-% @outlines.prompt
-% def prompt_fn(abs, cats):
-%     """<s>[INST] The following is a successful proposal abstract for the Hubble Space Telescope: "{{abs}}"
-
-% The following is a list of categories (astronomical concepts) that this abstract could correspond to.
-
-% {{cats}}
-
-% Please answer which of these listed concepts best describes this proposal, based on the objects and phenomena mentioned in the abstract.
-% The concept should meaningfully be present in the abstract and the eventual observation.
-
-% - For example, "The locations of supernovae {SNe} in the local stellar and gaseous environment in galaxies, as measured in high spatial resolution WFPC2 and ACS images, contain important clues to their progenitor stars." should return "supernova".
-% - If the abstract centers calibration and/or instrumentation efforts, return calibration or instrumention".
-
-% If no concept make sense, return "None". [/INST]
-% """
-% \end{lstlisting}
 
@@
   \end{tabular}
-  \caption{Text associations from a curated list most closely matching four image queries (first column, the same as in Tab.~\ref{tab:itt}), for the \textcolor{deeppurple}{base} (CLIP-ViT-B/16), \textcolor{deepblue}{abstract fine-tuned}, and \textcolor{deepred}{summary fine-tuned} models.}
+  \caption{Text associations from a curated list most closely matching four image queries (first column, the same as in Tab.~\ref{tab:itt}), for the \textcolor{deeppurple}{base} (\texttt{CLIP-ViT-B/16}), \textcolor{deepblue}{abstract fine-tuned}, and \textcolor{deepred}{summary fine-tuned} models.}
   \label{tab:itt_abs}
@@
 
-\section{\changes{Variations on Model and Training}}
+\section{\changes{Additional Variations on Model and Training}}
 \label{app:ablations}

### 2024-03-11 paper/hubble_paperclip.tex
Started adding related works

@@
 
-The CLIP family of foundation models, which in their original form embed images and associated captions into a common representation space via contrastive learning, have shown strong performance and generalization capabilities on a variety of downstream tasks including zero-shot classification and image retrieval.
+\paragraph*{Related work}
+
+% The CLIP family of foundation models, which in their original form embed images and associated captions into a common representation space via contrastive learning, have shown strong performance and generalization capabilities on a variety of downstream tasks including zero-shot classification and image retrieval~\citep{radford2021learning}.
 %
-The concept of associating diverse modalities via contrastive training has been employed in other scientific domains~\citep[e.g.,][]{liu2023text,Sanchez-Fernandez2022.11.17.516915,lanusse2023astroclip,cepeda2023geoclip}, and has been shown to be effective in learning semantically meaningful joint representations.
-Here, we present for the first time an application associating astronomical data with the text modality.
+The concept of learning task-agnostic representations via self-supervised learning has been applied within astrophysics \cite{slijepcevic2024radio,stein2021self,hayat2021self,slijepcevic2022learning} and used for downstream tasks like object similar search \citep{stein2021self}, gravitational lens finding \citep{stein2022mining}, estimation of Galactic distances \citep{hayat2021estimating}, and identification of rare galaxies \citep{walmsley2023rare}.
+%
+Associating different modalities via contrastive training has been employed in many other scientific domains~\citep[e.g.,][]{liu2023text,Sanchez-Fernandez2022.11.17.516915,lanusse2023astroclip,cepeda2023geoclip}, and has been shown to be effective in learning semantically meaningful joint representations.
+Here, we present for the first time an application associating diverse astronomical data with the text modality.
+%
+For a recent review of contrastive learning in astrophysics, see \citet{huertas2023brief}. 
+%
+% \citet{bowles2023radio,bowles2022new}
 
@@
 (Right) The fine-tuned model can then be used for downstream tasks such as observation retrieval (i.e., finding the observations most relevant to a given text query).
-The proposal abstract snippet here corresponds to proposal ID \href{https://archive.stsci.edu/proposal_search.php?id=16914&mission=hst}{16914}}}
+The proposal abstract snippet here corresponds to proposal ID \href{https://archive.stsci.edu/proposal_search.php?id=16914&mission=hst}{16914}}.}
 \label{fig:overview}
@@
 
-\subsection{Contrastive Language-Image Pre-training (CLIP)}
+\subsection{Contrastive Language-Image Pre-training}
 
-CLIP \citep[Contrastive Language-Image Pre-training;][]{radford2021learning} is a multi-modal neural network model pre-trained on a large corpus of image-text pairs via weak supervision using a contrastive loss.
+Contrastive Language-Image Pre-training \citep[CLIP;][]{radford2021learning} is a multi-modal neural network model pre-trained on a large corpus of image-text pairs via weak supervision using a contrastive loss.
 %
@@
 %
-For the description/text retrieval evaluation, we define a list of possible text associations (i.e., classes), which we show in App.~\ref{app:categories}, by querying the \textsc{Claude 2}\footnote{\url{https://claude.ai/}} large language model along with manual curation.
+For the description/text retrieval evaluation, we define a list of possible text associations (i.e., classes), which we show in App.~\ref{app:categories}, by querying the \textsc{Claude 2}\footnote{\url{https://claude.ai/}} large language followed by with manual curation.
 

### 2024-03-11 paper/hubble_paperclip.tex
Finished related works

@@
 
-\def\preprintno{XXXX} % Insert correct preprint number
-
 \usepackage[
@@
 % Define a new fancy page style
+% Insert correct preprint number
+\def\preprintno{5690} 
 \fancypagestyle{firstpage}{
@@
 
-% \SM{Add some qualitative evals with the abstract model.}
-
 Machine learning (ML) is starting to have a significant impact in the sciences, with astrophysics being no exception.
@@
 %
-\textsc{AstroLLaMA}~\citep{nguyen2023astrollama,perkowski2024astrollama} is another recent effort to fine-tune a publicly-available model (\textsc{Llama-2}) on astrophysics-specific textual data from the arXiv.
 
@@
 %
-
-\paragraph*{Related work}
-
-% The CLIP family of foundation models, which in their original form embed images and associated captions into a common representation space via contrastive learning, have shown strong performance and generalization capabilities on a variety of downstream tasks including zero-shot classification and image retrieval~\citep{radford2021learning}.
-%
-The concept of learning task-agnostic representations via self-supervised learning has been applied within astrophysics \cite{slijepcevic2024radio,stein2021self,hayat2021self,slijepcevic2022learning} and used for downstream tasks like object similar search \citep{stein2021self}, gravitational lens finding \citep{stein2022mining}, estimation of Galactic distances \citep{hayat2021estimating}, and identification of rare galaxies \citep{walmsley2023rare}.
-%
-Associating different modalities via contrastive training has been employed in many other scientific domains~\citep[e.g.,][]{liu2023text,Sanchez-Fernandez2022.11.17.516915,lanusse2023astroclip,cepeda2023geoclip}, and has been shown to be effective in learning semantically meaningful joint representations.
-Here, we present for the first time an application associating diverse astronomical data with the text modality.
-%
-For a recent review of contrastive learning in astrophysics, see \citet{huertas2023brief}. 
-%
-% \citet{bowles2023radio,bowles2022new}
 
@@
 (Left) A pre-trained CLIP model is fine-tuned using a dataset of \hubble observations and corresponding proposal abstracts.
+The proposal abstracts are optionally summarized using guided large language model generation.
 (Right) The fine-tuned model can then be used for downstream tasks such as observation retrieval (i.e., finding the observations most relevant to a given text query).
-The proposal abstract snippet here corresponds to proposal ID \href{https://archive.stsci.edu/proposal_search.php?id=16914&mission=hst}{16914}}.}
+The proposal abstract snippet shown here corresponds to proposal ID \href{https://archive.stsci.edu/proposal_search.php?id=16914&mission=hst}{16914}}.}
 \label{fig:overview}
 \end{figure*}
+
+\section{\changes{Related work}}
+
+% The CLIP family of foundation models, which in their original form embed images and associated captions into a common representation space via contrastive learning, have shown strong performance and generalization capabilities on a variety of downstream tasks including zero-shot classification and image retrieval~\citep{radford2021learning}.
+%
+The concept of learning task-agnostic representations via self-supervised and contrastive learning has been applied within astrophysics \citep{slijepcevic2024radio,stein2021self,hayat2021self,slijepcevic2022learning} and used for downstream tasks like object similarity search \citep{stein2021self}, gravitational lens finding \citep{stein2022mining}, estimation of Galactic distances \citep{hayat2021estimating}, identification of rare galaxies \citep{walmsley2023rare}, and data compression \citep{akhmetzhanova2024data}.
+For a recent review of contrastive learning in astrophysics, see \citet{huertas2023brief}.
+%
+
+Beyond applications to a single modality, \textsc{AstroCLIP}~\citep{lanusse2023astroclip} recently used contrastive learning to learn a joint representation between galaxy images and associated spectra, showing that the learned representation embodies relevant physical properties and can be effectively used for downstream tasks like redshift and mass estimation.
+%
+
+\citet{bowles2023radio,bowles2022new} introduced a method to associate radio galaxy images with a natural language description of their morphology by using human-generated descriptions, with the goal of deriving semantic morphology classes and using them for classification.
+
+%
+Associating diverse modalities via contrastive learning has been employed in many other scientific domains~\citep[e.g.,][]{liu2023text,Sanchez-Fernandez2022.11.17.516915,lanusse2023astroclip,cepeda2023geoclip}, and has been shown to be effective in learning semantically meaningful joint representations.
+
+% While the interaction of astrophysical data with large language models is fairly nascent, \textsc{AstroLLaMA}~\citep{nguyen2023astrollama,perkowski2024astrollama} is a recent effort to fine-tune a publicly-available model (\textsc{Llama-2}) on astrophysics-specific textual data from the arXiv.
+
+In this paper, we present for the first time an application associating target-agnostic astronomical data with the text modality, showing that this can be effectively accomplished through contrastive learning by leveraging observational proposals to inform text captions.
+%
 

### 2024-03-11 paper/hubble_paperclip.tex
Reference to overview figure
Spheres in paper

@@
 Our method opens up the possibility of interacting with astronomical survey data using free-form natural language as an interface, which is a cornerstone of the success of the modern foundation model paradigm.
+\changes{A high-level overview of the method is shown in Fig.~\ref{fig:overview}}.
 %
@@
 
-\section{\changes{Related work}}
+\section{\changes{Related Work}}
 
@@
           \color{black}\{
-            \color{codegreen}'objects_and_phenomena'\color{black}: [\color{codegreen}'star forming galaxy', 'lensed galaxy'\color{black}], 
-            \color{codegreen}'science_use_cases'\color{black}: [\color{codegreen}'measure lensing magnification'\color{black}]
+            \color{codegreen}'objects_and_phenomena'\color{black}: [\color{codegreen}'star forming galaxy', 'lensed galaxy'\color{black}, ...], 
+            \color{codegreen}'science_use_cases'\color{black}: [\color{codegreen}'measure lensing magnification'\color{black}, ...]
           \color{black}\}

### 2024-03-27 paper/colm_submission/colm2024_conference.tex
COLM draft

--- 2024-03-11 paper/hubble_paperclip.tex
@@
-\documentclass[10pt]{article} % For LaTeX2e
-\usepackage[preprint]{tmlr}
-% If accepted, instead use the following line for the camera-ready submission:
-%\usepackage[accepted]{tmlr}
-% To de-anonymize and remove mentions to TMLR (for example for posting to preprint servers), instead use the following:
-%\usepackage[preprint]{tmlr}
+\documentclass{article} % For LaTeX2e
+\usepackage{colm2024_conference}
 
-% Optional math commands from https://github.com/goodfeli/dlbook_notation.
-\input{math_commands.tex}
+\usepackage{microtype}
+% \usepackage{hyperref}
+\usepackage{url}
+\usepackage{booktabs}
 
-\usepackage{url}
 \usepackage{graphicx}
 \usepackage{array}
-\usepackage{booktabs}
 \usepackage{xspace}
@@
 
-% Custom colors
-\usepackage{color}
+\definecolor{darkblue}{rgb}{0, 0, 0.5}
+% \hypersetup{colorlinks=true, citecolor=darkblue, linkcolor=darkblue, urlcolor=darkblue}
 
@@
 
-% For model colours
+% For model colours and code
 \definecolor{deepblue}{rgb}{0.29411765 0.45882353 0.61960784}
@@
 \definecolor{deeppurple}{rgb}{0.52941176 0.32941176 0.56470588}
-
-% For code listings
-\definecolor{codegreen}{rgb}{0,0.4,0}
 \definecolor{codegray}{rgb}{0.5,0.5,0.5}
-\definecolor{codepurple}{rgb}{0.58,0,0.82}
 \definecolor{backcolour}{rgb}{0.95,0.95,0.92}
@@
 \newcommand{\package}[1]{\textsl{#1}\xspace}
-\newcommand{\eg}{{e.\,g.}\xspace}
-\newcommand{\ie}{{i.\,e.}\xspace}
 \newcommand{\hubble}{\emph{Hubble}\xspace}
@@
 \newcommand{\SM}[1]{\textcolor{blue}{[SM: #1]}}
-% Changes
 \newcommand{\changes}[1]{\textcolor{blue}{#1}}
@@
 
+% Listings style
 \lstdefinestyle{mystyle}{ 
@@
   baselinestretch=1.1,
-  formatcom=\color{codegreen},
+  formatcom=\color{deepgreen},
   xleftmargin=15pt,
@@
 
-\author{\name Siddharth Mishra-Sharma \email \href{mailto:smsharma@mit.edu}{smsharma@mit.edu} \\
-      \addr The NSF AI Institute for Artificial Intelligence and Fundamental Interactions\\
-      Center for Theoretical Physics, Massachusetts Institute of Technology, Cambridge, MA 02139, USA \\
-      Department of Physics, Harvard University, Cambridge, MA 02138, USA
+% Authors must not appear in the submitted version. They should be hidden
+% as long as the \colmfinalcopy macro remains commented out below.
+% Non-anonymous submissions will be rejected without review.
+
+\author{Antiquus S.~Hippocampus, Natalia Cerebro \& Amelie P. Amygdale \thanks{ Use footnote for providing further information
+about author (webpage, alternative address)---\emph{not} for acknowledging
+funding agencies.  Funding acknowledgements go at the end of the paper.} \\
+Department of Computer Science\\
+Cranberry-Lemon University\\
+Pittsburgh, PA 15213, USA \\
+\texttt{\{hippo,brain,jen\}@cs.cranberry-lemon.edu} \\
+\And
+Ji Q. Ren \& Yevgeny LeNet \\
+Department of Computational Neuroscience \\
+University of the Witwatersrand \\
+Joburg, South Africa \\
+\texttt{\{robot,net\}@wits.ac.za} \\
 \AND
-      \name Yiding Song \email \href{mailto:ydsong@mit.edu}{ydsong@mit.edu} \\
-      \addr The NSF AI Institute for Artificial Intelligence and Fundamental Interactions\\
-      Department of Physics, Massachusetts Institute of Technology, Cambridge, MA 02139, USA \\
-      \AND
-      \name Jesse Thaler \email \href{mailto:jthaler@mit.edu}{jthaler@mit.edu} \\
-      \addr The NSF AI Institute for Artificial Intelligence and Fundamental Interactions\\
-      Center for Theoretical Physics, Massachusetts Institute of Technology, Cambridge, MA 02139, USA \\
+Coauthor \\
+Affiliation \\
+Address \\
+\texttt{email}
 }
+
+% The \author macro works with any number of authors. There are two commands
+% used to separate the names and addresses of multiple authors: \And and \AND.
+%
+% Using \And between authors leaves it to \LaTeX{} to determine where to break
+% the lines. Using \AND forces a linebreak at that point. So, if \LaTeX{}
+% puts 3 of 4 authors names on the first line, and the last on the second
+% line, try using \AND instead of \And before the third author name.
 
@@
 
-\def\month{MM}  % Insert correct month for camera-ready version
-\def\year{YYYY} % Insert correct year for camera-ready version
-\def\openreview{\url{https://openreview.net/forum?id=XXXX}} % Insert correct link to OpenReview for camera-ready version
-
+% \colmfinalcopy % Uncomment for camera-ready version, but NOT for submission.
 \begin{document}
@@
 
-\thispagestyle{firstpage}
-
 \begin{abstract}
-We present PAPERCLIP (Proposal Abstracts Provide an Effective Representation for Contrastive Language-Image Pre-training), a method which associates astronomical observations imaged by surveys and telescopes with natural language  using a neural network model.
-The model is fine-tuned from a pre-trained Contrastive Language–Image Pre-training (CLIP) model using successful observing proposal abstracts, optionally summarized via guided generation using large language models (LLMs), and corresponding downstream observations.
-Using observations from the \hubble Space Telescope (HST) as an example, we show that the fine-tuned model embodies a meaningful joint representation between observations and text through tests targeting image retrieval (i.e., finding the most relevant observations using natural language queries) and description retrieval (i.e., querying for astrophysical object classes and science use cases most relevant to a given observation).
-Our study demonstrates the potential for using generalist \changes{foundation models} rather than task-specific models for \changes{interacting with} astronomical data by leveraging text as an interface.
-\githubmaster
+   We present PAPERCLIP (Proposal Abstracts Provide an Effective Representation for Contrastive Language-Image Pre-training), a method which associates astronomical observations imaged by telescopes with natural language using a neural network model.
+The model is fine-tuned from a pre-trained Contrastive Language–Image Pre-training (CLIP) model using successful observing proposal abstracts and corresponding downstream observations, with the abstracts optionally summarized via guided generation using large language models (LLMs).
+Using observations from the \hubble Space Telescope (HST) as an example, we show that the fine-tuned model embodies a meaningful joint representation between observations and natural language through quantitative evaluation as well as tests targeting image retrieval (i.e., finding the most relevant observations using natural language queries).
+  and description retrieval (i.e., querying for astrophysical object classes and use cases most relevant to a given observation). 
+  Our study demonstrates the potential for using generalist foundation models rather than task-specific models for interacting with astronomical data by leveraging text as an interface.
 \end{abstract}
-
-\tableofcontents
 
@@
 %
-Until recently, most applications of ML within astrophysics have focused on augmenting traditional techniques in order to improve performance on specific tasks.
+Most applications of ML within astrophysics have focused on augmenting traditional techniques in order to improve performance on specific tasks.
 %
-The {Foundation Model} paradigm, in contrast, seeks to develop generalist models which can be deployed to simultaneously tackle a wide range of tasks \citep{bommasani2021opportunities}.
+The {foundation model} paradigm, in contrast, seeks to develop generalist models which can be deployed to simultaneously tackle a wide range of tasks \citep{bommasani2021opportunities}.
 %
-These models are typically pre-trained on massive amounts of unlabeled data using self-supervised or weakly-supervised learning techniques, enabling them to learn powerful representations which can then be used downstream in different ways.
+These models are typically pre-trained on massive amounts of unlabeled data using self-supervised or weakly-supervised learning techniques, enabling them to learn powerful representations which can then be used downstream.
 %
-Foundation models can often benefit from additional training ({fine tuning}) using a relatively small amounts of domain-specific data in order to increase their usefulness when applied to specialized domains.
+Foundation models can often benefit from additional training ({fine-tuning}) using a relatively small amount of domain-specific data in order to increase their usefulness when applied to specialized domains.
 
-There is considerable interest in developing custom foundation models for the sciences \citep[e.g., ][]{batatia2023foundation,subramanian2023towards}, with astrophysics being ripe for such an effort given the large amounts of publicly-available data and diverse modes of interacting with it.
+There is considerable interest in developing custom foundation models for the sciences \citep[e.g., ][]{batatia2023foundation,subramanian2023towards,mccabe2023multiple,Birk:2024knn,vig2024finetuning,heinrich2024masked}, with astrophysics being ripe for such an effort given the large amounts of publicly-available data and diverse ways of interacting with it.
 %
@@
 %
-This multi-modality was recently exploited in \textsc{AstroCLIP}~\citep{lanusse2023astroclip} to construct a joint embedding space between multi-band images and optical spectra from the Dark Energy Spectroscopic Instrument (DESI).
-%
+% This multi-modality was recently exploited in \textsc{AstroCLIP}~\citep{lanusse2023astroclip} to construct a joint physically-informative embedding space between multi-band images and optical spectra from the Dark Energy Spectroscopic Instrument (DESI).
+% %
 
-In this paper, we describe \text{PAPERCLIP} (Proposal Abstracts Provide an Effective Representation for Contrastive Language-Image Pre-training\footnote{Technically, we fine tune rather than pre train, but ``PAPERCLIFT'' was rejected by the senior author of this paper.}), a method that connects astronomical image observations with natural language by leveraging the association between abstracts of successful observing proposals and images corresponding to downstream observations.
+In this paper, we describe \text{PAPERCLIP} (Proposal Abstracts Provide an Effective Representation for Contrastive Language-Image Pre-training),
+% \footnote{Technically, we fine tune rather than pre train, but ``PAPERCLIFT'' was rejected by the senior author of this paper.}), 
+a method that connects, for the first time, astronomical image observations with natural language by leveraging the association between abstracts of successful observing proposals written by astronomers and images corresponding to downstream observations imaged by telescopes.
 %
@@
 Our method opens up the possibility of interacting with astronomical survey data using free-form natural language as an interface, which is a cornerstone of the success of the modern foundation model paradigm.
+A high-level overview of the method is shown in Fig.~\ref{fig:overview}.
 %
+
+\begin{figure*}[!t]
+   \centering
+   \includegraphics[width=0.97\textwidth]{plots/figure.pdf}
+   \caption{Overview of the PAPERCLIP method.
+(Left) A pre-trained CLIP model is fine-tuned using a dataset of \hubble observations and corresponding proposal abstracts.
+The proposal abstracts are optionally summarized using guided large language model generation.
+(Right) The fine-tuned model can then be used for downstream tasks such as observation retrieval i.e., finding the observations most relevant to a given text query.
+The proposal abstract snippet shown here corresponds to proposal ID \href{https://archive.stsci.edu/proposal_search.php?id=16914&mission=hst}{16914}.}
+   \label{fig:overview}
+   \end{figure*}
+
+  %  \section{Related Work}
+  %  \label{sec:related}
+
+   \paragraph*{Related Work}
+
+   The concept of learning task-agnostic representations via self-supervised and contrastive learning has been applied within astrophysics \citep{slijepcevic2024radio,stein2021self,hayat2021self,slijepcevic2022learning} and used for downstream tasks like object similarity search \citep{stein2021self}, gravitational lens finding \citep{stein2022mining}, estimation of Galactic distances \citep{hayat2021estimating}, identification of rare galaxies \citep{walmsley2023rare}, and data compression \citep{akhmetzhanova2024data}.
+For a recent review of contrastive learning in astrophysics, see \citet{huertas2023brief}.
+   %
+   Beyond applications to a single modality, \textsc{AstroCLIP}~\citep{lanusse2023astroclip} recently used contrastive learning to learn a joint representation between galaxy images and associated spectra, showing that the learned representation embodies relevant physical properties and can be effectively used for downstream tasks like redshift and mass estimation.
+   %
+   \citet{bowles2023radio,bowles2022new} introduced a method to associate radio galaxy images with a natural language description of their morphology by using human-generated descriptions, with the goal of deriving semantic morphology classes and using them for classification.
+  %  %
+  In contrast with previous work, our application is the first to associate astronomical observation with the text modality in a generalist framework, showcasing the potential of language models in specialized scientific domains like astronomy.
+  %
+
+  %  Associating diverse modalities via contrastive learning has been employed in many other scientific domains~\citep[e.g.,][]{liu2023text,Sanchez-Fernandez2022.11.17.516915,lanusse2023astroclip,cepeda2023geoclip}, and has been shown to be effective in learning semantically meaningful joint representations.
+
+  %  In this paper, we present for the first time an application associating target-agnostic astronomical data with the text modality, showing that this can be effectively accomplished through contrastive learning by leveraging observing proposal abstracts to inform text captions.
+  %  %
 
@@
 %
+% We review related work in Sec.~\ref{sec:related}.
+% %
 In Sec.~\ref{sec:dataset}, we describe the \hubble dataset used in this work, including the curation and processing of observations as well as text captions.
@@
 We discuss future prospects and conclude in Sec.~\ref{sec:conclusion}.
-
-\begin{figure*}[!t]
-\centering
-\includegraphics[width=0.99\textwidth]{plots/figure.pdf}
-\caption{\changes{Overview of the PAPERCLIP method.
-(Left) A pre-trained CLIP model is fine-tuned using a dataset of \hubble observations and corresponding proposal abstracts.
-The proposal abstracts are optionally summarized using guided large language model generation.
-(Right) The fine-tuned model can then be used for downstream tasks such as observation retrieval (i.e., finding the observations most relevant to a given text query).
-The proposal abstract snippet shown here corresponds to proposal ID \href{https://archive.stsci.edu/proposal_search.php?id=16914&mission=hst}{16914}}.}
-\label{fig:overview}
-\end{figure*}
-
-\section{\changes{Related work}}
-
-% The CLIP family of foundation models, which in their original form embed images and associated captions into a common representation space via contrastive learning, have shown strong performance and generalization capabilities on a variety of downstream tasks including zero-shot classification and image retrieval~\citep{radford2021learning}.
-%
-The concept of learning task-agnostic representations via self-supervised and contrastive learning has been applied within astrophysics \citep{slijepcevic2024radio,stein2021self,hayat2021self,slijepcevic2022learning} and used for downstream tasks like object similarity search \citep{stein2021self}, gravitational lens finding \citep{stein2022mining}, estimation of Galactic distances \citep{hayat2021estimating}, identification of rare galaxies \citep{walmsley2023rare}, and data compression \citep{akhmetzhanova2024data}.
-For a recent review of contrastive learning in astrophysics, see \citet{huertas2023brief}.
-%
-
-Beyond applications to a single modality, \textsc{AstroCLIP}~\citep{lanusse2023astroclip} recently used contrastive learning to learn a joint representation between galaxy images and associated spectra, showing that the learned representation embodies relevant physical properties and can be effectively used for downstream tasks like redshift and mass estimation.
-%
-
-\citet{bowles2023radio,bowles2022new} introduced a method to associate radio galaxy images with a natural language description of their morphology by using human-generated descriptions, with the goal of deriving semantic morphology classes and using them for classification.
-
-%
-Associating diverse modalities via contrastive learning has been employed in many other scientific domains~\citep[e.g.,][]{liu2023text,Sanchez-Fernandez2022.11.17.516915,lanusse2023astroclip,cepeda2023geoclip}, and has been shown to be effective in learning semantically meaningful joint representations.
-
-% While the interaction of astrophysical data with large language models is fairly nascent, \textsc{AstroLLaMA}~\citep{nguyen2023astrollama,perkowski2024astrollama} is a recent effort to fine-tune a publicly-available model (\textsc{Llama-2}) on astrophysics-specific textual data from the arXiv.
-
-In this paper, we present for the first time an application associating target-agnostic astronomical data with the text modality, showing that this can be effectively accomplished through contrastive learning by leveraging observational proposals to inform text captions.
-%
 
@@
    %
-It can be seen that the images in this dataset exhibit specific characteristics as well as artifacts particular to HST data-taking and processing which distinguishes them from the distribution of natural images typically used for large-scale pre-training of foundation models \citep{deng2009imagenet}.
+   It can be seen that the images in this dataset exhibit specific characteristics as well as artifacts particular to HST data-taking and processing which distinguishes them from the distribution of natural images typically used for large-scale pre-training of foundation models.
    %
-This further motivates the need for fine tuning on domain-specific data.
+   This further motivates fine-tuning on domain-specific data.
 
@@
      \centering
-  \begin{tabular}{m{0.20\textwidth} p{1.9cm} p{1.9cm} m{7.5cm}}
+     \begin{tabular}{m{0.20\textwidth} p{1.9cm} p{1.9cm} m{5.5cm}}
          \toprule
@@
          \midrule
-      \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_1.pdf} & \centering \input{\thedatafolder/cycle_1.txt} \\ (2013) & \centering \input{\thedatafolder/id_1.txt} &  {\scriptsize \input{\thedatafolder/abs1_1.txt}} \tabularnewline
-      \midrule
+        %  \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_1.pdf} & \centering \input{\thedatafolder/cycle_1.txt} \\ (2013) & \centering \input{\thedatafolder/id_1.txt} &  {\scriptsize \input{\thedatafolder/abs1_1.txt}} \tabularnewline
+        %  \midrule
          \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_3.pdf} & \centering \input{\thedatafolder/cycle_3.txt} \\ (2016) & \centering \input{\thedatafolder/id_3.txt} &  {\scriptsize \input{\thedatafolder/abs1_3.txt}} \tabularnewline
-      \midrule
-      \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_0.pdf} & \centering \input{\thedatafolder/cycle_0.txt} \\ (2019) & \centering \input{\thedatafolder/id_0.txt} &  {\scriptsize \input{\thedatafolder/abs1_0.txt}} \tabularnewline
+        %  \midrule
+        %  \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_0.pdf} & \centering \input{\thedatafolder/cycle_0.txt} \\ (2019) & \centering \input{\thedatafolder/id_0.txt} &  {\scriptsize \input{\thedatafolder/abs1_0.txt}} \tabularnewline
          \bottomrule
@@
        \centering
-    \begin{tabular}{m{1.8cm} m{6.6cm} m{6.6cm}}
+       \begin{tabular}{m{1.8cm} m{3.6cm} m{7.2cm}}
            \toprule
@@
            \cmidrule(r){2-3}
-        & \centering\arraybackslash \bfseries Objects and phenomena & \centering\arraybackslash \bfseries Science use cases \tabularnewline
+           & \centering\arraybackslash \bfseries \small{Objects and phenomena} & \centering\arraybackslash \bfseries \small{Science use cases} \tabularnewline
            \midrule
@@
            \midrule
-        \input{\thedatafolder/id1_1.txt} & {\scriptsize \input{\thedatafolder/obj1_1.txt}} & {\scriptsize \input{\thedatafolder/sci1_1.txt}} \tabularnewline
-        \midrule
+          %  \input{\thedatafolder/id1_1.txt} & {\scriptsize \input{\thedatafolder/obj1_1.txt}} & {\scriptsize \input{\thedatafolder/sci1_1.txt}} \tabularnewline
+          %  \midrule
            \input{\thedatafolder/id1_3.txt} & {\scriptsize \input{\thedatafolder/obj1_3.txt}} & {\scriptsize \input{\thedatafolder/sci1_3.txt}} \tabularnewline
-        \midrule
-        \input{\thedatafolder/id1_0.txt} & {\scriptsize \input{\thedatafolder/obj1_0.txt}} & {\scriptsize \input{\thedatafolder/sci1_0.txt}} \tabularnewline
+          %  \midrule
+          %  \input{\thedatafolder/id1_0.txt} & {\scriptsize \input{\thedatafolder/obj1_0.txt}} & {\scriptsize \input{\thedatafolder/sci1_0.txt}} \tabularnewline
            \bottomrule
        \end{tabular}
-    \caption{\changes{For the \hubble proposal abstracts shown in Tab.~\ref{tab:dataset},} the LLM (\textsc{Mixtral-8x7B})-extracted summaries showing objects and phenomena (middle column) as well as potential downstream science use cases (last column) separately.}
+       \caption{For the \hubble proposal abstracts shown in Tab.~\ref{tab:dataset}, the LLM (\textsc{Mixtral-8x7B})-extracted summaries showing objects and phenomena (middle column) as well as potential downstream science use cases (last column) separately.
+The proposal IDs (left column) contain hyperlinks to the MAST page corresponding to the proposal.}
        \label{tab:datasetsumm}
@@
    %
-We note that these are not science-grade observations, but rather lower-resolution images useful for diagnostic purposes.
+   We note that these are not science-grade observations, but rather lower-resolution images useful for diagnostic or preview purposes.
    %
@@
    %
-Color previews (i.e., observations taken with multiple wavelength filters assigned to individual RGB channels) are manually excluded via a filename filter in order to maintain consistency across the dataset; models trained on datasets with color images included were observed to show worse performance on evaluation metrics.
+   Color previews (i.e., observations taken with multiple wavelength filters assigned to individual RGB channels) are manually excluded via a filename filter in order to maintain consistency across the samples.
+   % models trained on datasets with color images included were observed to show worse performance on evaluation metrics.
    %
@@
    3,194 images are held out for validation, with no abstract being common between training and validation sets in order to ensure an independent set of image-text pairs for evaluation.
-The held out images correspond to 438 unique abstracts.
+The held out images correspond to 429 unique abstracts.
 % Double check numbers in final version.
@@
    %
-We choose to not filter out these pairs from our dataset, in order to have a larger sample of HST observations that the model can leverage to adapt to the distinctive characteristics of \hubble images.
-
-% Mention that the signal is noisy!
+   We choose to not filter out these pairs, in order to have a larger sample of HST observations that the model can leverage to adapt to the distinctive characteristics of \hubble images.
 
@@
    The schema is designed to represent a list of the objects (e.g., `Type Ia supernova') and phenomena (e.g., `gravitational lensing'), as well as potential downstream science uses cases (e.g., `set constraints on supernova explosion models') that could correspond to the eventual imaged observation given the abstract text, with a minimum of 1 and a maximum of 5 elements per list.
-
+   %
    The procedure guides the generation of LLM outputs while ensuring that the schema is respected at every step in the generation process by masking out tokens that would violate the intended format.
    %
-By framing the problem in terms of transitions between a set of finite states (a finite-state machine), \cite{willard2023efficient} showed that guided generation can be performed with negligible overhead compared to unconstrained generation.
+   By framing the problem in terms of transitions between a set of finite states (i.e., a finite-state machine), \cite{willard2023efficient} showed that guided generation can be performed with negligible overhead compared to unconstrained generation.
    %
-See App.~\ref{app:guided-generation} for a more detailed description of the guidance generation method used here.
-
+   See App.~\ref{app:guided-generation} for a more detailed description of the guidance generation method used here, including an overview of technical details.
+   %
    While the schema-guided generation ensures the \emph{format} of the output, the prompt and choice of LLM will dictate the \emph{content} of the generated summaries.
@@
 
-The guided generation process ensures that, in this case, the output of the LLM strictly conforms to the format of the following example:
-\begin{center}
-  \begin{jsoncode}
-    \centering
-          \color{black}\{
-            \color{codegreen}'objects_and_phenomena'\color{black}: [\color{codegreen}'star forming galaxy', 'lensed galaxy'\color{black}], 
-            \color{codegreen}'science_use_cases'\color{black}: [\color{codegreen}'measure lensing magnification'\color{black}]
-          \color{black}\}
-  \end{jsoncode}
-  \end{center}
-which is then used to construct the summarized caption by combining the two key elements.
+  %  The guided generation process ensures that, in this case, the output of the generated output of the LLM strictly conforms to the format of the following example:\\
+  %  \begin{center}
+  %    \begin{jsoncode}
+  %      \centering
+  %            \color{black}\{
+  %              \color{deepgreen}'objects_and_phenomena'\color{black}: [\color{deepgreen}'star forming galaxy', 'lensed galaxy'\color{black}, ...], 
+  %              \color{deepgreen}'science_use_cases'\color{black}: [\color{deepgreen}'measure lensing magnification'\color{black}, ...]
+  %            \color{black}\}
+  %    \end{jsoncode}
+  %    \end{center}
+  %  which is then used to construct the summarized caption by combining the two key elements.
    %
-Examples of raw abstract snippets and corresponding LLM-generated summaries are shown in Tab.~\ref{tab:datasetsumm}.
+   Examples of LLM-generated abstract summaries are shown in Tab.~\ref{tab:datasetsumm}, for the same set of abstracts as shown in Tab.~\ref{tab:dataset}.
    %
@@
 
-Our goal is to learn a semantically meaningful joint representation between images corresponding to HST observation and natural (English) language.
-%
-With PAPERCLIP, we leverage the strong generalization capabilities demonstrated by pre-trained CLIP models and adapt these to work with domain-specific \hubble data via fine tuning.
+  %  Our goal is to learn a semantically meaningful joint representation between images corresponding to HST observation and natural (English) language.
+  %  %
+  %  With PAPERCLIP, we leverage the strong generalization capabilities demonstrated by pre-trained CLIP models and adapt these to work with domain-specific \hubble data via fine-tuning.
 
@@
    %
-We use the standard softmax-based bidirectional variant of the InfoNCE~\citep{oord2018representation} contrastive loss function introduced for training CLIP-style architectures \citep{radford2021learning}
+   We use the standard bidirectional variant of the InfoNCE~\citep{oord2018representation} contrastive loss function introduced for training CLIP-style architectures \citep{radford2021learning},
+  %  %
+  %  \begin{equation}
+  %    \label{eq:softmax_loss}
+  %    \mathcal{L}(\mathcal{B})=-\frac{1}{2|\mathcal{B}|} \sum_{i=1}^{|\mathcal{B}|}\left(\log \frac{e^{x_i \cdot y_i / \tau}}{\sum_{j=1}^{|\mathcal{B}|} e^{x_i \cdot y_j / \tau}}+\log \frac{e^{x_i \cdot y_i / \tau}}{\sum_{j=1}^{|\mathcal{B}|} e^{x_j \cdot y_i / \tau}}\right)
+  %  \end{equation}
+   $\mathcal{L}(\mathcal{B})=-\frac{1}{2|\mathcal{B}|} \sum_{i=1}^{|\mathcal{B}|}\left(\log \frac{e^{x_i \cdot y_i / \tau}}{\sum_{j=1}^{|\mathcal{B}|} e^{x_i \cdot y_j / \tau}}+\log \frac{e^{x_i \cdot y_i / \tau}}{\sum_{j=1}^{|\mathcal{B}|} e^{x_j \cdot y_i / \tau}}\right)$,
    %
-\begin{equation}
-  \label{eq:softmax_loss}
-  \mathcal{L}(\mathcal{B})=-\frac{1}{2|\mathcal{B}|} \sum_{i=1}^{|\mathcal{B}|}\left(\log \frac{e^{x_i \cdot y_i / \tau}}{\sum_{j=1}^{|\mathcal{B}|} e^{x_i \cdot y_j / \tau}}+\log \frac{e^{x_i \cdot y_i / \tau}}{\sum_{j=1}^{|\mathcal{B}|} e^{x_j \cdot y_i / \tau}}\right)
-\end{equation}
-%
-where ${x}_i={f\left(I_i\right)}/{\left\|f\left(I_i\right)\right\|}$ and ${y}_i={g\left(T_i\right)}/{\left\|g\left(T_i\right)\right\|}$ are the normalized representations of the $i$-th image and text, respectively, and $\tau$ is a learnable temperature hyperparameter.
+   where ${x}_i={f\left(I_i\right)}/{\left\|f\left(I_i\right)\right\|}$ and ${y}_i={g\left(T_i\right)}/{\left\|g\left(T_i\right)\right\|}$ are the normalized representations of the $i$-th image and text caption, respectively, and $\tau$ is a learnable temperature hyperparameter.
    %
@@
 
-We use the \texttt{CLIP-ViT-B/16} \citep{radford2021learning} variant as the base pre-trained CLIP model.
-%
-This model uses a 12-layer, 12-head, 768-embedding dimension vision transformer as the image encoder and a 12-layer, 8-head, 512-embedding dimension text sequence transformer as the text backbone.
-%
-The text encoder has a maximum length of 77 tokens and the image encoder a native resolution of $224\times224$ pixels.
-%
-Linear projection layers map the outputs of the image and text encoders to a common embedding space of dimension $n_\text{emb}=512$.
-%
-In total, the model contains $\sim 149$ million trainable parameters.
-%
-This model was originally pre-trained on 400 million image-text pairs from internet data \citep{radford2021learning}.
-%
+  %  We use the \texttt{CLIP-ViT-B/16} \citep{radford2021learning} variant as the base pre-trained CLIP model.
+  %  %
+  %  This model uses a 12-layer, 12-head, 768-embedding dimension vision transformer with patch size $16\times16$ as the image encoder \citep{dosovitskiy2020image} and a 12-layer, 8-head, 512-embedding dimension text sequence transformer as the text backbone \citep{vaswani2017attention}.
+  %  %
+  %  The text encoder has a maximum length of 77 tokens and the image encoder has a native resolution of $224\times224$ pixels.
+  %  %
+  %  Linear projection layers map the outputs of the image and text encoders to a common embedding space of dimension $n_\text{emb}=512$.
+  %  %
+  %  In total, the model contains $\sim 149$ million trainable parameters.
+  %  %
+  %  This model was originally pre-trained on $\sim 400$ million image-text pairs from internet data.
+  %  %
 
@@
 
+   We use the \texttt{CLIP-ViT-B/16} \citep{radford2021learning} variant as the base pre-trained CLIP model.
+   %
   We explore three different methods of training the model on our domain dataset: \emph{(1)} Fine-tuning the entire network starting from the pre-trained base model; \emph{(2)} Freezing the base image/text encoders and training a small projection head; and \emph{(3)} Training the entire model from scratch.
@@
    For \emph{(2)}, we use a 2-layer MLP with 1024 hidden units and a GELU activation layer, projecting onto the 512-dimensional common embedding space.
-
-All models were trained over 20,000 steps with 2000 linear warmup steps 
-% and cosine decay 
-using the AdamW optimizer \citep{DBLP:conf/iclr/LoshchilovH19,DBLP:journals/corr/KingmaB14} with  %peak 
-learning rate $10^{-5}$ and weight decay $10^{-3}$.
-%
-Training takes approximately 3 hours on 4 Nvidia A100 GPUs.
-Models were instantiated using the \package{Transformers} \citep{wolf2019huggingface} library and trained using packages from the \package{Jax} \citep{jax2018github} ecosystem.
-%
+Additional details on the CLIP model and fine-tuning procedure are provided in App.~\ref{app:model_details}.
 
@@
 
-The model is evaluated by tracking the contrastive loss in Eq.~\eqrefb{eq:softmax_loss} as well as the top-$k\%$ retrieval accuracy on the held out validation set over the course of training.
+   The model is evaluated by tracking the contrastive loss as well as the top-$k\%$ retrieval accuracy on the held out validation set over the course of training.
    %
-The retrieval accuracy is defined as the fraction of associated captions which fall within the top $k\%$ of captions by cosine similarity of the normalized image and caption embeddings, averaged over the images in the validation set:
-\begin{equation}
-\text{Retrieval accuracy}_k = \frac{1}{|\mathcal V|} \sum_{i=1}^{|\mathcal V|} \mathbbm{1}\left[\operatorname{rank}\left({x}_i \cdot {y}_{i}; \{{x}_i \cdot {y}_{j}\}_{j=1}^{|\mathcal V|}\right) \leq \left\lfloor\frac{k}{100}|\mathcal V|\right\rfloor\right]
-\label{eq:retrieval_accuracy}
-\end{equation}
+   The retrieval accuracy is defined as the fraction of associated captions (either raw or LLM-summarized abstracts) which fall within the top $k\%$ of captions by cosine similarity of the normalized image and caption embeddings, averaged over the images in the validation set:
+   $\frac{1}{|\mathcal V|} \sum_{i=1}^{|\mathcal V|} \mathbbm{1}\left[\operatorname{rank}\left({x}_i \cdot {y}_{i}; \{{x}_i \cdot {y}_{j}\}_{j=1}^{|\mathcal V|}\right) \leq \left\lfloor\frac{k}{100}|\mathcal V|\right\rfloor\right]$
+  %  \begin{equation}
+  %   \text{Retrieval accuracy}_k = \frac{1}{|\mathcal V|} \sum_{i=1}^{|\mathcal V|} \mathbbm{1}\left[\operatorname{rank}\left({x}_i \cdot {y}_{i}; \{{x}_i \cdot {y}_{j}\}_{j=1}^{|\mathcal V|}\right) \leq \left\lfloor\frac{k}{100}|\mathcal V|\right\rfloor\right]
+  %   \label{eq:retrieval_accuracy}
+  %   \end{equation} 
    where $|\mathcal V|$ is the total number of images in the validation set, $\mathbbm{1}[\cdot]$ is the indicator function that returns 1 if the condition inside the brackets is true and 0 otherwise, $\operatorname{rank}\left({x}_i \cdot {y}_{i}; \{{x}_i \cdot {y}_{j}\}_{j=1}^{|\mathcal V|}\right)$ is a function that returns the rank of the cosine similarity between ${x}_i$ and ${y}_{i}$ among the cosine similarities between ${x}_i$ and all captions ${y}_j$ in the validation set, and $k$ is the percentage of top captions considered for the retrieval accuracy.
@@
 
-We also qualitatively evaluate the learned embeddings through image retrieval (i.e., retrieving the most relevant images from the validation set using natural language queries) and description retrieval (i.e., querying the astrophysical object classes and science use cases most relevant to a given observation, akin to zero-shot classification) experiments.
+   We also qualitatively evaluate the learned embeddings through image retrieval (i.e., retrieving the most relevant images from the validation set using natural language queries).
+% and description retrieval (i.e., querying the astrophysical object classes and science use cases most relevant to a given observation, akin to zero-shot classification) experiments.
    %
-For the description/text retrieval evaluation, we define a list of possible text associations (i.e., classes), which we show in App.~\ref{app:categories}, by querying the \textsc{Claude 2}\footnote{\url{https://claude.ai/}} large language followed by with manual curation.
+  %  For the description/text retrieval evaluation, we define a list of possible text associations (i.e., classes), which we show in App.~\ref{app:categories}, by querying the \textsc{Claude 2}\footnote{\url{https://claude.ai/}} large language followed by manual curation.
 
@@
    %
-Interestingly, the base pre-trained model performs better than random expectation, with a top-10\% retrieval accuracy of $\sim 15\%$.
+   Interestingly, the base pre-trained model performs better than random expectation, with a top-10\% retrieval accuracy of $\sim 15\%$ (as see from the left-most datum in Fig.~\ref{fig:retrieval_acc} right, for the curves corresponding to fine-tuned models).
    %
@@
    \begin{figure*}[!h]
-\includegraphics[width=0.99\textwidth]{plots/val_metrics.pdf}
-\caption{The CLIP contrastive loss from Eq.~\eqrefb{eq:softmax_loss} (left) and the top-10\% retrieval accuracy from Eq.~\eqrefb{eq:retrieval_accuracy} (right) computed on the validation set over the course of training.
+   \centering
+   \includegraphics[width=0.85\textwidth]{plots/val_metrics.pdf}
+   \caption{The CLIP contrastive loss (left) and the top-10\% retrieval accuracy (right) computed on the validation set over the course of training.
 Shown for the dataset with summarized abstracts as captions (red), dataset using raw proposal abstracts as captions (blue), only fine-tuning a small MLP head (dotted green), training from scratch with summarized abstracts as captions (yellow), and trained with shuffled image-text pairs (dashed orange).}
@@
 
-Figure~\ref{fig:sim_valtrain} (right) shows the retrieval accuracy, as defined in Eq.~\eqrefb{eq:retrieval_accuracy}, as a function of the retrieval fraction $k\%$.
+   Figure~\ref{fig:sim_valtrain} (right) shows the retrieval accuracy as a function of the retrieval fraction $k\%$.
    %
-In this case, we evaluate all four models (fine tuned on raw abstracts (blue), fine-tuned on LLM-summarized abstracts (red), trained on LLM-summarized abstracts from scratch (yellow), and the base model (purple)) on the same captions dataset -- the summarized abstracts -- for a direct comparison.
+   In this case, we evaluate all four models (fine-tuned on raw abstracts (blue), fine-tuned on LLM-summarized abstracts (red), trained on LLM-summarized abstracts from scratch (yellow), and the base model (purple)) on the same captions dataset -- the summarized abstracts -- for a direct comparison.
    %
    Remarkably, the model trained on raw abstracts shows very similar performance when evaluated on the summarized abstracts compared to that trained on the summarized abstracts themselves, indicating that \emph{(1)} the image-text association signal is preserved in the summarization process, and \emph{(2)} the model is able to effectively leverage meaningful concepts in the noisy raw abstracts through weak supervision.
-\changes{The significantly worse performance of the model trained from scratch, compared to the fine-tuned models, highlights the crucial role of the inductive bias inherent in the base pretrained model, which effectively captures the rich associations between images and language.}
+The significantly worse performance of the model trained from scratch, compared to the fine-tuned models, highlights the crucial role of the inductive bias inherited from the base pre-trained model, which effectively captures rich associations between images and language.
 
    \begin{figure*}[!h]
-  \includegraphics[width=0.49\textwidth]{plots/sim_val.pdf}
-  \includegraphics[width=0.49\textwidth]{plots/retrieval.pdf}
+    \centering
+     \includegraphics[width=0.42\textwidth]{plots/sim_val.pdf}
+     \includegraphics[width=0.42\textwidth]{plots/retrieval.pdf}
      \caption{(Left) Distribution of cosine similarities between corresponding image and text embeddings, $x_i$ and $y_i$, shown when using the base CLIP model (purple lines), and the summary fine-tuned CLIP model (red line).
@@
 
-\changes{We show retrieval accuracy performance for additional variations on the model and training configuration in App.~\ref{app:ablations}.}
+   We show retrieval accuracy performance for additional variations on the model and training configuration in App.~\ref{app:ablations}.
 
@@
    Having aligned the image and text representations, we can embed a natural language query using the model and show the closest images by embedding from the validation set when ranked by cosine similarity.
+A sketch of this procedure is shown in Fig.~\ref{fig:overview} (right).
    %
-We show these in Tabs.~\ref{tab:tti_base} and \ref{tab:tti} for the base and fine-tuned models respectively using four simple curated queries: \texttt{dwarf galaxy} (small galaxies that typically orbit larger galaxies like the Milky Way), \texttt{Jupiter},  \texttt{SN1987A} (a specific, prominent supernova), and \texttt{strong lensing} (the phenomenon of bending of light due to the gravitational influence of a foreground distribution of matter).
+   We show these in Tabs.~\ref{tab:tti_base} and \ref{tab:tti} for the base and fine-tuned models respectively using two simple curated queries: \texttt{Jupiter} and  \texttt{SN1987A} (a specific supernova).
 The proposal ID corresponding to the retrieved images is shown below each image, and contains a hyperlink to the MAST page corresponding to the proposal for further details.
 
-While the base model shows some signs of meaningful retrieval (e.g., the image of Jupiter in the second row of Tab.~\ref{tab:tti_base}, and images of galaxies in first row), it is challenging to discern meaningful, strong associations between the retrieved images and corresponding query.
+   While the base model shows some signs of meaningful retrieval (e.g., the image of Jupiter in the first row of Tab.~\ref{tab:tti_base}), it is challenging to discern meaningful, strong associations between the retrieved images and corresponding query.
 
@@
    %
-The \texttt{dwarf galaxy}-queried images correspond to proposals aiming to measure the kinematics of the stellar cores of dwarf galaxies.
+  %  The \texttt{dwarf galaxy}-queried images correspond to proposals aiming to measure the kinematics of the stellar cores of dwarf galaxies.
+  %  %
+   Images looking like Jupiter are returned for the \texttt{Jupiter} query. 
    %
-Images of Jupiter are returned for the \texttt{Jupiter} query, with the planet clearly visible in all four images.
+   However, this example also illustrates the model's potential to misidentify objects, with the first and third image actually showing Saturn with artifacts on the planet and partially obscured rings.
    %
-Supernova SN1987 itself can be seen in the three closest images for the \texttt{SN1987A} query with the fourth image being a supernova remnant.
-%
-Cluster-scale as well as galaxy-scale gravitational lenses are returned by the \texttt{strong lensing} query, with lensing patterns visible in the images.
+   Supernova SN1987 itself can be seen in the three closest images for the \texttt{SN1987A} query.  % with the fourth image being a supernova remnant.
+  %  %
+  %  Cluster-scale as well as galaxy-scale gravitational lenses are returned by the \texttt{strong lensing} query, with lensing patterns visible in the images.
 
-\subsection{Text Retrieval}
+   Note that we chose to illustrate qualitative performance on image retrieval using the model fine-tuned on summarized abstracts, rather than raw abstracts. 
+  %  We show analogous results for the model fine-tuned on raw abstracts in App.~\ref{app:eval_raw}. 
+   Although the two models show very similar quantitative performance on retrieval metrics (as shown in Fig.~\ref{fig:sim_valtrain}), they exhibit characteristically different behaviors in terms of images retrieved, with the summary fine-tuned models generally retrieving images that look more visually ``relevant'' to a domain expert.
+We emphasize that for scientific usefulness, the goal is not necessarily to correctly retrieve the most relevant objects, but rather to identify a diverse set of interesting candidates for manual follow-up and further analysis.
+% ; both models are seen to perform sensibly, even if differently, in this regard.
 
-We can use images from the validation set as queries and retrieve the most relevant text chunks (e.g., objects and use cases) from a curated list as described in Sec.~\ref{sec:eval}.
-%
-We show the result of image-to-text retrieval in Tab.~\ref{tab:itt}, for the base (second column) as well as summary fine-tuned (third column) models, using four observations (left-most column) from the validation set.
-%
-
-The top four text associations are shown for each image query.
-%
-The `ground truth' summarized abstract is shown in the right column.
-% 
-The base as well as fine-tuned models are seen to return a mix of relevant and less-relevant associations, although showing different qualitative behavior.
-Purely qualitatively, the fine-tuned model is seen to consistently return more relevant associations compared to the base model.
-%
-
-The second row (an image of supernova 1987A) highlights an interesting pattern -- the base model erroneously attributes the object at the center of the image to a gravitational lens, while the fine-tuned model correctly identifies it as a supernova remnant.
-This kind of reasonable misattribution is common when querying the base model, and largely absent in the fine-tuned model.
-
-Note that we chose to illustrate qualitative performance on text and image retrieval using the model fine-tuned on summarized abstracts, rather than raw abstracts.
-We show analogous results for the model fine-tuned on raw abstracts in App.~\ref{app:eval_raw}.
-Although the two models show very similar quantitative performance on retrieval metrics (as shown in Fig.~\ref{fig:sim_valtrain}), they exhibit characteristically different behaviors in terms of objects (images/text) retrieved.
-We emphasize that for scientific usefulness, the goal is not necessarily to correctly retrieve the most ``relevant'' objects, but rather to identify a diverse set of interesting candidates for manual follow-up and further analysis; both models are seen to perform sensibly, even if differently, in this regard.
+  The fine-tuned model can similarly be used for description/text retrieval, akin to the traditional zero-shot classification setting, where the closest text snippets from a curated list are returned given an observed astronomical image.
+We show examples of the text retrieval task in App.~\ref{app:text_retrieval}.
 
@@
      \centering
-  \begin{tabular}{m{2.7cm} p{2.9cm} p{2.9cm} p{2.9cm} p{2.9cm}}
+     \begin{tabular}{m{2.7cm} p{2.9cm} p{2.9cm} p{2.9cm}}
          \toprule
-      \centering \bfseries Query & \multicolumn{4}{c}{\bfseries{Top-4 most similar images using \textcolor{deeppurple}{base CLIP model}}} \tabularnewline
+         \centering \bfseries Query & \multicolumn{3}{c}{\bfseries{Top-3 most similar images using \textcolor{deeppurple}{base off-the-shelf CLIP model}}} \tabularnewline
+        %  \midrule
+        %  \texttt{\input{\thedatafolder/query_tti_base_1.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_1_0.pdf} \\ \input{\thedatafolder/propid_tti_base_1_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_1_1.pdf} \\ \input{\thedatafolder/propid_tti_base_1_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_1_2.pdf} \\ \input{\thedatafolder/propid_tti_base_1_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_1_3.pdf} \\ \input{\thedatafolder/propid_tti_base_1_3.txt}  \tabularnewline
          \midrule
-      \texttt{\input{\thedatafolder/query_tti_base_1.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_1_0.pdf} \\ \input{\thedatafolder/propid_tti_base_1_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_1_1.pdf} \\ \input{\thedatafolder/propid_tti_base_1_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_1_2.pdf} \\ \input{\thedatafolder/propid_tti_base_1_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_1_3.pdf} \\ \input{\thedatafolder/propid_tti_base_1_3.txt}  \tabularnewline
+          \texttt{\input{\thedatafolder/query_tti_base_0.txt}} \vspace{20mm} & \centering \includegraphics[width=0.16\textwidth]{\thedatafolder/img_tti_base_0_0.pdf} \\ \input{\thedatafolder/propid_tti_base_0_0.txt} & \centering \includegraphics[width=0.16\textwidth]{\thedatafolder/img_tti_base_0_1.pdf} \\ \input{\thedatafolder/propid_tti_base_0_1.txt} & \centering \includegraphics[width=0.16\textwidth]{\thedatafolder/img_tti_base_0_2.pdf} \\ \input{\thedatafolder/propid_tti_base_0_2.txt} \tabularnewline
          \midrule
-       \texttt{\input{\thedatafolder/query_tti_base_0.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_0_0.pdf} \\ \input{\thedatafolder/propid_tti_base_0_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_0_1.pdf} \\ \input{\thedatafolder/propid_tti_base_0_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_0_2.pdf} \\ \input{\thedatafolder/propid_tti_base_0_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_0_3.pdf} \\ \input{\thedatafolder/propid_tti_base_0_3.txt}  \tabularnewline
-      \midrule
-      \texttt{\input{\thedatafolder/query_tti_base_2.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_2_0.pdf} \\ \input{\thedatafolder/propid_tti_base_2_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_2_1.pdf} \\ \input{\thedatafolder/propid_tti_base_2_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_2_2.pdf} \\ \input{\thedatafolder/propid_tti_base_2_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_2_3.pdf} \\ \input{\thedatafolder/propid_tti_base_2_3.txt}  \tabularnewline
-      \midrule
-      \texttt{\input{\thedatafolder/query_tti_base_3.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_3_0.pdf} \\ \input{\thedatafolder/propid_tti_base_3_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_3_1.pdf} \\ \input{\thedatafolder/propid_tti_base_3_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_3_2.pdf} \\ \input{\thedatafolder/propid_tti_base_3_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_3_3.pdf} \\ \input{\thedatafolder/propid_tti_base_3_3.txt}  \tabularnewline
+         \texttt{\input{\thedatafolder/query_tti_base_2.txt}} \vspace{20mm} & \centering \includegraphics[width=0.16\textwidth]{\thedatafolder/img_tti_base_2_0.pdf} \\ \input{\thedatafolder/propid_tti_base_2_0.txt} & \centering \includegraphics[width=0.16\textwidth]{\thedatafolder/img_tti_base_2_1.pdf} \\ \input{\thedatafolder/propid_tti_base_2_1.txt} & \centering \includegraphics[width=0.16\textwidth]{\thedatafolder/img_tti_base_2_2.pdf} \\ \input{\thedatafolder/propid_tti_base_2_2.txt} \tabularnewline
+        %  \midrule
+        %  \texttt{\input{\thedatafolder/query_tti_base_3.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_3_0.pdf} \\ \input{\thedatafolder/propid_tti_base_3_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_3_1.pdf} \\ \input{\thedatafolder/propid_tti_base_3_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_3_2.pdf} \\ \input{\thedatafolder/propid_tti_base_3_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_3_3.pdf} \\ \input{\thedatafolder/propid_tti_base_3_3.txt}  \tabularnewline
          \bottomrule
      \end{tabular}
-  \caption{For four text queries (left-most column), the four most similar images from the validation dataset by cosine similarity when using the \textbf{\textcolor{deeppurple}{base CLIP model}} (\texttt{CLIP-ViT-B/16}).
+     \caption{For two text queries (left-most column), the three most similar images from the validation dataset by cosine similarity when using the \textbf{\textcolor{deeppurple}{base (off-the-shelf) CLIP model}} (\texttt{CLIP-ViT-B/16}).
 The proposal ID associated with each image is given below the image and contains a hyperlink to the MAST page corresponding to the proposal.}
@@
      \centering
-  \begin{tabular}{m{2.7cm} p{2.9cm} p{2.9cm} p{2.9cm} p{2.9cm}}
+     \begin{tabular}{m{2.7cm} p{2.9cm} p{2.9cm} p{2.9cm}}
          \toprule
-      \centering \bfseries Query & \multicolumn{4}{c}{\bfseries{Top-4 most similar images using \textcolor{deepred}{summary fine-tuned CLIP model}}} \tabularnewline
+         \centering \bfseries Query & \multicolumn{3}{c}{\bfseries{Top-3 most similar images using \textcolor{deepred}{summary fine-tuned CLIP model}}} \tabularnewline
          \midrule
-      \texttt{\input{\thedatafolder/query_tti_1.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_1_0.pdf} \\ \input{\thedatafolder/propid_tti_1_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_1_1.pdf} \\ \input{\thedatafolder/propid_tti_1_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_1_2.pdf} \\ \input{\thedatafolder/propid_tti_1_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_1_3.pdf} \\ \input{\thedatafolder/propid_tti_1_3.txt}  \tabularnewline
+        %  \texttt{\input{\thedatafolder/query_tti_1.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_1_0.pdf} \\ \input{\thedatafolder/propid_tti_1_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_1_1.pdf} \\ \input{\thedatafolder/propid_tti_1_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_1_2.pdf} \\ \input{\thedatafolder/propid_tti_1_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_1_3.pdf} \\ \input{\thedatafolder/propid_tti_1_3.txt}  \tabularnewline
+        %  \midrule
+          \texttt{\input{\thedatafolder/query_tti_0.txt}} \vspace{20mm} & \centering \includegraphics[width=0.16\textwidth]{\thedatafolder/img_tti_0_0.pdf} \\ \input{\thedatafolder/propid_tti_0_0.txt} & \centering \includegraphics[width=0.16\textwidth]{\thedatafolder/img_tti_0_1.pdf} \\ \input{\thedatafolder/propid_tti_0_1.txt} & \centering \includegraphics[width=0.16\textwidth]{\thedatafolder/img_tti_0_2.pdf} \\ \input{\thedatafolder/propid_tti_0_2.txt}  \tabularnewline
          \midrule
-       \texttt{\input{\thedatafolder/query_tti_0.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_0_0.pdf} \\ \input{\thedatafolder/propid_tti_0_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_0_1.pdf} \\ \input{\thedatafolder/propid_tti_0_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_0_2.pdf} \\ \input{\thedatafolder/propid_tti_0_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_0_3.pdf} \\ \input{\thedatafolder/propid_tti_0_3.txt}  \tabularnewline
-      \midrule
-      \texttt{\input{\thedatafolder/query_tti_2.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_2_0.pdf} \\ \input{\thedatafolder/propid_tti_2_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_2_1.pdf} \\ \input{\thedatafolder/propid_tti_2_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_2_2.pdf} \\ \input{\thedatafolder/propid_tti_2_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_2_3.pdf} \\ \input{\thedatafolder/propid_tti_2_3.txt}  \tabularnewline
-      \midrule
-      \texttt{\input{\thedatafolder/query_tti_3.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_3_0.pdf} \\ \input{\thedatafolder/propid_tti_3_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_3_1.pdf} \\ \input{\thedatafolder/propid_tti_3_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_3_2.pdf} \\ \input{\thedatafolder/propid_tti_3_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_3_3.pdf} \\ \input{\thedatafolder/propid_tti_3_3.txt}  \tabularnewline
+         \texttt{\input{\thedatafolder/query_tti_2.txt}} \vspace{20mm} & \centering \includegraphics[width=0.16\textwidth]{\thedatafolder/img_tti_2_0.pdf} \\ \input{\thedatafolder/propid_tti_2_0.txt} & \centering \includegraphics[width=0.16\textwidth]{\thedatafolder/img_tti_2_1.pdf} \\ \input{\thedatafolder/propid_tti_2_1.txt} & \centering \includegraphics[width=0.16\textwidth]{\thedatafolder/img_tti_2_2.pdf} \\ \input{\thedatafolder/propid_tti_2_2.txt}  \tabularnewline
+        %  \midrule
+        %  \texttt{\input{\thedatafolder/query_tti_3.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_3_0.pdf} \\ \input{\thedatafolder/propid_tti_3_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_3_1.pdf} \\ \input{\thedatafolder/propid_tti_3_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_3_2.pdf} \\ \input{\thedatafolder/propid_tti_3_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_3_3.pdf} \\ \input{\thedatafolder/propid_tti_3_3.txt}  \tabularnewline
          \bottomrule
@@
 
-\begin{table}[h!]
-  \centering
-  \renewcommand{\arraystretch}{0.1}
-  \begin{tabular}{m{3cm} m{3.6cm} m{3.6cm} m{4cm}}
-      \toprule
-      \centering \bfseries \hubble image & \centering \textbf{Top-4 text} \\ \small\textbf{\textcolor{deeppurple}{(base)}} & \centering  \textbf{Top-4 text} \\ \small\textbf{\textcolor{deepred}{(summary fine-tuned)}} & \centering \textbf{Summarized abstract} \\ (objects; `ground truth') \tabularnewline
-      \midrule
-      \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_0.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_0.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_0.txt} &  {\scriptsize \input{\thedatafolder/abs_itt_0.txt}} \tabularnewline
-      \midrule
-      \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_1.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_1.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_1.txt} &  {\scriptsize \input{\thedatafolder/abs_itt_1.txt}} \tabularnewline
-      \midrule
-      \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_2.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_2.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_2.txt} &  {\scriptsize \input{\thedatafolder/abs_itt_2.txt}} \tabularnewline
-      \midrule
-      \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_3.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_3.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_3.txt} &  {\scriptsize \input{\thedatafolder/abs_itt_3.txt}} \tabularnewline
-      \bottomrule
-  \end{tabular}
-  \caption{Text snippets from a curated list most closely matching a given image query (left-most column) by cosine similarity of respective embeddings, shown for the \textcolor{deeppurple}{base} (\texttt{CLIP-ViT-B/16}) and \textcolor{deepred}{summary fine-tuned} models.
-The `ground truth' LLM-summarized abstract (only objects/phenomena) is shown in the right-most column.}
-  \label{tab:itt}
-\end{table}
-
    \section{Outlook and Conclusions}
@@
 
-In this paper, we present \textsc{PAPERCLIP}, a method for training domain-specific multi-modal models for astrophysics that associates observations imaged by telescopes with natural language in a common, semantically-meaningful embedding space.
+   We present \textsc{PAPERCLIP}, a method for training domain-specific multi-modal models for astrophysics that associates observations imaged by telescopes with natural language in a common embedding space.
    %
@@
    %
-Overall, the procedure demonstrates the efficacy of fine-tuning generalist pre-trained models on small amounts of domain-specific data, in particular astronomical datasets, and leveraging text as an interface.
+   Overall, the procedure demonstrates the efficacy of fine-tuning generalist pre-trained models on small amounts of domain-specific data, in particular astronomical datasets, and leveraging text as an interface for interacting with the data.
 
@@
    %
-A model trained using weakly-supervised image-text pairs can be used to query survey data e.g., PHANGS~\citep{lee2022phangs}, COSMOS~\citep{scoville2007cosmic} using natural language, as well as to efficiently find patterns in such data that may not be apparent using specialized models or manual inspection.
+   A model trained using weakly-supervised image-text pairs can be used to query large amounts of unlabeled survey data e.g., PHANGS~\citep{lee2022phangs}, COSMOS~\citep{scoville2007cosmic} for objects or use-cases of interest using natural language, as well as to efficiently find patterns in such data that may not be apparent using specialized models or manual inspection.
    %
@@
 
-\subsubsection*{Code and Data Availability}
+% \subsubsection*{Author Contributions}
+% If you'd like to, you may include  a section for author contributions as is done
+% in many journals. This is optional and at the discretion of the authors.
 
-The code, dataset, and fine-tuned models used in this work are available at \url{https://www.github.com/smsharma/PAPERCLIP-Hubble}.
-
-\subsubsection*{Software}
+\subsubsection*{Acknowledgments}
+% Use unnumbered third level headings for the acknowledgments. All
+% acknowledgments, including those to funding agencies, go at the end of the paper.
 
@@
 
-% Code repo and package cites.
-% Dataset availability
+This research is based on observations made with the NASA/ESA Hubble Space Telescope obtained from the Space Telescope Science Institute, which is operated by the Association of Universities for Research in Astronomy, Inc., under NASA contract NAS 5-26555.
+%
+Based on observations made with the NASA/ESA Hubble Space Telescope, and obtained from the Hubble Legacy Archive, which is a collaboration between the Space Telescope Science Institute (STScI/NASA), the Space Telescope European Coordinating Facility (ST-ECF/ESAC/ESA) and the Canadian Astronomy Data Centre (CADC/NRC/CSA).
 
-\subsubsection*{Broader Impact}
-
+\section{Ethics Statement}
 This work relies on using abstracts from successful \hubble Space Telescope observing proposals as part of a dataset for training and evaluating machine learning models.
 While these abstracts are publicly available, the authors likely did not anticipate their text being used in this manner, raising questions around consent, attribution, and appropriate use of data.
-Since this research intends to develop methods to aid astronomical research and does not use sensitive personal information or target commercial gain, we believe that the scientific benefits outweigh the potential concerns in this case while acknowledging good-faith arguments to the contrary.
+Since this research intends to develop methods to aid astronomical research and does not use sensitive personal information or target commercial gain, we believe that the scientific benefits outweigh the potential concerns in this case, while acknowledging good-faith arguments to the contrary.
 As the use of foundation models in the sciences increases, it will be important for the community to consider norms and guidelines around the appropriate use and attribution of various data sources for model training and evaluation, including qualitative textual data, to ensure transparency and maintain trust.
 
-\subsubsection*{Acknowledgments}
-
-We thank Michael Brenner for helpful conversations.
-%
-This work is supported by the National Science Foundation under Cooperative Agreement PHY-2019786 (The NSF AI Institute for Artificial Intelligence and Fundamental Interactions, \url{http://iaifi.org/}).
-%
-This material is based upon work supported by the U.S. Department of Energy, Office of Science, Office of High Energy Physics of U.S. Department of Energy under grant Contract Number  DE-SC0012567. 
-%
-YS was supported by the Research Science Institute (RSI) program at MIT.
-%
-This research was supported by an award from Google,  ``Interpretation of Multimodal Images from Astronomy''.
-%
-This research was supported by the Munich Institute for Astro-, Particle and BioPhysics (MIAPbP), which is funded by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) under Germany's Excellence Strategy – EXC-2094 – 390783311.
-%
-The computations in this paper were run on the FASRC Cannon cluster supported by the FAS Division of Science Research Computing Group at Harvard University.
-
-This research is based on observations made with the NASA/ESA Hubble Space Telescope obtained from the Space Telescope Science Institute, which is operated by the Association of Universities for Research in Astronomy, Inc., under NASA contract NAS 5-26555.
-
-Based on observations made with the NASA/ESA Hubble Space Telescope, and obtained from the Hubble Legacy Archive, which is a collaboration between the Space Telescope Science Institute (STScI/NASA), the Space Telescope European Coordinating Facility (ST-ECF/ESAC/ESA) and the Canadian Astronomy Data Centre (CADC/NRC/CSA).
-
 \bibliography{hubble_paperclip}
-\bibliographystyle{tmlr}
+\bibliographystyle{colm2024_conference}
 
@@
 \section{Details on the Abstract Summarization Procedure}
+
+We provide additional details of the abstract summarization procedure, including a brief overview of the guided generation method used, as well as the prompts and schema used for the summarization task.
 
@@
 
-Formally, let $\mathcal{M} = (Q, \Sigma, \delta, q_0, F)$ be the FSM representing the regular expression, where $Q$ is the set of states, $\Sigma$ is the alphabet of the regular expression, $\delta: Q \times \Sigma \rightarrow Q$ is the transition function between states, $q_0$ is the start state, and $F$ is the set of accept states which terminate the generation.
+Formally, let $\mathcal{M} = (Q, \Sigma, \delta, q_0, F)$ be the FSM representing the regular expression, where $Q$ is the set of states, $\Sigma$ is the alphabet of the regular expression, $\delta: Q \times \Sigma \rightarrow Q$ is the transition function between states, $q_0$ is the start state, and $F\subseteq Q$ is the set of accept states which terminate the generation.
 An index $\sigma: Q \rightarrow \mathcal{P}(V)$ is first constructed, where $V$ is the LLM's token vocabulary and $\mathcal{P}(V)$ denotes the power set of $V$.
@@
 
-During the token-by-token generation process, we keep track of the current FSM state $q_t$ after sampling each token $v_t$.
-At each step $t$, we mask the LLM's output logits based on the valid next tokens $\sigma(q_t)$, setting the logits of invalid tokens to $-\infty$.
+During the sequential generation process, the current FSM state $q_t$ is kept track of after sampling each token $v_t$.
+At each step $t$, the LLM's output logits are masked based on the valid next tokens $\sigma(q_t)$, setting the logits of invalid tokens to $-\infty$.
 The next token is then sampled from the categorical distribution defined by the unmasked logits, and the FSM transitions to the next state $q_{t+1} = \delta(q_t, v_{t+1})$, where $v_{t+1} \in \Sigma$ is the token in the regular expression alphabet corresponding to the sampled token.
@@
 
-We list here the prompts and schema (i.e., desired output formats) used for guided text generation via \package{Outlines} package interfacing with the \textsc{Mixtral-8x7B-Instruct} open-weights LLM.
+We list here the prompts and schema (i.e., desired output formats) used for guided text generation via \package{Outlines} package interfacing with the \textsc{Mixtral-8x7B-Instruct} open-weights large language model.
 
-The following schema is used to guide the generation of the summaries, intended to produce between one and five objects and hypotheses, as well as science use cases. \\
+The following schema, specified using the data-validation package \package{Pydantic}, is used to guide the generation of the summaries, intended to produce between one and five objects and hypotheses, as well as science use cases, given a raw proposal abstract.
+Both fields are of type \texttt{conlist}, a \package{Pydantic} type that represents a constrained list.
+\\
 
@@
 
-The following prompt function is used to produce a list of possible objects and phenomena shown in HST observations downstream of a proposal abstract, as well as one to five possible science use cases.
-\\
+The following prompt function is used to produce a list of one to five possible objects and phenomena shown in HST observations downstream of a proposal abstract, as well as one to five possible science use cases, in the format native to \package{Outlines}.
+\textcolor{deepgreen}{\lstinline{"<s>[INST]"}} and \textcolor{deepgreen}{\lstinline{"[/INST]"}} are start and end instruction delimiters, respectively, for the \textsc{Mixtral-8x7B} model.\\
 
@@
 
-\section{List of Categories for Text Retrieval Task}
-\label{app:categories}
+% \section{Evaluation of Model Trained on Raw Abstracts}
+% \label{app:eval_raw}
 
-The following curated categories are used in the text retrieval experiment in Sec.~\ref{sec:results}.
-%
-These are derived by initially prompting \textsc{Claude 2}, having attached a subsample of 30 proposal abstracts in the online interface to be used as context, to produce a list of categories corresponding to typical HST observations.
-The list is then manually curated to remove similar entries and ensure a representative sample of categories.
-\\
+% In the main text, we illustrated qualitative evaluation (image and text retrieval) for the model fine-tuned on summarized abstracts.
+Here, we show the same for the model fine-tuned on raw proposal abstracts.
 
-\begin{lstlisting}[language=Python]
-["star forming galaxies", "lyman alpha", "dust", "crowded stellar field", "core-collapse supernova", "cosmology", "gravitational lensing", "supernovae", "diffuse galaxies", "globular clusters", "stellar populations", "interstellar medium", "black holes", "dark matter", "galaxy clusters", "galaxy evolution", "galaxy formation", "quasars", "circumstellar disks", "exoplanets", "Kuiper Belt objects", "solar system objects", "cosmic web structure", "distant galaxies", "galaxy mergers", "galaxy interactions", "star formation", "stellar winds", "brown dwarfs", "white dwarfs", "nebulae", "star clusters", "galaxy archeology", "galactic structure", "active galactic nuclei", "gamma-ray bursts", "stellar nurseries", "intergalactic medium", "dark energy", "dwarf galaxies", "barred spiral galaxies", "irregular galaxies", "starburst galaxies", "low surface brightness galaxies", "ultra diffuse galaxies", "circumgalactic medium", "intracluster medium", "cosmic dust", "interstellar chemistry", "star formation histories", "initial mass function", "stellar proper motions", "binary star systems", "open clusters", "pre-main sequence stars", "protostars", "protoplanetary disks", "jets and outflows", "interstellar shocks", "planetary nebulae", "supernova remnants", "red giants", "Cepheid variables", "RR Lyrae variables", "stellar abundances", "stellar dynamics", "compact stellar remnants", "Einstein rings", "trans-Neptunian objects", "cosmic microwave background", "reionization epoch", "first stars", "first galaxies", "high-redshift quasars", "primordial black holes", "resolved binaries", "binary stars"]
-\end{lstlisting}
-
-The following prompt is used to generate the initial list before manual curation: \emph{``Here is a list of Hubble proposals.
-Base on this, please provide a list of about 100 strings, each describing a science target or use case for observations imaged by the Hubble Space Telescope.
-You may use these proposals and also rely on your general knowledge.
-For example, ["gravitational lensing", "supernovae", "diffuse galaxies", ...]''}
-
-% \section{Additional Evaluation Metrics and Ablations}
-% \label{app:ablations}
-
-\section{Evaluation of Model Trained on Raw Abstracts}
-\label{app:eval_raw}
-
-In the main text, we illustrated qualitative evaluation (image and text retrieval) for the model fine-tuned on summarized abstracts.
-Here, we show the same for the model fine-tuned on raw proposal abstracts.
-Table~\ref{tab:tti_abs} shows the top-4 most similar images for the abstract fine-tuned CLIP model on the same curated queries as in Tab.~\ref{tab:tti} for the summary fine-tuned model.
+% Table~\ref{tab:tti_abs} shows the Top-3 most similar images for the abstract fine-tuned CLIP model on the same curated queries as in Tab.~\ref{tab:tti} for the summary fine-tuned model.
 Table~\ref{tab:itt_abs} shows text associations from the curated list most closely matching the image queries, for the base and abstract fine-tuned models, as well as the summary fine-tuned model, for comparison.
@@
 
-\begin{table}[h!]
-  \centering
-  \begin{tabular}{m{2.7cm} p{2.9cm} p{2.9cm} p{2.9cm} p{2.9cm}}
-    \toprule
-      \centering \bfseries Query & \multicolumn{4}{c}{\bfseries{Top-4 most similar images using \textcolor{deepblue}{abstract fine-tuned CLIP model}}} \tabularnewline
-      \midrule
-      \texttt{\input{\thedatafolder/query_tti_abs_1.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_1_0.pdf} \\ \input{\thedatafolder/propid_tti_abs_1_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_1_1.pdf} \\ \input{\thedatafolder/propid_tti_abs_1_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_1_2.pdf} \\ \input{\thedatafolder/propid_tti_abs_1_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_1_3.pdf} \\ \input{\thedatafolder/propid_tti_abs_1_3.txt}  \tabularnewline
-      \midrule
-      \texttt{\input{\thedatafolder/query_tti_abs_0.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_0_0.pdf} \\ \input{\thedatafolder/propid_tti_abs_0_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_0_1.pdf} \\ \input{\thedatafolder/propid_tti_abs_0_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_0_2.pdf} \\ \input{\thedatafolder/propid_tti_abs_0_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_0_3.pdf} \\ \input{\thedatafolder/propid_tti_abs_0_3.txt}  \tabularnewline
-      \midrule
-      \texttt{\input{\thedatafolder/query_tti_abs_2.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_2_0.pdf} \\ \input{\thedatafolder/propid_tti_abs_2_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_2_1.pdf} \\ \input{\thedatafolder/propid_tti_abs_2_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_2_2.pdf} \\ \input{\thedatafolder/propid_tti_abs_2_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_2_3.pdf} \\ \input{\thedatafolder/propid_tti_abs_2_3.txt}  \tabularnewline
-      \midrule
-      \texttt{\input{\thedatafolder/query_tti_abs_3.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_3_0.pdf} \\ \input{\thedatafolder/propid_tti_abs_3_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_3_1.pdf} \\ \input{\thedatafolder/propid_tti_abs_3_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_3_2.pdf} \\ \input{\thedatafolder/propid_tti_abs_3_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_3_3.pdf} \\ \input{\thedatafolder/propid_tti_abs_3_3.txt}  \tabularnewline
-      \bottomrule
-  \end{tabular}
-  \caption{Same as Tabs.~\ref{tab:tti_base} and \ref{tab:tti}, but using the \textbf{\textcolor{deepblue}{abstract fine-tuned CLIP model}}.}
-  \label{tab:tti_abs}
-\end{table}
+% \begin{table}[h!]
+%   \centering
+%   \begin{tabular}{m{2.7cm} p{2.9cm} p{2.9cm} p{2.9cm}}
+%     \toprule
+%       \centering \bfseries Query & \multicolumn{3}{c}{\bfseries{Top-3 most similar images using \textcolor{deepblue}{abstract fine-tuned CLIP model}}} \tabularnewline
+%       \midrule
+%       % \texttt{\input{\thedatafolder/query_tti_abs_1.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_1_0.pdf} \\ \input{\thedatafolder/propid_tti_abs_1_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_1_1.pdf} \\ \input{\thedatafolder/propid_tti_abs_1_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_1_2.pdf} \\ \input{\thedatafolder/propid_tti_abs_1_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_1_3.pdf} \\ \input{\thedatafolder/propid_tti_abs_1_3.txt}  \tabularnewline
+%       % \midrule
+%       \texttt{\input{\thedatafolder/query_tti_abs_0.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_0_0.pdf} \\ \input{\thedatafolder/propid_tti_abs_0_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_0_1.pdf} \\ \input{\thedatafolder/propid_tti_abs_0_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_0_2.pdf} \\ \input{\thedatafolder/propid_tti_abs_0_2.txt}  \tabularnewline
+%       \midrule
+%       \texttt{\input{\thedatafolder/query_tti_abs_2.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_2_0.pdf} \\ \input{\thedatafolder/propid_tti_abs_2_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_2_1.pdf} \\ \input{\thedatafolder/propid_tti_abs_2_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_2_2.pdf} \\ \input{\thedatafolder/propid_tti_abs_2_2.txt}   \tabularnewline
+%       % \midrule
+%       % \texttt{\input{\thedatafolder/query_tti_abs_3.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_3_0.pdf} \\ \input{\thedatafolder/propid_tti_abs_3_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_3_1.pdf} \\ \input{\thedatafolder/propid_tti_abs_3_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_3_2.pdf} \\ \input{\thedatafolder/propid_tti_abs_3_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_abs_3_3.pdf} \\ \input{\thedatafolder/propid_tti_abs_3_3.txt}  \tabularnewline
+%       \bottomrule
+%   \end{tabular}
+%   \caption{Same as Tabs.~\ref{tab:tti_base} and \ref{tab:tti}, but using the \textbf{\textcolor{deepblue}{abstract fine-tuned CLIP model}}.}
+%   \label{tab:tti_abs}
+% \end{table}
 
-\begin{table}[t!]
-  \centering
-  \renewcommand{\arraystretch}{0.1}
-  \begin{tabular}{m{3cm} m{3.9cm} m{3.9cm} m{3.9cm}}
-      \toprule
-      \centering \bfseries \hubble image & \centering \textbf{Top-4 text} \\ \textbf{\textcolor{deeppurple}{(base)}} & \centering  \textbf{Top-4 text} \\ \textbf{\textcolor{deepblue}{(abstract fine-tuned)}} & \centering  \textbf{Top-4 text} \\ \textbf{\textcolor{deepred}{(summary fine-tuned)}} \tabularnewline
-      \midrule
-      \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_abs_0.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_0.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_abs_0.txt} &  {\scriptsize \verbatiminput{\thedatafolder/sci_itt_0.txt}} \tabularnewline
-      \midrule
-      \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_abs_1.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_1.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_abs_1.txt} &  {\scriptsize \verbatiminput{\thedatafolder/sci_itt_1.txt}} \tabularnewline
-      \midrule
-      \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_abs_2.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_2.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_abs_2.txt} &  {\scriptsize \verbatiminput{\thedatafolder/sci_itt_2.txt}} \tabularnewline
-      \midrule
-      \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_abs_3.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_3.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_abs_3.txt} &  {\scriptsize \verbatiminput{\thedatafolder/sci_itt_3.txt}} \tabularnewline
-      \bottomrule
-  \end{tabular}
-  \caption{Text associations from a curated list most closely matching four image queries (first column, the same as in Tab.~\ref{tab:itt}), for the \textcolor{deeppurple}{base} (\texttt{CLIP-ViT-B/16}), \textcolor{deepblue}{abstract fine-tuned}, and \textcolor{deepred}{summary fine-tuned} models.}
-  \label{tab:itt_abs}
-\end{table}
+% \begin{table}[t!]
+%   \centering
+%   \renewcommand{\arraystretch}{0.1}
+%   \begin{tabular}{m{3cm} m{3.9cm} m{3.9cm} m{3.9cm}}
+%       \toprule
+%       \centering \bfseries \hubble image & \centering \textbf{Top-3 text} \\ \textbf{\textcolor{deeppurple}{(base off-the-shelf)}} & \centering  \textbf{Top-3 text} \\ \textbf{\textcolor{deepblue}{(abstract fine-tuned)}} & \centering  \textbf{Top-3 text} \\ \textbf{\textcolor{deepred}{(summary fine-tuned)}} \tabularnewline
+%       \midrule
+%       \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_abs_0.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_0.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_abs_0.txt} &  {\scriptsize \verbatiminput{\thedatafolder/sci_itt_0.txt}} \tabularnewline
+%       \midrule
+%       \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_abs_1.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_1.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_abs_1.txt} &  {\scriptsize \verbatiminput{\thedatafolder/sci_itt_1.txt}} \tabularnewline
+%       \midrule
+%       \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_abs_2.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_2.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_abs_2.txt} &  {\scriptsize \verbatiminput{\thedatafolder/sci_itt_2.txt}} \tabularnewline
+%       \midrule
+%       \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_abs_3.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_3.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_abs_3.txt} &  {\scriptsize \verbatiminput{\thedatafolder/sci_itt_3.txt}} \tabularnewline
+%       \bottomrule
+%   \end{tabular}
+%   \caption{Text associations from a curated list most closely matching four image queries (first column, the same as in Tab.~\ref{tab:itt}), for the \textcolor{deeppurple}{base off-the-shelf} (\texttt{CLIP-ViT-B/16}), \textcolor{deepblue}{abstract fine-tuned}, and \textcolor{deepred}{summary fine-tuned} models.}
+%   \label{tab:itt_abs}
+% \end{table}
 
-\section{\changes{Additional Variations on Model and Training}}
+\section{Additional Model and Training Details}
+\label{app:model_details}
+
+We use the \texttt{CLIP-ViT-B/16} \citep{radford2021learning} variant as the base pre-trained CLIP model.
+%
+This model uses a 12-layer, 12-head, 768-embedding dimension vision transformer with patch size $16\times16$ as the image encoder \citep{dosovitskiy2020image} and a 12-layer, 8-head, 512-embedding dimension text sequence transformer as the text backbone \citep{vaswani2017attention}.
+%
+The text encoder has a maximum length of 77 tokens and the image encoder has a native resolution of $224\times224$ pixels.
+%
+Linear projection layers map the outputs of the image and text encoders to a common embedding space of dimension $n_\text{emb}=512$.
+%
+In total, the model contains $\sim 149$ million trainable parameters.
+%
+This model was originally pre-trained on $\sim 400$ million image-text pairs from internet data.
+%
+
+All models were trained over 20,000 steps with 2000 linear warmup steps 
+using the AdamW optimizer \citep{DBLP:conf/iclr/LoshchilovH19,DBLP:journals/corr/KingmaB14} with  %peak 
+learning rate $10^{-5}$ and weight decay $10^{-3}$.
+%
+Training takes approximately 3 hours on 4 Nvidia A100 GPUs.
+Models were instantiated using the \package{Transformers} \citep{wolf2019huggingface} library and trained using packages from the \package{Jax} \citep{jax2018github} ecosystem.
+%
+
+\section{Additional Variations on Model and Training}
 \label{app:ablations}
 
-Figure~\ref{fig:sim_app} shows the retrieval accuracy as defined in Eq.~\eqrefb{eq:retrieval_accuracy} as a function of the retrieval fraction for further variations of the model or training, evaluated and trained on summarized abstracts.
+Figure~\ref{fig:sim_app} shows the retrieval accuracy as a function of the retrieval fraction for further variations of the model or training, evaluated and trained on summarized abstracts.
 The red line corresponds to the model trained on summarized abstract described in the main text (fine-tuned on \texttt{CLIP-ViT-B/16} with constant learning rate $\mathrm{LR}=10^{-5}$ after linear warmup).
@@
 
+  \section{Text Retrieval Task}
+  \label{app:text_retrieval}
+
+  We can use images from the validation set as queries and retrieve the most relevant text chunks (e.g., objects and use cases) from a curated list.
+  %
+
+  The following curated categories are used in the text retrieval experiment in Sec.~\ref{sec:results}.
+%
+These are derived by initially prompting \textsc{Claude 2}\footnote{\url{https://claude.ai/}}, having attached a subsample of 30 proposal abstracts in the online interface to be used as context, to produce a list of categories corresponding to typical HST observations.
+The list is then manually curated to remove similar entries and ensure a representative sample of categories.
+\\
+
+\begin{lstlisting}[language=Python]
+["star forming galaxies", "lyman alpha", "dust", "crowded stellar field", "core-collapse supernova", "cosmology", "gravitational lensing", "supernovae", "diffuse galaxies", "globular clusters", "stellar populations", "interstellar medium", "black holes", "dark matter", "galaxy clusters", "galaxy evolution", "galaxy formation", "quasars", "circumstellar disks", "exoplanets", "Kuiper Belt objects", "solar system objects", "cosmic web structure", "distant galaxies", "galaxy mergers", "galaxy interactions", "star formation", "stellar winds", "brown dwarfs", "white dwarfs", "nebulae", "star clusters", "galaxy archeology", "galactic structure", "active galactic nuclei", "gamma-ray bursts", "stellar nurseries", "intergalactic medium", "dark energy", "dwarf galaxies", "barred spiral galaxies", "irregular galaxies", "starburst galaxies", "low surface brightness galaxies", "ultra diffuse galaxies", "circumgalactic medium", "intracluster medium", "cosmic dust", "interstellar chemistry", "star formation histories", "initial mass function", "stellar proper motions", "binary star systems", "open clusters", "pre-main sequence stars", "protostars", "protoplanetary disks", "jets and outflows", "interstellar shocks", "planetary nebulae", "supernova remnants", "red giants", "Cepheid variables", "RR Lyrae variables", "stellar abundances", "stellar dynamics", "compact stellar remnants", "Einstein rings", "trans-Neptunian objects", "cosmic microwave background", "reionization epoch", "first stars", "first galaxies", "high-redshift quasars", "primordial black holes", "resolved binaries", "binary stars"]
+\end{lstlisting}
+
+The following prompt is used to generate the initial list before manual curation: \emph{``Here is a list of Hubble proposals.
+Base on this, please provide a list of about 100 strings, each describing a science target or use case for observations imaged by the Hubble Space Telescope.
+You may use these proposals and also rely on your general knowledge.
+For example, ["gravitational lensing", "supernovae", "diffuse galaxies", ...]''}
+
+  We show the result of image-to-text retrieval in Tab.~\ref{tab:itt}, for the base (second column) as well as summary fine-tuned (third column) models, using four observations (left-most column) from the validation set.
+  %
+
+  The top four text associations are shown for each image query.
+  %
+  The `ground truth' summarized abstract is shown in the right column.
+  % 
+  The base as well as fine-tuned models are seen to return a mix of relevant and less-relevant associations, although showing different qualitative behavior.
+Purely qualitatively, the fine-tuned model is seen to consistently return more relevant associations compared to the base model.
+  %
+
+  The second row (an image of supernova 1987A) highlights an interesting pattern -- the base model erroneously attributes the object at the center of the image to a gravitational lens, while the fine-tuned model correctly identifies it as a supernova remnant.
+This kind of reasonable misattribution is common when querying the base model, and largely absent in the fine-tuned model.
+
+  \begin{table}[h!]
+   \centering
+   \renewcommand{\arraystretch}{0.1}
+   \begin{tabular}{m{3cm} m{3.0cm} m{3.0cm} m{3.2cm}}
+       \toprule
+       \centering \bfseries \hubble image & \centering \textbf{Top-4 text} \\ \small\textbf{\textcolor{deeppurple}{(base off-the-shelf)}} & \centering  \textbf{Top43 text} \\ \small\textbf{\textcolor{deepred}{(summary fine-tuned)}} & \centering \textbf{Summarized abstract} \\ (objects; `ground truth') \tabularnewline
+       \midrule
+       \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_0.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_0.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_0.txt} &  {\scriptsize \input{\thedatafolder/abs_itt_0.txt}} \tabularnewline
+       \midrule
+       \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_1.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_1.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_1.txt} &  {\scriptsize \input{\thedatafolder/abs_itt_1.txt}} \tabularnewline
+       \midrule
+       \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_2.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_2.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_2.txt} &  {\scriptsize \input{\thedatafolder/abs_itt_2.txt}} \tabularnewline
+       \midrule
+       \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_3.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_3.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_3.txt} &  {\scriptsize \input{\thedatafolder/abs_itt_3.txt}} \tabularnewline
+       \bottomrule
+   \end{tabular}
+   \caption{Text snippets from a curated list most closely matching a given image query (left-most column) by cosine similarity of respective embeddings, shown for the \textcolor{deeppurple}{base off-the-shelf} (\texttt{CLIP-ViT-B/16}) and \textcolor{deepred}{summary fine-tuned} models.
+The `ground truth' LLM-summarized abstract (only objects/phenomena) is shown in the right-most column.}
+   \label{tab:itt}
+ \end{table}
+
+%  \section{List of Categories for Text Retrieval Task}
+% \label{app:categories}
+
 \end{document}

### 2024-03-27 paper/colm_submission/colm2024_conference.tex
Tweaking COLM

@@
   %  %
-  In contrast with previous work, our application is the first to associate astronomical observation with the text modality in a generalist framework, showcasing the potential of language models in specialized scientific domains like astronomy.
+  In contrast with previous work, our application is the first to associate astronomical observation with the text modality in a task-agnostic manner, showcasing the potential of language models in specialized scientific domains like astronomy.
   %
@@
 The list is then manually curated to remove similar entries and ensure a representative sample of categories.
-\\
+\\ \\
 

### 2024-07-29 paper/colm_camera/colm2024_conference.tex
Camera-ready draft

--- 2024-03-27 paper/colm_submission/colm2024_conference.tex
@@
 \newcommand{\SM}[1]{\textcolor{blue}{[SM: #1]}}
-\newcommand{\changes}[1]{\textcolor{blue}{#1}}
+\newcommand{\changes}[1]{\textcolor{red}{#1}}
 
@@
 
-\author{Antiquus S.~Hippocampus, Natalia Cerebro \& Amelie P. Amygdale \thanks{ Use footnote for providing further information
-about author (webpage, alternative address)---\emph{not} for acknowledging
-funding agencies.  Funding acknowledgements go at the end of the paper.} \\
-Department of Computer Science\\
-Cranberry-Lemon University\\
-Pittsburgh, PA 15213, USA \\
-\texttt{\{hippo,brain,jen\}@cs.cranberry-lemon.edu} \\
-\And
-Ji Q. Ren \& Yevgeny LeNet \\
-Department of Computational Neuroscience \\
-University of the Witwatersrand \\
-Joburg, South Africa \\
-\texttt{\{robot,net\}@wits.ac.za} \\
-\AND
-Coauthor \\
-Affiliation \\
-Address \\
-\texttt{email}
+\author{Siddharth Mishra-Sharma, Yiding Song, \& Jesse Thaler \\
+MIT \& IAIFI \\
+\texttt{\{smsharma,ydsong,jthaler\}@mit.edu}
 }
+
+%\author{Siddharth Mishra-Sharma \\
+%The NSF AI Institute for Artificial Intelligence and Fundamental Interactions \\
+%Center for Theoretical Physics, Massachusetts Institute of Technology, Cambridge, MA 02139, USA \\
+%Department of Physics, Harvard University, Cambridge, MA 02138, USA \\
+%\texttt{\href{mailto:smsharma@mit.edu}{smsharma@mit.edu}}
+%\And
+%Yiding Song \\
+%The NSF AI Institute for Artificial Intelligence and Fundamental Interactions \\
+%Department of Physics, Massachusetts Institute of Technology, Cambridge, MA 02139, USA \\
+%\texttt{\href{mailto:ydsong@mit.edu}{ydsong@mit.edu}}
+%\And
+%Jesse Thaler \\
+%The NSF AI Institute for Artificial Intelligence and Fundamental Interactions \\
+%Center for Theoretical Physics, Massachusetts Institute of Technology, Cambridge, MA 02139, USA \\
+%\texttt{\href{mailto:jthaler@mit.edu}{jthaler@mit.edu}}
+%}
 
@@
 
-% \colmfinalcopy % Uncomment for camera-ready version, but NOT for submission.
+\colmfinalcopy % Uncomment for camera-ready version, but NOT for submission.
 \begin{document}
@@
 a method that connects, for the first time, astronomical image observations with natural language by leveraging the association between abstracts of successful observing proposals written by astronomers and images corresponding to downstream observations imaged by telescopes.
+%
+\changes{This approach demonstrates the potential of adapting generalist multi-modal foundation models to astronomy, complementing task-specific models in the domain by providing a flexible, language-based interface for interacting with observational data.}
 %
@@
 % Double check numbers in final version.
+   %
+   \changes{Due to practical limitations associated with the small size of the fine-tuning dataset, we did not use different datasets for validation and testing, deeming the current approach sufficient for a proof-of-principle exposition.}
 
@@
 
-   While the base model shows some signs of meaningful retrieval (e.g., the image of Jupiter in the first row of Tab.~\ref{tab:tti_base}), it is challenging to discern meaningful, strong associations between the retrieved images and corresponding query.
+   While the base model shows some signs of meaningful retrieval (e.g., the image of Jupiter in the first row of Tab.~\ref{tab:tti_base}), 
+  %  it is challenging to discern meaningful, strong associations between the retrieved images and corresponding query.
+   \changes{associations between the retrieved images and corresponding queries are not consistent.}
 
@@
 
+   \changes{We also evaluate the observation retrieval task more quantitatively.
+We design a prompt which lets us evaluate whether the abstract corresponding to a retrieved observation is relevant or not, with the output constrained to be a boolean using \package{Outlines}.
+We then evaluated this prompt, for the base as well as fine-tuned models, on the top 10 closest images by cosine similarity returned for 10 different queries.
+38\% and 77\% of the retrieved observations are deemed relevant when using the base and fine-tuned models, respectively.
+The fine-tuned model is thus significantly more likely to return images relevant to the query.
+The prompt and curated queries for this test are described in App.~\ref{app:quanteval_prompt}.}
+
    Note that we chose to illustrate qualitative performance on image retrieval using the model fine-tuned on summarized abstracts, rather than raw abstracts. 
@@
 % ; both models are seen to perform sensibly, even if differently, in this regard.
+   %
+   \changes{By diverse, we mean that retrieved observations may contain different types of objects or phenomena, which may be relevant to the query in distinct ways.}
 
@@
 
-This work relied on the \package{Astroquery} \citep{2019AJ....157...98G}, \package{BitsAndBytes} \citep{dettmers2022llmint8}, \package{Flax} \citep{flax2020github}, \package{Jax} \citep{jax2018github}, \package{Jupyter} \citep{Kluyver2016jupyter}, \package{Matplotlib} \citep{Hunter:2007}, \package{Numpy} \citep{harris2020array}, \package{Optax} \citep{deepmind2020jax}, \package{Outlines}, \package{Pandas} \citep{2020SciPy-NMeth}, \package{Pydantic}, \package{PyTorch} \citep{paszke2019pytorch}, \package{SciPy} \citep{2020SciPy-NMeth}, \package{Transformers} \citep{wolf2019huggingface}, and \package{Wandb} \citep{wandb} software packages.
+We thank Michael Brenner, Fran\c{c}ois Lanusse, and Julian Mu\~{n}oz for helpful conversations.
+%
+This work is supported by the National Science Foundation under Cooperative Agreement PHY-2019786 (The NSF AI Institute for Artificial Intelligence and Fundamental Interactions, \url{http://iaifi.org/}).
+%
+This material is based upon work supported by the U.S. Department of Energy, Office of Science, Office of High Energy Physics of U.S. Department of Energy under grant Contract Number  DE-SC0012567. 
+%
+YS was supported by the Research Science Institute (RSI) program at MIT.
+%
+This research was supported by an award from Google,  ``Interpretation of Multimodal Images from Astronomy''.
+%
+This research was supported by the Munich Institute for Astro-, Particle and BioPhysics (MIAPbP), which is funded by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) under Germany's Excellence Strategy – EXC-2094 – 390783311.
+%
+The computations in this paper were run on the FASRC Cannon cluster supported by the FAS Division of Science Research Computing Group at Harvard University.
 
@@
 
-\section{Ethics Statement}
+This work relied on the \package{Astroquery} \citep{2019AJ....157...98G}, \package{BitsAndBytes} \citep{dettmers2022llmint8}, \package{Flax} \citep{flax2020github}, \package{Jax} \citep{jax2018github}, \package{Jupyter} \citep{Kluyver2016jupyter}, \package{Matplotlib} \citep{Hunter:2007}, \package{Numpy} \citep{harris2020array}, \package{Optax} \citep{deepmind2020jax}, \package{Outlines}, \package{Pandas} \citep{2020SciPy-NMeth}, \package{Pydantic}, \package{PyTorch} \citep{paszke2019pytorch}, \package{SciPy} \citep{2020SciPy-NMeth}, \package{Transformers} \citep{wolf2019huggingface}, and \package{Wandb} \citep{wandb} software packages.
+
+ \subsubsection*{Reproducibility Statement}
+
+ Code used to reproduce the results in this work is available at \url{https://github.com/smsharma/PAPERCLIP-Hubble/tree/main}.
+
+\subsubsection*{Ethics Statement}
 This work relies on using abstracts from successful \hubble Space Telescope observing proposals as part of a dataset for training and evaluating machine learning models.
@@
 \end{lstlisting}
+
+\subsection{Prompt Used for Quantitative Evaluation of Observation Retrieval}
+\label{app:quanteval_prompt}
+
+\changes{The following prompt was used to evaluate the relevance of abstracts corresponding to retrieved images to a query, when quantitatively assessing the observation retrieval task.}
+
+\begin{lstlisting}[language=Python]
+import outlines
+
+@outlines.prompt
+def prompt_fn(abstract, query):
+     """[INST]
+You are an expert astrophysicist, with broad expertise across observational and theoretical astrophysics.
+
+Abstract: "{{abstract}}"
+Query: "{{query}}"
+
+The above is an abstract for a proposed observation taken by the Hubble Space Telescope (labeled "Abstract"), and an object or concept (labeled "Query").
+
+Could the observations corresponding to the abstract contain the query? Be precise, and do not contain related concepts or objects.
+
+Your response should be either True or False. Only return True if the query is closely related to the abstract, and the downstream observation could be relevant to the query.
+[/INST]
+"""
+\end{lstlisting}
+
+\changes{The queries used in the evaluation were \texttt{["globular cluster", "dwarf galaxy", "SN1987A", "strong lensing", "galaxy clusters", "interstellar medium", "dark matter", "spiral galaxies", "lyman alpha", "comets"].}}
 

### 2024-08-01 paper/colm_camera/colm2024_conference.tex
Typos

@@
 %
-\changes{This approach demonstrates the potential of adapting generalist multi-modal foundation models to astronomy, complementing task-specific models in the domain by providing a flexible, language-based interface for interacting with observational data.}
+\changes{This approach demonstrates the potential of adapting generalist multi-modal foundation models to astronomy, complementing task-specific models by providing a flexible, language-based interface for interacting with observational data.}
 %

### 2024-08-02 paper/colm_submission/colm2024_conference.tex
update pdf

@@
 
-This research is based on observations made with the NASA/ESA Hubble Space Telescope obtained from the Space Telescope Science Institute, which is operated by the Association of Universities for Research in Astronomy, Inc., under NASA contract NAS 5-26555.
-%
-Based on observations made with the NASA/ESA Hubble Space Telescope, and obtained from the Hubble Legacy Archive, which is a collaboration between the Space Telescope Science Institute (STScI/NASA), the Space Telescope European Coordinating Facility (ST-ECF/ESAC/ESA) and the Canadian Astronomy Data Centre (CADC/NRC/CSA).
+% This research is based on observations made with the NASA/ESA Hubble Space Telescope obtained from the Space Telescope Science Institute, which is operated by the Association of Universities for Research in Astronomy, Inc., under NASA contract NAS 5-26555.
+% %
+% Based on observations made with the NASA/ESA Hubble Space Telescope, and obtained from the Hubble Legacy Archive, which is a collaboration between the Space Telescope Science Institute (STScI/NASA), the Space Telescope European Coordinating Facility (ST-ECF/ESAC/ESA) and the Canadian Astronomy Data Centre (CADC/NRC/CSA).
 
-\section{Ethics Statement}
+% \subsubsection*{Reproducibility Statement}
+
+% Code used to reproduce the results in this work is available in the Supplementary Material associated with the paper.
+
+\subsubsection*{Ethics Statement}
 This work relies on using abstracts from successful \hubble Space Telescope observing proposals as part of a dataset for training and evaluating machine learning models.

### 2024-08-02 paper/hubble_paperclip.tex
update pdf

--- 2024-03-11 paper/hubble_paperclip.tex
@@
 \begin{abstract}
-We present PAPERCLIP (Proposal Abstracts Provide an Effective Representation for Contrastive Language-Image Pre-training), a method which associates astronomical observations imaged by surveys and telescopes with natural language  using a neural network model.
-The model is fine-tuned from a pre-trained Contrastive Language–Image Pre-training (CLIP) model using successful observing proposal abstracts, optionally summarized via guided generation using large language models (LLMs), and corresponding downstream observations.
-Using observations from the \hubble Space Telescope (HST) as an example, we show that the fine-tuned model embodies a meaningful joint representation between observations and text through tests targeting image retrieval (i.e., finding the most relevant observations using natural language queries) and description retrieval (i.e., querying for astrophysical object classes and science use cases most relevant to a given observation).
-Our study demonstrates the potential for using generalist \changes{foundation models} rather than task-specific models for \changes{interacting with} astronomical data by leveraging text as an interface.
+We present PAPERCLIP (Proposal Abstracts Provide an Effective Representation for Contrastive Language-Image Pre-training), a method which associates astronomical observations imaged by telescopes with natural language using a neural network model.
+The model is fine-tuned from a pre-trained Contrastive Language–Image Pre-training (CLIP) model using successful observing proposal abstracts and corresponding downstream observations, with the abstracts optionally summarized via guided generation using large language models (LLMs).
+Using observations from the \hubble Space Telescope (HST) as an example, we show that the fine-tuned model embodies a meaningful joint representation between observations and natural language through tests targeting image retrieval (i.e., finding the most relevant observations using natural language queries) and description retrieval (i.e., querying for astrophysical object classes and use cases most relevant to a given observation).
+Our study demonstrates the potential for using generalist foundation models rather than task-specific models for interacting with astronomical data by leveraging text as an interface.
 \githubmaster
@@
 %
-The {Foundation Model} paradigm, in contrast, seeks to develop generalist models which can be deployed to simultaneously tackle a wide range of tasks \citep{bommasani2021opportunities}.
+The \emph{foundation model} paradigm, in contrast, seeks to develop generalist models which can be deployed to simultaneously tackle a wide range of tasks \citep{bommasani2021opportunities}.
 %
@@
 %
-Foundation models can often benefit from additional training ({fine tuning}) using a relatively small amounts of domain-specific data in order to increase their usefulness when applied to specialized domains.
+Foundation models can often benefit from additional training ({fine-tuning}) using a relatively small amount of domain-specific data in order to increase their usefulness when applied to specialized domains.
 
-There is considerable interest in developing custom foundation models for the sciences \citep[e.g., ][]{batatia2023foundation,subramanian2023towards}, with astrophysics being ripe for such an effort given the large amounts of publicly-available data and diverse modes of interacting with it.
+There is considerable interest in developing custom foundation models for the sciences \citep[e.g., ][]{batatia2023foundation,subramanian2023towards,mccabe2023multiple,Birk:2024knn,vig2024finetuning,heinrich2024masked}, with astrophysics being ripe for such an effort given the large amounts of publicly-available data and diverse modes of interacting with it.
 %
@@
 %
-This multi-modality was recently exploited in \textsc{AstroCLIP}~\citep{lanusse2023astroclip} to construct a joint embedding space between multi-band images and optical spectra from the Dark Energy Spectroscopic Instrument (DESI).
+This multi-modality was recently exploited in \textsc{AstroCLIP}~\citep{lanusse2023astroclip} to construct a joint physically-informative embedding space between multi-band images and optical spectra from the Dark Energy Spectroscopic Instrument (DESI).
 %
@@
 Our method opens up the possibility of interacting with astronomical survey data using free-form natural language as an interface, which is a cornerstone of the success of the modern foundation model paradigm.
+A high-level overview of the method is shown in Fig.~\ref{fig:overview}.
 %
@@
 The rest of this paper is organized as follows.
+%
+We review related work in Sec.~\ref{sec:related}.
 %
@@
 \includegraphics[width=0.99\textwidth]{plots/figure.pdf}
-\caption{\changes{Overview of the PAPERCLIP method.
+\caption{Overview of the PAPERCLIP method.
 (Left) A pre-trained CLIP model is fine-tuned using a dataset of \hubble observations and corresponding proposal abstracts.
@@
 (Right) The fine-tuned model can then be used for downstream tasks such as observation retrieval (i.e., finding the observations most relevant to a given text query).
-The proposal abstract snippet shown here corresponds to proposal ID \href{https://archive.stsci.edu/proposal_search.php?id=16914&mission=hst}{16914}}.}
+The proposal abstract snippet shown here corresponds to proposal ID \href{https://archive.stsci.edu/proposal_search.php?id=16914&mission=hst}{16914}.}
 \label{fig:overview}
@@
 
-\section{\changes{Related work}}
+\section{Related Work}
+\label{sec:related}
 
-% The CLIP family of foundation models, which in their original form embed images and associated captions into a common representation space via contrastive learning, have shown strong performance and generalization capabilities on a variety of downstream tasks including zero-shot classification and image retrieval~\citep{radford2021learning}.
-%
 The concept of learning task-agnostic representations via self-supervised and contrastive learning has been applied within astrophysics \citep{slijepcevic2024radio,stein2021self,hayat2021self,slijepcevic2022learning} and used for downstream tasks like object similarity search \citep{stein2021self}, gravitational lens finding \citep{stein2022mining}, estimation of Galactic distances \citep{hayat2021estimating}, identification of rare galaxies \citep{walmsley2023rare}, and data compression \citep{akhmetzhanova2024data}.
@@
 
-% While the interaction of astrophysical data with large language models is fairly nascent, \textsc{AstroLLaMA}~\citep{nguyen2023astrollama,perkowski2024astrollama} is a recent effort to fine-tune a publicly-available model (\textsc{Llama-2}) on astrophysics-specific textual data from the arXiv.
-
-In this paper, we present for the first time an application associating target-agnostic astronomical data with the text modality, showing that this can be effectively accomplished through contrastive learning by leveraging observational proposals to inform text captions.
+In this paper, we present for the first time an application associating target-agnostic astronomical data with the text modality, showing that this can be effectively accomplished through contrastive learning by leveraging observing proposal abstracts to inform text captions.
 %
@@
 %
-It can be seen that the images in this dataset exhibit specific characteristics as well as artifacts particular to HST data-taking and processing which distinguishes them from the distribution of natural images typically used for large-scale pre-training of foundation models \citep{deng2009imagenet}.
+It can be seen that the images in this dataset exhibit specific characteristics as well as artifacts particular to HST data-taking and processing which distinguishes them from the distribution of natural images typically used for large-scale pre-training of foundation models.
 %
-This further motivates the need for fine tuning on domain-specific data.
+This further motivates the need for fine-tuning on domain-specific data.
 
@@
     \end{tabular}
-    \caption{\changes{For the \hubble proposal abstracts shown in Tab.~\ref{tab:dataset},} the LLM (\textsc{Mixtral-8x7B})-extracted summaries showing objects and phenomena (middle column) as well as potential downstream science use cases (last column) separately.}
+    \caption{For the \hubble proposal abstracts shown in Tab.~\ref{tab:dataset}, the LLM (\textsc{Mixtral-8x7B})-extracted summaries showing objects and phenomena (middle column) as well as potential downstream science use cases (last column) separately.
+The proposal IDs (left column) contain hyperlinks to the MAST page corresponding to the proposal.}
     \label{tab:datasetsumm}
@@
 %
-We note that these are not science-grade observations, but rather lower-resolution images useful for diagnostic purposes.
+We note that these are not science-grade observations, but rather lower-resolution images useful for diagnostic or preview purposes.
 %
@@
 3,194 images are held out for validation, with no abstract being common between training and validation sets in order to ensure an independent set of image-text pairs for evaluation.
-The held out images correspond to 438 unique abstracts.
+The held out images correspond to 429 unique abstracts.
 % Double check numbers in final version.
@@
 We choose to not filter out these pairs from our dataset, in order to have a larger sample of HST observations that the model can leverage to adapt to the distinctive characteristics of \hubble images.
-
-% Mention that the signal is noisy!
 
@@
 %
-By framing the problem in terms of transitions between a set of finite states (a finite-state machine), \cite{willard2023efficient} showed that guided generation can be performed with negligible overhead compared to unconstrained generation.
+By framing the problem in terms of transitions between a set of finite states (i.e., a finite-state machine), \cite{willard2023efficient} showed that guided generation can be performed with negligible overhead compared to unconstrained generation.
 %
-See App.~\ref{app:guided-generation} for a more detailed description of the guidance generation method used here.
+See App.~\ref{app:guided-generation} for a more detailed description of the guidance generation method used here, including an overview of technical details.
 
@@
 
-The guided generation process ensures that, in this case, the output of the LLM strictly conforms to the format of the following example:
+The guided generation process ensures that, in this case, the output of the generated output of the LLM strictly conforms to the format of the following example:\\
 \begin{center}
@@
           \color{black}\{
-            \color{codegreen}'objects_and_phenomena'\color{black}: [\color{codegreen}'star forming galaxy', 'lensed galaxy'\color{black}], 
-            \color{codegreen}'science_use_cases'\color{black}: [\color{codegreen}'measure lensing magnification'\color{black}]
+            \color{deepgreen}'objects_and_phenomena'\color{black}: [\color{deepgreen}'star forming galaxy', 'lensed galaxy'\color{black}, ...], 
+            \color{deepgreen}'science_use_cases'\color{black}: [\color{deepgreen}'measure lensing magnification'\color{black}, ...]
           \color{black}\}
@@
 %
-Examples of raw abstract snippets and corresponding LLM-generated summaries are shown in Tab.~\ref{tab:datasetsumm}.
+Examples of LLM-generated abstract summaries are shown in Tab.~\ref{tab:datasetsumm}, for the same set of abstracts as shown in Tab.~\ref{tab:dataset}.
 %
@@
 %
-With PAPERCLIP, we leverage the strong generalization capabilities demonstrated by pre-trained CLIP models and adapt these to work with domain-specific \hubble data via fine tuning.
+With PAPERCLIP, we leverage the strong generalization capabilities demonstrated by pre-trained CLIP models and adapt these to work with domain-specific \hubble data via fine-tuning.
 
@@
 %
-where ${x}_i={f\left(I_i\right)}/{\left\|f\left(I_i\right)\right\|}$ and ${y}_i={g\left(T_i\right)}/{\left\|g\left(T_i\right)\right\|}$ are the normalized representations of the $i$-th image and text, respectively, and $\tau$ is a learnable temperature hyperparameter.
+where ${x}_i={f\left(I_i\right)}/{\left\|f\left(I_i\right)\right\|}$ and ${y}_i={g\left(T_i\right)}/{\left\|g\left(T_i\right)\right\|}$ are the normalized representations of the $i$-th image and text caption, respectively, and $\tau$ is a learnable temperature hyperparameter.
 %
@@
 %
-This model uses a 12-layer, 12-head, 768-embedding dimension vision transformer as the image encoder and a 12-layer, 8-head, 512-embedding dimension text sequence transformer as the text backbone.
+This model uses a 12-layer, 12-head, 768-embedding dimension vision transformer with patch size $16\times16$ as the image encoder \citep{dosovitskiy2020image} and a 12-layer, 8-head, 512-embedding dimension text sequence transformer as the text backbone \citep{vaswani2017attention}.
 %
-The text encoder has a maximum length of 77 tokens and the image encoder a native resolution of $224\times224$ pixels.
+The text encoder has a maximum length of 77 tokens and the image encoder has a native resolution of $224\times224$ pixels.
 %
@@
 %
-This model was originally pre-trained on 400 million image-text pairs from internet data \citep{radford2021learning}.
+This model was originally pre-trained on $\sim 400$ million image-text pairs from internet data.
 %
@@
 All models were trained over 20,000 steps with 2000 linear warmup steps 
-% and cosine decay 
 using the AdamW optimizer \citep{DBLP:conf/iclr/LoshchilovH19,DBLP:journals/corr/KingmaB14} with  %peak 
@@
 %
-The retrieval accuracy is defined as the fraction of associated captions which fall within the top $k\%$ of captions by cosine similarity of the normalized image and caption embeddings, averaged over the images in the validation set:
+The retrieval accuracy is defined as the fraction of associated captions (either raw or LLM-summarized abstracts) which fall within the top $k\%$ of captions by cosine similarity of the normalized image and caption embeddings, averaged over the images in the validation set:
 \begin{equation}
@@
 %
-For the description/text retrieval evaluation, we define a list of possible text associations (i.e., classes), which we show in App.~\ref{app:categories}, by querying the \textsc{Claude 2}\footnote{\url{https://claude.ai/}} large language followed by with manual curation.
+For the description/text retrieval evaluation, we define a list of possible text associations (i.e., classes), which we show in App.~\ref{app:categories}, by querying the \textsc{Claude 2}\footnote{\url{https://claude.ai/}} large language followed by manual curation.
 
@@
 %
-Interestingly, the base pre-trained model performs better than random expectation, with a top-10\% retrieval accuracy of $\sim 15\%$.
+Interestingly, the base pre-trained model performs better than random expectation, with a top-10\% retrieval accuracy of $\sim 15\%$ (as see from the left-most datum in Fig.~\ref{eq:retrieval_accuracy} right, for the curves corresponding to fine-tuned models).
 %
@@
 %
-In this case, we evaluate all four models (fine tuned on raw abstracts (blue), fine-tuned on LLM-summarized abstracts (red), trained on LLM-summarized abstracts from scratch (yellow), and the base model (purple)) on the same captions dataset -- the summarized abstracts -- for a direct comparison.
+In this case, we evaluate all four models (fine-tuned on raw abstracts (blue), fine-tuned on LLM-summarized abstracts (red), trained on LLM-summarized abstracts from scratch (yellow), and the base model (purple)) on the same captions dataset -- the summarized abstracts -- for a direct comparison.
 %
 Remarkably, the model trained on raw abstracts shows very similar performance when evaluated on the summarized abstracts compared to that trained on the summarized abstracts themselves, indicating that \emph{(1)} the image-text association signal is preserved in the summarization process, and \emph{(2)} the model is able to effectively leverage meaningful concepts in the noisy raw abstracts through weak supervision.
-\changes{The significantly worse performance of the model trained from scratch, compared to the fine-tuned models, highlights the crucial role of the inductive bias inherent in the base pretrained model, which effectively captures the rich associations between images and language.}
+The significantly worse performance of the model trained from scratch, compared to the fine-tuned models, highlights the crucial role of the inductive bias inherited from the base pre-trained model, which effectively captures rich associations between images and language.
 
@@
 
-\changes{We show retrieval accuracy performance for additional variations on the model and training configuration in App.~\ref{app:ablations}.}
+We show retrieval accuracy performance for additional variations on the model and training configuration in App.~\ref{app:ablations}.
 
@@
 Having aligned the image and text representations, we can embed a natural language query using the model and show the closest images by embedding from the validation set when ranked by cosine similarity.
+A sketch of this procedure is shown in Fig.~\ref{fig:overview} (right).
 %
-We show these in Tabs.~\ref{tab:tti_base} and \ref{tab:tti} for the base and fine-tuned models respectively using four simple curated queries: \texttt{dwarf galaxy} (small galaxies that typically orbit larger galaxies like the Milky Way), \texttt{Jupiter},  \texttt{SN1987A} (a specific, prominent supernova), and \texttt{strong lensing} (the phenomenon of bending of light due to the gravitational influence of a foreground distribution of matter).
+We show these in Tabs.~\ref{tab:tti_base} and \ref{tab:tti} for the base and fine-tuned models respectively using four simple curated queries: \texttt{dwarf galaxy} (small galaxies that typically orbit larger galaxies like the Milky Way), \texttt{Jupiter},  \texttt{SN1987A} (a specific supernova), and \texttt{strong lensing} (the phenomenon of bending of light due to the gravitational influence of a foreground distribution of matter).
 The proposal ID corresponding to the retrieved images is shown below each image, and contains a hyperlink to the MAST page corresponding to the proposal for further details.
@@
 %
-Images of Jupiter are returned for the \texttt{Jupiter} query, with the planet clearly visible in all four images.
+Images looking like Jupiter are returned for the \texttt{Jupiter} query. 
+%
+However, this example also illustrates the model's potential to misidentify objects, with the first and third image actually showing Saturn with artifacts on the planet and partially obscured rings.
 %
@@
       \toprule
-      \centering \bfseries Query & \multicolumn{4}{c}{\bfseries{Top-4 most similar images using \textcolor{deeppurple}{base CLIP model}}} \tabularnewline
+      \centering \bfseries Query & \multicolumn{4}{c}{\bfseries{Top-4 most similar images using \textcolor{deeppurple}{base off-the-shelf CLIP model}}} \tabularnewline
       \midrule
@@
   \end{tabular}
-  \caption{For four text queries (left-most column), the four most similar images from the validation dataset by cosine similarity when using the \textbf{\textcolor{deeppurple}{base CLIP model}} (\texttt{CLIP-ViT-B/16}).
+  \caption{For four text queries (left-most column), the four most similar images from the validation dataset by cosine similarity when using the \textbf{\textcolor{deeppurple}{base (off-the-shelf) CLIP model}} (\texttt{CLIP-ViT-B/16}).
 The proposal ID associated with each image is given below the image and contains a hyperlink to the MAST page corresponding to the proposal.}
@@
       \toprule
-      \centering \bfseries \hubble image & \centering \textbf{Top-4 text} \\ \small\textbf{\textcolor{deeppurple}{(base)}} & \centering  \textbf{Top-4 text} \\ \small\textbf{\textcolor{deepred}{(summary fine-tuned)}} & \centering \textbf{Summarized abstract} \\ (objects; `ground truth') \tabularnewline
+      \centering \bfseries \hubble image & \centering \textbf{Top-4 text} \\ \small\textbf{\textcolor{deeppurple}{(base off-the-shelf)}} & \centering  \textbf{Top-4 text} \\ \small\textbf{\textcolor{deepred}{(summary fine-tuned)}} & \centering \textbf{Summarized abstract} \\ (objects; `ground truth') \tabularnewline
       \midrule
@@
   \end{tabular}
-  \caption{Text snippets from a curated list most closely matching a given image query (left-most column) by cosine similarity of respective embeddings, shown for the \textcolor{deeppurple}{base} (\texttt{CLIP-ViT-B/16}) and \textcolor{deepred}{summary fine-tuned} models.
+  \caption{Text snippets from a curated list most closely matching a given image query (left-most column) by cosine similarity of respective embeddings, shown for the \textcolor{deeppurple}{base off-the-shelf} (\texttt{CLIP-ViT-B/16}) and \textcolor{deepred}{summary fine-tuned} models.
 The `ground truth' LLM-summarized abstract (only objects/phenomena) is shown in the right-most column.}
@@
 
-In this paper, we present \textsc{PAPERCLIP}, a method for training domain-specific multi-modal models for astrophysics that associates observations imaged by telescopes with natural language in a common, semantically-meaningful embedding space.
+In this paper, we present \textsc{PAPERCLIP}, a method for training domain-specific multi-modal models for astrophysics that associates observations imaged by telescopes with natural language in a common embedding space.
 %
@@
 %
-Overall, the procedure demonstrates the efficacy of fine-tuning generalist pre-trained models on small amounts of domain-specific data, in particular astronomical datasets, and leveraging text as an interface.
+Overall, the procedure demonstrates the efficacy of fine-tuning generalist pre-trained models on small amounts of domain-specific data, in particular astronomical datasets, and leveraging text as an interface for interacting with the data.
 
@@
 %
-A model trained using weakly-supervised image-text pairs can be used to query survey data e.g., PHANGS~\citep{lee2022phangs}, COSMOS~\citep{scoville2007cosmic} using natural language, as well as to efficiently find patterns in such data that may not be apparent using specialized models or manual inspection.
+A model trained using weakly-supervised image-text pairs can be used to query large amounts of unlabeled survey data e.g., PHANGS~\citep{lee2022phangs}, COSMOS~\citep{scoville2007cosmic} for objects or use-cases of interest using natural language, as well as to efficiently find patterns in such data that may not be apparent using specialized models or manual inspection.
 %
@@
 While these abstracts are publicly available, the authors likely did not anticipate their text being used in this manner, raising questions around consent, attribution, and appropriate use of data.
-Since this research intends to develop methods to aid astronomical research and does not use sensitive personal information or target commercial gain, we believe that the scientific benefits outweigh the potential concerns in this case while acknowledging good-faith arguments to the contrary.
+Since this research intends to develop methods to aid astronomical research and does not use sensitive personal information or target commercial gain, we believe that the scientific benefits outweigh the potential concerns in this case, while acknowledging good-faith arguments to the contrary.
 As the use of foundation models in the sciences increases, it will be important for the community to consider norms and guidelines around the appropriate use and attribution of various data sources for model training and evaluation, including qualitative textual data, to ensure transparency and maintain trust.
@@
 
-We thank Michael Brenner for helpful conversations.
+We thank Michael Brenner, Fran\c{c}ois Lanusse, and Julian Mu\~{n}oz for helpful conversations.
 %
@@
 This research is based on observations made with the NASA/ESA Hubble Space Telescope obtained from the Space Telescope Science Institute, which is operated by the Association of Universities for Research in Astronomy, Inc., under NASA contract NAS 5-26555.
-
+%
 Based on observations made with the NASA/ESA Hubble Space Telescope, and obtained from the Hubble Legacy Archive, which is a collaboration between the Space Telescope Science Institute (STScI/NASA), the Space Telescope European Coordinating Facility (ST-ECF/ESAC/ESA) and the Canadian Astronomy Data Centre (CADC/NRC/CSA).
@@
 
-% \section{Additional Evaluation Metrics and Ablations}
-% \label{app:ablations}
-
 \section{Evaluation of Model Trained on Raw Abstracts}
@@
 Here, we show the same for the model fine-tuned on raw proposal abstracts.
+
 Table~\ref{tab:tti_abs} shows the top-4 most similar images for the abstract fine-tuned CLIP model on the same curated queries as in Tab.~\ref{tab:tti} for the summary fine-tuned model.
@@
       \toprule
-      \centering \bfseries \hubble image & \centering \textbf{Top-4 text} \\ \textbf{\textcolor{deeppurple}{(base)}} & \centering  \textbf{Top-4 text} \\ \textbf{\textcolor{deepblue}{(abstract fine-tuned)}} & \centering  \textbf{Top-4 text} \\ \textbf{\textcolor{deepred}{(summary fine-tuned)}} \tabularnewline
+      \centering \bfseries \hubble image & \centering \textbf{Top-4 text} \\ \textbf{\textcolor{deeppurple}{(base off-the-shelf)}} & \centering  \textbf{Top-4 text} \\ \textbf{\textcolor{deepblue}{(abstract fine-tuned)}} & \centering  \textbf{Top-4 text} \\ \textbf{\textcolor{deepred}{(summary fine-tuned)}} \tabularnewline
       \midrule
@@
   \end{tabular}
-  \caption{Text associations from a curated list most closely matching four image queries (first column, the same as in Tab.~\ref{tab:itt}), for the \textcolor{deeppurple}{base} (\texttt{CLIP-ViT-B/16}), \textcolor{deepblue}{abstract fine-tuned}, and \textcolor{deepred}{summary fine-tuned} models.}
+  \caption{Text associations from a curated list most closely matching four image queries (first column, the same as in Tab.~\ref{tab:itt}), for the \textcolor{deeppurple}{base off-the-shelf} (\texttt{CLIP-ViT-B/16}), \textcolor{deepblue}{abstract fine-tuned}, and \textcolor{deepred}{summary fine-tuned} models.}
   \label{tab:itt_abs}
@@
 
-\section{\changes{Additional Variations on Model and Training}}
+\section{Additional Variations on Model and Training}
 \label{app:ablations}

### 2024-08-02 models/losses.py

@@ -1,0 +1,104 @@
+import jax
+import jax.numpy as np
+
+# Enable float64 for numerical stability
+jax.config.update("jax_enable_x64", True)
+
+def mini_batch_sigmoid_loss(text_embeds, image_embeds, logit_scale, logit_bias, negative_samples):
+    """Positive samples are on the diagonal"""
+    bs = text_embeds.shape[0]
+    if negative_samples:  # All negative samples 
+        labels = -np.ones((bs, bs))
+    else:  # Positive samples are on the diagonal
+        labels = 2 * np.eye(bs) - np.ones((bs, bs))
+    logits = np.matmul(text_embeds, image_embeds.T) * logit_scale + logit_bias
+    
+    # Increase precision for large batches
+    logits = logits.astype(np.float64)
+    return -np.mean(jax.nn.log_sigmoid(labels * logits))
+
+def sigmoid_loss(outputs):
+    """ SigLIP loss (https://arxiv.org/abs/2303.15343); 
+        Adapted from https://github.com/borisdayma/clip-jax/blob/main/training/train.py
+    """
+
+    # Get outputs
+    ztxt = outputs["text_embeds"]
+    zimg = outputs["image_embeds"]
+    logit_scale = outputs["logit_scale"]  # Same as temperature
+    logit_bias = outputs.get("logit_bias", 0.)
+
+    # Number of chunks (devices)
+    axis_size = jax.lax.psum(1, axis_name="batch")
+
+    # Calculate local device loss
+    loss = mini_batch_sigmoid_loss(ztxt, zimg, logit_scale, logit_bias, negative_samples=False)
+
+    # Add negative losses
+    def add_negative_loss(i, carrys):
+        cumul_loss, zimg = carrys
+        
+        # Shift image_embeds
+        zimg = jax.lax.ppermute(
+            zimg, axis_name="batch", perm=[(j, (j - 1) % axis_size) for j in range(axis_size)]
+        )
+        # Add loss (all negative samples)
+        cumul_loss += mini_batch_sigmoid_loss(
+            ztxt, zimg, logit_scale, logit_bias, negative_samples=True
+        )
+        
+        return cumul_loss, zimg
+
+    loss, _ = jax.lax.fori_loop(0, axis_size - 1, add_negative_loss, (loss, zimg))
+    loss = loss / axis_size
+
+    loss = loss.reshape((-1,))
+
+    # Average loss across devices
+    loss = np.mean(loss)
+    return loss
+
[…]

### 2024-08-02 scripts/submit_summarize.sh

@@ -1,0 +1,29 @@
+#!/bin/bash
+
+#SBATCH --job-name=summarize_mixtral
+#SBATCH --nodes=1
+#SBATCH --mem=80GB
+#SBATCH --time=48:00:00
+#SBATCH --gres=gpu:4
+#SBATCH --account=iaifi_lab
+#SBATCH -p iaifi_gpu_priority
+
+export TF_CPP_MIN_LOG_LEVEL="2"
+
+# Load modules
+module load python/3.10.9-fasrc01
+module load cuda/12.2.0-fasrc01
+module load gcc/12.2.0-fasrc01
+module load openmpi/4.1.4-fasrc01
+
+export ENV=outlines
+
+# Activate env
+mamba activate $ENV
+
+alias pip=/n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/pip
+alias python=/n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python
+alias jupyter=/n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/jupyter
+
+cd /n/holystore01/LABS/iaifi_lab/Users/smsharma/multimodal-data/
+/n/holystore01/LABS/iaifi_lab/Users/smsharma/envs/$ENV/bin/python summarize.py --data_dir /n/holystore01/LABS/iaifi_lab/Users/smsharma/multimodal-data/data --observations_dir observations_v1 --save_filename summary_v2 --verbose

### 2024-08-02 summarize.py

@@ -1,0 +1,155 @@
+import sys, os
+
+sys.path.append("../")
+
+import outlines
+import outlines.models as models
+import outlines.text as text
+
+import argparse
+from functools import partial
+from tqdm import tqdm
+import pandas as pd
+
+import transformers
+from transformers import BitsAndBytesConfig
+import torch
+from utils.summarize_utils import prompt_fn, ConstrainedResponseHST
+from utils.abstract_utils import read_abstracts_file
+
+
+def summarize(
+    data_dir: str = "./data/",
+    observations_dir: str = "observations_v1",
+    i_start: int = 0,
+    n_max_abstracts: int = 9999999,
+    abstract_filename: str = "abstracts.cat",
+    model_name: str = "mistralai/Mixtral-8x7B-Instruct-v0.1",  # "TheBloke/OpenHermes-2.5-Mistral-7B-AWQ",
+    save_filename: str = "summary_v1",
+    seed: int = 42,
+    verbose: bool = False,
+    n_max_tries: int = 5,
+):
+    if model_name == "mistralai/Mixtral-8x7B-Instruct-v0.1":
+        config = transformers.AutoConfig.from_pretrained(
+            model_name,
+            trust_remote_code=True,
+            asd=True,
+        )
+
+        bnb_config = BitsAndBytesConfig(load_in_4bit=True, bnb_4bit_use_double_quant=True, bnb_4bit_quant_type="nf4", bnb_4bit_compute_dtype=torch.bfloat16)
+
+        model = models.transformers(
+            model_name=model_name,
+            model_kwargs={"config": config, "quantization_config": bnb_config, "trust_remote_code": True, "device_map": "auto", "load_in_4bit": True, "cache_dir": "/n/holystore01/LABS/iaifi_lab/Users/smsharma/hf_cache/"},
+        )
+
+    elif model_name == "TheBloke/OpenHermes-2.5-Mistral-7B-AWQ":
+        model = models.awq(model_name)
+    else:
+        raise ValueError("Invalid model name.")
+
+    # Load the abstracts
+    abstract_file = f"{data_dir}/{abstract_filename}"
+    abstracts_df = read_abstracts_file(abstract_file)
+
+    # Drop rows with missing Cycle
+    abstracts_df = abstracts_df.dropna(subset=["Cycle"])
+    abstracts_df = abstracts_df[abstracts_df["Cycle"] != ""]
+
+    # Convert Cycle and ID to int
[…]

### 2024-08-02 utils/download_data.py

@@ -1,0 +1,116 @@
+import os
+import numpy as np
+from astroquery.mast import Observations
+from PIL import Image
+
+
+def downsample_image(image_path, max_size):
+    """
+    Downsample an image to have a maximum side length of max_size.
+
+    Parameters:
+    - image_path: Path to the image to be downsampled.
+    - max_size: The maximum side length for the output image.
+    """
+    with Image.open(image_path) as img:
+        img_ratio = img.width / img.height
+        if img.width > img.height:
+            new_width = max_size
+            new_height = int(max_size / img_ratio)
+        else:
+            new_height = max_size
+            new_width = int(max_size * img_ratio)
+
+        img = img.resize((new_width, new_height), Image.LANCZOS)
+        img.save(image_path)
+
+
+def download_data(proposal_id, n_max_images, max_resolution, seed, data_dir="../data", exclude_color=False):
+    """Download data from MAST.
+
+    Args:
+        proposal_id (int): Hubble proposal ID.
+        n_max_images (int): Maximum number of images per proposal (selected randomly).
+        max_resolution (int): Maximum resolution for the downloaded images.
+        seed (int): RNG seed for taking a random sample of images up to n_max_images.
+        data_dir (str, optional): Directory where data will be downloaded to. Defaults to "../data".
+        exclude_color (bool, optional): Whether to exclude color images. Defaults to False.
+    """
+    download_dir = f"{data_dir}/proposal_{proposal_id}"
+
+    # Query MAST for observations
+    obs_table = Observations.query_criteria(
+        obs_collection="HST",
+        proposal_id=[f"{proposal_id}"],
+        dataRights="PUBLIC",
+    )
+
+    # If no observations found, return
+    if len(obs_table) == 0:
+        print(f"No observations found for proposal {proposal_id}")
+        return
+
+    # Get preview products
+    products = Observations.get_product_list(obs_table)
+    products = products[products["productType"] == "PREVIEW"]
+
+    # Only total preview image
+    match = "total"
+
+    # Potentially exclude images with "color" in the filename
+    if exclude_color:
+        exclude = "color"
+    else:
+        exclude = "dont_exclude_anything"  # Dummy
+
+    # Filter for total preview images and potentially exclude color
+    match_mask = [match in row["productFilename"] for row in products]
+    exclude_mask = [exclude not in row["productFilename"] for row in products]
+
+    mask = [m and e for m, e in zip(match_mask, exclude_mask)]
+
+    # Check if there are any images
+    if not np.any(mask):
[…]
